### 2014-11-12 form.tex
Expiring initial spec

\documentclass{article}
\usepackage{dsfont}
\usepackage[margin=1.5cm]{geometry}
\renewcommand{\o}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\p}[1]{\ensuremath{\mathit{#1}}}
\newcommand{\s}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\nintset}{\ensuremath{\mathds{Z}_{\ge 0}}}
\newcommand{\moreset}[1]{\ensuremath{\mathds{Z}_{\ge #1}}}

\begin{document}

An ``expiring map'' $m \in M[K,V]$ is an abstraction that maps keys $k \in K$ to values $v \in V$ and allows individual mappings to expire at an indicated time.

\paragraph{Abstract state $S(m\in M)$}
\begin{itemize}
   \item $m$.\s{mapping}: a set of entries $[k,v,t]$, with each entry mapping key $k$ to value $v$ until time $t$, after which the mapping expires
   \item $m$.\s{size}: the number of keys in m.\s{mapping} (for now, we assume at most one entry per key)
   \item $m$.\s{maxSize}: the maximum number of keys that this map $m$ can accommodate
\end{itemize}

\paragraph{Abstract operations}
\begin{itemize}
   \item $m \gets \o{constructor}(s \in \nintset$) constructs a map $m\in M$ that holds at most $s$ entries
   \item $p\gets \o{has}(m \in M, k \in K)$ indicates whether key $k$ exists in $m$ or not
   \item $(v\in V, e\in \nintset)\gets \o{get}(m \in M, k \in K)$ returns the pair of value $v$ to which $k$ is mapped and time $e$ until this mapping is kept in $m$. If the key $k$ is not present, can return $fail$.
   \item $x\gets \o{put}(m \in M, k \in K, v \in V, e \in \nintset)$ adds to $m$ a mapping from $k$ to $v$ that is valid until time $e$. Returns boolean $x$ - whether the operation was successfull.
   \item $x\gets \o{erase}(m\in M, k \in K)$ removes the entry from $m$ that maps $k$ to a value. Returns boolean $x$ - whether the operation was successfull.
   %\item $\o{nop}(m\in M)$ does nothing, returns nothing.
\end{itemize}

Consider a map $m\in M$.  We model the use of $m$ as a finite sequence 
$\tau = <o_0, o_1, ..., o_n>$
of operations $o_t$ on $m$ indexed by time $t \in \nintset$. 
In our model, an operation completes fully before the next one commences (i.e., there is no concurrency).
%A continuous subsequence $<o_a, o_{a+1}, ..., o_{b-1}, o_b> (0\le a\le b\le n)$ of $\tau$ is designated as $\tau_a^b$
For simplicity, we assume that each operation takes exactly one unit of time.
We use the shorthand $\o{operation}_t(...)$ to indicate the operation that occurs at time $t$.

An execution is an interleaving of operations and ``snapshots'' of the abstract state
$\epsilon = <o_0, S_0, o_1, S_1, ..., o_n, S_n>$, with
each subsequence $<S_{i-1}, o_i, S_i>$ indicating that operation $o_i$ transforms abstract state $S_{i-1}$ into $S_i$.
An accessor does not mutate state, so it leaves the abstract state unchanged.
We use the shorthand $\s{component}_t$ to indicate the value of $S.$\s{component} immediately after $o_t$ completes and before $o_{t+1}$ commences.

\paragraph{Helper predicates}
\begin{itemize}
   \item An ``empty'' map is one without any keys \\
            $\p{empty}_t(m\in M) \Rightarrow (\forall k\in K \Rightarrow \neg \o{has}_t(m, k))$ \\
            (it can be easily shown that $\p{empty}_t(m) \Leftrightarrow m.\s{size}_t=0$)
   \item A ``const'' operation does not modify nonexpired mappings:\\
     $\p{const}_t(m\in M) ~\Leftrightarrow~\\ \Big(\forall k\in K,v \in V, e\in \nintset : \o{has}_{t-1}(m, k), (v,e)=\o{get}_{t-1}(m, k), e>t \Rightarrow \o{has}_t(m, k),\o{get}_t(m, k)=(v,e)\Big)\wedge\\ \Big(\forall k\in K: \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$
   \item An ``isolated'' operation does not modify nonexpired mappings, except for a single one:\\
     $\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~\\ \Big(\forall k\in K,v \in V, e\in \nintset : k\neq k', \o{has}_{t-1}(m, k), (v,e)=\o{get}_{t-1}(m, k), e>t \Rightarrow \o{has}_t(m, k),\o{get}_t(m, k)=(v,e)\Big)\wedge\\ \Big(\forall k\in K:k\neq k', \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$
   \item ...

\end{itemize}

\paragraph{Semantics for accessors (may mutate $S$, do not change the nonexpired mappings)}
\begin{itemize}
   \item[has]
     \begin{itemize}
       \item $\o{has}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t$\\
       \item If the map is empty, there no keys in it:\\
         $$m.\s{size}_t=0 \Rightarrow \forall k\in K \o{has}_t(m, k) = false $$
       \item It does not change nonexpired mappings:\\
         $$\o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
     \end{itemize}
     \item[get]
       \begin{itemize}
         \item One can get a mapping only if it exists in the map:\\
           $$\o{get}_t(m \in M, k \in K) \rightarrow fail ~\Leftrightarrow~ \neg\o{has}_t(m \in M, k \in K)$$
           $$(v\in V,e\in \nintset)\gets \o{get}_t(m \in M, k \in K) ~\Leftrightarrow~ \o{has}_t(m \in M, k \in K) \wedge \exists [k,v,e] \in m.\s{mapping}_t$$
         \item In anyway, it does not change existing unexpired mappings:\\
           $$\o{get}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
       \end{itemize}
\end{itemize}

\paragraph{Semantics for operations that mutate state}
\begin{itemize}
  \item[constructor]
    \begin{itemize}
      \item The constructor forms a new empty map \\
        $m=\o{constructor}_0(s \in \nintset) \rightarrow \p{empty}_0(m) \wedge m.\s{maxSize}_t=s$
      \item The constructor must be called once and only once, at the very beginning of $\tau$ (i.e., $\tau$ contains no other constructor call after the first one) \\
        $o_0 \equiv \o{constructor}_0 \wedge (\forall t>0 \Rightarrow o_t \neq \o{constructor}_t)$ 
    \end{itemize}
  \item[erase]
    \begin{itemize}
      \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \p{mapping}, it fails. It does not change other mappings if they are not expired.
        $$\Big(x=\o{erase}_t(m\in M, k\in K) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
      \item If there were such a mapping for $k$, ``erase'' will remove it.
        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
      \item If there were not --- it will fail. 
        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
    \end{itemize}
    \item[put]
      \begin{itemize}
        \item ``Put'' inserts a new mapping, if the number of unexpired elements is less than \s{maxSize} and there is no unexpired entry with the same key.
          This function does not touch any unexpired entry with key other than $k$.
          $$\Big(x=\o{put}_t(m\in M, k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
        \item If there were not such mupping before, and there is enough room, it will succeed.
          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
        \item If there is no room, but some mappings have expired, one still can put.
          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
        \item If there is an unexpired mapping with the same key, put will fail.
          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge (v, e)=\o{get}_{t-1}(m, k) \wedge e \ge t \Leftrightarrow$$
          $$x=fail \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge \o{has}_t(m, k) \wedge \o{get}_t(m, k) = (v, e)\Big)\wedge$$
        \item Into full map, one can not put any more keys.
          $$\Big(x=\o{put}_t(m, k, v, e) \wedge$$
          $$\big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(m, k')\wedge(\cdot,e')=\o{get}_{t-1}(m,k') \Rightarrow e' \ge t \big) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \Leftrightarrow$$
          $$x=fail \wedge m.\s{size}_t = m.\s{size}_{t-1} \wedge \p{const}_t(m, k)\Big)\wedge$$
      \end{itemize}

\end{itemize}

\paragraph{Properties to be proven}
\begin{itemize}
 \item The size of the map is always properly bounded 
   $$\forall t\in \nintset: t>0  \wedge m=\o{constructor}_0(s \in \nintset) \Rightarrow 0 \le m.\s{size}_t \le m.\s{maxSize}$$
 \item No operation except for $\o{constructor}_0$ can change map capacity:
   $$\forall t\in \nintset: t > 0 \Rightarrow m.\s{maxSize}_t=m.\s{maxSize}_0$$
 \item For any subsequence of operations other than \o{put}, size of the map may only decrease:
   $$\forall a\in \nintset,b\in \nintset: a<b \wedge \big(\forall t\in \nintset:a<t<b \Rightarrow o_t \neq \o{put}_t \big) \Rightarrow \s{size}_a \ge \s{size}_b$$
 \item If a subsequence of operations does not include \o{erase} for key $k\in K$, and there is a mapping, it will be kept until the end of the subsequence, if not expired:
   $$\forall a\in \nintset,b, k\in K, m\in M, v\in V, e\in \moreset{b}: a < b \wedge \o{has}_a(m, k) \wedge (v, e)=\o{get}_a{m, k} \wedge e \ge b ~ \wedge$$
   $$\big(\forall t\in \nintset: a<t<b \Rightarrow o_t\neq \o{erase}_t(m, k)\big) \Rightarrow$$
   $$\o{has}_b(m, k) \wedge \o{get}_b(m, k) = (v,e)$$
\end{itemize}

\end{document}

### 2014-11-12 form.tex
Rewriting in terms of abstract state. pure API predicates should go to theorems

@@
    \item $m$.\s{mapping}: a set of entries $[k,v,t]$, with each entry mapping key $k$ to value $v$ until time $t$, after which the mapping expires
-   \item $m$.\s{size}: the number of keys in m.\s{mapping} (for now, we assume at most one entry per key)
+   \item $m$.\s{size}: the number of unexpired keys in $m$.\s{mapping} (for now, we assume at most one entry per key)
    \item $m$.\s{maxSize}: the maximum number of keys that this map $m$ can accommodate
@@
 
+{\it Note:} Symbol $\simeq$ in the following context: $o_t\simeq ...$ means that lefthand {\it operation} ($o_t$) can be equal to the righthand operation.
+E.g.
+$\forall k \in K, m \in M o_t\simeq \o{has}_t(m, k) \Leftrightarrow \p{const}_t(m)$ means, that the operation number t ($o_t$) may be $\o{has}(m, k)$ (not the result of such function call) if and only if holds $\p{const}_t(m)$.
+
+By default, if symbols $m, e, k, v$ are used in a formula, and not mentioned before in that formula, the universal quantifier $\forall$ is implied: $\forall m\in M, k \in K, v\in V, e\in \nintset$.
+
 \paragraph{Helper predicates}
@@
    \item An ``empty'' map is one without any keys \\
-            $\p{empty}_t(m\in M) \Rightarrow (\forall k\in K \Rightarrow \neg \o{has}_t(m, k))$ \\
-            (it can be easily shown that $\p{empty}_t(m) \Leftrightarrow m.\s{size}_t=0$)
+            $\p{empty}_t(m\in M) \Leftrightarrow [k, v, e] \not\in m.\s{mapping}_t$
    \item A ``const'' operation does not modify nonexpired mappings:\\
-     $\p{const}_t(m\in M) ~\Leftrightarrow~\\ \Big(\forall k\in K,v \in V, e\in \nintset : \o{has}_{t-1}(m, k), (v,e)=\o{get}_{t-1}(m, k), e>t \Rightarrow \o{has}_t(m, k),\o{get}_t(m, k)=(v,e)\Big)\wedge\\ \Big(\forall k\in K: \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$
+     $\p{const}_t(m\in M) ~\Leftrightarrow~\\ \Big([k, v, e] \in m.\s{mapping}_{t-1} \wedge e>t \Rightarrow [k, v, e] \in m.\s{mapping}_t\Big)\wedge\\ \Big([k, v, e] \not\in m.\s{mapping}_{t-1} \Rightarrow [k, v, e] \not\in m.\s{mapping}_t\Big)\wedge m.\s{size}_{t-1} \ge m.\s{size}_t$
    \item An ``isolated'' operation does not modify nonexpired mappings, except for a single one:\\
-     $\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~\\ \Big(\forall k\in K,v \in V, e\in \nintset : k\neq k', \o{has}_{t-1}(m, k), (v,e)=\o{get}_{t-1}(m, k), e>t \Rightarrow \o{has}_t(m, k),\o{get}_t(m, k)=(v,e)\Big)\wedge\\ \Big(\forall k\in K:k\neq k', \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$
-   \item ...
+     $\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~\\ \Big(k\neq k' \wedge [k, v, e] \in m.\s{mapping}_{t-1} \wedge e>t \Rightarrow [k, v, e] \in m.\s{mapping}_t\Big)\wedge\\ \Big(k\neq k' \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \Rightarrow [k, v, e] \not\in m.\s{mapping}_t\Big)$
 
@@
      \begin{itemize}
-       \item $\o{has}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t$\\
-       \item If the map is empty, there no keys in it:\\
-         $$m.\s{size}_t=0 \Rightarrow \forall k\in K \o{has}_t(m, k) = false $$
-       \item It does not change nonexpired mappings:\\
-         $$\o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
+       \item $\o{has}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t \wedge o_t=\o{has}_t(m, k)$\\
+       \item It does not change nonexpired mappings:
+         $$o_t \simeq \o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
      \end{itemize}
@@
        \begin{itemize}
-         \item One can get a mapping only if it exists in the map:\\
-           $$\o{get}_t(m \in M, k \in K) \rightarrow fail ~\Leftrightarrow~ \neg\o{has}_t(m \in M, k \in K)$$
-           $$(v\in V,e\in \nintset)\gets \o{get}_t(m \in M, k \in K) ~\Leftrightarrow~ \o{has}_t(m \in M, k \in K) \wedge \exists [k,v,e] \in m.\s{mapping}_t$$
-         \item In anyway, it does not change existing unexpired mappings:\\
-           $$\o{get}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
+         \item One can get a mapping only if it exists in the map:
+           $$\o{get}_t(m \in M, k \in K) \rightarrow fail ~\Leftrightarrow~ [k,v,e] \not\in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k)$$
+           $$(v\in V,e\in \nintset)\gets \o{get}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k)$$
+         \item In anyway, it does not change existing unexpired mappings:
+           $$o_t \simeq \o{get}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
        \end{itemize}
@@
       \item The constructor forms a new empty map \\
-        $m=\o{constructor}_0(s \in \nintset) \rightarrow \p{empty}_0(m) \wedge m.\s{maxSize}_t=s$
+        $m \gets \o{constructor}_0(s \in \nintset) \Leftrightarrow \p{empty}_0(m) \wedge \forall t \in \nintset m.\s{maxSize}_t=s$
       \item The constructor must be called once and only once, at the very beginning of $\tau$ (i.e., $\tau$ contains no other constructor call after the first one) \\
-        $o_0 \equiv \o{constructor}_0 \wedge (\forall t>0 \Rightarrow o_t \neq \o{constructor}_t)$ 
+        $o_0 \equiv \o{constructor}_0 \wedge (\forall t\in \moreset{1} \Rightarrow o_t \neq \o{constructor}_t)$ 
     \end{itemize}
@@
       \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \p{mapping}, it fails. It does not change other mappings if they are not expired.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
-      \item If there were such a mapping for $k$, ``erase'' will remove it.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
-      \item If there were not --- it will fail. 
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
+        $$\Big(o_t \simeq \o{erase}_t(m\in M, k\in K) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
+      \item If there was a such mapping for $k$, ``erase'' will remove it.
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \exists [k, v, e] \in m.\s{mapping}_{t-1} ~\Leftrightarrow~$$
+        $$x=success \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_t\big) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
+      \item If there was not --- it will fail. 
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge [k,v,e]\not\in m.\s{mapping}_{t-1} ~\Leftrightarrow~$$
+        $$x=fail \wedge \p{const}_t(m)\Big)$$
     \end{itemize}
@@
           This function does not touch any unexpired entry with key other than $k$.
-          $$\Big(x=\o{put}_t(m\in M, k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
-        \item If there were not such mupping before, and there is enough room, it will succeed.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
-          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
-        \item If there is no room, but some mappings have expired, one still can put.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
-          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
-          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
-        \item If there is an unexpired mapping with the same key, put will fail.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge (v, e)=\o{get}_{t-1}(m, k) \wedge e \ge t \Leftrightarrow$$
-          $$x=fail \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge \o{has}_t(m, k) \wedge \o{get}_t(m, k) = (v, e)\Big)\wedge$$
-        \item Into full map, one can not put any more keys.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge$$
-          $$\big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(m, k')\wedge(\cdot,e')=\o{get}_{t-1}(m,k') \Rightarrow e' \ge t \big) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \Leftrightarrow$$
-          $$x=fail \wedge m.\s{size}_t = m.\s{size}_{t-1} \wedge \p{const}_t(m, k)\Big)\wedge$$
+          $$\Big(o_t \simeq \o{put}_t(m\in M, k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
+        \item If there was no such mupping before, and there is enough room, it will succeed.
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_{t-1}\big) \Leftrightarrow $$
+          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge$$
+        \item If a mapping with key $k$ allready exists, and it does not expires on $t$, put will fail.
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in m.\s{mapping}_{t-1} \wedge e > t \big) \Leftrightarrow$$
+          $$x=fail \wedge \p{const}_t(m)\Big)\wedge$$
+        \item If a mapping with key $k$ expires on $t$, put will replace it:
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V: [k, v', t]\in m.\s{mapping}_{t-1}\big) \Leftrightarrow$$
+          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge$$
+        \item If there is no room, but some mappings expire on $t$, put will succed:
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \wedge $$
+          $$\big(\exists k'\in K, v'\in V: [k', v', t] \in m.\s{mapping}_{t-1}\big) \Leftrightarrow$$
+          $$x=success \wedge m.\s{state}_t \le m.\s{state}_{t-1} \wedge [k, v, t] \in m.\s{mapping}_t\Big)\wedge$$
+        \item If there is no room, and no mappings expire on $t$, put will fail:
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge $$
+          $$\big(\forall k'\in K, v'\in V, e' \in \nintset: [k', v', e'] \in m.\s{mapping}_{t-1} \Rightarrow e' > t \big) \Leftrightarrow$$
+          $$x=fail \wedge \p{const}_t(m)\Big)$$
+%        \item If there is no room, but some mappings have expired, one still can put.
+%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
+%          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
+%          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
+
+%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
+%          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
+%          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
+%        \item If there is an unexpired mapping with the same key, put will fail.
+%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge (v, e)=\o{get}_{t-1}(m, k) \wedge e \ge t \Leftrightarrow$$
+%          $$x=fail \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge \o{has}_t(m, k) \wedge \o{get}_t(m, k) = (v, e)\Big)\wedge$$
+%        \item Into full map, one can not put any more keys.
+%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge$$
+%          $$\big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(m, k')\wedge(\cdot,e')=\o{get}_{t-1}(m,k') \Rightarrow e' \ge t \big) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \Leftrightarrow$$
+%          $$x=fail \wedge m.\s{size}_t = m.\s{size}_{t-1} \wedge \p{const}_t(m, k)\Big)\wedge$$
       \end{itemize}
@@
 \begin{itemize}
+  \item $\p{empty}_t(m) \Leftrightarrow m.\s{size}_t=0$
+  \item If the map is empty, there no keys in it:
+    $$m.\s{size}_t=0 \Rightarrow \forall k\in K \o{has}_t(m, k) = false $$
+  \item $\o{has}_t(m, k) \Leftrightarrow \o{get}_t(m, k) \not\rightarrow fail$
+
+      \item If there was a such mapping for $k$, ``erase'' will remove it.
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
+      \item If there were not --- it will fail. 
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
+
+      \item 
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
+          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
+
   \item The size of the map is always properly bounded 
@@
    $$\o{has}_b(m, k) \wedge \o{get}_b(m, k) = (v,e)$$
+  \item $$\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~$$
+     $$\Big(\forall k\in K,v \in V, e\in \nintset : k\neq k' \wedge \o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge$$
+     $$\Big(\forall k\in K:k\neq k' \wedge \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$$
+  \item $$\p{const}_t(m\in M) ~\Leftrightarrow~$$
+     $$\Big(\forall k\in K,v \in V, e\in \nintset : \o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge$$
+     $$\Big(\forall k\in K: \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$$
+
 \end{itemize}

### 2014-11-12 form.tex
before replacement math mode by equation environments

@@
        \item It does not change nonexpired mappings:
-         $$o_t \simeq \o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
+         \begin{equation}
+           o_t \simeq \o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)
+         \end{equation}
      \end{itemize}
@@
           $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge$$
-        \item If there is no room, but some mappings expire on $t$, put will succed:
+        \item If there is no room, but some mappings expire on $t$, put will succeed:
           $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \wedge $$
@@
           $$x=fail \wedge \p{const}_t(m)\Big)$$
+      \end{itemize}
+
+\end{itemize}
+
+\paragraph{Properties to be proven}
+\begin{itemize}
+  \item $\p{empty}_t(m) \Leftrightarrow m.\s{size}_t=0$
+  \item If the map is empty, there no keys in it:
+    $$m.\s{size}_t=0 \Leftrightarrow \forall k\in K \o{has}_t(m, k) = false $$
+  \item $\o{has}_t(m, k) \Leftrightarrow \o{get}_t(m, k) \not\rightarrow fail$
+
+      \item If there was a such mapping for $k$, ``erase'' will remove it.
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
+      \item If there were not --- it will fail. 
+        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
+
+      \item 
+          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
+          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
+
 %        \item If there is no room, but some mappings have expired, one still can put.
@@
 %          $$x=fail \wedge m.\s{size}_t = m.\s{size}_{t-1} \wedge \p{const}_t(m, k)\Big)\wedge$$
-      \end{itemize}
-
-\end{itemize}
-
-\paragraph{Properties to be proven}
-\begin{itemize}
-  \item $\p{empty}_t(m) \Leftrightarrow m.\s{size}_t=0$
-  \item If the map is empty, there no keys in it:
-    $$m.\s{size}_t=0 \Rightarrow \forall k\in K \o{has}_t(m, k) = false $$
-  \item $\o{has}_t(m, k) \Leftrightarrow \o{get}_t(m, k) \not\rightarrow fail$
-
-      \item If there was a such mapping for $k$, ``erase'' will remove it.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
-      \item If there were not --- it will fail. 
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
-
-      \item 
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
-          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
 

### 2014-11-12 form.tex
Rewritten in terms of abstract state. Old representation is kept in 'to be proved' paragraph

--- 2014-11-12 form.tex
@@
 \usepackage{dsfont}
+\usepackage{amsmath}
 \usepackage[margin=1.5cm]{geometry}
@@
 
-By default, if symbols $m, e, k, v$ are used in a formula, and not mentioned before in that formula, the universal quantifier $\forall$ is implied: $\forall m\in M, k \in K, v\in V, e\in \nintset$.
+By default, if symbols $m, e, k, v, t$ are used in a formula, and not mentioned before in that formula, the universal quantifier $\forall$ is implied: $\forall m\in M, k \in K, v\in V, e\in \nintset, t\in \nintset$.
 
@@
   $\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~\\ \Big(k\neq k' \wedge [k, v, e] \in m.\s{mapping}_{t-1} \wedge e>t \Rightarrow [k, v, e] \in m.\s{mapping}_t\Big)\wedge\\ \Big(k\neq k' \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \Rightarrow [k, v, e] \not\in m.\s{mapping}_t\Big)$
+\end{itemize}
 
+\paragraph{Invariants}
+These predicates allways hold:
+\begin{itemize}
+\item Map does not keep expired items: $\forall [k, v, e] \in m.\s{mapping}_t \Rightarrow e > t$
+\item No operation except for $\o{constructor}_0$ can change map capacity:
+  \begin{gather*}
+    \forall t\in \nintset: t > 0 \Rightarrow m.\s{maxSize}_t=m.\s{maxSize}_0
+  \end{gather*}
 \end{itemize}
@@
   \item $\o{has}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t \wedge o_t=\o{has}_t(m, k)$\\
-       \item It does not change nonexpired mappings:
-         $$o_t \simeq \o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
+  \item It does not change mappings:
+    \begin{gather*}
+      o_t \simeq \o{has}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)
+    \end{gather*}
   \end{itemize}
@@
   \item One can get a mapping only if it exists in the map:
-           $$\o{get}_t(m \in M, k \in K) \rightarrow fail ~\Leftrightarrow~ [k,v,e] \not\in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k)$$
-           $$(v\in V,e\in \nintset)\gets \o{get}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k)$$
+      \begin{gather}
+        \o{get}_t(m \in M, k \in K) \rightarrow fail ~\Leftrightarrow~ [k,v,e] \not\in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k)\label{eq:get1}\\
+        (v\in V,e\in \nintset)\gets \o{get}_t(m \in M, k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in m.\s{mapping}_t \wedge o_t=\o{get}_t(m, k) \label{eq:get2}
+      \end{gather}
   \item In anyway, it does not change existing unexpired mappings:
-           $$o_t \simeq \o{get}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)$$
+    \begin{gather*}
+      o_t \simeq \o{get}_t(m, k) ~\Leftrightarrow~ \p{const}_t(m)
+    \end{gather*}
   \end{itemize}
@@
   \begin{itemize}
-      \item The constructor forms a new empty map \\
-        $m \gets \o{constructor}_0(s \in \nintset) \Leftrightarrow \p{empty}_0(m) \wedge \forall t \in \nintset m.\s{maxSize}_t=s$
-      \item The constructor must be called once and only once, at the very beginning of $\tau$ (i.e., $\tau$ contains no other constructor call after the first one) \\
-        $o_0 \equiv \o{constructor}_0 \wedge (\forall t\in \moreset{1} \Rightarrow o_t \neq \o{constructor}_t)$ 
+  \item The constructor forms a new empty map
+    \begin{gather*}
+      m \gets \o{constructor}_0(s \in \nintset) \Leftrightarrow \p{empty}_0(m) \wedge \big(\forall t \in \nintset \Rightarrow m.\s{maxSize}_t=s\big)
+    \end{gather*}
+  \item The constructor must be called once and only once, at the very beginning of $\tau$ (i.e., $\tau$ contains no other constructor call after the first one)
+    \begin{gather*}
+      o_0 \equiv \o{constructor}_0 \wedge \big(\forall t\in \moreset{1} \Rightarrow o_t \neq \o{constructor}_t\big)
+    \end{gather*}
   \end{itemize}
@@
   \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \p{mapping}, it fails. It does not change other mappings if they are not expired.
-        $$\Big(o_t \simeq \o{erase}_t(m\in M, k\in K) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
+    \begin{gather*}
+      \Big(o_t \simeq \o{erase}_t(m\in M, k\in K) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge
+    \end{gather*}
   \item If there was a such mapping for $k$, ``erase'' will remove it.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \exists [k, v, e] \in m.\s{mapping}_{t-1} ~\Leftrightarrow~$$
-        $$x=success \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_t\big) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
+    \begin{gather}
+      \Big(x=\o{erase}_t(m\in M, k\in K) \wedge \exists [k, v, e] \in m.\s{mapping}_{t-1} ~\Leftrightarrow~\nonumber\\
+      x=success \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_t\big) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge
+      \label{eq:erase1}
+    \end{gather}
   \item If there was not --- it will fail.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge [k,v,e]\not\in m.\s{mapping}_{t-1} ~\Leftrightarrow~$$
-        $$x=fail \wedge \p{const}_t(m)\Big)$$
+    \begin{gather}
+      \Big( x=\o{erase}_t(m\in M, k\in K) \wedge [k,v,e]\not\in m.\s{mapping}_{t-1} ~\Leftrightarrow~ \nonumber\\
+      x=fail \wedge \p{const}_t(m)\Big)
+      \label{eq:erase2}
+    \end{gather}
   \end{itemize}
@@
     This function does not touch any unexpired entry with key other than $k$.
-          $$\Big(o_t \simeq \o{put}_t(m\in M, k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge$$
+    \begin{gather*}
+      \Big(o_t \simeq \o{put}_t(m\in M, k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(m, k)\Big) \wedge
+    \end{gather*}
   \item If there was no such mupping before, and there is enough room, it will succeed.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_{t-1}\big) \Leftrightarrow $$
-          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge$$
+    \begin{gather}
+      \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in m.\s{mapping}_{t-1}\big) \Leftrightarrow \nonumber\\
+      x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge
+      \label{eq:put1}
+    \end{gather}
   \item If a mapping with key $k$ allready exists, and it does not expires on $t$, put will fail.
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in m.\s{mapping}_{t-1} \wedge e > t \big) \Leftrightarrow$$
-          $$x=fail \wedge \p{const}_t(m)\Big)\wedge$$
+    \begin{gather}
+      \Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in m.\s{mapping}_{t-1} \wedge e > t \big) \Leftrightarrow\nonumber\\
+      x=fail \wedge \p{const}_t(m)\Big)\wedge
+      \label{eq:put2}
+    \end{gather}
   \item If a mapping with key $k$ expires on $t$, put will replace it:
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V: [k, v', t]\in m.\s{mapping}_{t-1}\big) \Leftrightarrow$$
-          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge$$
-        \item If there is no room, but some mappings expire on $t$, put will succed:
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \wedge $$
-          $$\big(\exists k'\in K, v'\in V: [k', v', t] \in m.\s{mapping}_{t-1}\big) \Leftrightarrow$$
-          $$x=success \wedge m.\s{state}_t \le m.\s{state}_{t-1} \wedge [k, v, t] \in m.\s{mapping}_t\Big)\wedge$$
+    \begin{gather}
+      \Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V: [k, v', t]\in m.\s{mapping}_{t-1}\big) \Leftrightarrow\nonumber\\
+      x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge [k, v, e] \in m.\s{mapping}_t\Big)\wedge
+      \label{eq:put3}
+    \end{gather}
+  \item If there is no room, but some mappings expire on $t$, put will succeed:
+    \begin{gather}
+      \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge [k, v, e] \not\in m.\s{mapping}_{t-1} \wedge \nonumber\\
+      \big(\exists k'\in K, v'\in V: [k', v', t] \in m.\s{mapping}_{t-1}\big) \Leftrightarrow\nonumber\\
+      x=success \wedge m.\s{state}_t \le m.\s{state}_{t-1} \wedge [k, v, t] \in m.\s{mapping}_t\Big)\wedge
+      \label{eq:put4}
+    \end{gather}
   \item If there is no room, and no mappings expire on $t$, put will fail:
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge $$
-          $$\big(\forall k'\in K, v'\in V, e' \in \nintset: [k', v', e'] \in m.\s{mapping}_{t-1} \Rightarrow e' > t \big) \Leftrightarrow$$
-          $$x=fail \wedge \p{const}_t(m)\Big)$$
-%        \item If there is no room, but some mappings have expired, one still can put.
-%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
-%          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
-%          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
-
-%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge$$
-%          $$\big(\exists k'\in K, e'\in \nintset: \o{has}_{t-1}(m,k')\wedge (\cdot,e')=\o{get}_{t-1}(m, k) \wedge e'<t \big) \Leftrightarrow$$
-%          $$x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
-%        \item If there is an unexpired mapping with the same key, put will fail.
-%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge (v, e)=\o{get}_{t-1}(m, k) \wedge e \ge t \Leftrightarrow$$
-%          $$x=fail \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge \o{has}_t(m, k) \wedge \o{get}_t(m, k) = (v, e)\Big)\wedge$$
-%        \item Into full map, one can not put any more keys.
-%          $$\Big(x=\o{put}_t(m, k, v, e) \wedge$$
-%          $$\big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(m, k')\wedge(\cdot,e')=\o{get}_{t-1}(m,k') \Rightarrow e' \ge t \big) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \Leftrightarrow$$
-%          $$x=fail \wedge m.\s{size}_t = m.\s{size}_{t-1} \wedge \p{const}_t(m, k)\Big)\wedge$$
+    \begin{gather}
+      \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \nonumber\\
+      \big(\forall k'\in K, v'\in V, e' \in \nintset: [k', v', e'] \in m.\s{mapping}_{t-1} \Rightarrow e' > t \big) \Leftrightarrow\nonumber\\
+      x=fail \wedge \p{const}_t(m)\Big)
+      \label{eq:put5}
+    \end{gather}
   \end{itemize}
@@
 \item If the map is empty, there no keys in it:
-    $$m.\s{size}_t=0 \Rightarrow \forall k\in K \o{has}_t(m, k) = false $$
-  \item $\o{has}_t(m, k) \Leftrightarrow \o{get}_t(m, k) \not\rightarrow fail$
+  \begin{gather*}
+    \p{empty}_t(m) \Leftrightarrow \o{has}_t(m, k) = false
+  \end{gather*}
+\item Equivalent helper predicate \p{isolated} definition
+  \begin{gather*}
+    \p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~\nonumber\\
+    \Big(k\neq k' \wedge \o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge\nonumber\\
+    \Big(k\neq k' \wedge \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)
+  \end{gather*}
+\item Equivalent helper predicate \p{const} definition
+  \begin{gather*}
+    \p{const}_t(m\in M) ~\Leftrightarrow~\nonumber\\
+    \Big(\o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge\nonumber\\
+    \Big(\forall k\in K: \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)
+  \end{gather*}
 
-      \item If there was a such mapping for $k$, ``erase'' will remove it.
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big) \wedge$$
-      \item If there were not --- it will fail. 
-        $$\Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)$$
+\item Equivalent of \eqref{eq:get1}:
+  \[
+  \o{has}_t(m, k) \Leftrightarrow \o{get}_t(m, k) \not\rightarrow fail
+  \]
+\item If there was a mapping for $k$, ``erase'' removes it. Equivalent of \eqref{eq:erase1}:
+  \begin{gather*}
+    \Big(x=\o{erase}_t(m\in M, k\in K) \wedge \o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(m, k) \wedge m.\s{size}_t \le m.\s{size}_{t-1}-1 \Big)
+  \end{gather*}
+\item If there were not --- ``erase'' will fail. Equivalent of \eqref{eq:erase2}:
+  \begin{gather*}
+    \Big(x=\o{erase}_t(m\in M, k\in K) \wedge \neg\o{has}_{t-1}(m, k) ~\Leftrightarrow~ x=fail\wedge \p{const}_t(m)\Big)
+  \end{gather*}
 
-      \item 
-          $$\Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow $$
-          $$ x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)\wedge$$
+\item Equivalent of \eqref{eq:put1}
+  \begin{gather*}
+    \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} < m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \Leftrightarrow \nonumber\\
+    x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1}+1 \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)
+  \end{gather*}
+
+\item If there is an unexpired mapping with the same key, put will fail. Equivalent of \eqref{eq:put2}
+  \begin{gather*}
+    \Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge \exists e'\in \nintset: (\cdot, e')=\o{get}_{t-1}(m, k) \wedge e' > t \Leftrightarrow\\
+    x=fail \wedge \p{const}_t(m)\Big)
+  \end{gather*}
+
+\item Equivalent of \eqref{eq:put3}. Expired mapping replaced:
+  \begin{gather*}
+    \Big(x=\o{put}_t(m, k, v, e) \wedge \o{has}_{t-1}(m, k) \wedge (\cdot, t)=\o{get}_{t-1}(m, k) \Leftrightarrow\\
+    x=success \wedge m.\s{size}_t \le m.\s{size}_{t-1} \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e)\Big)
+  \end{gather*}
+
+\item Equivalent of \eqref{eq:put4}. If there is no room, but some mappings have expired, one still can put.
+  \begin{gather*}
+    \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(m, k) \wedge\\
+    \big(\exists k'\in K: \o{has}_{t-1}(m,k')\wedge (\cdot,t)=\o{get}_{t-1}(m, k) \big) \Leftrightarrow\\
+    x=success \wedge \o{has}_t(m, k)\wedge\o{get}_t(m, k) = (v, e) \wedge m.\s{size}_t \le m.\s{size}_{t-1}\Big)
+\end{gather*}
+
+\item Into full map, one can not put any more keys. Equivalent of \eqref{eq:put5}:
+  \begin{gather*}
+    \Big(x=\o{put}_t(m, k, v, e) \wedge m.\s{size}_{t-1} = m.\s{maxSize}_{t-1} \wedge  \\
+    \big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(m, k')\wedge(\cdot,e')=\o{get}_{t-1}(m,k') \Rightarrow e' > t \big) \Leftrightarrow\\
+    x=fail \wedge m.size_t = m.size_{t-1} \wedge \p{const}_t(m, k)\Big)
+  \end{gather*}
 
 \item The size of the map is always properly bounded
-   $$\forall t\in \nintset: t>0  \wedge m=\o{constructor}_0(s \in \nintset) \Rightarrow 0 \le m.\s{size}_t \le m.\s{maxSize}$$
-  \item No operation except for $\o{constructor}_0$ can change map capacity:
-   $$\forall t\in \nintset: t > 0 \Rightarrow m.\s{maxSize}_t=m.\s{maxSize}_0$$
+  \begin{gather*}
+    \forall t\in \nintset: t>0  \wedge m=\o{constructor}_0(s \in \nintset) \Rightarrow 0 \le m.\s{size}_t \le m.\s{maxSize}
+  \end{gather*}
+
 \item For any subsequence of operations other than \o{put}, size of the map may only decrease:
-   $$\forall a\in \nintset,b\in \nintset: a<b \wedge \big(\forall t\in \nintset:a<t<b \Rightarrow o_t \neq \o{put}_t \big) \Rightarrow \s{size}_a \ge \s{size}_b$$
-  \item If a subsequence of operations does not include \o{erase} for key $k\in K$, and there is a mapping, it will be kept until the end of the subsequence, if not expired:
-   $$\forall a\in \nintset,b, k\in K, m\in M, v\in V, e\in \moreset{b}: a < b \wedge \o{has}_a(m, k) \wedge (v, e)=\o{get}_a{m, k} \wedge e \ge b ~ \wedge$$
-   $$\big(\forall t\in \nintset: a<t<b \Rightarrow o_t\neq \o{erase}_t(m, k)\big) \Rightarrow$$
-   $$\o{has}_b(m, k) \wedge \o{get}_b(m, k) = (v,e)$$
-  \item $$\p{isolated}_t(m\in M, k'\in K) ~\Leftrightarrow~$$
-     $$\Big(\forall k\in K,v \in V, e\in \nintset : k\neq k' \wedge \o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge$$
-     $$\Big(\forall k\in K:k\neq k' \wedge \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$$
-  \item $$\p{const}_t(m\in M) ~\Leftrightarrow~$$
-     $$\Big(\forall k\in K,v \in V, e\in \nintset : \o{has}_{t-1}(m, k) \wedge (v,e)=\o{get}_{t-1}(m, k) \wedge e>t \Rightarrow \o{has}_t(m, k) \wedge \o{get}_t(m, k)=(v,e)\Big)\wedge$$
-     $$\Big(\forall k\in K: \neg \o{has}_{t-1}(m, k) \Rightarrow \neg\o{has}_t(m, k)\Big)$$
+  \begin{gather*}
+    \forall a\in \nintset,b\in \nintset: a<b \wedge \big(\forall t\in \nintset:a<t<b \Rightarrow o_t \neq \o{put}_t \big) \Rightarrow \s{size}_a \ge \s{size}_b
+  \end{gather*}
+
+\item If a subsequence of operations does not include \o{erase} for key $k\in K$, and there is a mapping, it will be kept until the end of the subsequence, or until expired:
+  \begin{gather*}
+    \forall a\in \nintset,b, k\in K, m\in M, v\in V, e\in \moreset{b}: a < b \wedge \o{has}_a(m, k) \wedge (v, e)=\o{get}_a{m, k} \wedge e > b ~ \wedge\nonumber\\
+    \big(\forall t\in \nintset: a<t<b \Rightarrow o_t\neq \o{erase}_t(m, k)\big) \Rightarrow\nonumber\\
+    \o{has}_b(m, k) \wedge \o{get}_b(m, k) = (v,e)
+  \end{gather*}
 

### 2014-11-13 form.tex
some corrections

@@
 each subsequence $<S_{i-1}, o_i, S_i>$ indicating that operation $o_i$ transforms abstract state $S_{i-1}$ into $S_i$.
-An accessor does not mutate state, so it leaves the abstract state unchanged.
+An accessor does not mutate state, so it leaves the abstract state unchanged, but the state can change due to continuous expiration process.
 We use the shorthand $\s{component}_t$ to indicate the value of $S.$\s{component} immediately after $o_t$ completes and before $o_{t+1}$ commences.
@@
 E.g.
-$\forall k \in K, m \in M o_t\simeq \o{has}_t(m, k) \Leftrightarrow \p{const}_t(m)$ means, that the operation number t ($o_t$) may be $\o{has}(m, k)$ (not the result of such function call) if and only if holds $\p{const}_t(m)$.
+$\forall k \in K, m \in M o_t\simeq \o{has}_t(m, k) \Leftrightarrow \p{const}_t(m)$ means, that the operation number t ($o_t$) may be the $\o{has}(m, k)$ call (not the result of it) if and only if holds $\p{const}_t(m)$.
+\underline{Attention}: $x=\o{has}_t(m, k)$ has different meaning then $o_t=\o{has}_t(m, k)$.
 
@@
   \begin{itemize}
-  \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \p{mapping}, it fails. It does not change other mappings if they are not expired.
+  \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \s{mapping}, it fails. It does not change other mappings if they are not expired.
     \begin{gather*}
@@
       \Big( x=\o{erase}_t(m\in M, k\in K) \wedge [k,v,e]\not\in m.\s{mapping}_{t-1} ~\Leftrightarrow~ \nonumber\\
-      x=fail \wedge \p{const}_t(m)\Big)
+      x=fail \wedge \p{const}_t(m) \wedge o_t=\o{erase}_t(m, k)\Big)
       \label{eq:erase2}
@@
     \begin{gather}
-      \Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in m.\s{mapping}_{t-1} \wedge e > t \big) \Leftrightarrow\nonumber\\
-      x=fail \wedge \p{const}_t(m)\Big)\wedge
+      \Big(x=\o{put}_t(m, k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in m.\s{mapping}_{t-1} \wedge e > t\big) \Leftrightarrow\nonumber\\
+      x=fail \wedge \p{const}_t(m) \wedge o_t = \o{put}_t(m,k,v,e)\Big)\wedge
       \label{eq:put2}
@@
       \big(\forall k'\in K, v'\in V, e' \in \nintset: [k', v', e'] \in m.\s{mapping}_{t-1} \Rightarrow e' > t \big) \Leftrightarrow\nonumber\\
-      x=fail \wedge \p{const}_t(m)\Big)
+      x=fail \wedge \p{const}_t(m) \wedge o_t = \o{put}_t(m,k,v,e)\Big)
       \label{eq:put5}

### 2016-09-22 doc/paper/paper.tex
the submitted for review paper and related materials

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage{changepage}

\definecolor{mygreen}{rgb}{0.1,0.5,0.1}
\definecolor{myblue}{rgb}{0.1,0.1,0.7}

\lstset{language=C,
    basicstyle=\footnotesize\ttfamily,
    keywordstyle=\color{myblue}\ttfamily,
    stringstyle=\color{mygreen}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    keepspaces=true,
    tabsize=2,
    otherkeywords={bool}
}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\vignat}{{\sc VigNAT}\xspace}
\newcommand{\vignata}{{\sc VigNAT$\alpha$}\xspace}
\newcommand{\vignatb}{{\sc VigNAT$\beta$}\xspace}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

\newcounter{solOverviewSteps}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Vigor: Toward Carrier-Grade Software Network Functions}

%\author{
%    {\rm Arseniy Zaostrovnykh}\\
%    EPFL
%    \and
%    {\rm Katerina Argyraki}\\
%    EPFL
%    \and
%    {\rm George Candea}\\
%    EPFL
%}
\author{
    {\rm Paper \#113}\\
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
The objective of this work is to enable developers to productively write software-based data plane network functions that simultaneously meet the highest standards of reliability, security, and performance. We present a verification approach for formally proving properties of stateful network functions, in a way that is practical both in terms of developer effort and the performance of the resulting data plane software. Unlike alternative approaches that require drastic changes in the way network functions are developed, our approach allows the use of C/C++ together with popular frameworks like DPDK. To the best of our knowledge, our work is the first to both be practical and formally verify stateful network functions. We introduce a new ``speculative'' form of symbolic execution, which achieves scalability to real-world network functions by leveraging insights related to the structure and cyclomatic complexity of data plane software. This technique dovetails with a new library of verified components that are commonly used in network functions. We illustrate our approach by developing a middlebox that is verified to never crash, to not leak memory, be free of several classes of security vulnerabilities, and whose performance is on par with other NATs. %\footnote{Disclaimer: we use acronym NAT in this paper to refer to a more narrow kind - Dynamic Source NAPT (network address-port translation~\cite{rfc3022}).}. 
We view this as an important milestone on the road to complete verification of all software-based network functions.

\section{Introduction}

Product development in the telecommunication industry has traditionally followed rigorous standards for quality, security, and protocol adherence, with the term carrier-grade designating network equipment that meets these high standards. While this model worked well in the past, it typically entailed reliance on proprietary hardware and long product cycles, which slowed progress. To meet the growing demand for new network functionality and faster deployment cycles, network service providers started turning to software.

In recent years, new platforms emerged for developing network functionality in software~\cite{kohler2000click,bosshart2014p4,mccanne1993bsd,song2015unified,shah2003np,shahbaz2016pisces}, enabling the rapid prototyping and deployment of new routing protocols~\cite{sivaraman2016packet}, QoS policies~\cite{sivaraman2016programmable}, load balancing algorithms~\cite{zhang2013steering}, intrusion prevention systems~\cite{xing2014sdnips}, etc. The agility that software engenders determined even traditional hardware-based network device manufacturers to invest in software-based approaches~\cite{ciscodevelopssdn}.

But this raises the question of how to develop carrier-grade network functions (NF) in software? Since NFs are often part of critical infrastructure, they must simultaneously be secure, reliable, and high-performance, i.e., be carrier-grade. The Achilles' heel of software has always been the (perceived) lack of reliability and security, particularly when it is written in C. This has led researchers to advocate using safe languages~\cite{clipsham2015safe, kourtis2015intelligent, fowler2014verified} or relatively heavyweight approaches~\cite{madhavapeddy2007melange, madhavapeddy2005splat, musuvathi2004model, alur2001verifying, ennals2004linear, ridge2008rigorous, qadir2015applying}. Yet many frameworks (e.g., P4~\cite{bosshart2014p4}, DPDK~\cite{intel2014data}, PF\_RING~\cite{pfring}, NetMap~\cite{rizzo2012revisiting}) and standalone implementations (Open vSwitch~\cite{pfaff2009extending}, NetFilter~\cite{boye2013netfilter}) use C. % These approaches require developers to drastically change or augment the way in which they write these network apps: massively annotate their code with logical derivations, use ``pure'' (side-effect-free) functions, abandon manual memory management and layout, and/or learn a new language.

Developers do not want to change the way they write code, they want to be able to convince their users that the code they wrote is carrier-grade (and, if they cannot prove correctness, to then fix the corner cases that impede the proof). We conducted a small survey~\cite{verifsurvey} among NF developers from both industry and academia, and almost 70\% insist on continuing to use C. % instead of a higher level alternative. With current common practices and languages, it is not possible to achieve the desired level of confidence. The continued perception of unreliability and insecurity hinders adoption and therefore slows down progress (can we quote an operator who doesn't want to deploy stuff out of fear?).

The goal of our work is to enable developers to provide  carrier-grade network functionality in software without having to do a forklift upgrade of their programming practices. In this paper we present a technique and a tool suite (a) for formally verifying crash- and leak-freedom as well as the absence of several classes of security vulnerabilities in NFs written in C/C++ (b) in a way that is practical both in terms of developer effort and the performance of the resulting software.

There exist approaches for providing formal guarantees for {\em stateless} network packet processing code~\cite{dobrescu2014software}; to the best of our knowledge, we are the first to provide such guarantees for real-world {\em stateful} software NFs without impacting performance. There are many such NFs: all dynamic routing protocols (RIP, BGP, STP), NAT, BRAS~\cite{wiki:bras}, CDN, IDS, DPI boxes, etc.

Practical verification of stateful software NFs is a difficult research problem. Hence none of the verification techniques proposed so far can scale anywhere close to verifying real-world stateful NFs. For the purpose of this paper, we take as an objective to develop a NAT~\cite{rfc3022} box that is proven to not have certain classes of vulnerabilities or crash or hang under any input. Even though NAT is a classic network function, it has proven hard to get right over time: the NAT on various Cisco devices can be crashed~\cite{cve-2015-6271} or hung~\cite{cve-2013-1138} by carefully crafted inputs, with the same type of problems existing in Juniper's NAT~ \cite{cve-2014-3817}, NATs based on NetFilter~\cite{cve-2014-9715}, and the NAT in Windows Server~\cite{ms13-064}. Thus, a NAT function that is formally verified to not crash, while providing production-class performance, is an important milestone in this line of work.

We achieve both scalable verification and minimal developer effort by leveraging two insights into the verification of data plane software:
\begin{itemize}
    \item {\bf Canonical data structures}: Unlike general-purpose software, data plane software employs a small number of commonly used network data structures (hash table, least-prefix-match, rings, etc.) and components. These lend themselves naturally to being abstracted into a library. They can be formally verified a priori, and their proofs then automatically leveraged for proving properties of the code that uses them.
    \item {\bf Hybrid verification}: While typically one picks a verification technique and applies it throughout, any {\em one} technique works well for {\em some} of the code and not for other. However, data plane software offers us the opportunity to advantageously employ symbolic execution for high-level code (which is mostly stateless and has short execution paths) and verification based on separation logic to the library of components/data structures (which is highly stateful).

\end{itemize}

In this paper we present work that builds upon these insights and makes three contributions:
\begin{itemize}
    \item {\bf Speculative data plane verification} (SDSE), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution. % It uses formal verification for data structures. It replaces them with simple models that encode some assumptions about the behaviour of the data structures. It then runs SE on the application code and later validates the used models. This approach postpones high-effort/low-return work and discards it when it turns out to be irrelevant.
    \item {\bf The Vigor library of verified components} along with a tool chain that makes SDSE available to those who write software NFs in C/C++. The library consists of data structures (Hash Map, Batcher, Static Array) and a Port Allocator that are common for software NF. % The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
    \item {\bf The \vignat verified NAT box} that offers both high performance and is formally proven to not crash or leak memory. We estimate that it would take an NF developer about 1 week to develop and verify a NAT using Vigor.
\end{itemize}

In the rest of the paper, we give a short overview of Vigor (\S\ref{sec:solution-overview}), describe our design (\S\ref{sec:design}), and present our implementation (\S\ref{sec:implementation}). We then evaluate the technique for two NAT prototypes (\S\ref{sec:evaluation}), discuss limitations (\S\ref{sec:limitations}), present related work (\S\ref{sec:related-work}), and conclude (\S\ref{sec:conclusion}).

%==============================================

\section{Solution Overview}
\label{sec:solution-overview}

This section motivates and explains at a high level what it's like to use Vigor when developing NFs. For the sake of illustration, say we wish to develop a trivial stateful NF that implements the discard protocol~\cite{rfc863}: an infinite loop receives packets from one interface, discards the ones that go to port 9 (``discard port''), and forwards the others through another interface.

\begin{lstlisting}
#define CAP 512

int main() {
  struct packet packets[CAP] = {};
  int begin = 0, end = 0;

  while(1) {
    if (end != (begin - 1) || 
        !(end == CAP - 1 && begin == 0)) {
      if (receive_packet(&packets[end]) && 
          packets[end].port != 9)
        end = (end + 1)%CAP;
    }
    if (end != begin && can_send_packet()) {
      send_packet(&packets[begin]);
      begin = (begin + 1)%CAP;
    }
  }
  return 0;
}

\end{lstlisting}

In order to absorb bursts, the code employs a ring buffer. The interface to the network is: {\em bool receive\_packet(struct packet* dst)}, which non-blockingly reads an inbound packet and stores it at {\em dst}, returning success or failure; {\em bool can\_send\_packet()}, which checks if a new packet can be sent; and {\em void send\_packet(struct packet* src)}, which sends the packet stored at {\em src}. The application follows common practice: it consists of a main event loop, it has no other complex loops, and it uses a ring, a popular data structure in network programming.

To verify that this NF does not crash, one could use the approach in Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The symbolic execution (SE) technique entails using a symbolic execution engine (SEE) to explore all paths through the code and check whether any failure conditions hold --- in our case crash conditions. The inputs to the program are marked ``symbolic,'' meaning that they subsume all possible input values, and then the SEE simulates the execution of the program while propagating the symbolic inputs along the dataflow paths into the program's variables. When facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the program state and takes each feasible path separately. Branches impose constraints on variable values (e.g., an \code{if} statement imposes restrictions on the values of the condition variable along the \code{else} branch); the SEE collects these constraints along each path. When evaluating a branch condition, the SEE asks an SMT constraint solver which of the branches are feasible, given the constraints collected so far on that path prefix.

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the code above exhaustively --- once all paths are found to be crash-free, the crash-freedom proof is complete. To perform SE, it is necessary to provide symbolic models for the three network calls. Something simple suffices, such as \code{receive\_packet} generating a new symbolic packet in \code{dst}; \code{can\_send\_packet} forking program state into two states, one for each possible return value; and \code{send\_packet} doing nothing.

Alas, symbolic execution would never terminate. The reason is ``path explosion:'' there is a large (often infinite) number of paths to explore, so the SEE cannot finish.  Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, three \code{if} conditions inside an infinite loop give rise to an infinite number of paths.

{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.  We added to KLEE support for havoc-ing~\cite{boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.

This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10 sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for \code{CAP=60} it takes 3 hours, and so on. It takes long due to another classic SE challenge, ``symbolic indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset}: when indexing an array with a symbolic variable (or when dereferencing a symbolic pointer), the constraints on program state become extremely complex, because the symbolic pointer or index could be referencing many different possible locations. The complexity of the constraints chokes the SMT solver, and the SEE grinds to a halt. In general, SE doesn't work for code with symbolic pointers or offsets. Alas, stateful C code almost always has some of these.

{\bf Step 2: Use abstract data structures.} In Vigor, we eliminate the problem of symbolic indexing by encapsulating program state into well-defined data structures and putting them into a ``library of components''. This works well for NF development, because the NF code uses a relatively small set of standard structures, such as ring buffers, hash tables, least-prefix-match tables, etc. We provide to the SEE symbolic models of these abstracted components that ``hide'' symbolic indexes and pointers from the SEE.

For our example, the library provides a \code{ring} with a simple interface: {\em bool ring\_empty()} checks whether there is a packet in the ring, {\em bool ring\_full()} checks whether the ring has run out of space, {\em void ring\_push\_back(struct packet* src)} adds packet {\em src} to the ring, and {\em void ring\_pop\_front(struct packet* dst)} removes a packet from the ring and places it in {\em dst}. We then write a symbolic model for the \code{ring} and provide it to the SEE instead of the actual implementation. The NF now becomes:

% \begin{lstlisting}
% bool ring_full() {return SYMBOLIC("full");}

% bool ring_empty() {return SYMBOLIC("empty");}

% void ring_push_back(struct packet* p) {
%   ASSERT(p->port != 9); //Extra check.
% }

% void ring_pop_front(struct packet* p) {
%   FILL_SYMBOLIC(p, sizeof(struct packet), 
%                 "popped_packet");
%   ASSUME(p->port != 9);
% }
% \end{lstlisting}

\begin{lstlisting}
#define CAP 512

int main() {
  struct packet p;
  while(LOOP(1)) {
    if (!ring_full())
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(&p);
    if (!ring_empty() && can_send_packet()) {
      ring_pop_front(&p);
      ASSERT(p.port != 9);
      send_packet(&p);
    }
  }
  return 0;
}
\end{lstlisting}

Now exhaustive SE of the code takes less than 100 millisec with our modified KLEE, regardless of the size of \code{CAP}.  The fact that the code does not crash on any path amounts to a proof of crash freedom. Or does it? The validity of the proof depends on (1)  our library implementation of \code{ring} not crashing; and (2) the symbolic model of \code{ring} accurately capturing how the \code{ring} implementation works.

The Vigor library comes with a formal specification and a formal proof that the implementation satisfies that specification (we show neither of them here, due to space constraints). In other words, there is a formal spec for \code{ring} and a proof that the \code{ring} implementation satisfies the spec. This is how we enable SE to verify the high level code, while using separation logic to verify the library of components---two different verifications techniques for two different parts of the NF. Doing the library verification requires deep expertise in program verification, but our effort is amortized across all NFs that use the library. Vigor makes it possible for NF developers to verify their code without having to understand verification. Note that stating invariants for dataplane software turns out to require no verification expertise.

Vigor must also validate that the model used for SE matches the library spec. Some symbolic models work well for some code, while others work well for other code, so the Vigor library comes with {\em multiple} simple, overapproximating models for each component, and Vigor automatically chooses combinations of models for symbolically executing the NF code. Once it finds a combination of models with which the desired property (crash freedom, no memory leaks, absence of certain security vulnerabilities) can be proven, Vigor validates that the behaviors exposed by the model during verification are consistent with the component's formal specification.  Details on how it does this appear in \S\ref{sec:design}.

{\bf Step 3: Link the NF code with the verified library} of components. The resulting executable is guaranteed to be free of crashes, subject to the assumptions listed in~\S\ref{sec:limitations}.

%Note, that the \code{ring\_pop} model does not even check whether it is called with an empty ring or not. The example application makes sure to call pop only if the ring is not empty. Therefore the check is not necessary.
%This model clearly overapproximates a ``ring'', e.g. \code{ring\_empty} may return true even right after \code{ring\_push}. This means, it may have caused false positives, but as long as it does not, no need to complicate it.
% For the sake of the example conciseness, we hard coded the invariant ``every packet has port != 9'' and the packet structure into the ring component itself. In a real component this detail should be abstracted and supplied as a parameter.

% Now a SE can analyze the code in an instant. As you can see the model does not maintain any state. It only \code{ASSERT}s that any pushed packet satisfy the invariant, and \code{ASSUME}s that every popped packet does. Last steps:

%========================================
\section{Design}
\label{sec:design}

In this section we describe the design of the Vigor technique and system. Figure~\ref{fig:arch} provides a high-level overview of the three main components of the tool chain. In order to prove properties of a network function that uses the Vigor component library, Vigor symbolically execute the NF using a symbolic execution engine (SEE); it then uses a Validator to combine the result of the SEE with the formal spec and proof of the component library.

\begin{figure}[h!]
    \includegraphics[width=\columnwidth]{figures/impl_overview.png}
    \caption{The Vigor tool chain for automatic verification.}
    \label{fig:arch}
\end{figure}

%---------------
\subsection{Speculative Verification}
\label{sec:our-approach}

Vigor views a network function as a combination of stateless application logic tied to components and data structures (collectively called ``components'') that are stateful, such as hash tables, arrays, least-prefix-match tables, etc. In the example of \S\ref{sec:solution-overview}, we encapsulated the \code{packets} array into the \code{ring} component. Singular variables of course need not be abstracted away, just large or variable-sized structures potentially accessed with symbolic indexes. The role of the library is to group stateful components into a unit that can be formally verified {\em once} an then {\em reused} across all NFs---we can view the Vigor library as a domain-specific extension of the C language. This makes sense because there is a relatively low number of components needed to support a wide range of NFs. Figure~ \ref{fig:algo} summarizes how Vigor works.

\begin{figure}[]
    \includegraphics[width=\columnwidth]{figures/vigor_algorithm.png}
    \caption{The Vigor verification workflow. Rectangles represent data, hexagons are the Vigor tools, and arrows indicate the flow of data.}
    \label{fig:algo}
\end{figure}

It is feasible for NF developers to develop verified components and add them to the library. However, this requires a fair amount of verification expertise: formally specify the component interface, implement it in a way that can be verified but also has good performance, then produce a mechanically checkable proof, write one or more symbolic models for it, and finally a plug-in for the Validator. Vigor therefore adopts a usage model in which the vast majority of NF developers just use the components in the Vigor library and run Vigor to prove (or debug) their code.

%We had to find our own way to support high level data structures in SE, because the popular approach does not suit our goal. Many systems based on SE extend the target language with specification syntax allowing developers to formulate complex conditions. For instance, Frama-C/Value~\cite{kirchner2015frama} and VeriFast~\cite{jacobs2008verifast} use their own abstract specification language. We strive to make the verification simple and automated, but unfortunately current state of the art systems cannot reason automatically in such sophisticated terms and require manual small-step derivations. Developers end up spending more effort on the correctness proof than on the code itself.

%We explore an alternative way:

For each component of the library, Vigor provides several symbolic models. A model is a concise over-approximating abstraction of how the specific component behaves when invoked by NF code. A good model has very little state. When symbolically executing an NF, the SEE uses one model for each component, and explores all execution paths through the NF together with the models. If every execution path has the desired property, then the NF has that property at all times. Since models are relatively easy to write, it is conceivable for NF developers to also write symbolic models.

A desired property is formulated as an assert. ``No division by zero''  is stated as an assert that the denominator is non-zero; ``safe memory access'' is stated as an assert that accessed memory is ``owned'' by the accessor. More details appear in \S\ref{sec:properties}. If a symbolically executed code path completes without any assert failing, we say the execution path was successful.

For each execution path, the Vigor SEE produces a trace of how the NF code interacted with the components it uses (we define these traces formally in \S\ref{sec:model-validation-criteria}). This is  {\em speculative} in that we don't have proof that the models are accurate abstractions of the components. The question the Validator answers is whether these traces are consistent with the formal specs of the components, i.e., could the interaction the NF code had with the model have been a legitimate interaction with an implementation of the component's specification. This is a weaker property than the model being a correct abstraction of the component, but it is sufficient for our purposes.

Of course, execution paths may not always succeed during SE. While this may be indicative of a bug in the NF, it could also be simply because the model is too approximate. For example, in \S\ref{sec:solution-overview} a model of \code{ring\_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring\_empty} to return false right after pushing an element. A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring\_empty} return true only when this counter is zero. The above NF code would not crash with this model; on the downside, the more stateful model makes SE take longer. Therefore, Vigor starts with the least-state models, and tries alternates until it either finds a model with which SE succeeds on all paths or not.

If Vigor does not find any symbolic model for which all execution paths through the NF code succeed, then it points the developers to the places where the properties fail to hold, and helps the debug.

%    \item The application developer provides loop invariants for long/unbounded loops (e.g. event loop in the example). The invariants are mostly a conjunction of the corresponding invariants for the symbolic models, but they are likely to have an application specific part.

% a number of speculative assumptions that we believe will hold for the real implementations. In the example above some of these assumptions are: ``the \code{ring\_empty} function may return either 0 or 1'', ``if the \code{ring\_pop\_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring\_push\_back} never crashes''.

% ^^^ do not forget to uncomment the references together with the figure
% \begin{figure}[h]
%     \centering
%     \def\svgwidth{\columnwidth}
%     \input{pie.pdf_tex}
%     \caption{Our conceptual division of the code into stateless and stateful parts with a strict interface between them.}
%     \label{fig:pie}
% \end{figure}

%-----------------
\subsection{Assumptions}
\label{sec:assumptions}

Our approach makes several assumptions. First, we assume that our symbolic models are over-approximations of the environment in which the NF runs (network stack, OS, etc.). In devising these models, we aim to include every possible behavior, but the absence of a formal specification of the environment makes it impossible to validate the models. Second, our guarantees rely on the hardware running the software dataplane correctly. Third, we assume the compiler implements the same language semantics employed by our tools (e.g., same byte length for C primitive types).

Thus, the trusted computing base for a Vigor-verified NF consists of the compiler, the Vigor tool chain (including its formal verification engine and SEE), the network stack, device drivers, OS kernel, BIOS, and hardware.

% We distinguish of the verified component library from development of an application using this library. These two activities require different skills and ideally should be done by different people.

% This paper proposes the following approach expressed in a number of steps. We use three tools: ``SE verifier'' stands for our SEE, ``formal verifier'' is our formal verification proof checker and assistant and a ``validator'' --- the tool that validates legitimacy and exhaustiveness of the model behaviour post factum. We differentiate human work between two developers: a component developer and an application developer. During the development of our NAT box they were the same person.

%------------------------------
\subsection{Handling pointers}
\label{sec:handling-pointers}

Pointers are the soul of the C language---they embody the power and the danger of low-level systems code. A NF focused on performance can not avoid them. We embrace them and support the most used idioms in our tool.

An important choice is the memory allocation strategy. Dynamic allocation offers more flexibility but incurs performance penalties. Static allocation should be preferred because it allows better control over layout, making it possible to optimize for cache and same the allocation overhead at runtime. However static allocation reduces flexibility, e.g. it fixes the component capacity at the compile time.

Among all the pointers appearing in the execution context, only the pointers crossing API boundary are important for the model validation. We divide these pointer into two kinds: argument pointers and return pointers.

An argument pointer may be either:
\begin{itemize}
    \item The address of the component state structure. It is an equivalent of ``this'' pointer for C++/Java or ``self'' reference for Rust/Python. We do not inspect the memory by such pointers, we only note the aliasing between them. The component owns this memory and only the formally verified code can access it.
    \item Output parameter. The application owns this memory. Therefore we have to trace the evolution of the pointee value. The pointee value before the call, which we trace as the function input, and the pointee value after the call, which goes to the function output. A special case of output pointer is a double pointer to the component state structure (X** p), set by the component allocation function. The app then owns the pointee (*p), and the SE verifier traces it, but the pointee of the pointee (**p) is the component internal state, so no need to trace it. We currently do not support other cases of double or deeper pointers (e.g. an output parameter to be set to point to a piece of the internal state).
\end{itemize}

A return pointer for our components always points to a piece of the internal state (e.g. an array cell). We trace the pointee and also make sure the application does not access it when the component regains the ownership.

%--------------------
\subsection{Property Formulation}
\label{sec:properties}

Vigor can verify any property that can be stated as a safety property, i.e., formulated as an \code{assert} condition. This includes various sources of crashes (dereferencing a null pointer, reading from an unmapped memory page, dividing by zero, etc.) as well as several classes of security bugs related to memory safety (writing to unallocated memory, writing to the wrong memory object, use-after-free, double-free, etc.) and related to integer operations (e.g., integer overflow).

When symbolically executing a code path, the SEE checks whether any asserts fail.

To check for memory leaks (or leaked resources in general) is a bit more complicated, because that is a global property. Vigor therefore must keep track of memory ownership. In the application logic we forbid dynamic memory allocation, and all allocated memory is controlled in the library, thereby giving the formal verifier explicit visibility into memory ownership. The temporary pointer access control enables the SE verifier to ensure the application returns all the memory chunks taken from a component and never accesses them after that.

Vigor does not specifically deal with semantic properties (e.g., dropping a packet, using an incorrect MAC address as a destination), but many of these properties can be formulated by NF developers as \code{assert} statements in their code. Vigor will then determine whether those \code{assert}s hold on all paths.

%----------------------
\subsection{Validating a Symbolic Model}
\label{sec:model-validation-criteria}

To validate a model behavior under certain circumstances means to ensure the behavior agrees with the specification, or is more relaxed (i.e., admits more behaviors than the spec). Suppose that, for the input \(I\) to the prefix of an execution trace (a ``run''), the specification produces the output $O_s$, and the model produces the case output \(O_m\). Then the model is valid iff for all call sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).

More formally:
\begin{enumerate}
    \item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and run output. It summarizes a particular interaction scenario with the model. It consists of the calls to component API functions (with function names, arguments and return values), and leaves out all other details like other calls or unrelated path conditions.
    \item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of: function names, function inputs(\(Fi\)) for all the functions in the sequence and function outputs(\(Fo\)) for all but the last call.
    \item \(Co\): The run output (\(Co = Fon\)) is the output of the last API function call in the sequence.
    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments). For a pointer argument (which must point to a concrete address --- practical limitation), the pointee must be considered according to the \S\ref{sec:handling-pointers}. We use the numerical value of the pointer to detect aliasing between the variables, and then throw it away, since we forbid pointer comparison anyway.
    \item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after + constraints\)) is its return value and the pointee values for the output pointer arguments, as well as the constraints for all the symbols involved (transitive closure, same as above).
\end{enumerate}

Let us represent the formal specification and the symbolic model total states by large sets. A total state consists of variables V, bound by logical formulas F. So, the specification and model total states are:
\[
\{V_s : F_s(V_s)\} and \{V_m : F_m(V_m)\}
\]
The output is then simply a projection of the corresponding state. The projection (P) gives the output variables (return value and output pointers of the tip call): 
\[
OV_s = P_s(V_s): OV_m = P_m(V_m).
\]

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/projections.png}
    \caption{The outputs \(O_m, O_s\) represented as projections of the total formal specification and model states.}
    \label{projections}
\end{figure}

With the introduced definitions, the following formulas define the outputs:
\[
O_s = \{P_s(V_s): \forall V_s F_s(V_s)\} \\
O_m = \{P_m(V_m) : \forall V_m F_m(V_m)\}
\]
Then, the model exhibits all the specified behaviours (and may be some more) iff \(O_s \in O_m\), or: 
\[
\forall V_s : F_s(V_s) \exists V_m : F_m (V_m) \cap P_s(V_s) = P_m(V_m)
\]
This formula in general is too difficult for the existing SMT solvers because of the quantifier. However, due to similarities in the model and the specification it is easy to find appropriate values for \(V_m\) --- they happen to be some values from Vs. Let us call this assignment \(A\). \(A\) by construction satisfies the equality \(P_s(V_s)=P_m(A(V_s))\). So the condition becomes:
\[
\forall V_s : F_s(V_s) \rightarrow F_m(A(V_s))
\]
  --- a typical validity query, which we pass  to an SMT solver.

This validation criteria guarantees that the formally verified implementation will never do anything that the model did not explore during verification. So each possible behaviour of the real execution (and may be some extra) was checked during the SE, and it is 1) allowed by the specification and 2) does not cause the application code to crash.

\subsection{Algorithmic Requirements}
\label{sec:algorithmic-requirements}

For a NF to be verifiable by Vigor, it must fulfill several requirements. These are fully aligned with best practices in developing NFs.

First, the application must contain no deep recursion and only use well structured loops. For long or unbounded loops, a developer must supply an invariant. Most NFs naturally satisfy this requirement due to performance considerations.

Second, program state must be encapsulated in a set of formally specified components (e.g., port allocator) or data structures (e.g., hashtable). Vigor uses the component API as a boundary between the code that is automatically verified using SE and the code that is manually verified.% (Figure~\ref{fig:pie}).

Third, there must be a well defined OS interface. As it is infeasible to symbolically execute the whole OS, the SEE stops at the OS interface boundary. Behind this interface we trust the correct implementation and we use symbolic models of the OS API functions for SE. It is conceivable to use a verified OS kernel, like SeL4~\cite{klein2009sel4}, and thus have a fully verified stack.

Finally, the NF must implement a classic event loop; most NFs already do so. For Vigor specifically, there must be a single infinite loop, with each iteration dedicated to processing a single event (an incoming packet). Within each iteration, the sequence of interactions between the code and the library components and the OS must be relatively short --- the complexity of models and the amount of symbolic state are often proportional to the length of the API call sequence it needs to support.

\subsection{Language Restrictions}
\label{sec:language-restrictions}

When writing NFs to be verified with Vigor, developers are free to use C. However, we impose a few restrictions that ensure good isolation between application code and the library components: We require that the component-related memory be accessed only through the library API. There should be no pointer comparisons except for equality --- the C99 standard anyway prohibits most pointer comparisons, and we extend this restriction in order to better isolate the component implementation from the code that uses it. Finally, pointer accesses must be both domain-limited and time-bounded: certain pointers may access only predefined range (i.e. a single unit of the declared type) and only on a restricted segment of the execution, between the corresponding API calls.

These restrictions do not apply to the formally verified component library.

% \subsection{Component Design Choice}

% A component designer faces a choice between conciseness and performance. Low expressive power of C forces the component developer to implement an abstraction either through typeless pointers or macros. Pointer implementation introduces an extra dereference, which may hurt performance. Macros are a zero cost abstraction. They perform better as they avoid extra pointer dereference and inline-friendly. Unfortunately, they hurt conciseness and readability of the code. In the long stretch a generics language extension such as C++99 templates could alleviate this trade-off.

\section{Implementation}
\label{sec:implementation}

In this section we describe the prototypes that embody the ideas presented in the previous section. We then use these prototypes to evaluate Vigor on the development of a NAT middlebox, described in \S\ref{sec:evaluation}.

\subsection{Vigor Tool Chain}

To assess the practicality of Vigor technique, we implemented the three tools described in the section~\ref{sec:our-approach}. The SEE is responsible for the application verification, the Formal Verifier checks correctness of the components’ implementation in the library, and the Validator ensures the two verifiers agree on the component APIs used.

\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE Klee~\cite{cadar2008klee}. By default Klee terminates execution paths that are too long or take too much memory. We disabled this feature to ensure it explores all the paths. We also extended KLee with three new features:
Partial loop invariant support. The analysis finds automatically the variables that may change in the loop, and havocs them. Currently, a developer have to manually insert the assertions and assumptions according to the loop invariant definition.
Dynamic pointer access control. The patch provides primitive intrinsics that allow a component developer to open and close access to a pointer between the API calls.
API call sequence dump. This patch summarizes each control path in the program extracting the component API calls and dumps them for the Validator.

\paragraph{Formal verifier} Our verifier is essentially VeriFast~\cite{jacobs2008verifast} extended with the ability to export the abstract state (\(O_s\)) for the Validator. Specifically we export the accumulated conditions for the output variables at the end of the API calls sequence.

\paragraph{Validator} We wrote this third tool completely from scratch. Validator independently analyzes each API calls sequence. We exploit the data parallelism using GNU parallel tool~\cite{Tange2011a}. For each API call sequence in the Klee dump, Validator first transforms it into VeriFast task. A VeriFast task is essentially a C source file, containing the sequence of calls. The Validator enriches the sequence with:
\begin{itemize}
    \item Symbolic variables used in the sequence.
    \item The constraints that form the sequence context.
    \item Lemmas to help formal verification.
\end{itemize}
Next, it runs VeriFast to check that the specification permits this particular sequence of calls under the provided constraints. Finally, the Validator imports the specification output state \(O_s\) and checks the inclusion of the specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3} SMT solver (see Section~\ref{sec:model-validation-criteria}).

Figure~\ref{fig:validator} shows the data flow between the Validator elements.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/validator.png}
    \caption{The internal Validator dataflow.}
    \label{fig:validator}
\end{figure}

\subsection{Verified Component Library}

For our Vigor library prototype, we implemented four stateful components. These are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation): A hash map keeps current flows, and allows lookup by two keys corresponding to the packets coming from internal and external interfaces; the Port allocator is responsible for allocating new ports and expiring inactive entries; the Batcher allows batch processing, keeping outbound packets until the NIC is ready to send them; and the Static Array instances keep various configuration values. We are careful in how we abstract components like the Static Array, because NFs are expected to access them on hot paths, and so it is important that the data reside in the processor cache.

\paragraph{Hash Map}

We implement a two-key hash map using three parts: integer indexed array of flows and two single-key hash maps. The array stores flows together with the keys. The hash maps internally are fixed size arrays of tuples $<$busy-bit, key-hash, key-ptr, flow-index$>$:
\begin{itemize}
    \item Key-hash is the value of the hash function for the key.
    \item Key-ptr is the pointer to the full key for precise comparison.
    \item Value-index is the index in the array of values.
    \item Busy-bit flags the tuples that are not garbage.
\end{itemize}

Flow insertion linearly searches a free cell (with busy-bit = 0) from an index equal to the hash of the corresponding key. To insert a flow, user must also supply an unused index. We use port allocator to allocate these indexes.

Flow search and removal also starts from the hash-induced index looks for a cell with the same hash, with confirmation by a total equality of the stored and searched keys.

This implementation is not efficient for some cases. For instance, if the flow is not present in the table, the algorithm will traverse the whole table and the search time will be proportional to the capacity.

To address this inefficiency, we optimize the single-key maps. We add another bit into the tuple. These bits group cells into hash-collision chains, and the map searches for a flow only across that chain, not across the entire table. Unfortunately, we did not yet verify the optimized implementation. We estimate re-verification to take about half of the initial verification effort (see section~\ref{sec:efficiency})

A hash map is a widely used data structure. It allows to store and retrieve entries by arbitrary complex keys. Such functions as L2 switching, ACL filtering, IPS, etc. all use hash tables.

\paragraph{Port Allocator}

Port allocator keeps two chains \emph{alloc} and \emph{free} packed into an array. The size of the array equal to the port range available to allocator, and each cell corresponds to a port to be allocated. \emph{Alloc} chain is a double linked list. It contains all the allocated ports with last access timestamps sorted by these timestamps. \emph{Free} chain is a single linked list. It keeps track of the available ports.

All operations take constant time, and only manipulate the cells of the array. Allocate operation moves head of the free chain to the head of the alloc chain. Access moves the element from the middle of alloc chain to the head. Expire takes the tail element from the alloc chain and moves it to the head of the free chain.

The port allocator is essentially a number allocator from a predefined range. Constant time allocation and expiration makes it a basic component for any application that manages resources in run time, such as L2 switch or L3 router.

\paragraph{Batcher}

Batcher is an array with a counter. It allows to store elements one by one (up to a preset capacity), and to release them all at once. Almost every DPDK example uses a batcher. It is necessary to achieve high throughput, because NICs work better on batches, rather than on individual packets.

\paragraph{Static Array}

Every DPDK example uses multiple static arrays for various configuration parameters, counters and other statistics. Our static array is a thin wrapper around a regular C array. This wrapper adds zero overhead, but enables our tool-chain to control access to the array and to substitute the whole array with just a couple of bytes during SE.

%===================

\section{Evaluation}

\label{sec:evaluation}

In this section we answer three questions:
\begin{enumerate}
    \item Is our approach effective at verifying real-world stateful NFs? In \S\ref{sec:effectiveness} we provide a qualitative answer to this question for a DPDK-based NAT middlebox.
    \item How efficient is our approach in terms of human and compute effort (\S\ref{sec:efficiency})?
    \item Is the performance of the verified NAT box good enough for practical real-world use by sysadmins and network operators (\S\ref{sec:performance})?
    %\item What are the trade-offs inherent to Vigor, and how do they impact verifiability and performance? In \S\ref{sec:tradeoffs} we give the reader a deeper understanding on how the various parts of the technique contribute to its scalability and cost.
\end{enumerate}

As a proof of concept, we implemented and verified \vignat, a NAT middlebox based on the DPDK framework~\cite{intel2014data}. We use NAT to refer specifically to Dynamic Source Network Address-Port Translation~\cite{rfc3022}. We reused boilerplate code from the DPDK L3 forwarding example. \vignat has 1,175 LOC, and it links with the Vigor library that has an additional 941 LOC. \vignat supports TCP and UDP packet forwarding with dynamic source address--port translation between multiple internal and one external interfaces. It supports connection initiation from an internal interface and refuses connections from the external one. It supports concurrent flows up to a configurable maximum number; flows expire after a configurable period of inactivity.

% The biggest difference of our application code from a usual code you can see in network applications is array access. To get an element from our Static Array component, due to our memory ownership technique, a developer have to explicitly borrow it through API call and then return the borrowed element to access another one.

Our evaluation uses two versions of this NAT box. During our experiments we discovered a silly performance bug in our hash table implementation that led to excessive hash collisions under certain circumstances. We fixed it, but did not complete the re-verification by the submission deadline. We therefore report results for both \vignata (100\% verified but slow) and \vignatb (verified, except for the hash table).

%---------------------------
\subsection{Vigor's Effectiveness}
\label{sec:effectiveness}

The ultimate goal of our work is to give operators the confidence that software-based NFs will not undermine the robustness and security of their network. It took about 12 minutes on a laptop for Vigor to prove that \vignat is free of crashes, free of memory leaks, free of memory overflows, and free of integer overflows. We verified \vignat only for one internal and one external interface, but do not foresee any challenges in doing so for multiple interfaces.

Beyond merely providing a binary yes/no answer to the question of whether an NF has a given property or not, Vigor also helps developers find those corner cases that impede the proof. There are three classes of bugs that Vigor uncovers in the verification process, and as the citations show, these appeared in the released Open vSwitch and DPDK code:
\begin{enumerate}

    \item Security vulnerabilities~\cite{bug:ovs-uaf, bug:dpdk-uaf1, bug:dpdk-uaf2, bug:ovs-faf, bug:ovs-overflow, bug:dpdk-overflow1, bug:dpdk-overflow2, bug:ovs-int-overflow, bug:dpdk-int-overflow} in the form of memory overflows or integer overflows. These can enable an attacker to gain remote control of the middlebox.

    \item Crashes~\cite{bug:ovs-0ptr, bug:dpdk-0ptr1, bug:dpdk-0ptr2, bug:dpdk-0ptr3} destabilize the network. Furthermore, many of these can be triggered by a remote attacker with a carefully crafted packet (e.g., trigger a bad pointer dereference or a division by zero), thereby opening up the middlebox to denial-of-service attacks.

    \item Memory leaks~\cite{bug:ovs-mem-leak1, bug:ovs-mem-leak2, bug:ovs-mem-leak3, bug:ovs-mem-leak4, bug:dpdk-mem-leak1, bug:dpdk-mem-leak2} are often exploited for denial-of-service by exhausting the middlebox's memory. Vigor's memory ownership model makes them both easy to find and to debug.

\end{enumerate}

Vigor does not support today proving concurrency properties or higher level semantic properties. Nevertheless, we consider the verifiable absence of low-level flaws to be a solid step toward complete verification.

%---------------------------
\subsection{Efficiency of Using Vigor}
\label{sec:efficiency}

A key question when using Vigor is how much longer does it take a developer to produce verified code vs. non-verified, and how much harder is it.  Another question is how much effort it takes to develop and verify the component library, but that effort gets amortized across many uses and is therefore less important.

\vignat's 1,175 LOC require about 30 lines of developer annotations (3\% overhead) in the form of loop invariants and loop labels like \code{LOOP(1)} in \S\ref{sec:solution-overview}). Static analysis can infer all the other information needed for the proofs. In our current Vigor prototype we have not implemented this static analysis, and so our \vignat prototype has 176 lines of annotations (15\% overhead)in the form of C structure layout description, conditional compilation constructs, etc. that make up for simple but unimplemented analyses in Vigor.

Our estimate is that it would take an NF developer about 1 week to develop and verify \vignat based on reading this paper (in particular \S\ref{sec:algorithmic-requirements} and \S\ref{sec:language-restrictions}). Our own development time is less representative, both because of our understanding of the system and the fact that we co-developed Vigor with \vignat.

Once the code is developed and annotated, verification is fast. On an Ubuntu 16.04 Linux laptop with an Intel I7-6600U CPU and 16GB LPDDR3 RAM, exhaustive SE of \vignat takes 2 minutes and 650 MB of RAM, and the Validator takes 10 minutes to validate the 700 call sequences that result from SE. Verification of the library's correctness proof takes 15 seconds.

\medskip

Developing and verifying the library took about 5 person-weeks, not counting our learning curve and various setbacks. Detailed data appears in Appendix B.

Table~\ref{tab:loc} shows the size of each of the library's four components.  As expected, the formal specification and proof dwarf by a 10-20$\times$ margin the implementation.  This is why we consider Vigor's design---using SE for NF code on the one hand, and heavy verification for the reusable library on the other hand---to provide a favorable balance for real-world NF development.  The symbolic model and the validator plug-in are more concise than the component implementation, which makes it feasible to provide several of them for each component, to suit different usage patterns.

\begin{table*}[h!]
\centering
\caption{Size (in LOC) of the parts required to develop and prove each  library component and data structure.}
\label{tab:loc}
\begin{tabular}{lllllll}
\hline
  Component             & Interface & Implementation & Formal spec & Symbolic model & Proof & Validator plug-in \\ \hline
{\bf Hash Map}        & 29         & 688            & 596                  & 281            & 4,078         & 450              \\
{\bf Port Allocator} & 22         & 210            & 310                  & 95             & 4,439         & 129              \\
{\bf Static Array}   & 8          & 15             & 31                   & 145            & 155          & 39               \\
{\bf Batcher}        & 22         & 28             & 27                   & 60             & 65           & 23               \\ \hline
\end{tabular}
\end{table*}

%---------------------------
\subsection{\vignat Performance}
\label{sec:performance}

In this section we address the question of whether verifiability comes at the price of performance. We evaluate \vignat's forwarding throughput and latency, comparing it to a NAT based on  NetFilter~\cite{boye2013netfilter}, which is widespread and used in all the Linux-based home and small-enterprise routers. We set up the NetFilter masquerade rules as shown in Appendix A, and we tune the Linux kernel for optimal performance as described in~\cite{nf-testing}. We label it {\em LinuxNAT} in our plots.

As mentioned earlier, we report on two versions of our NAT: \vignata has a performance bug in the hash table, whereas \vignatb does not, but its hash table is not yet verified. We configure both to have a maximum capacity of 61,000 flows. We use three identical machines, each a 32-core Intel Xeon E5-2667 v2 @ 3.30GHz with 32GB RAM with i350 1 Gbps DPDK-compatible NICs.

% The throughput results are stable, so we do not collect statistical significance. However latency varies noticeably and we measure repeat the experiment ten times to reveal the variability.

%-------------------------
\paragraph{Throughput.}

We set up two nodes in the topology shown in Figure~\ref{fig:throughput-setup}, as per RFC 2647~\cite{rfc2544}: a tester machine connects to the middlebox over two interfaces. We use \code{DPDK-Pktgen}, a high-performance traffic generator; it sends packets through {\em outbound} and counts them on the {\em inbound} interface. It assigns each packet a port from a given range to generate the corresponding number of distinct flows. We run 40-second measurement intervals.

\begin{figure}[h!]
    \includegraphics[width=\columnwidth]{figures/throughput_setup.png}
    \caption{Network topology for throughput experiment.}
    \label{fig:throughput-setup}
\end{figure}

In Figure~\ref{fig:throughput-overall} we show the maximum packet rate achieved with a maximum of 1\% packet loss, for different numbers of concurrent flows. At the outset, \vignat outperforms LinuxNAT; this is in no way related to Vigor but rather is due to DPDK and to LinuxNAT being interrupt driven. % Our prototype is not optimized either, but the fact that it uses DPDK makes it much more efficient. %; compared to a trivial DPDK application that merely forwards packets, \vignat does not add

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/brpt.pdf}
    \caption{Throughput as a function of the number of concurrent flows, with maximum 1\% loss rate.}
    \label{fig:throughput-overall}
\end{figure}

\vignata does well up until $\approx$28K connections, at which point it collapses due to the performance bug. In contrast, \vignatb sticks to the zero-loss line up until $\approx$59K connections, at which point our 61K-entry hashtable experiences too many collisions to keep up---increasing the size of the hash table would move this point further to the right.

To zoom in on what happens around those two inflection points, we plot two throughput graphs. Figure~\ref{fig:throughput-27900} shows what happens for 27,800 concurrent flows: the hash collision bug in \vignata starts causing it to grind to a halt at specific send rates; this behavior gets rapidly worse as the number of concurrent flows increases. Figure~\ref{fig:throughput-59k} shows what happens for 59,000 concurrent flows: \vignatb's flow table fills to 95\% and thus slows down, starting to perform comparably to LinuxNAT on average, but with significant jitter. As the number of flows increases, \vignatb slows down further. % Reducing jitter is part of future work.

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/rt27900.pdf}
    \caption{Packet rates for 27,900 concurrent flows.}
    \label{fig:throughput-27900}
\end{figure}

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/rt59k.pdf}
    \caption{Packet rates for 59,000 concurrent flows.}
    \label{fig:throughput-59k}
\end{figure}

%---------------
\paragraph{Latency}

To assess \vignat's latency characteristics, we use the network topology in Figure~\ref{fig:latency-setup}. We use separate ports on the tester for flow generation and latency sampling. We use again DPDK's \code{PktGen} for packet generation. We sample latency with netperf~\cite{jones1996netperf} \code{TCP\_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/latency_setup.png}
    \caption{Network topology for latency measurements.}
    \label{fig:latency-setup}
\end{figure}

We also wrote a trivial DPDK application, derived from the DPDK l2fwd example~\cite{dpdk-l2fwd} which only forwards packets. To eliminate the routing overhead, we replaced L2 routing with a simple MAC address swap. This application is stateless and performs only a couple writes to the packet header, so we consider it the lowest possible latency one could achieve with DPDK; we label it {\em Plain forwarding} in our plots.

Figure~\ref{fig:latency-plot} shows one-way latency as the number of concurrent connections through the middlebox increases. The performance issue in \vignata causes its latency to increase with the number of flows, but \vignatb stays flat. There is increasing jitter as the number of flows increases, and investigating this is part of future work (we believe it's due to DPDK). \vignat has on average 23\% higher latency than LinuxNAT, but nevertheless stays under 111~$\mathrm{\mu sec}$, providing sufficient performance for many settings. Note that DPDK is designed to sustain high throughput, so it uses batching aggressively, which hurts latency---\vignat introduces a mere 5\% overhead compared to {\em Plain forwarding}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/lat.pdf}
    \caption{Latency vs. number of concurrent flows.}
    \label{fig:latency-plot}
\end{figure}

%---------------
\bigskip

If the reader believes our assertion that we can verify \vignatb's hash table in a few person-days, then the results in this section imply that developing a verified NAT middlebox does not come at the expense of performance. NF developer effort is relatively low, which makes this a practical approach. We therefore conclude that Vigor can succeed in its objective of enabling developers to productively write software-based data plane NFs that simultaneously meet the highest standards of reliability, security, and performance.

\section{Limitations}

\label{sec:limitations}

Vigor's greatest limitation is that it is not fully automated yet. In particular, we expect NF developers have to write loop invariants for long and infinite loops. These loop invariants are essential to enabling scalable exhaustive SE with our modified KLEE. As mentioned before, coming up with these invariants for NF code is typically not difficult. Nevertheless, one of our immediate future work items is to leverage invariant induction techniques~\cite{perkins2004efficient, yang2004automatically} to automate at least part of this task.

Application developers may also encounter ``false positives'', i.e., Vigor fails to prove the desired property. As mentioned before, this can be due to the property indeed not holding, or because the symbolic models are too over-approximate. The NF developer can either develop their own more detailed model, or request the component library developers to provide a suitable model.

Besides the trusted computing base described in \S\ref{sec:assumptions}, Vigor relies on the NF developer to state correct invariants, as well as the library developers to correctly annotate the symbolic models with KLEE intrinsics (so that they produce correct traces). Finally, we trust the NF developer to not link their application with unverified components. If any of these assumptions fail to hold, Vigor cannot guarantee any properties.

With respect to our NAT use case, as already mentioned we have two versions, one of which is not fully verified (but will be shortly). Also, our NAT does not participate in any routing protocols; it needs an L3 router on each interface to dispatch the packets to the rest of the network. We supply the MAC addresses of these routers as a runtime argument to the NAT, and can not update it on-the-fly.

\section{Related Work}

\label{sec:related-work}

%In this section we review some of the similar projects undertaken in the past.

Dobrescu et al.~\cite{dobrescu2014software} analyzed modular NFs. They managed to analyze many stateless Click elements using exhaustive SE, and exploiting modularity and low cyclomatic complexity of such applications. However, their original approach is very limited in verification of stateful elements. In addition to crash freedom they managed to verify bounded execution and filtering properties. Doing this in Vigor for stateful NFs is future work.

SeL4~\cite{klein2009sel4}, CompCert~\cite{stewart2015compositional}, and FSCQ~\cite{chen2015using} show how to do functional verification of practical systems. All of them prove the semantics of the system. Unfortunately they all require the use of a high-level programming language and deep expertise in the verification domain, which is unacceptable for our problem.

Beringer et al~\cite{beringer2015verified} applied formal verification to an OpenSSL implementation, proving its functional correctness and cryptographic properties. They used certified tool chain, reducing trusted code base just to just a small CoQ proof checker core (~8k LOC). While it paves the road for formal verification towards real systems, it is not yet suitable as a common development practice.

SymNet~\cite{stoenescu2013symnet} uses SE to reason about whole network configuration with per-flow state. They do not verify the middle box implementations, instead parse the configuration files and use special models in place of the components. Their work naturally complements ours.

IronFleet~\cite{hawblitzel2015ironfleet} also aims for a verification of applications distributed over a network. They use a different approach, though. Habwlitzel et al develop IronFleet from scratch in Dafny --- a high level language with built-in verification support.

Others checked existing distributed systems. Pedrosa~\cite{pedrosa2015analysis} and Guerraoui~\cite{guerraoui2011model} explored the global state tree down to a certain depth. Guerrauoi et al. used a similar approach to ours: they made reasonable assumptions that simplify verification and validated the ones that turned out important.

CEGAR~\cite{clarke2000counterexample} is similar to our technique in the spirit of optimizing out the high effort/low return work. It only refines the models when given a clear evidence of the their insufficiency. Clarke at al. demonstrated a very generic approach. We focus on the networking domain, and exploit the features inaccessible for CEGAR, like the low number of data structures and simplicity of the application logic.

\section{Conclusion}

\label{sec:conclusion}

We presented a technique for proving properties of software NFs. We exploit the specifics of NFs to combine the benefits of symbolic execution with formal verification. Our experimental results demonstrate the practicality of the approach for the development of NAT box that is guaranteed to have no memory leaks, never crash, be free of certain security vulnerabilities, yet still performs on par with NetFilter.

{\footnotesize \bibliographystyle{acm}
\bibliography{bibliography}}

\section*{Appendix A: NetFilter rules}

{\tt \small
\begin{verbatim}
iptables -t nat -A POSTROUTING -o eth0 -j MASQUERADE
iptables -A FORWARD -i eth0 -o eth1 -m state \
    --state NEW,RELATED,ESTABLISHED -j ACCEPT
iptables -A FORWARD -i eth0 -o eht1 -j ACCEPT
\end{verbatim}
}

\section*{Appendix B: Person-hours to verify \vignat}

Table~\ref{tab:man-hours} shows the number of person-hours spent to

\begin{table*}[h!]
\centering
\caption{Developer person-hours spent on each component part in hours.}
\label{tab:man-hours}
\begin{tabular}{lllllll}
\hline
               & Interface design & Implementation & Formal spec & Symbolic model & Verification & Validator plug-in \\ \hline
Hash Map        & 2.5*       & 2*             & 4.5*                 & 5.5*           & 73.5         & 11.5             \\
Port Allocator & 1*         & 1.5*           & 3.5*                 & 1.5*           & 70           & 7.5              \\
Static Array   & 1.5        & 0.5            & 2.5                  & 2              & 6            & 2                \\
Batcher        & 1          & 0.2            & 0.5                  & 1              & 0.5          & 0.2              \\ \hline
\end{tabular}
\end{table*}

\end{document}

### 2016-10-24 doc/paper/paper.tex
Bugfix in the paper latex source.

@@
 For our Vigor library prototype, we implemented four stateful components.
-These are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation): A hash map keeps current flows, and allows lookup by two keys corresponding to the packets coming from internal and external interfaces; the Port allocator is responsible for allocating new ports and expiring inactive entries; the Batcher allows batch processing, keeping outbound packets until the NIC is ready to send them; and the Static Array instances keep various configuration values.
+These are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation}: A hash map keeps current flows, and allows lookup by two keys corresponding to the packets coming from internal and external interfaces; the Port allocator is responsible for allocating new ports and expiring inactive entries; the Batcher allows batch processing, keeping outbound packets until the NIC is ready to send them; and the Static Array instances keep various configuration values.
 We are careful in how we abstract components like the Static Array, because NFs are expected to access them on hot paths, and so it is important that the data reside in the processor cache.

### 2016-10-24 doc/paper/paper.tex
paper typesetting: inline code snippets; hyperlinks.

@@
 \usepackage{changepage}
-
+\usepackage{hyperref}
 \definecolor{mygreen}{rgb}{0.1,0.5,0.1}
 \definecolor{myblue}{rgb}{0.1,0.1,0.7}
+
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black}
+}
 
@@
 
-\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\code}[1]{\lstinline{#1}}
 \newcommand{\vignat}{{\sc VigNAT}\xspace}
@@
 \begin{itemize}
-    \item {\bf Speculative data plane verification} (SDSE), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution.
+    \item {\bf Speculative data plane verification} (SDPV), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution.
 % It uses formal verification for data structures.
@@
 This approach postpones high-effort/low-return work and discards it when it turns out to be irrelevant.
-    \item {\bf The Vigor library of verified components} along with a tool chain that makes SDSE available to those who write software NFs in C/C++.
+    \item {\bf The Vigor library of verified components} along with a tool chain
+      that makes SDPV available to those who write software NFs in C/C++.
       The library consists of data structures (Hash Map, Batcher, Static Array) and a Port Allocator that are common for software NF.
@@
 
-In order to absorb bursts, the code employs a ring buffer.
-The interface to the network is: {\em bool receive\_packet(struct packet* dst)}, which non-blockingly reads an inbound packet and stores it at {\em dst}, returning success or failure; {\em bool can\_send\_packet()}, which checks if a new packet can be sent; and {\em void send\_packet(struct packet* src)}, which sends the packet stored at {\em src}.
+In order to absorb bursts, the code employs a ring buffer. The interface to the
+network is: \code{bool receive_packet(struct packet* dst)}, which
+non-blockingly reads an inbound packet and stores it at \code{dst}, returning success or failure; \code{bool can_send_packet()}, which checks if a new packet can be sent; and \code{void send_packet(struct packet* src)}, which sends the packet stored at \code{src}.
 The application follows common practice: it consists of a main event loop, it has no other complex loops, and it uses a ring, a popular data structure in network programming.
@@
 To perform SE, it is necessary to provide symbolic models for the three network calls.
-Something simple suffices, such as \code{receive\_packet} generating a new symbolic packet in \code{dst}; \code{can\_send\_packet} forking program state into two states, one for each possible return value; and \code{send\_packet} doing nothing.
+Something simple suffices, such as \code{receive_packet} generating a new symbolic packet in \code{dst}; \code{can_send_packet} forking program state into two states, one for each possible return value; and \code{send_packet} doing nothing.
 
@@
 
-For our example, the library provides a \code{ring} with a simple interface: {\em bool ring\_empty()} checks whether there is a packet in the ring, {\em bool ring\_full()} checks whether the ring has run out of space, {\em void ring\_push\_back(struct packet* src)} adds packet {\em src} to the ring, and {\em void ring\_pop\_front(struct packet* dst)} removes a packet from the ring and places it in {\em dst}.
+For our example, the library provides a \code{ring} with a simple interface: \code{bool ring_empty()} checks whether there is a packet in the ring, \code{bool ring_full()} checks whether the ring has run out of space, \code{void ring_push_back(struct packet* src)} adds packet \code{src} to the ring, and \code{void ring_pop_front(struct packet* dst)} removes a packet from the ring and places it in \code{dst}.
 We then write a symbolic model for the \code{ring} and provide it to the SEE instead of the actual implementation.
@@
 
-%Note, that the \code{ring\_pop} model does not even check whether it is called with an empty ring or not.
+%Note, that the \code{ring_pop} model does not even check whether it is called with an empty ring or not.
 The example application makes sure to call pop only if the ring is not empty.
@@
 %This model clearly overapproximates a ``ring'', e.g.
-\code{ring\_empty} may return true even right after \code{ring\_push}.
+\code{ring_empty} may return true even right after \code{ring_push}.
 This means, it may have caused false positives, but as long as it does not, no need to complicate it.
@@
 While this may be indicative of a bug in the NF, it could also be simply because the model is too approximate.
-For example, in \S\ref{sec:solution-overview} a model of \code{ring\_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring\_empty} to return false right after pushing an element.
-A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring\_empty} return true only when this counter is zero.
+For example, in \S\ref{sec:solution-overview} a model of \code{ring_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring_empty} to return false right after pushing an element.
+A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring_empty} return true only when this counter is zero.
 The above NF code would not crash with this model; on the downside, the more stateful model makes SE take longer.
@@
 % a number of speculative assumptions that we believe will hold for the real implementations.
-In the example above some of these assumptions are: ``the \code{ring\_empty} function may return either 0 or 1'', ``if the \code{ring\_pop\_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring\_push\_back} never crashes''.
+In the example above some of these assumptions are: ``the \code{ring_empty} function may return either 0 or 1'', ``if the \code{ring_pop_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring_push_back} never crashes''.
 
@@
 We use again DPDK's \code{PktGen} for packet generation.
-We sample latency with netperf~\cite{jones1996netperf} \code{TCP\_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.
+We sample latency with netperf~\cite{jones1996netperf} \code{TCP_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.
 

### 2016-10-24 doc/paper/paper.tex
Continuation of the paper example (draft, of course).

@@
 
+Suppose we had an overapproximate model:
+\begin{lstlisting}
+void ring_pop_front(struct ring* r,
+                    struct packet* p) {
+  FILL_SYMBOLIC(p, sizeof(struct packet),
+                "popped_packet");
+}
+\end{lstlisting}
+Then the \code{ASSERT} would fail.
+
 Now exhaustive SE of the code takes less than 100 millisec with our modified KLEE, regardless of the size of \code{CAP}.
@@
 
-The Vigor library comes with a formal specification and a formal proof that the implementation satisfies that specification (we show neither of them here, due to space constraints).
+The Vigor library comes with a formal specification and a formal proof that the implementation satisfies that specification.
 In other words, there is a formal spec for \code{ring} and a proof that the \code{ring} implementation satisfies the spec.
@@
 
+Here is the example of the proven specification for the function
+\code{ring_pop_front}. We do not include the proof of this specification due to
+space constraints:
+
+\begin{lstlisting}
+void ring_pop_front(struct ring* r,
+                    struct packet* p);
+/*@ requires ringp(r, ?lst) &*&
+             lst != nil &*&
+             p->port |-> _ &*&
+             struct_packet_padding(p); @*/
+/*@ ensures ringp(r, tail(lst)) &*&
+            pktp(p, head(lst)); @*/
+\end{lstlisting}
+
+Note \code{pktp(p, head(lst));} predicate includes the \code{p->port != 9} property.
+
 Vigor must also validate that the model used for SE matches the library spec.
@@
 Details on how it does this appear in \S\ref{sec:design}.
+
+To validate the model fit for the particular use case, SEE dumps
+all the call sequences (30 in this case) that may happen during the execution.
+Vigor checks whether the symbolic model enumerated all the possible
+outputs for each sequence.
+
+You can see the example call sequence below. It starts with
+\code{loop_invariant_produce} that marks the beginning of a loop iteration. It
+continues then with calls to \code{ring_full} \code{ring_empty}, noting that
+both return \code{false}. The seqeunce ends with a call to
+\code{ring_pop_front}, formally specified above. The last line checks that the
+value returned by model (\code{arg2.port != 9}) covers all the cases prescribed
+by the specification.
+
+\begin{lstlisting}
+int port_0;
+struct ring* arg1;
+struct packet arg2;
+//@ assume((9 == port_0));
+arg2.port = port_0;
+loop_invariant_produce(&(arg1));
+bool ret1 = ring_full(arg1);
+//@ assume(ret1 == false);
+bool ret2 = ring_empty(arg1);
+//@ assume(ret2 == false);
+
+ring_pop_front(arg1, &(arg2));
+//@ open pktp(&(arg2), _);
+/*@ assert((false == (9 == arg2.port))); @*/
+\end{lstlisting}
+
+Suppose we had an underapproximate model:
+\begin{lstlisting}
+void ring_pop_front(struct ring* r,
+                    struct packet* p) {
+  p->port = 0;
+}
+\end{lstlisting}
+then the last line in the call sequence check will look like:
+\begin{lstlisting}
+/*@ assert(0 == arg2.port); @*/
+\end{lstlisting}
+And the formal verification will fail, because specification admits a wider
+range of possible outputs.
 

### 2016-10-26 doc/paper/paper.tex
Some corrections for the example section in the paper.

@@
 Alas, symbolic execution would never terminate.
-The reason is ``path explosion:'' there is a large (often infinite) number of paths to explore, so the SEE cannot finish.
+The reason is ``path explosion'': there is a large (often infinite) number of paths to explore, so the SEE cannot finish.
 Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, three \code{if} conditions inside an infinite loop give rise to an infinite number of paths.
@@
 
+Now exhaustive SE of the code takes less than 100 millisec with our modified
+KLEE, regardless of the size of \code{CAP}.  The fact that the code does not
+crash on any path amounts to a proof of crash freedom. Or does it? The validity
+of the proof depends on (1)  our library implementation of \code{ring} not
+crashing; and (2) the symbolic model of \code{ring} accurately capturing how the
+\code{ring} implementation works ({\em Step 3}).
+
 Suppose we had an overapproximate model:
@@
 \end{lstlisting}
-Then the \code{ASSERT} would fail.
-
-Now exhaustive SE of the code takes less than 100 millisec with our modified KLEE, regardless of the size of \code{CAP}.
-The fact that the code does not crash on any path amounts to a proof of crash freedom.
-Or does it?
-The validity of the proof depends on (1)  our library implementation of \code{ring} not crashing; and (2) the symbolic model of \code{ring} accurately capturing how the \code{ring} implementation works.
+Then the \code{ASSERT} from the previous listing would fail.
 
@@
 Here is the example of the proven specification for the function
-\code{ring_pop_front}. We do not include the proof of this specification due to
-space constraints:
+\code{ring_pop_front}. We do not include the certified implementation of this
+specification due to space constraints:
 
@@
 
-Vigor must also validate that the model used for SE matches the library spec.
+{\bf Step 3: Validate the model used for SE.}
+The model must match the library spec.
 Some symbolic models work well for some code, while others work well for other code, so the Vigor library comes with {\em multiple} simple, overapproximating models for each component, and Vigor automatically chooses combinations of models for symbolically executing the NF code.
@@
 To validate the model fit for the particular use case, SEE dumps
-all the call sequences (30 in this case) that may happen during the execution.
+all the call sequences (30 in this example) that may happen during the execution.
 Vigor checks whether the symbolic model enumerated all the possible
@@
 
-{\bf Step 3: Link the NF code with the verified library} of components. The resulting executable is guaranteed to be free of crashes, subject to the assumptions listed in~\S\ref{sec:limitations}.
+{\bf Step 4: Link the NF code with the verified library} of components. The resulting executable is guaranteed to be free of crashes, subject to the assumptions listed in~\S\ref{sec:limitations}.
 

### 2016-10-26 doc/paper/paper.tex
Put the new example into the paper.

@@
 
-% \begin{lstlisting}
-% bool ring_full() {return SYMBOLIC("full");}
-
-% bool ring_empty() {return SYMBOLIC("empty");}
-
-% void ring_push_back(struct packet* p) {
-%   ASSERT(p->port != 9); //Extra check.
-% }
-
-% void ring_pop_front(struct packet* p) {
-%   FILL_SYMBOLIC(p, sizeof(struct packet), 
-%                 "popped_packet");
-%   ASSUME(p->port != 9);
-% }
-% \end{lstlisting}
+\begin{lstlisting}
+int main(int argc, char** argv) {
+  struct packet p;
+  struct ring *r = ring_create(CAP);
+  if (!r) return 1;
+  while(LOOP(1))
+  {
+    loop_iteration_begin(&r);
+    if (!ring_full(r))
+      if (receive_packet(&p) && p.port != 9)
+        ring_push_back(r, &p);
+    if (!ring_empty(r) && can_send_packet()) {
+      ring_pop_front(r, &p);
+      VIGOR_CHECK(p.port != 9);
+      send_packet(&p);
+    }
+    loop_iteration_end(&r);
+  }
+  return 0;
+}
+\end{lstlisting}
 
@@
 #define CAP 512
-
-int main() {
-  struct packet p;
-  while(LOOP(1)) {
-    if (!ring_full())
-      if (receive_packet(&p) && p.port != 9)
-        ring_push_back(&p);
-    if (!ring_empty() && can_send_packet()) {
-      ring_pop_front(&p);
-      ASSERT(p.port != 9);
-      send_packet(&p);
-    }
-  }
-  return 0;
+#define RING_CAPACITY 512
+
+/*@
+  fixpoint bool user_packet_property(packet p) {
+    switch(p) { case packet(port): return port != 9; }
+  }
+  @*/
+
+static bool packet_constraints(struct packet* p) {
+  return p->port != 9;
 }
@@
                     struct packet* p);
-/*@ requires ringp(r, ?lst) &*&
+/*@ requires ringp(r, ?lst,
+                   ?property, ?capacity) &*&
              lst != nil &*&
-             p->port |-> _ &*&
-             struct_packet_padding(p); @*/
-/*@ ensures ringp(r, tail(lst)) &*&
-            pktp(p, head(lst)); @*/
+             packetp(p, _); @*/
+/*@ ensures ringp(r, tail(lst),
+                  property, capacity) &*&
+            packetp(p, head(lst)) &*&
+            true == property(head(lst)); @*/
 \end{lstlisting}
 
-Note \code{pktp(p, head(lst));} predicate includes the \code{p->port != 9} property.
+Note that \code{property(head(lst));} in the post condition ensures \code{p->port != 9} property.
 
@@
 \begin{lstlisting}
-int port_0;
 struct ring* arg1;
 struct packet arg2;
-//@ assume((9 == port_0));
-arg2.port = port_0;
+
 loop_invariant_produce(&(arg1));
+/*@ open loop_invariant(_); @*/
 bool ret1 = ring_full(arg1);
@@
 //@ assume(ret2 == false);
+//@ close packetp(&(arg2), packet((&(arg2))->port));
 
 ring_pop_front(arg1, &(arg2));
-//@ open pktp(&(arg2), _);
+//@ open packetp(&(arg2), _);
+
 /*@ assert((false == (9 == arg2.port))); @*/

### 2016-10-26 doc/paper/example/model.c

@@ -4,8 +4,6 @@
 #include "loop-model.h"
 #include "user-params.h"
 
-#define CAP RING_CAPACITY
-
 int main(int argc, char** argv) {
   struct packet p;
   struct ring *r = ring_create(CAP);

### 2016-10-27 doc/paper/paper.tex
Missing reference; wrap code listings into figures with captions.

@@
     tabsize=2,
-    otherkeywords={bool}
+    otherkeywords={bool},
+    numbers=left,
+    numbersep=1pt,
+    numberstyle=\tiny\color{gray}
 }
@@
 \subsection*{Abstract}
-The objective of this work is to enable developers to productively write software-based data plane network functions that simultaneously meet the highest standards of reliability, security, and performance.
-We present a verification approach for formally proving properties of stateful network functions, in a way that is practical both in terms of developer effort and the performance of the resulting data plane software.
-Unlike alternative approaches that require drastic changes in the way network functions are developed, our approach allows the use of C/C++ together with popular frameworks like DPDK.
-To the best of our knowledge, our work is the first to both be practical and formally verify stateful network functions.
-We introduce a new ``speculative'' form of symbolic execution, which achieves scalability to real-world network functions by leveraging insights related to the structure and cyclomatic complexity of data plane software.
-This technique dovetails with a new library of verified components that are commonly used in network functions.
-We illustrate our approach by developing a middlebox that is verified to never crash, to not leak memory, be free of several classes of security vulnerabilities, and whose performance is on par with other NATs.
+The objective of this work is to enable developers to productively write
+software-based data plane network functions that simultaneously meet the highest
+standards of reliability, security, and performance. We present a verification
+approach for formally proving properties of stateful network functions, in a way
+that is practical both in terms of developer effort and the performance of the
+resulting data plane software. Unlike alternative approaches that require
+drastic changes in the way network functions are developed, our approach allows
+the use of C/C++ together with popular frameworks like DPDK. To the best of our
+knowledge, our work is the first to both be practical and formally verify
+stateful network functions. We introduce a new ``speculative'' form of symbolic
+execution, which achieves scalability to real-world network functions by
+leveraging insights related to the structure and cyclomatic complexity of data
+plane software. This technique dovetails with a new library of verified
+components that are commonly used in network functions. We illustrate our
+approach by developing a NAT box that is verified to never crash, to not leak
+memory, be free of several classes of security vulnerabilities, and whose
+performance is on par with other NATs.
 %\footnote{Disclaimer: we use acronym NAT in this paper to refer to a more narrow kind - Dynamic Source NAPT (network address-port translation~\cite{rfc3022}).}.
@@
 They can be formally verified a priori, and their proofs then automatically leveraged for proving properties of the code that uses them.
-    \item {\bf Hybrid verification}: While typically one picks a verification technique and applies it throughout, any {\em one} technique works well for {\em some} of the code and not for other.
+    \item {\bf Hybrid verification}: While typically one picks a verification
+      technique and applies it throughout, no {\em single} technique works well for {\em all} of the code.
 However, data plane software offers us the opportunity to advantageously employ symbolic execution for high-level code (which is mostly stateless and has short execution paths) and verification based on separation logic to the library of components/data structures (which is highly stateful).
@@
 
+\begin{figure}[h!]
 \begin{lstlisting}
@@
 }
-
 \end{lstlisting}
+  \caption{Original implementation of the Drop protocol}
+  \label{lst:orig}
+\end{figure}
 
@@
 These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.
-We added to KLEE support for havoc-ing~\cite{boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.
+We added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.
 
@@
 \end{lstlisting}
-
+  \caption{Vigor implementation of Drop protocol.}
+  \label{lst:vigor}
+\end{figure}
+
+For the symbolic execution and future validation, we need also to customize the
+general-purpose packet ring with the following definitions:
+
+\begin{figure}
 \begin{lstlisting}
 #define CAP 512
-#define RING_CAPACITY 512
+#define RING_CAPACITY CAP
 
@@
 \end{lstlisting}
+  \caption{Custom definitions capturing the desired property.}
+  \label{lst:customizations}
+\end{figure}
 
@@
 Suppose we had an overapproximate model:
+\begin{figure}[h!]
 \begin{lstlisting}
@@
 \end{lstlisting}
+  \caption{An extract from an overapproximate ring model.}
+  \label{lst:overapproximate}
+\end{figure}
 Then the \code{ASSERT} from the previous listing would fail.
@@
 \end{lstlisting}
+  \caption{A formal contract for the ring\_pop\_front function.}
+  \label{lst:contract}
+\end{figure}
 
@@
 \end{lstlisting}
+  \caption{A call sequence, produced by SEE.}
+  \label{lst:call-sequence}
+\end{figure}
 
 Suppose we had an underapproximate model:
+\begin{figure}[h!]
 \begin{lstlisting}
@@
 \end{lstlisting}
+  \caption{An underapproximate model fragment.}
+  \label{lst:underapproximate}
+\end{figure}
 then the last line in the call sequence check will look like:

### 2016-10-27 doc/paper/paper.tex
Fix the ring implementation in the original code version of the example.

@@
 This section motivates and explains at a high level what it's like to use Vigor when developing NFs.
-For the sake of illustration, say we wish to develop a trivial stateful NF that implements the discard protocol~\cite{rfc863}: an infinite loop receives packets from one interface, discards the ones that go to port 9 (``discard port''), and forwards the others through another interface.
+For the sake of illustration, say we wish to develop a trivial stateful NF that implements the discard protocol~\cite{rfc863}: an infinite loop receives packets from one interface, discards the ones that target port 9 (``discard port''), and forwards the others through another interface.
 
@@
   struct packet packets[CAP] = {};
-  int begin = 0, end = 0;
+  int begin = 0, queue_len = 0;
 
   while(1) {
-    if (end != (begin - 1) || 
-        !(end == CAP - 1 && begin == 0)) {
+    if (queue_len < CAP) {
+      int end = (begin + queue_len)%CAP;
       if (receive_packet(&packets[end]) &&
           packets[end].port != 9)
-        end = (end + 1)%CAP;
-    }
-    if (end != begin && can_send_packet()) {
+        queue_len += 1;
+    }
+    if (0 < queue_len) {
       send_packet(&packets[begin]);

### 2016-10-27 doc/paper/example/orig.c

@@ -5,15 +5,16 @@
 
 int main() {
   struct packet packets[CAP] = {};
-  int begin = 0, end = 0;
+  int begin = 0, queue_len = 0;
 
-  while(LOOP(1)) {
-    if (end != (begin - 1) || !(end == CAP - 1 && begin == 0)) {
-      if (receive_packet(&packets[end]) && packets[end].port != 9)
-        end = (end + 1)%CAP;
+  while(1) {
+    if (queue_len < CAP) {
+      int end = (begin + queue_len)%CAP;
+      if (receive_packet(&packets[end]) &&
+          packets[end].port != 9)
+        queue_len += 1;
     }
-    if (end != begin && can_send_packet()) {
-      ASSERT(packets[begin].port != 9);
+    if (0 < queue_len) {
       send_packet(&packets[begin]);
       begin = (begin + 1)%CAP;
     }

### 2016-10-27 doc/paper/paper.tex
Fix typos.

@@
 
-To verify that this NF does not crash, one could use the approach in Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution.
+To verify that this NF does not crash, one could use the approach of Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution.
 The symbolic execution (SE) technique entails using a symbolic execution engine (SEE) to explore all paths through the code and check whether any failure conditions hold --- in our case crash conditions.
@@
 
-One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the code above exhaustively --- once all paths are found to be crash-free, the crash-freedom proof is complete.
-To perform SE, it is necessary to provide symbolic models for the three network calls.
-Something simple suffices, such as \code{receive_packet} generating a new symbolic packet in \code{dst}; \code{can_send_packet} forking program state into two states, one for each possible return value; and \code{send_packet} doing nothing.
+One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the
+code above exhaustively --- once all paths are found to be crash-free, the
+crash-freedom proof is complete. To perform SE, it is necessary to provide
+symbolic models for the three network calls. Something simple suffices, such as
+\code{receive_packet} generating a new symbolic packet in \code{dst};
+\code{can_send_packet} forking program state into two states for busy and idle
+interface correspondingly; and \code{send_packet} doing nothing.
 
@@
 The reason is ``path explosion'': there is a large (often infinite) number of paths to explore, so the SEE cannot finish.
-Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, three \code{if} conditions inside an infinite loop give rise to an infinite number of paths.
-
-{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}.
-These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.
-We added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.
+Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, the infinite loop gives rise to an infinite number of paths.
+
+{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be
+simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These
+are annotations that inform the SEE about invariant properties of the loop, such
+as ``no element of \code{packets} between \code{begin} and \code{begin +
+  queue_len} has destination port equal to 9'', and are a good match for
+NF-style loops.  We added to KLEE support for havoc-ing~\cite{barnett2005boogie}
+automatically, to take advantage of such invariants and essentially reduce the
+infinite loop to a single abstract iteration that can be reasoned about in one
+pass.
 
@@
     \item \(Co\): The run output (\(Co = Fon\)) is the output of the last API function call in the sequence.
-    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments).
+    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals\_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments).
 For a pointer argument (which must point to a concrete address --- practical limitation), the pointee must be considered according to the \S\ref{sec:handling-pointers}.
@@
 
-\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE Klee~\cite{cadar2008klee}.
-By default Klee terminates execution paths that are too long or take too much memory.
+\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE KLEE~\cite{cadar2008klee}.
+By default KLEE terminates execution paths that are too long or take too much memory.
 We disabled this feature to ensure it explores all the paths.
-We also extended KLee with three new features:
+We also extended KLEE with three new features:
 Partial loop invariant support.
@@
 We exploit the data parallelism using GNU parallel tool~\cite{Tange2011a}.
-For each API call sequence in the Klee dump, Validator first transforms it into VeriFast task.
+For each API call sequence in the KLEE dump, Validator first transforms it into VeriFast task.
 A VeriFast task is essentially a C source file, containing the sequence of calls.

### 2016-10-27 doc/paper/paper.tex
Break down the longlines into paragraph in the LaTeX source.

@@
 
-Product development in the telecommunication industry has traditionally followed rigorous standards for quality, security, and protocol adherence, with the term carrier-grade designating network equipment that meets these high standards.
-While this model worked well in the past, it typically entailed reliance on proprietary hardware and long product cycles, which slowed progress.
-To meet the growing demand for new network functionality and faster deployment cycles, network service providers started turning to software.
-
-In recent years, new platforms emerged for developing network functionality in software~\cite{kohler2000click,bosshart2014p4,mccanne1993bsd,song2015unified,shah2003np,shahbaz2016pisces}, enabling the rapid prototyping and deployment of new routing protocols~\cite{sivaraman2016packet}, QoS policies~\cite{sivaraman2016programmable}, load balancing algorithms~\cite{zhang2013steering}, intrusion prevention systems~\cite{xing2014sdnips}, etc.
-The agility that software engenders determined even traditional hardware-based network device manufacturers to invest in software-based approaches~\cite{ciscodevelopssdn}.
-
-But this raises the question of how to develop carrier-grade network functions (NF) in software?
-Since NFs are often part of critical infrastructure, they must simultaneously be secure, reliable, and high-performance, i.e., be carrier-grade.
-The Achilles' heel of software has always been the (perceived) lack of reliability and security, particularly when it is written in C.
-This has led researchers to advocate using safe languages~\cite{clipsham2015safe, kourtis2015intelligent, fowler2014verified} or relatively heavyweight approaches~\cite{madhavapeddy2007melange, madhavapeddy2005splat, musuvathi2004model, alur2001verifying, ennals2004linear, ridge2008rigorous, qadir2015applying}.
-Yet many frameworks (e.g., P4~\cite{bosshart2014p4}, DPDK~\cite{intel2014data}, PF\_RING~\cite{pfring}, NetMap~\cite{rizzo2012revisiting}) and standalone implementations (Open vSwitch~\cite{pfaff2009extending}, NetFilter~\cite{boye2013netfilter}) use C.
+Product development in the telecommunication industry has traditionally followed
+rigorous standards for quality, security, and protocol adherence, with the term
+carrier-grade designating network equipment that meets these high standards.
+While this model worked well in the past, it typically entailed reliance on
+proprietary hardware and long product cycles, which slowed progress. To meet the
+growing demand for new network functionality and faster deployment cycles,
+network service providers started turning to software.
+
+In recent years, new platforms emerged for developing network functionality in
+software~\cite{kohler2000click,bosshart2014p4,mccanne1993bsd,song2015unified,shah2003np,shahbaz2016pisces},
+enabling the rapid prototyping and deployment of new routing
+protocols~\cite{sivaraman2016packet}, QoS
+policies~\cite{sivaraman2016programmable}, load balancing
+algorithms~\cite{zhang2013steering}, intrusion prevention
+systems~\cite{xing2014sdnips}, etc. The agility that software engenders
+determined even traditional hardware-based network device manufacturers to
+invest in software-based approaches~\cite{ciscodevelopssdn}.
+
+But this raises the question of how to develop carrier-grade network functions
+(NF) in software? Since NFs are often part of critical infrastructure, they must
+simultaneously be secure, reliable, and high-performance, i.e., be
+carrier-grade. The Achilles' heel of software has always been the (perceived)
+lack of reliability and security, particularly when it is written in C. This has
+led researchers to advocate using safe languages~\cite{clipsham2015safe,
+  kourtis2015intelligent, fowler2014verified} or relatively heavyweight
+approaches~\cite{madhavapeddy2007melange, madhavapeddy2005splat,
+  musuvathi2004model, alur2001verifying, ennals2004linear, ridge2008rigorous,
+  qadir2015applying}. Yet many frameworks (e.g., P4~\cite{bosshart2014p4},
+DPDK~\cite{intel2014data}, PF\_RING~\cite{pfring},
+NetMap~\cite{rizzo2012revisiting}) and standalone implementations (Open
+vSwitch~\cite{pfaff2009extending}, NetFilter~\cite{boye2013netfilter}) use
+C.
 % These approaches require developers to drastically change or augment the way in which they write these network apps: massively annotate their code with logical derivations, use ``pure'' (side-effect-free) functions, abandon manual memory management and layout, and/or learn a new language.
 
-Developers do not want to change the way they write code, they want to be able to convince their users that the code they wrote is carrier-grade (and, if they cannot prove correctness, to then fix the corner cases that impede the proof).
-We conducted a small survey~\cite{verifsurvey} among NF developers from both industry and academia, and almost 70\% insist on continuing to use C.
+Developers do not want to change the way they write code, they want to be able
+to convince their users that the code they wrote is carrier-grade (and, if they
+cannot prove correctness, to then fix the corner cases that impede the proof).
+We conducted a small survey~\cite{verifsurvey} among NF developers from both
+industry and academia, and almost 70\% insist on continuing to use
+C.
 % instead of a higher level alternative.
@@
 
-The goal of our work is to enable developers to provide  carrier-grade network functionality in software without having to do a forklift upgrade of their programming practices.
-In this paper we present a technique and a tool suite (a) for formally verifying crash- and leak-freedom as well as the absence of several classes of security vulnerabilities in NFs written in C/C++ (b) in a way that is practical both in terms of developer effort and the performance of the resulting software.
-
-There exist approaches for providing formal guarantees for {\em stateless} network packet processing code~\cite{dobrescu2014software}; to the best of our knowledge, we are the first to provide such guarantees for real-world {\em stateful} software NFs without impacting performance.
-There are many such NFs: all dynamic routing protocols (RIP, BGP, STP), NAT, BRAS~\cite{wiki:bras}, CDN, IDS, DPI boxes, etc.
+The goal of our work is to enable developers to provide carrier-grade network
+functionality in software without having to do a forklift upgrade of their
+programming practices. In this paper we present a technique and a tool suite (a)
+for formally verifying crash- and leak-freedom as well as the absence of several
+classes of security vulnerabilities in NFs written in C/C++ (b) in a way that is
+practical both in terms of developer effort and the performance of the resulting
+software.
+
+There exist approaches for providing formal guarantees for {\em stateless}
+network packet processing code~\cite{dobrescu2014software}; to the best of our
+knowledge, we are the first to provide such guarantees for real-world {\em
+  stateful} software NFs without impacting performance. There are many such NFs:
+all dynamic routing protocols (RIP, BGP, STP), NAT, BRAS~\cite{wiki:bras}, CDN,
+IDS, DPI boxes, etc.
 
 Practical verification of stateful software NFs is a difficult research problem.
-Hence none of the verification techniques proposed so far can scale anywhere close to verifying real-world stateful NFs.
-For the purpose of this paper, we take as an objective to develop a NAT~\cite{rfc3022} box that is proven to not have certain classes of vulnerabilities or crash or hang under any input.
-Even though NAT is a classic network function, it has proven hard to get right over time: the NAT on various Cisco devices can be crashed~\cite{cve-2015-6271} or hung~\cite{cve-2013-1138} by carefully crafted inputs, with the same type of problems existing in Juniper's NAT~ \cite{cve-2014-3817}, NATs based on NetFilter~\cite{cve-2014-9715}, and the NAT in Windows Server~\cite{ms13-064}.
-Thus, a NAT function that is formally verified to not crash, while providing production-class performance, is an important milestone in this line of work.
+Hence none of the verification techniques proposed so far can scale anywhere
+close to verifying real-world stateful NFs. For the purpose of this paper, we
+take as an objective to develop a NAT~\cite{rfc3022} box that is proven to not
+have certain classes of vulnerabilities or crash or hang under any input. Even
+though NAT is a classic network function, it has proven hard to get right over
+time: the NAT on various Cisco devices can be crashed~\cite{cve-2015-6271} or
+hung~\cite{cve-2013-1138} by carefully crafted inputs, with the same type of
+problems existing in Juniper's NAT~ \cite{cve-2014-3817}, NATs based on
+NetFilter~\cite{cve-2014-9715}, and the NAT in Windows Server~\cite{ms13-064}.
+Thus, a NAT function that is formally verified to not crash, while providing
+production-class performance, is an important milestone in this line of work.
 
@@
 \begin{itemize}
-    \item {\bf Canonical data structures}: Unlike general-purpose software, data plane software employs a small number of commonly used network data structures (hash table, least-prefix-match, rings, etc.) and components.
-These lend themselves naturally to being abstracted into a library.
-They can be formally verified a priori, and their proofs then automatically leveraged for proving properties of the code that uses them.
+\item {\bf Canonical data structures}: Unlike general-purpose software, data
+  plane software employs a small number of commonly used network data structures
+  (hash table, least-prefix-match, rings, etc.) and components. These lend
+  themselves naturally to being abstracted into a library. They can be formally
+  verified a priori, and their proofs then automatically leveraged for proving
+  properties of the code that uses them.
 \item {\bf Hybrid verification}: While typically one picks a verification
-      technique and applies it throughout, no {\em single} technique works well for {\em all} of the code.
-However, data plane software offers us the opportunity to advantageously employ symbolic execution for high-level code (which is mostly stateless and has short execution paths) and verification based on separation logic to the library of components/data structures (which is highly stateful).
+  technique and applies it throughout, no {\em single} technique works well for
+  {\em all} of the code. However, data plane software offers us the opportunity
+  to advantageously employ symbolic execution for high-level code (which is
+  mostly stateless and has short execution paths) and verification based on
+  separation logic to the library of components/data structures (which is highly
+  stateful).
 
@@
 
-In this paper we present work that builds upon these insights and makes three contributions:
+In this paper we present work that builds upon these insights and makes three
+contributions:
 \begin{itemize}
-    \item {\bf Speculative data plane verification} (SDPV), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution.
+\item {\bf Speculative data plane verification} (SDPV), a technique for
+  verifying stateful software NFs by fusing powerful formal verification with
+  automatic symbolic
+  execution.
 % It uses formal verification for data structures.
@@
 \item {\bf The Vigor library of verified components} along with a tool chain
-      that makes SDPV available to those who write software NFs in C/C++.
-      The library consists of data structures (Hash Map, Batcher, Static Array) and a Port Allocator that are common for software NF.
-% The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
-    \item {\bf The \vignat verified NAT box} that offers both high performance and is formally proven to not crash or leak memory.
-We estimate that it would take an NF developer about 1 week to develop and verify a NAT using Vigor.
+  that makes SDPV available to those who write software NFs in C/C++. The
+  library consists of data structures (Hash Map, Batcher, Static Array) and a
+  Port Allocator that are common for software
+  NF. % The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
+\item {\bf The \vignat verified NAT box} that offers both high performance and
+  is formally proven to not crash or leak memory. We estimate that it would take
+  an NF developer about 1 week to develop and verify a NAT using Vigor.
 \end{itemize}
 
-In the rest of the paper, we give a short overview of Vigor (\S\ref{sec:solution-overview}), describe our design (\S\ref{sec:design}), and present our implementation (\S\ref{sec:implementation}).
-We then evaluate the technique for two NAT prototypes (\S\ref{sec:evaluation}), discuss limitations (\S\ref{sec:limitations}), present related work (\S\ref{sec:related-work}), and conclude (\S\ref{sec:conclusion}).
+In the rest of the paper, we give a short overview of Vigor
+(\S\ref{sec:solution-overview}), describe our design (\S\ref{sec:design}), and
+present our implementation (\S\ref{sec:implementation}). We then evaluate the
+technique for two NAT prototypes (\S\ref{sec:evaluation}), discuss limitations
+(\S\ref{sec:limitations}), present related work (\S\ref{sec:related-work}), and
+conclude (\S\ref{sec:conclusion}).
 
@@
 
-This section motivates and explains at a high level what it's like to use Vigor when developing NFs.
-For the sake of illustration, say we wish to develop a trivial stateful NF that implements the discard protocol~\cite{rfc863}: an infinite loop receives packets from one interface, discards the ones that target port 9 (``discard port''), and forwards the others through another interface.
+This section motivates and explains at a high level what it's like to use Vigor
+when developing NFs. For the sake of illustration, say we wish to develop a
+trivial stateful NF that implements the discard protocol~\cite{rfc863}: an
+infinite loop receives packets from one interface, discards the ones that target
+port 9 (``discard port''), and forwards the others through another interface.
 
@@
 In order to absorb bursts, the code employs a ring buffer. The interface to the
-network is: \code{bool receive_packet(struct packet* dst)}, which
-non-blockingly reads an inbound packet and stores it at \code{dst}, returning success or failure; \code{bool can_send_packet()}, which checks if a new packet can be sent; and \code{void send_packet(struct packet* src)}, which sends the packet stored at \code{src}.
-The application follows common practice: it consists of a main event loop, it has no other complex loops, and it uses a ring, a popular data structure in network programming.
-
-To verify that this NF does not crash, one could use the approach of Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution.
-The symbolic execution (SE) technique entails using a symbolic execution engine (SEE) to explore all paths through the code and check whether any failure conditions hold --- in our case crash conditions.
-The inputs to the program are marked ``symbolic,'' meaning that they subsume all possible input values, and then the SEE simulates the execution of the program while propagating the symbolic inputs along the dataflow paths into the program's variables.
-When facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the program state and takes each feasible path separately.
-Branches impose constraints on variable values (e.g., an \code{if} statement imposes restrictions on the values of the condition variable along the \code{else} branch); the SEE collects these constraints along each path.
-When evaluating a branch condition, the SEE asks an SMT constraint solver which of the branches are feasible, given the constraints collected so far on that path prefix.
+network is: \code{bool receive_packet(struct packet* dst)}, which non-blockingly
+reads an inbound packet and stores it at \code{dst}, returning success or
+failure; \code{bool can_send_packet()}, which checks if a new packet can be
+sent; and \code{void send_packet(struct packet* src)}, which sends the packet
+stored at \code{src}. The application follows common practice: it consists of a
+main event loop, it has no other complex loops, and it uses a ring, a popular
+data structure in network programming.
+
+To verify that this NF does not crash, one could use the approach of Dobrescu et
+al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The
+symbolic execution (SE) technique entails using a symbolic execution engine
+(SEE) to explore all paths through the code and check whether any failure
+conditions hold --- in our case crash conditions. The inputs to the program are
+marked ``symbolic,'' meaning that they subsume all possible input values, and
+then the SEE simulates the execution of the program while propagating the
+symbolic inputs along the dataflow paths into the program's variables. When
+facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the
+program state and takes each feasible path separately. Branches impose
+constraints on variable values (e.g., an \code{if} statement imposes
+restrictions on the values of the condition variable along the \code{else}
+branch); the SEE collects these constraints along each path. When evaluating a
+branch condition, the SEE asks an SMT constraint solver which of the branches
+are feasible, given the constraints collected so far on that path prefix.
 
@@
 
-Alas, symbolic execution would never terminate.
-The reason is ``path explosion'': there is a large (often infinite) number of paths to explore, so the SEE cannot finish.
-Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, the infinite loop gives rise to an infinite number of paths.
+Alas, symbolic execution would never terminate. The reason is ``path
+explosion'': there is a large (often infinite) number of paths to explore, so
+the SEE cannot finish. Path explosion results from cyclomatic
+complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number
+of paths, and every loop multiplies it by a factor; in our case, the infinite
+loop gives rise to an infinite number of paths.
 
@@
 
-This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10 sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for \code{CAP=60} it takes 3 hours, and so on.
-It takes long due to another classic SE challenge, ``symbolic indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset}: when indexing an array with a symbolic variable (or when dereferencing a symbolic pointer), the constraints on program state become extremely complex, because the symbolic pointer or index could be referencing many different possible locations.
-The complexity of the constraints chokes the SMT solver, and the SEE grinds to a halt.
-In general, SE doesn't work for code with symbolic pointers or offsets.
-Alas, stateful C code almost always has some of these.
-
-{\bf Step 2: Use abstract data structures.} In Vigor, we eliminate the problem of symbolic indexing by encapsulating program state into well-defined data structures and putting them into a ``library of components''.
-This works well for NF development, because the NF code uses a relatively small set of standard structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
-We provide to the SEE symbolic models of these abstracted components that ``hide'' symbolic indexes and pointers from the SEE.
-
-For our example, the library provides a \code{ring} with a simple interface: \code{bool ring_empty()} checks whether there is a packet in the ring, \code{bool ring_full()} checks whether the ring has run out of space, \code{void ring_push_back(struct packet* src)} adds packet \code{src} to the ring, and \code{void ring_pop_front(struct packet* dst)} removes a packet from the ring and places it in \code{dst}.
-We then write a symbolic model for the \code{ring} and provide it to the SEE instead of the actual implementation.
-The NF now becomes:
+This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
+sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
+\code{CAP=60} it takes 3 hours, and so on. It takes long due to another classic
+SE challenge, ``symbolic
+indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset}: when
+indexing an array with a symbolic variable (or when dereferencing a symbolic
+pointer), the constraints on program state become extremely complex, because the
+symbolic pointer or index could be referencing many different possible
+locations. The complexity of the constraints chokes the SMT solver, and the SEE
+grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
+offsets. Alas, stateful C code almost always has some of these.
+
+{\bf Step 2: Use abstract data structures.} In Vigor, we eliminate the problem
+of symbolic indexing by encapsulating program state into well-defined data
+structures and putting them into a ``library of components''. This works well
+for NF development, because the NF code uses a relatively small set of standard
+structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
+We provide to the SEE symbolic models of these abstracted components that
+``hide'' symbolic indexes and pointers from the SEE.
+
+For our example, the library provides a \code{ring} with a simple interface:
+\code{bool ring_empty()} checks whether there is a packet in the ring,
+\code{bool ring_full()} checks whether the ring has run out of space, \code{void
+  ring_push_back(struct packet* src)} adds packet \code{src} to the ring, and
+\code{void ring_pop_front(struct packet* dst)} removes a packet from the ring
+and places it in \code{dst}. We then write a symbolic model for the \code{ring}
+and provide it to the SEE instead of the actual implementation. The NF now
+becomes:
 
@@
 
-The Vigor library comes with a formal specification and a formal proof that the implementation satisfies that specification.
-In other words, there is a formal spec for \code{ring} and a proof that the \code{ring} implementation satisfies the spec.
-This is how we enable SE to verify the high level code, while using separation logic to verify the library of components---two different verifications techniques for two different parts of the NF.
-Doing the library verification requires deep expertise in program verification, but our effort is amortized across all NFs that use the library.
-Vigor makes it possible for NF developers to verify their code without having to understand verification.
-Note that stating invariants for dataplane software turns out to require no verification expertise.
+The Vigor library comes with a formal specification and a formal proof that the
+implementation satisfies that specification. In other words, there is a formal
+spec for \code{ring} and a proof that the \code{ring} implementation satisfies
+the spec. This is how we enable SE to verify the high level code, while using
+separation logic to verify the library of components---two different
+verifications techniques for two different parts of the NF. Doing the library
+verification requires deep expertise in program verification, but our effort is
+amortized across all NFs that use the library. Vigor makes it possible for NF
+developers to verify their code without having to understand verification. Note
+that stating invariants for dataplane software turns out to require no
+verification expertise.
 
@@
 
-Note that \code{property(head(lst));} in the post condition ensures \code{p->port != 9} property.
-
-{\bf Step 3: Validate the model used for SE.}
-The model must match the library spec.
-Some symbolic models work well for some code, while others work well for other code, so the Vigor library comes with {\em multiple} simple, overapproximating models for each component, and Vigor automatically chooses combinations of models for symbolically executing the NF code.
-Once it finds a combination of models with which the desired property (crash freedom, no memory leaks, absence of certain security vulnerabilities) can be proven, Vigor validates that the behaviors exposed by the model during verification are consistent with the component's formal specification.
-Details on how it does this appear in \S\ref{sec:design}.
+Note that \code{property(head(lst));} in the post condition ensures
+\code{p->port != 9} property.
+
+{\bf Step 3: Validate the model used for SE.} The model must match the library
+spec. Some symbolic models work well for some code, while others work well for
+other code, so the Vigor library comes with {\em multiple} simple,
+overapproximating models for each component, and Vigor automatically chooses
+combinations of models for symbolically executing the NF code. Once it finds a
+combination of models with which the desired property (crash freedom, no memory
+leaks, absence of certain security vulnerabilities) can be proven, Vigor
+validates that the behaviors exposed by the model during verification are
+consistent with the component's formal specification. Details on how it does
+this appear in \S\ref{sec:design}.
 
@@
 
-{\bf Step 4: Link the NF code with the verified library} of components. The resulting executable is guaranteed to be free of crashes, subject to the assumptions listed in~\S\ref{sec:limitations}.
-
-%Note, that the \code{ring_pop} model does not even check whether it is called with an empty ring or not.
-The example application makes sure to call pop only if the ring is not empty.
-Therefore the check is not necessary.
-%This model clearly overapproximates a ``ring'', e.g.
-\code{ring_empty} may return true even right after \code{ring_push}.
-This means, it may have caused false positives, but as long as it does not, no need to complicate it.
-% For the sake of the example conciseness, we hard coded the invariant ``every packet has port != 9'' and the packet structure into the ring component itself.
-In a real component this detail should be abstracted and supplied as a parameter.
-
-% Now a SE can analyze the code in an instant.
-As you can see the model does not maintain any state.
-It only \code{ASSERT}s that any pushed packet satisfy the invariant, and \code{ASSUME}s that every popped packet does.
-Last steps:
+{\bf Step 4: Link the NF code with the verified library} of components. The
+resulting executable is guaranteed to be free of crashes, subject to the
+assumptions listed in~\S\ref{sec:limitations}.
+
+% Note, that the \code{ring_pop} model does not even check whether it is called
+% with an empty ring or not. The example application makes sure to call pop only
+% if the ring is not empty. Therefore the check is not necessary. This model
+% clearly overapproximates a ``ring'', e.g. \code{ring_empty} may return true
+% even right after \code{ring_push}. This means, it may have caused false
+% positives, but as long as it does not, no need to complicate it. For the sake
+% of the example conciseness, we hard coded the invariant ``every packet has
+% port != 9'' and the packet structure into the ring component itself. In a real
+% component this detail should be abstracted and supplied as a parameter.
+
+% Now a SE can analyze the code in an instant. As you can see the model does not
+% maintain any state. It only \code{ASSERT}s that any pushed packet satisfy the
+% invariant, and \code{ASSUME}s that every popped packet does. Last steps:
 
@@
 In this section we describe the design of the Vigor technique and system.
-Figure~\ref{fig:arch} provides a high-level overview of the three main components of the tool chain.
-In order to prove properties of a network function that uses the Vigor component library, Vigor symbolically execute the NF using a symbolic execution engine (SEE); it then uses a Validator to combine the result of the SEE with the formal spec and proof of the component library.
+Figure~\ref{fig:arch} provides a high-level overview of the three main
+components of the tool chain. In order to prove properties of a network function
+that uses the Vigor component library, Vigor symbolically execute the NF using a
+symbolic execution engine (SEE); it then uses a Validator to combine the result
+of the SEE with the formal spec and proof of the component library.
 
@@
 
-Vigor views a network function as a combination of stateless application logic tied to components and data structures (collectively called ``components'') that are stateful, such as hash tables, arrays, least-prefix-match tables, etc.
-In the example of \S\ref{sec:solution-overview}, we encapsulated the \code{packets} array into the \code{ring} component.
-Singular variables of course need not be abstracted away, just large or variable-sized structures potentially accessed with symbolic indexes.
-The role of the library is to group stateful components into a unit that can be formally verified {\em once} an then {\em reused} across all NFs---we can view the Vigor library as a domain-specific extension of the C language.
-This makes sense because there is a relatively low number of components needed to support a wide range of NFs.
-Figure~ \ref{fig:algo} summarizes how Vigor works.
+Vigor views a network function as a combination of stateless application logic
+tied to components and data structures (collectively called ``components'') that
+are stateful, such as hash tables, arrays, least-prefix-match tables, etc. In
+the example of \S\ref{sec:solution-overview}, we encapsulated the \code{packets}
+array into the \code{ring} component. Singular variables of course need not be
+abstracted away, just large or variable-sized structures potentially accessed
+with symbolic indexes. The role of the library is to group stateful components
+into a unit that can be formally verified {\em once} an then {\em reused} across
+all NFs---we can view the Vigor library as a domain-specific extension of the C
+language. This makes sense because there is a relatively low number of
+components needed to support a wide range of NFs. Figure~ \ref{fig:algo}
+summarizes how Vigor works.
 
@@
 
-It is feasible for NF developers to develop verified components and add them to the library.
-However, this requires a fair amount of verification expertise: formally specify the component interface, implement it in a way that can be verified but also has good performance, then produce a mechanically checkable proof, write one or more symbolic models for it, and finally a plug-in for the Validator.
-Vigor therefore adopts a usage model in which the vast majority of NF developers just use the components in the Vigor library and run Vigor to prove (or debug) their code.
-
-%We had to find our own way to support high level data structures in SE, because the popular approach does not suit our goal.
-Many systems based on SE extend the target language with specification syntax allowing developers to formulate complex conditions.
-For instance, Frama-C/Value~\cite{kirchner2015frama} and VeriFast~\cite{jacobs2008verifast} use their own abstract specification language.
-We strive to make the verification simple and automated, but unfortunately current state of the art systems cannot reason automatically in such sophisticated terms and require manual small-step derivations.
-Developers end up spending more effort on the correctness proof than on the code itself.
+It is feasible for NF developers to develop verified components and add them to
+the library. However, this requires a fair amount of verification expertise:
+formally specify the component interface, implement it in a way that can be
+verified but also has good performance, then produce a mechanically checkable
+proof, write one or more symbolic models for it, and finally a plug-in for the
+Validator. Vigor therefore adopts a usage model in which the vast majority of NF
+developers just use the components in the Vigor library and run Vigor to prove
+(or debug) their code.
+
+% We had to find our own way to support high level data structures in SE,
+% because the popular approach does not suit our goal. Many systems based on SE
+% extend the target language with specification syntax allowing developers to
+% formulate complex conditions. For instance,
+% Frama-C/Value~\cite{kirchner2015frama} and VeriFast~\cite{jacobs2008verifast}
+% use their own abstract specification language. We strive to make the
+% verification simple and automated, but unfortunately current state of the art
+% systems cannot reason automatically in such sophisticated terms and require
+% manual small-step derivations. Developers end up spending more effort on the
+% correctness proof than on the code itself.
 
@@
 
-For each component of the library, Vigor provides several symbolic models.
-A model is a concise over-approximating abstraction of how the specific component behaves when invoked by NF code.
-A good model has very little state.
-When symbolically executing an NF, the SEE uses one model for each component, and explores all execution paths through the NF together with the models.
-If every execution path has the desired property, then the NF has that property at all times.
-Since models are relatively easy to write, it is conceivable for NF developers to also write symbolic models.
-
-A desired property is formulated as an assert.
-``No division by zero''  is stated as an assert that the denominator is non-zero; ``safe memory access'' is stated as an assert that accessed memory is ``owned'' by the accessor.
-More details appear in \S\ref{sec:properties}.
-If a symbolically executed code path completes without any assert failing, we say the execution path was successful.
-
-For each execution path, the Vigor SEE produces a trace of how the NF code interacted with the components it uses (we define these traces formally in \S\ref{sec:model-validation-criteria}).
-This is  {\em speculative} in that we don't have proof that the models are accurate abstractions of the components.
-The question the Validator answers is whether these traces are consistent with the formal specs of the components, i.e., could the interaction the NF code had with the model have been a legitimate interaction with an implementation of the component's specification.
-This is a weaker property than the model being a correct abstraction of the component, but it is sufficient for our purposes.
-
-Of course, execution paths may not always succeed during SE.
-While this may be indicative of a bug in the NF, it could also be simply because the model is too approximate.
-For example, in \S\ref{sec:solution-overview} a model of \code{ring_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring_empty} to return false right after pushing an element.
-A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring_empty} return true only when this counter is zero.
-The above NF code would not crash with this model; on the downside, the more stateful model makes SE take longer.
-Therefore, Vigor starts with the least-state models, and tries alternates until it either finds a model with which SE succeeds on all paths or not.
-
-If Vigor does not find any symbolic model for which all execution paths through the NF code succeed, then it points the developers to the places where the properties fail to hold, and helps the debug.
-
-%    \item The application developer provides loop invariants for long/unbounded loops (e.g.
-event loop in the example).
-The invariants are mostly a conjunction of the corresponding invariants for the symbolic models, but they are likely to have an application specific part.
-
-% a number of speculative assumptions that we believe will hold for the real implementations.
-In the example above some of these assumptions are: ``the \code{ring_empty} function may return either 0 or 1'', ``if the \code{ring_pop_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring_push_back} never crashes''.
+For each component of the library, Vigor provides several symbolic models. A
+model is a concise over-approximating abstraction of how the specific component
+behaves when invoked by NF code. A good model has very little state. When
+symbolically executing an NF, the SEE uses one model for each component, and
+explores all execution paths through the NF together with the models. If every
+execution path has the desired property, then the NF has that property at all
+times. Since models are relatively easy to write, it is conceivable for NF
+developers to also write symbolic models.
+
+A desired property is formulated as an assert. ``No division by zero'' is stated
+as an assert that the denominator is non-zero; ``safe memory access'' is stated
+as an assert that accessed memory is ``owned'' by the accessor. More details
+appear in \S\ref{sec:properties}. If a symbolically executed code path completes
+without any assert failing, we say the execution path was successful.
+
+For each execution path, the Vigor SEE produces a trace of how the NF code
+interacted with the components it uses (we define these traces formally in
+\S\ref{sec:model-validation-criteria}). This is {\em speculative} in that we
+don't have proof that the models are accurate abstractions of the components.
+The question the Validator answers is whether these traces are consistent with
+the formal specs of the components, i.e., could the interaction the NF code had
+with the model have been a legitimate interaction with an implementation of the
+component's specification. This is a weaker property than the model being a
+correct abstraction of the component, but it is sufficient for our purposes.
+
+Of course, execution paths may not always succeed during SE. While this may be
+indicative of a bug in the NF, it could also be simply because the model is too
+approximate. For example, in \S\ref{sec:solution-overview} a model of
+\code{ring_empty} may be as simple as randomly returning true or false---for
+some NF code, this behavior could crash it, if it assumes \code{ring_empty} to
+return false right after pushing an element. A different model, however, keeps a
+counter of how many elements are in the ring, and makes \code{ring_empty} return
+true only when this counter is zero. The above NF code would not crash with this
+model; on the downside, the more stateful model makes SE take longer. Therefore,
+Vigor starts with the least-state models, and tries alternates until it either
+finds a model with which SE succeeds on all paths or not.
+
+If Vigor does not find any symbolic model for which all execution paths through
+the NF code succeed, then it points the developers to the places where the
+properties fail to hold, and helps the debug.
+
+% \item The application developer provides loop invariants for long/unbounded
+%   loops (e.g. event loop in the example). The invariants are mostly a
+%   conjunction of the corresponding invariants for the symbolic models, but
+%   they are likely to have an application specific part.
+
+%   a number of speculative assumptions that we believe will hold for the real
+%   implementations. In the example above some of these assumptions are: ``the
+%   \code{ring_empty} function may return either 0 or 1'', ``if the
+%   \code{ring_pop_front} returns success the popped packet’s port will never be
+%   equal to 9'', and ``\code{ring_push_back} never crashes''.
 
@@
 
-Our approach makes several assumptions.
-First, we assume that our symbolic models are over-approximations of the environment in which the NF runs (network stack, OS, etc.).
-In devising these models, we aim to include every possible behavior, but the absence of a formal specification of the environment makes it impossible to validate the models.
-Second, our guarantees rely on the hardware running the software dataplane correctly.
-Third, we assume the compiler implements the same language semantics employed by our tools (e.g., same byte length for C primitive types).
-
-Thus, the trusted computing base for a Vigor-verified NF consists of the compiler, the Vigor tool chain (including its formal verification engine and SEE), the network stack, device drivers, OS kernel, BIOS, and hardware.
-
-% We distinguish of the verified component library from development of an application using this library. These two activities require different skills and ideally should be done by different people.
-
-% This paper proposes the following approach expressed in a number of steps.
-We use three tools: ``SE verifier'' stands for our SEE, ``formal verifier'' is our formal verification proof checker and assistant and a ``validator'' --- the tool that validates legitimacy and exhaustiveness of the model behaviour post factum.
-We differentiate human work between two developers: a component developer and an application developer.
-During the development of our NAT box they were the same person.
+Our approach makes several assumptions. First, we assume that our symbolic
+models are over-approximations of the environment in which the NF runs (network
+stack, OS, etc.). In devising these models, we aim to include every possible
+behavior, but the absence of a formal specification of the environment makes it
+impossible to validate the models. Second, our guarantees rely on the hardware
+running the software dataplane correctly. Third, we assume the compiler
+implements the same language semantics employed by our tools (e.g., same byte
+length for C primitive types).
+
+Thus, the trusted computing base for a Vigor-verified NF consists of the
+compiler, the Vigor tool chain (including its formal verification engine and
+SEE), the network stack, device drivers, OS kernel, BIOS, and hardware.
+
+% We distinguish of the verified component library from development of an
+% application using this library. These two activities require different skills
+% and ideally should be done by different people.
+
+% This paper proposes the following approach expressed in a number of steps. We
+% use three tools: ``SE verifier'' stands for our SEE, ``formal verifier'' is
+% our formal verification proof checker and assistant and a ``validator'' ---
+% the tool that validates legitimacy and exhaustiveness of the model behaviour
+% post factum. We differentiate human work between two developers: a component
+% developer and an application developer. During the development of our NAT box
+% they were the same person.
 
@@
 
-Pointers are the soul of the C language---they embody the power and the danger of low-level systems code.
-A NF focused on performance can not avoid them.
-We embrace them and support the most used idioms in our tool.
-
-An important choice is the memory allocation strategy.
-Dynamic allocation offers more flexibility but incurs performance penalties.
-Static allocation should be preferred because it allows better control over layout, making it possible to optimize for cache and same the allocation overhead at runtime.
-However static allocation reduces flexibility, e.g.
-it fixes the component capacity at the compile time.
-
-Among all the pointers appearing in the execution context, only the pointers crossing API boundary are important for the model validation.
-We divide these pointer into two kinds: argument pointers and return pointers.
+Pointers are the soul of the C language---they embody the power and the danger
+of low-level systems code. A NF focused on performance can not avoid them. We
+embrace them and support the most used idioms in our tool.
+
+An important choice is the memory allocation strategy. Dynamic allocation offers
+more flexibility but incurs performance penalties. Static allocation should be
+preferred because it allows better control over layout, making it possible to
+optimize for cache and same the allocation overhead at runtime. However static
+allocation reduces flexibility, e.g. it fixes the component capacity at the
+compile time.
+
+Among all the pointers appearing in the execution context, only the pointers
+crossing API boundary are important for the model validation. We divide these
+pointer into two kinds: argument pointers and return pointers.
 
@@
 \begin{itemize}
-    \item The address of the component state structure.
-It is an equivalent of ``this'' pointer for C++/Java or ``self'' reference for Rust/Python.
-We do not inspect the memory by such pointers, we only note the aliasing between them.
-The component owns this memory and only the formally verified code can access it.
-    \item Output parameter.
-The application owns this memory.
-Therefore we have to trace the evolution of the pointee value.
-The pointee value before the call, which we trace as the function input, and the pointee value after the call, which goes to the function output.
-A special case of output pointer is a double pointer to the component state structure (X** p), set by the component allocation function.
-The app then owns the pointee (*p), and the SE verifier traces it, but the pointee of the pointee (**p) is the component internal state, so no need to trace it.
-We currently do not support other cases of double or deeper pointers (e.g.
-an output parameter to be set to point to a piece of the internal state).
+\item The address of the component state structure. It is an equivalent of
+  ``this'' pointer for C++/Java or ``self'' reference for Rust/Python. We do not
+  inspect the memory by such pointers, we only note the aliasing between them.
+  The component owns this memory and only the formally verified code can access
+  it.
+\item Output parameter. The application owns this memory. Therefore we have to
+  trace the evolution of the pointee value. The pointee value before the call,
+  which we trace as the function input, and the pointee value after the call,
+  which goes to the function output. A special case of output pointer is a
+  double pointer to the component state structure (X** p), set by the component
+  allocation function. The app then owns the pointee (*p), and the SE verifier
+  traces it, but the pointee of the pointee (**p) is the component internal
+  state, so no need to trace it. We currently do not support other cases of
+  double or deeper pointers (e.g. an output parameter to be set to point to a
+  piece of the internal state).
 \end{itemize}
 
-A return pointer for our components always points to a piece of the internal state (e.g.
-an array cell).
-We trace the pointee and also make sure the application does not access it when the component regains the ownership.
+A return pointer for our components always points to a piece of the internal
+state (e.g. an array cell). We trace the pointee and also make sure the
+application does not access it when the component regains the ownership.
 
@@
 
-Vigor can verify any property that can be stated as a safety property, i.e., formulated as an \code{assert} condition.
-This includes various sources of crashes (dereferencing a null pointer, reading from an unmapped memory page, dividing by zero, etc.) as well as several classes of security bugs related to memory safety (writing to unallocated memory, writing to the wrong memory object, use-after-free, double-free, etc.) and related to integer operations (e.g., integer overflow).
+Vigor can verify any property that can be stated as a safety property, i.e.,
+formulated as an \code{assert} condition. This includes various sources of
+crashes (dereferencing a null pointer, reading from an unmapped memory page,
+dividing by zero, etc.) as well as several classes of security bugs related to
+memory safety (writing to unallocated memory, writing to the wrong memory
+object, use-after-free, double-free, etc.) and related to integer operations
+(e.g., integer overflow).
 
@@
 
-To check for memory leaks (or leaked resources in general) is a bit more complicated, because that is a global property.
-Vigor therefore must keep track of memory ownership.
-In the application logic we forbid dynamic memory allocation, and all allocated memory is controlled in the library, thereby giving the formal verifier explicit visibility into memory ownership.
-The temporary pointer access control enables the SE verifier to ensure the application returns all the memory chunks taken from a component and never accesses them after that.
-
-Vigor does not specifically deal with semantic properties (e.g., dropping a packet, using an incorrect MAC address as a destination), but many of these properties can be formulated by NF developers as \code{assert} statements in their code.
-Vigor will then determine whether those \code{assert}s hold on all paths.
+To check for memory leaks (or leaked resources in general) is a bit more
+complicated, because that is a global property. Vigor therefore must keep track
+of memory ownership. In the application logic we forbid dynamic memory
+allocation, and all allocated memory is controlled in the library, thereby
+giving the formal verifier explicit visibility into memory ownership. The
+temporary pointer access control enables the SE verifier to ensure the
+application returns all the memory chunks taken from a component and never
+accesses them after that.
+
+Vigor does not specifically deal with semantic properties (e.g., dropping a
+packet, using an incorrect MAC address as a destination), but many of these
+properties can be formulated by NF developers as \code{assert} statements in
+their code. Vigor will then determine whether those \code{assert}s hold on all
+paths.
 
@@
 
-To validate a model behavior under certain circumstances means to ensure the behavior agrees with the specification, or is more relaxed (i.e., admits more behaviors than the spec).
-Suppose that, for the input \(I\) to the prefix of an execution trace (a ``run''), the specification produces the output $O_s$, and the model produces the case output \(O_m\).
-Then the model is valid iff for all call sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).
+To validate a model behavior under certain circumstances means to ensure the
+behavior agrees with the specification, or is more relaxed (i.e., admits more
+behaviors than the spec). Suppose that, for the input \(I\) to the prefix of an
+execution trace (a ``run''), the specification produces the output $O_s$, and
+the model produces the case output \(O_m\). Then the model is valid iff for all
+call sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).
 
@@
 \begin{enumerate}
-    \item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and run output.
-It summarizes a particular interaction scenario with the model.
-It consists of the calls to component API functions (with function names, arguments and return values), and leaves out all other details like other calls or unrelated path conditions.
-    \item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of: function names, function inputs(\(Fi\)) for all the functions in the sequence and function outputs(\(Fo\)) for all but the last call.
-    \item \(Co\): The run output (\(Co = Fon\)) is the output of the last API function call in the sequence.
-    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals\_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments).
-For a pointer argument (which must point to a concrete address --- practical limitation), the pointee must be considered according to the \S\ref{sec:handling-pointers}.
-We use the numerical value of the pointer to detect aliasing between the variables, and then throw it away, since we forbid pointer comparison anyway.
-    \item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after + constraints\)) is its return value and the pointee values for the output pointer arguments, as well as the constraints for all the symbols involved (transitive closure, same as above).
+\item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and
+  run output. It summarizes a particular interaction scenario with the model. It
+  consists of the calls to component API functions (with function names,
+  arguments and return values), and leaves out all other details like other
+  calls or unrelated path conditions.
+\item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of:
+  function names, function inputs(\(Fi\)) for all the functions in the sequence
+  and function outputs(\(Fo\)) for all but the last call.
+\item \(Co\): The run output (\(Co = Fon\)) is the output of the last API
+  function call in the sequence.
+\item \(Fi\): The function input (\(Fi = fun\_name + args +
+  arg\_pointee\_vals\_before + constraints\)) consists of the function arguments
+  (possibly symbolic), and constraints bounding these arguments (specifically,
+  the transitive closure of all the constraints involving the symbols mentioned
+  in the formulas for the arguments). For a pointer argument (which must point
+  to a concrete address --- practical limitation), the pointee must be
+  considered according to the \S\ref{sec:handling-pointers}. We use the
+  numerical value of the pointer to detect aliasing between the variables, and
+  then throw it away, since we forbid pointer comparison anyway.
+\item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after +
+  constraints\)) is its return value and the pointee values for the output
+  pointer arguments, as well as the constraints for all the symbols involved
+  (transitive closure, same as above).
 \end{enumerate}
 
-Let us represent the formal specification and the symbolic model total states by large sets.
-A total state consists of variables V, bound by logical formulas F.
+Let us represent the formal specification and the symbolic model total states by
+large sets. A total state consists of variables V, bound by logical formulas F.
 So, the specification and model total states are:
@@
 \]
-The output is then simply a projection of the corresponding state. The projection (P) gives the output variables (return value and output pointers of the tip call): 
+The output is then simply a projection of the corresponding state. The
+projection (P) gives the output variables (return value and output pointers of
+the tip call):
 \[
@@
     \includegraphics[width=\columnwidth]{figures/projections.png}
-    \caption{The outputs \(O_m, O_s\) represented as projections of the total formal specification and model states.}
+    \caption{The outputs \(O_m, O_s\) represented as projections of the total
+      formal specification and model states.}
     \label{projections}
@@
 \]
-Then, the model exhibits all the specified behaviours (and may be some more) iff \(O_s \in O_m\), or: 
+Then, the model exhibits all the specified behaviours (and may be some more) iff
+\(O_s \in O_m\), or:
 \[
@@
 \]
-This formula in general is too difficult for the existing SMT solvers because of the quantifier.
-However, due to similarities in the model and the specification it is easy to find appropriate values for \(V_m\) --- they happen to be some values from Vs.
-Let us call this assignment \(A\).
-\(A\) by construction satisfies the equality \(P_s(V_s)=P_m(A(V_s))\).
-So the condition becomes:
+This formula in general is too difficult for the existing SMT solvers because of
+the quantifier. However, due to similarities in the model and the specification
+it is easy to find appropriate values for \(V_m\) --- they happen to be some
+values from Vs. Let us call this assignment \(A\). \(A\) by construction
+satisfies the equality \(P_s(V_s)=P_m(A(V_s))\). So the condition becomes:
 \[
@@
 
-This validation criteria guarantees that the formally verified implementation will never do anything that the model did not explore during verification.
-So each possible behaviour of the real execution (and may be some extra) was checked during the SE, and it is 1) allowed by the specification and 2) does not cause the application code to crash.
+This validation criteria guarantees that the formally verified implementation
+will never do anything that the model did not explore during verification. So
+each possible behaviour of the real execution (and may be some extra) was
+checked during the SE, and it is 1) allowed by the specification and 2) does not
+cause the application code to crash.
 
@@
 
-For a NF to be verifiable by Vigor, it must fulfill several requirements. These are fully aligned with best practices in developing NFs.
-
-First, the application must contain no deep recursion and only use well structured loops.
-For long or unbounded loops, a developer must supply an invariant.
-Most NFs naturally satisfy this requirement due to performance considerations.
-
-Second, program state must be encapsulated in a set of formally specified components (e.g., port allocator) or data structures (e.g., hashtable).
-Vigor uses the component API as a boundary between the code that is automatically verified using SE and the code that is manually verified.% (Figure~\ref{fig:pie}).
-
-Third, there must be a well defined OS interface.
-As it is infeasible to symbolically execute the whole OS, the SEE stops at the OS interface boundary.
-Behind this interface we trust the correct implementation and we use symbolic models of the OS API functions for SE.
-It is conceivable to use a verified OS kernel, like SeL4~\cite{klein2009sel4}, and thus have a fully verified stack.
-
-Finally, the NF must implement a classic event loop; most NFs already do so.
-For Vigor specifically, there must be a single infinite loop, with each iteration dedicated to processing a single event (an incoming packet).
-Within each iteration, the sequence of interactions between the code and the library components and the OS must be relatively short --- the complexity of models and the amount of symbolic state are often proportional to the length of the API call sequence it needs to support.
+For a NF to be verifiable by Vigor, it must fulfill several requirements. These
+are fully aligned with best practices in developing NFs.
+
+First, the application must contain no deep recursion and only use well
+structured loops. For long or unbounded loops, a developer must supply an
+invariant. Most NFs naturally satisfy this requirement due to performance
+considerations.
+
+Second, program state must be encapsulated in a set of formally specified
+components (e.g., port allocator) or data structures (e.g., hashtable). Vigor
+uses the component API as a boundary between the code that is automatically
+verified using SE and the code that is manually
+verified.% (Figure~\ref{fig:pie}).
+
+Third, there must be a well defined OS interface. As it is infeasible to
+symbolically execute the whole OS, the SEE stops at the OS interface boundary.
+Behind this interface we trust the correct implementation and we use symbolic
+models of the OS API functions for SE. It is conceivable to use a verified OS
+kernel, like SeL4~\cite{klein2009sel4}, and thus have a fully verified stack.
+
+Finally, the NF must implement a classic event loop; most NFs already do so. For
+Vigor specifically, there must be a single infinite loop, with each iteration
+dedicated to processing a single event (an incoming packet). Within each
+iteration, the sequence of interactions between the code and the library
+components and the OS must be relatively short --- the complexity of models and
+the amount of symbolic state are often proportional to the length of the API
+call sequence it needs to support.
 
@@
 When writing NFs to be verified with Vigor, developers are free to use C.
-However, we impose a few restrictions that ensure good isolation between application code and the library components: We require that the component-related memory be accessed only through the library API.
-There should be no pointer comparisons except for equality --- the C99 standard anyway prohibits most pointer comparisons, and we extend this restriction in order to better isolate the component implementation from the code that uses it.
-Finally, pointer accesses must be both domain-limited and time-bounded: certain pointers may access only predefined range (i.e.
-a single unit of the declared type) and only on a restricted segment of the execution, between the corresponding API calls.
+However, we impose a few restrictions that ensure good isolation between
+application code and the library components: We require that the
+component-related memory be accessed only through the library API. There should
+be no pointer comparisons except for equality --- the C99 standard anyway
+prohibits most pointer comparisons, and we extend this restriction in order to
+better isolate the component implementation from the code that uses it. Finally,
+pointer accesses must be both domain-limited and time-bounded: certain pointers
+may access only predefined range (i.e. a single unit of the declared type) and
+only on a restricted segment of the execution, between the corresponding API
+calls.
 
@@
 
-% A component designer faces a choice between conciseness and performance.
-Low expressive power of C forces the component developer to implement an abstraction either through typeless pointers or macros.
-Pointer implementation introduces an extra dereference, which may hurt performance.
-Macros are a zero cost abstraction.
-They perform better as they avoid extra pointer dereference and inline-friendly.
-Unfortunately, they hurt conciseness and readability of the code.
-In the long stretch a generics language extension such as C++99 templates could alleviate this trade-off.
+% A component designer faces a choice between conciseness and performance. Low
+% expressive power of C forces the component developer to implement an
+% abstraction either through typeless pointers or macros. Pointer implementation
+% introduces an extra dereference, which may hurt performance. Macros are a zero
+% cost abstraction. They perform better as they avoid extra pointer dereference
+% and inline-friendly. Unfortunately, they hurt conciseness and readability of
+% the code. In the long stretch a generics language extension such as C++99
+% templates could alleviate this trade-off.
 
@@
 
-In this section we describe the prototypes that embody the ideas presented in the previous section.
-We then use these prototypes to evaluate Vigor on the development of a NAT middlebox, described in \S\ref{sec:evaluation}.
+In this section we describe the prototypes that embody the ideas presented in
+the previous section. We then use these prototypes to evaluate Vigor on the
+development of a NAT middlebox, described in \S\ref{sec:evaluation}.
 
@@
 
-To assess the practicality of Vigor technique, we implemented the three tools described in the section~\ref{sec:our-approach}.
-The SEE is responsible for the application verification, the Formal Verifier checks correctness of the components’ implementation in the library, and the Validator ensures the two verifiers agree on the component APIs used.
-
-\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE KLEE~\cite{cadar2008klee}.
-By default KLEE terminates execution paths that are too long or take too much memory.
-We disabled this feature to ensure it explores all the paths.
-We also extended KLEE with three new features:
-Partial loop invariant support.
-The analysis finds automatically the variables that may change in the loop, and havocs them.
-Currently, a developer have to manually insert the assertions and assumptions according to the loop invariant definition.
-Dynamic pointer access control. The patch provides primitive intrinsics that allow a component developer to open and close access to a pointer between the API calls.
-API call sequence dump. This patch summarizes each control path in the program extracting the component API calls and dumps them for the Validator.
-
-\paragraph{Formal verifier} Our verifier is essentially VeriFast~\cite{jacobs2008verifast} extended with the ability to export the abstract state (\(O_s\)) for the Validator.
-Specifically we export the accumulated conditions for the output variables at the end of the API calls sequence.
+To assess the practicality of Vigor technique, we implemented the three tools
+described in the section~\ref{sec:our-approach}. The SEE is responsible for the
+application verification, the Formal Verifier checks correctness of the
+components’ implementation in the library, and the Validator ensures the two
+verifiers agree on the component APIs used.
+
+\paragraph{SE verifier} We built our verifier for speculative SE by modifying an
+open source C/C++ SEE KLEE~\cite{cadar2008klee}. By default KLEE terminates
+execution paths that are too long or take too much memory. We disabled this
+feature to ensure it explores all the paths. We also extended KLEE with three
+new features: Partial loop invariant support. The analysis finds automatically
+the variables that may change in the loop, and havocs them. Currently, a
+developer have to manually insert the assertions and assumptions according to
+the loop invariant definition. Dynamic pointer access control. The patch
+provides primitive intrinsics that allow a component developer to open and close
+access to a pointer between the API calls. API call sequence dump. This patch
+summarizes each control path in the program extracting the component API calls
+and dumps them for the Validator.
+
+\paragraph{Formal verifier} Our verifier is essentially
+VeriFast~\cite{jacobs2008verifast} extended with the ability to export the
+abstract state (\(O_s\)) for the Validator. Specifically we export the
+accumulated conditions for the output variables at the end of the API calls
+sequence.
 
 \paragraph{Validator} We wrote this third tool completely from scratch.
-Validator independently analyzes each API calls sequence.
-We exploit the data parallelism using GNU parallel tool~\cite{Tange2011a}.
-For each API call sequence in the KLEE dump, Validator first transforms it into VeriFast task.
-A VeriFast task is essentially a C source file, containing the sequence of calls.
+Validator independently analyzes each API calls sequence. We exploit the data
+parallelism using GNU parallel tool~\cite{Tange2011a}. For each API call
+sequence in the KLEE dump, Validator first transforms it into VeriFast task. A
+VeriFast task is essentially a C source file, containing the sequence of calls.
 The Validator enriches the sequence with:
@@
 \end{itemize}
-Next, it runs VeriFast to check that the specification permits this particular sequence of calls under the provided constraints.
-Finally, the Validator imports the specification output state \(O_s\) and checks the inclusion of the specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3} SMT solver (see Section~\ref{sec:model-validation-criteria}).
+Next, it runs VeriFast to check that the specification permits this particular
+sequence of calls under the provided constraints. Finally, the Validator imports
+the specification output state \(O_s\) and checks the inclusion of the
+specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3}
+SMT solver (see Section~\ref{sec:model-validation-criteria}).
 
@@
 
-For our Vigor library prototype, we implemented four stateful components.
-These are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation}: A hash map keeps current flows, and allows lookup by two keys corresponding to the packets coming from internal and external interfaces; the Port allocator is responsible for allocating new ports and expiring inactive entries; the Batcher allows batch processing, keeping outbound packets until the NIC is ready to send them; and the Static Array instances keep various configuration values.
-We are careful in how we abstract components like the Static Array, because NFs are expected to access them on hot paths, and so it is important that the data reside in the processor cache.
+For our Vigor library prototype, we implemented four stateful components. These
+are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation}: A hash
+map keeps current flows, and allows lookup by two keys corresponding to the
+packets coming from internal and external interfaces; the Port allocator is
+responsible for allocating new ports and expiring inactive entries; the Batcher
+allows batch processing, keeping outbound packets until the NIC is ready to send
+them; and the Static Array instances keep various configuration values. We are
+careful in how we abstract components like the Static Array, because NFs are
+expected to access them on hot paths, and so it is important that the data
+reside in the processor cache.
 
@@
 
-We implement a two-key hash map using three parts: integer indexed array of flows and two single-key hash maps.
-The array stores flows together with the keys.
-The hash maps internally are fixed size arrays of tuples $<$busy-bit, key-hash, key-ptr, flow-index$>$:
+We implement a two-key hash map using three parts: integer indexed array of
+flows and two single-key hash maps. The array stores flows together with the
+keys. The hash maps internally are fixed size arrays of tuples $<$busy-bit,
+key-hash, key-ptr, flow-index$>$:
 \begin{itemize}
@@
 
-Flow insertion linearly searches a free cell (with busy-bit = 0) from an index equal to the hash of the corresponding key.
-To insert a flow, user must also supply an unused index.
-We use port allocator to allocate these indexes.
-
-Flow search and removal also starts from the hash-induced index looks for a cell with the same hash, with confirmation by a total equality of the stored and searched keys.
-
-This implementation is not efficient for some cases.
-For instance, if the flow is not present in the table, the algorithm will traverse the whole table and the search time will be proportional to the capacity.
-
-To address this inefficiency, we optimize the single-key maps.
-We add another bit into the tuple.
-These bits group cells into hash-collision chains, and the map searches for a flow only across that chain, not across the entire table.
-Unfortunately, we did not yet verify the optimized implementation.
-We estimate re-verification to take about half of the initial verification effort (see section~\ref{sec:efficiency})
-
-A hash map is a widely used data structure. It allows to store and retrieve entries by arbitrary complex keys. Such functions as L2 switching, ACL filtering, IPS, etc. all use hash tables.
+Flow insertion linearly searches a free cell (with busy-bit = 0) from an index
+equal to the hash of the corresponding key. To insert a flow, user must also
+supply an unused index. We use port allocator to allocate these indexes.
+
+Flow search and removal also starts from the hash-induced index looks for a cell
+with the same hash, with confirmation by a total equality of the stored and
+searched keys.
+
+This implementation is not efficient for some cases. For instance, if the flow
+is not present in the table, the algorithm will traverse the whole table and the
+search time will be proportional to the capacity.
+
+To address this inefficiency, we optimize the single-key maps. We add another
+bit into the tuple. These bits group cells into hash-collision chains, and the
+map searches for a flow only across that chain, not across the entire table.
+Unfortunately, we did not yet verify the optimized implementation. We estimate
+re-verification to take about half of the initial verification effort (see
+section~\ref{sec:efficiency})
+
+A hash map is a widely used data structure. It allows to store and retrieve
+entries by arbitrary complex keys. Such functions as L2 switching, ACL
+filtering, IPS, etc. all use hash tables.
 
@@
 
-Port allocator keeps two chains \emph{alloc} and \emph{free} packed into an array.
-The size of the array equal to the port range available to allocator, and each cell corresponds to a port to be allocated.
-\emph{Alloc} chain is a double linked list.
-It contains all the allocated ports with last access timestamps sorted by these timestamps.
-\emph{Free} chain is a single linked list.
-It keeps track of the available ports.
+Port allocator keeps two chains \emph{alloc} and \emph{free} packed into an
+array. The size of the array equal to the port range available to allocator, and
+each cell corresponds to a port to be allocated. \emph{Alloc} chain is a double
+linked list. It contains all the allocated ports with last access timestamps
+sorted by these timestamps. \emph{Free} chain is a single linked list. It keeps
+track of the available ports.
 
@@
 Allocate operation moves head of the free chain to the head of the alloc chain.
-Access moves the element from the middle of alloc chain to the head.
-Expire takes the tail element from the alloc chain and moves it to the head of the free chain.
+Access moves the element from the middle of alloc chain to the head. Expire
+takes the tail element from the alloc chain and moves it to the head of the free
+chain.
 
 The port allocator is essentially a number allocator from a predefined range.
-Constant time allocation and expiration makes it a basic component for any application that manages resources in run time, such as L2 switch or L3 router.
+Constant time allocation and expiration makes it a basic component for any
+application that manages resources in run time, such as L2 switch or L3 router.
 
@@
 
-Batcher is an array with a counter.
-It allows to store elements one by one (up to a preset capacity), and to release them all at once.
-Almost every DPDK example uses a batcher.
-It is necessary to achieve high throughput, because NICs work better on batches, rather than on individual packets.
+Batcher is an array with a counter. It allows to store elements one by one (up
+to a preset capacity), and to release them all at once. Almost every DPDK
+example uses a batcher. It is necessary to achieve high throughput, because NICs
+work better on batches, rather than on individual packets.
 
@@
 
-Every DPDK example uses multiple static arrays for various configuration parameters, counters and other statistics.
-Our static array is a thin wrapper around a regular C array.
-This wrapper adds zero overhead, but enables our tool-chain to control access to the array and to substitute the whole array with just a couple of bytes during SE.
+Every DPDK example uses multiple static arrays for various configuration
+parameters, counters and other statistics. Our static array is a thin wrapper
+around a regular C array. This wrapper adds zero overhead, but enables our
+tool-chain to control access to the array and to substitute the whole array with
+just a couple of bytes during SE.
 
@@
 \begin{enumerate}
-    \item Is our approach effective at verifying real-world stateful NFs? In \S\ref{sec:effectiveness} we provide a qualitative answer to this question for a DPDK-based NAT middlebox.
-    \item How efficient is our approach in terms of human and compute effort (\S\ref{sec:efficiency})?
-    \item Is the performance of the verified NAT box good enough for practical real-world use by sysadmins and network operators (\S\ref{sec:performance})?
-    %\item What are the trade-offs inherent to Vigor, and how do they impact verifiability and performance?
-In \S\ref{sec:tradeoffs} we give the reader a deeper understanding on how the various parts of the technique contribute to its scalability and cost.
+\item Is our approach effective at verifying real-world stateful NFs? In
+  \S\ref{sec:effectiveness} we provide a qualitative answer to this question for
+  a DPDK-based NAT middlebox.
+\item How efficient is our approach in terms of human and compute effort
+  (\S\ref{sec:efficiency})?
+\item Is the performance of the verified NAT box good enough for practical
+  real-world use by sysadmins and network operators (\S\ref{sec:performance})?
+  % \item What are the trade-offs inherent to Vigor, and how do they impact
+  %   verifiability and performance? In \S\ref{sec:tradeoffs} we give the reader
+  %   a deeper understanding on how the various parts of the technique
+  %   contribute to its scalability and cost.
 \end{enumerate}
 
-As a proof of concept, we implemented and verified \vignat, a NAT middlebox based on the DPDK framework~\cite{intel2014data}.
-We use NAT to refer specifically to Dynamic Source Network Address-Port Translation~\cite{rfc3022}.
-We reused boilerplate code from the DPDK L3 forwarding example.
-\vignat has 1,175 LOC, and it links with the Vigor library that has an additional 941 LOC.
-\vignat supports TCP and UDP packet forwarding with dynamic source address--port translation between multiple internal and one external interfaces.
-It supports connection initiation from an internal interface and refuses connections from the external one.
-It supports concurrent flows up to a configurable maximum number; flows expire after a configurable period of inactivity.
-
-% The biggest difference of our application code from a usual code you can see in network applications is array access.
-To get an element from our Static Array component, due to our memory ownership technique, a developer have to explicitly borrow it through API call and then return the borrowed element to access another one.
-
-Our evaluation uses two versions of this NAT box.
-During our experiments we discovered a silly performance bug in our hash table implementation that led to excessive hash collisions under certain circumstances.
-We fixed it, but did not complete the re-verification by the submission deadline.
-We therefore report results for both \vignata (100\% verified but slow) and \vignatb (verified, except for the hash table).
+As a proof of concept, we implemented and verified \vignat, a NAT middlebox
+based on the DPDK framework~\cite{intel2014data}. We use NAT to refer
+specifically to Dynamic Source Network Address-Port Translation~\cite{rfc3022}.
+We reused boilerplate code from the DPDK L3 forwarding example. \vignat has
+1,175 LOC, and it links with the Vigor library that has an additional 941 LOC.
+\vignat supports TCP and UDP packet forwarding with dynamic source address--port
+translation between multiple internal and one external interfaces. It supports
+connection initiation from an internal interface and refuses connections from
+the external one. It supports concurrent flows up to a configurable maximum
+number; flows expire after a configurable period of inactivity.
+
+% The biggest difference of our application code from a usual code you can see
+% in network applications is array access. To get an element from our Static
+% Array component, due to our memory ownership technique, a developer have to
+% explicitly borrow it through API call and then return the borrowed element to
+% access another one.
+
+Our evaluation uses two versions of this NAT box. During our experiments we
+discovered a silly performance bug in our hash table implementation that led to
+excessive hash collisions under certain circumstances. We fixed it, but did not
+complete the re-verification by the submission deadline. We therefore report
+results for both \vignata (100\% verified but slow) and \vignatb (verified,
+except for the hash table).
 
@@
 
-The ultimate goal of our work is to give operators the confidence that software-based NFs will not undermine the robustness and security of their network.
-It took about 12 minutes on a laptop for Vigor to prove that \vignat is free of crashes, free of memory leaks, free of memory overflows, and free of integer overflows.
-We verified \vignat only for one internal and one external interface, but do not foresee any challenges in doing so for multiple interfaces.
-
-Beyond merely providing a binary yes/no answer to the question of whether an NF has a given property or not, Vigor also helps developers find those corner cases that impede the proof.
-There are three classes of bugs that Vigor uncovers in the verification process, and as the citations show, these appeared in the released Open vSwitch and DPDK code:
+The ultimate goal of our work is to give operators the confidence that
+software-based NFs will not undermine the robustness and security of their
+network. It took about 12 minutes on a laptop for Vigor to prove that \vignat is
+free of crashes, free of memory leaks, free of memory overflows, and free of
+integer overflows. We verified \vignat only for one internal and one external
+interface, but do not foresee any challenges in doing so for multiple
+interfaces.
+
+Beyond merely providing a binary yes/no answer to the question of whether an NF
+has a given property or not, Vigor also helps developers find those corner cases
+that impede the proof. There are three classes of bugs that Vigor uncovers in
+the verification process, and as the citations show, these appeared in the
+released Open vSwitch and DPDK code:
 \begin{enumerate}
 
-    \item Security vulnerabilities~\cite{bug:ovs-uaf, bug:dpdk-uaf1, bug:dpdk-uaf2, bug:ovs-faf, bug:ovs-overflow, bug:dpdk-overflow1, bug:dpdk-overflow2, bug:ovs-int-overflow, bug:dpdk-int-overflow} in the form of memory overflows or integer overflows.
-These can enable an attacker to gain remote control of the middlebox.
-
-    \item Crashes~\cite{bug:ovs-0ptr, bug:dpdk-0ptr1, bug:dpdk-0ptr2, bug:dpdk-0ptr3} destabilize the network.
-Furthermore, many of these can be triggered by a remote attacker with a carefully crafted packet (e.g., trigger a bad pointer dereference or a division by zero), thereby opening up the middlebox to denial-of-service attacks.
-
-    \item Memory leaks~\cite{bug:ovs-mem-leak1, bug:ovs-mem-leak2, bug:ovs-mem-leak3, bug:ovs-mem-leak4, bug:dpdk-mem-leak1, bug:dpdk-mem-leak2} are often exploited for denial-of-service by exhausting the middlebox's memory.
-Vigor's memory ownership model makes them both easy to find and to debug.
+\item Security vulnerabilities~\cite{bug:ovs-uaf, bug:dpdk-uaf1, bug:dpdk-uaf2,
+    bug:ovs-faf, bug:ovs-overflow, bug:dpdk-overflow1, bug:dpdk-overflow2,
+    bug:ovs-int-overflow, bug:dpdk-int-overflow} in the form of memory overflows
+  or integer overflows. These can enable an attacker to gain remote control of
+  the middlebox.
+
+\item Crashes~\cite{bug:ovs-0ptr, bug:dpdk-0ptr1, bug:dpdk-0ptr2,
+    bug:dpdk-0ptr3} destabilize the network. Furthermore, many of these can be
+  triggered by a remote attacker with a carefully crafted packet (e.g., trigger
+  a bad pointer dereference or a division by zero), thereby opening up the
+  middlebox to denial-of-service attacks.
+
+\item Memory leaks~\cite{bug:ovs-mem-leak1, bug:ovs-mem-leak2,
+    bug:ovs-mem-leak3, bug:ovs-mem-leak4, bug:dpdk-mem-leak1,
+    bug:dpdk-mem-leak2} are often exploited for denial-of-service by exhausting
+  the middlebox's memory. Vigor's memory ownership model makes them both easy to
+  find and to debug.
 
@@
 
-Vigor does not support today proving concurrency properties or higher level semantic properties.
-Nevertheless, we consider the verifiable absence of low-level flaws to be a solid step toward complete verification.
+Vigor does not support today proving concurrency properties or higher level
+semantic properties. Nevertheless, we consider the verifiable absence of
+low-level flaws to be a solid step toward complete verification.
 
@@
 
-A key question when using Vigor is how much longer does it take a developer to produce verified code vs.
-non-verified, and how much harder is it.
-Another question is how much effort it takes to develop and verify the component library, but that effort gets amortized across many uses and is therefore less important.
-
-\vignat's 1,175 LOC require about 30 lines of developer annotations (3\% overhead) in the form of loop invariants and loop labels like \code{LOOP(1)} in \S\ref{sec:solution-overview}).
-Static analysis can infer all the other information needed for the proofs.
-In our current Vigor prototype we have not implemented this static analysis, and so our \vignat prototype has 176 lines of annotations (15\% overhead)in the form of C structure layout description, conditional compilation constructs, etc.
-that make up for simple but unimplemented analyses in Vigor.
-
-Our estimate is that it would take an NF developer about 1 week to develop and verify \vignat based on reading this paper (in particular \S\ref{sec:algorithmic-requirements} and \S\ref{sec:language-restrictions}).
-Our own development time is less representative, both because of our understanding of the system and the fact that we co-developed Vigor with \vignat.
-
-Once the code is developed and annotated, verification is fast.
-On an Ubuntu 16.04 Linux laptop with an Intel I7-6600U CPU and 16GB LPDDR3 RAM, exhaustive SE of \vignat takes 2 minutes and 650 MB of RAM, and the Validator takes 10 minutes to validate the 700 call sequences that result from SE.
-Verification of the library's correctness proof takes 15 seconds.
+A key question when using Vigor is how much longer does it take a developer to
+produce verified code vs. non-verified, and how much harder is it. Another
+question is how much effort it takes to develop and verify the component
+library, but that effort gets amortized across many uses and is therefore less
+important.
+
+\vignat's 1,175 LOC require about 30 lines of developer annotations (3\%
+overhead) in the form of loop invariants and loop labels like \code{LOOP(1)} in
+\S\ref{sec:solution-overview}). Static analysis can infer all the other
+information needed for the proofs. In our current Vigor prototype we have not
+implemented this static analysis, and so our \vignat prototype has 176 lines of
+annotations (15\% overhead)in the form of C structure layout description,
+conditional compilation constructs, etc. that make up for simple but
+unimplemented analyses in Vigor.
+
+Our estimate is that it would take an NF developer about 1 week to develop and
+verify \vignat based on reading this paper (in particular
+\S\ref{sec:algorithmic-requirements} and \S\ref{sec:language-restrictions}). Our
+own development time is less representative, both because of our understanding
+of the system and the fact that we co-developed Vigor with \vignat.
+
+Once the code is developed and annotated, verification is fast. On an Ubuntu
+16.04 Linux laptop with an Intel I7-6600U CPU and 16GB LPDDR3 RAM, exhaustive SE
+of \vignat takes 2 minutes and 650 MB of RAM, and the Validator takes 10 minutes
+to validate the 700 call sequences that result from SE. Verification of the
+library's correctness proof takes 15 seconds.
 
@@
 
-Developing and verifying the library took about 5 person-weeks, not counting our learning curve and various setbacks. Detailed data appears in Appendix B.
-
-Table~\ref{tab:loc} shows the size of each of the library's four components.
-As expected, the formal specification and proof dwarf by a 10-20$\times$ margin the implementation.
-This is why we consider Vigor's design---using SE for NF code on the one hand, and heavy verification for the reusable library on the other hand---to provide a favorable balance for real-world NF development.
-The symbolic model and the validator plug-in are more concise than the component implementation, which makes it feasible to provide several of them for each component, to suit different usage patterns.
+Developing and verifying the library took about 5 person-weeks, not counting our
+learning curve and various setbacks. Detailed data appears in Appendix B.
+
+Table~\ref{tab:loc} shows the size of each of the library's four components. As
+expected, the formal specification and proof dwarf by a 10-20$\times$ margin the
+implementation. This is why we consider Vigor's design---using SE for NF code on
+the one hand, and heavy verification for the reusable library on the other
+hand---to provide a favorable balance for real-world NF development. The
+symbolic model and the validator plug-in are more concise than the component
+implementation, which makes it feasible to provide several of them for each
+component, to suit different usage patterns.
 
@@
 
-In this section we address the question of whether verifiability comes at the price of performance.
-We evaluate \vignat's forwarding throughput and latency, comparing it to a NAT based on  NetFilter~\cite{boye2013netfilter}, which is widespread and used in all the Linux-based home and small-enterprise routers.
-We set up the NetFilter masquerade rules as shown in Appendix A, and we tune the Linux kernel for optimal performance as described in~\cite{nf-testing}.
-We label it {\em LinuxNAT} in our plots.
-
-As mentioned earlier, we report on two versions of our NAT: \vignata has a performance bug in the hash table, whereas \vignatb does not, but its hash table is not yet verified.
-We configure both to have a maximum capacity of 61,000 flows.
-We use three identical machines, each a 32-core Intel Xeon E5-2667 v2 @ 3.30GHz with 32GB RAM with i350 1 Gbps DPDK-compatible NICs.
-
-% The throughput results are stable, so we do not collect statistical significance. However latency varies noticeably and we measure repeat the experiment ten times to reveal the variability.
+In this section we address the question of whether verifiability comes at the
+price of performance. We evaluate \vignat's forwarding throughput and latency,
+comparing it to a NAT based on NetFilter~\cite{boye2013netfilter}, which is
+widespread and used in all the Linux-based home and small-enterprise routers. We
+set up the NetFilter masquerade rules as shown in Appendix A, and we tune the
+Linux kernel for optimal performance as described in~\cite{nf-testing}. We label
+it {\em LinuxNAT} in our plots.
+
+As mentioned earlier, we report on two versions of our NAT: \vignata has a
+performance bug in the hash table, whereas \vignatb does not, but its hash table
+is not yet verified. We configure both to have a maximum capacity of 61,000
+flows. We use three identical machines, each a 32-core Intel Xeon E5-2667 v2 @
+3.30GHz with 32GB RAM with i350 1 Gbps DPDK-compatible NICs.
+
+% The throughput results are stable, so we do not collect statistical
+% significance. However latency varies noticeably and we measure repeat the
+% experiment ten times to reveal the variability.
 
@@
 
-We set up two nodes in the topology shown in Figure~\ref{fig:throughput-setup}, as per RFC 2647~\cite{rfc2544}: a tester machine connects to the middlebox over two interfaces.
-We use \code{DPDK-Pktgen}, a high-performance traffic generator; it sends packets through {\em outbound} and counts them on the {\em inbound} interface.
-It assigns each packet a port from a given range to generate the corresponding number of distinct flows.
-We run 40-second measurement intervals.
+We set up two nodes in the topology shown in Figure~\ref{fig:throughput-setup},
+as per RFC 2647~\cite{rfc2544}: a tester machine connects to the middlebox over
+two interfaces. We use \code{DPDK-Pktgen}, a high-performance traffic generator;
+it sends packets through {\em outbound} and counts them on the {\em inbound}
+interface. It assigns each packet a port from a given range to generate the
+corresponding number of distinct flows. We run 40-second measurement intervals.
 
@@
 
-In Figure~\ref{fig:throughput-overall} we show the maximum packet rate achieved with a maximum of 1\% packet loss, for different numbers of concurrent flows.
-At the outset, \vignat outperforms LinuxNAT; this is in no way related to Vigor but rather is due to DPDK and to LinuxNAT being interrupt driven.
+In Figure~\ref{fig:throughput-overall} we show the maximum packet rate achieved
+with a maximum of 1\% packet loss, for different numbers of concurrent flows. At
+the outset, \vignat outperforms LinuxNAT; this is in no way related to Vigor but
+rather is due to DPDK and to LinuxNAT being interrupt
+driven.
 % Our prototype is not optimized either, but the fact that it uses DPDK makes it much more efficient.
@@
 
-\vignata does well up until $\approx$28K connections, at which point it collapses due to the performance bug.
-In contrast, \vignatb sticks to the zero-loss line up until $\approx$59K connections, at which point our 61K-entry hashtable experiences too many collisions to keep up---increasing the size of the hash table would move this point further to the right.
-
-To zoom in on what happens around those two inflection points, we plot two throughput graphs.
-Figure~\ref{fig:throughput-27900} shows what happens for 27,800 concurrent flows: the hash collision bug in \vignata starts causing it to grind to a halt at specific send rates; this behavior gets rapidly worse as the number of concurrent flows increases.
-Figure~\ref{fig:throughput-59k} shows what happens for 59,000 concurrent flows: \vignatb's flow table fills to 95\% and thus slows down, starting to perform comparably to LinuxNAT on average, but with significant jitter.
-As the number of flows increases, \vignatb slows down further.
-% Reducing jitter is part of future work.
+\vignata does well up until $\approx$28K connections, at which point it
+collapses due to the performance bug. In contrast, \vignatb sticks to the
+zero-loss line up until $\approx$59K connections, at which point our 61K-entry
+hashtable experiences too many collisions to keep up---increasing the size of
+the hash table would move this point further to the right.
+
+To zoom in on what happens around those two inflection points, we plot two
+throughput graphs. Figure~\ref{fig:throughput-27900} shows what happens for
+27,800 concurrent flows: the hash collision bug in \vignata starts causing it to
+grind to a halt at specific send rates; this behavior gets rapidly worse as the
+number of concurrent flows increases. Figure~\ref{fig:throughput-59k} shows what
+happens for 59,000 concurrent flows: \vignatb's flow table fills to 95\% and
+thus slows down, starting to perform comparably to LinuxNAT on average, but with
+significant jitter. As the number of flows increases, \vignatb slows down
+further. % Reducing jitter is part of future work.
 
@@
 
-To assess \vignat's latency characteristics, we use the network topology in Figure~\ref{fig:latency-setup}.
-We use separate ports on the tester for flow generation and latency sampling.
-We use again DPDK's \code{PktGen} for packet generation.
-We sample latency with netperf~\cite{jones1996netperf} \code{TCP_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.
+To assess \vignat's latency characteristics, we use the network topology in
+Figure~\ref{fig:latency-setup}. We use separate ports on the tester for flow
+generation and latency sampling. We use again DPDK's \code{PktGen} for packet
+generation. We sample latency with netperf~\cite{jones1996netperf}
+\code{TCP_RR}, which measures the request-response delay to \code{netserver} and
+computes the average over 80 seconds.
 
@@
 
-We also wrote a trivial DPDK application, derived from the DPDK l2fwd example~\cite{dpdk-l2fwd} which only forwards packets.
-To eliminate the routing overhead, we replaced L2 routing with a simple MAC address swap.
-This application is stateless and performs only a couple writes to the packet header, so we consider it the lowest possible latency one could achieve with DPDK; we label it {\em Plain forwarding} in our plots.
-
-Figure~\ref{fig:latency-plot} shows one-way latency as the number of concurrent connections through the middlebox increases.
-The performance issue in \vignata causes its latency to increase with the number of flows, but \vignatb stays flat.
-There is increasing jitter as the number of flows increases, and investigating this is part of future work (we believe it's due to DPDK).
-\vignat has on average 23\% higher latency than LinuxNAT, but nevertheless stays under 111~$\mathrm{\mu sec}$, providing sufficient performance for many settings.
-Note that DPDK is designed to sustain high throughput, so it uses batching aggressively, which hurts latency---\vignat introduces a mere 5\% overhead compared to {\em Plain forwarding}.
+We also wrote a trivial DPDK application, derived from the DPDK l2fwd
+example~\cite{dpdk-l2fwd} which only forwards packets. To eliminate the routing
+overhead, we replaced L2 routing with a simple MAC address swap. This
+application is stateless and performs only a couple writes to the packet header,
+so we consider it the lowest possible latency one could achieve with DPDK; we
+label it {\em Plain forwarding} in our plots.
+
+Figure~\ref{fig:latency-plot} shows one-way latency as the number of concurrent
+connections through the middlebox increases. The performance issue in \vignata
+causes its latency to increase with the number of flows, but \vignatb stays
+flat. There is increasing jitter as the number of flows increases, and
+investigating this is part of future work (we believe it's due to DPDK). \vignat
+has on average 23\% higher latency than LinuxNAT, but nevertheless stays under
+111~$\mathrm{\mu sec}$, providing sufficient performance for many settings. Note
+that DPDK is designed to sustain high throughput, so it uses batching
+aggressively, which hurts latency---\vignat introduces a mere 5\% overhead
+compared to {\em Plain forwarding}.
 
@@
 
-If the reader believes our assertion that we can verify \vignatb's hash table in a few person-days, then the results in this section imply that developing a verified NAT middlebox does not come at the expense of performance.
-NF developer effort is relatively low, which makes this a practical approach.
-We therefore conclude that Vigor can succeed in its objective of enabling developers to productively write software-based data plane NFs that simultaneously meet the highest standards of reliability, security, and performance.
+If the reader believes our assertion that we can verify \vignatb's hash table in
+a few person-days, then the results in this section imply that developing a
+verified NAT middlebox does not come at the expense of performance. NF developer
+effort is relatively low, which makes this a practical approach. We therefore
+conclude that Vigor can succeed in its objective of enabling developers to
+productively write software-based data plane NFs that simultaneously meet the
+highest standards of reliability, security, and performance.
 
@@
 
-Vigor's greatest limitation is that it is not fully automated yet.
-In particular, we expect NF developers have to write loop invariants for long and infinite loops.
-These loop invariants are essential to enabling scalable exhaustive SE with our modified KLEE.
-As mentioned before, coming up with these invariants for NF code is typically not difficult.
-Nevertheless, one of our immediate future work items is to leverage invariant induction techniques~\cite{perkins2004efficient, yang2004automatically} to automate at least part of this task.
-
-Application developers may also encounter ``false positives'', i.e., Vigor fails to prove the desired property.
-As mentioned before, this can be due to the property indeed not holding, or because the symbolic models are too over-approximate.
-The NF developer can either develop their own more detailed model, or request the component library developers to provide a suitable model.
-
-Besides the trusted computing base described in \S\ref{sec:assumptions}, Vigor relies on the NF developer to state correct invariants, as well as the library developers to correctly annotate the symbolic models with KLEE intrinsics (so that they produce correct traces).
-Finally, we trust the NF developer to not link their application with unverified components.
-If any of these assumptions fail to hold, Vigor cannot guarantee any properties.
-
-With respect to our NAT use case, as already mentioned we have two versions, one of which is not fully verified (but will be shortly).
-Also, our NAT does not participate in any routing protocols; it needs an L3 router on each interface to dispatch the packets to the rest of the network.
-We supply the MAC addresses of these routers as a runtime argument to the NAT, and can not update it on-the-fly.
+Vigor's greatest limitation is that it is not fully automated yet. In
+particular, we expect NF developers have to write loop invariants for long and
+infinite loops. These loop invariants are essential to enabling scalable
+exhaustive SE with our modified KLEE. As mentioned before, coming up with these
+invariants for NF code is typically not difficult. Nevertheless, one of our
+immediate future work items is to leverage invariant induction
+techniques~\cite{perkins2004efficient, yang2004automatically} to automate at
+least part of this task.
+
+Application developers may also encounter ``false positives'', i.e., Vigor fails
+to prove the desired property. As mentioned before, this can be due to the
+property indeed not holding, or because the symbolic models are too
+over-approximate. The NF developer can either develop their own more detailed
+model, or request the component library developers to provide a suitable model.
+
+Besides the trusted computing base described in \S\ref{sec:assumptions}, Vigor
+relies on the NF developer to state correct invariants, as well as the library
+developers to correctly annotate the symbolic models with KLEE intrinsics (so
+that they produce correct traces). Finally, we trust the NF developer to not
+link their application with unverified components. If any of these assumptions
+fail to hold, Vigor cannot guarantee any properties.
+
+With respect to our NAT use case, as already mentioned we have two versions, one
+of which is not fully verified (but will be shortly). Also, our NAT does not
+participate in any routing protocols; it needs an L3 router on each interface to
+dispatch the packets to the rest of the network. We supply the MAC addresses of
+these routers as a runtime argument to the NAT, and can not update it
+on-the-fly.
 
@@
 
-Dobrescu et al.~\cite{dobrescu2014software} analyzed modular NFs.
-They managed to analyze many stateless Click elements using exhaustive SE, and exploiting modularity and low cyclomatic complexity of such applications.
-However, their original approach is very limited in verification of stateful elements.
-In addition to crash freedom they managed to verify bounded execution and filtering properties.
-Doing this in Vigor for stateful NFs is future work.
-
-SeL4~\cite{klein2009sel4}, CompCert~\cite{stewart2015compositional}, and FSCQ~\cite{chen2015using} show how to do functional verification of practical systems.
-All of them prove the semantics of the system.
-Unfortunately they all require the use of a high-level programming language and deep expertise in the verification domain, which is unacceptable for our problem.
-
-Beringer et al~\cite{beringer2015verified} applied formal verification to an OpenSSL implementation, proving its functional correctness and cryptographic properties.
-They used certified tool chain, reducing trusted code base just to just a small CoQ proof checker core (~8k LOC).
-While it paves the road for formal verification towards real systems, it is not yet suitable as a common development practice.
-
-SymNet~\cite{stoenescu2013symnet} uses SE to reason about whole network configuration with per-flow state.
-They do not verify the middle box implementations, instead parse the configuration files and use special models in place of the components.
-Their work naturally complements ours.
-
-IronFleet~\cite{hawblitzel2015ironfleet} also aims for a verification of applications distributed over a network.
-They use a different approach, though.
-Habwlitzel et al develop IronFleet from scratch in Dafny --- a high level language with built-in verification support.
-
-Others checked existing distributed systems.
-Pedrosa~\cite{pedrosa2015analysis} and Guerraoui~\cite{guerraoui2011model} explored the global state tree down to a certain depth.
-Guerrauoi et al.
-used a similar approach to ours: they made reasonable assumptions that simplify verification and validated the ones that turned out important.
-
-CEGAR~\cite{clarke2000counterexample} is similar to our technique in the spirit of optimizing out the high effort/low return work.
-It only refines the models when given a clear evidence of the their insufficiency.
-Clarke at al.
-demonstrated a very generic approach.
-We focus on the networking domain, and exploit the features inaccessible for CEGAR, like the low number of data structures and simplicity of the application logic.
+Dobrescu et al.~\cite{dobrescu2014software} analyzed modular NFs. They managed
+to analyze many stateless Click elements using exhaustive SE, and exploiting
+modularity and low cyclomatic complexity of such applications. However, their
+original approach is very limited in verification of stateful elements. In
+addition to crash freedom they managed to verify bounded execution and filtering
+properties. Doing this in Vigor for stateful NFs is future work.
+
+SeL4~\cite{klein2009sel4}, CompCert~\cite{stewart2015compositional}, and
+FSCQ~\cite{chen2015using} show how to do functional verification of practical
+systems. All of them prove the semantics of the system. Unfortunately they all
+require the use of a high-level programming language and deep expertise in the
+verification domain, which is unacceptable for our problem.
+
+Beringer et al~\cite{beringer2015verified} applied formal verification to an
+OpenSSL implementation, proving its functional correctness and cryptographic
+properties. They used certified tool chain, reducing trusted code base just to
+just a small CoQ proof checker core (~8k LOC). While it paves the road for
+formal verification towards real systems, it is not yet suitable as a common
+development practice.
+
+SymNet~\cite{stoenescu2013symnet} uses SE to reason about whole network
+configuration with per-flow state. They do not verify the middle box
+implementations, instead parse the configuration files and use special models in
+place of the components. Their work naturally complements ours.
+
+IronFleet~\cite{hawblitzel2015ironfleet} also aims for a verification of
+applications distributed over a network. They use a different approach, though.
+Habwlitzel et al develop IronFleet from scratch in Dafny --- a high level
+language with built-in verification support.
+
+Others checked existing distributed systems. Pedrosa~\cite{pedrosa2015analysis}
+and Guerraoui~\cite{guerraoui2011model} explored the global state tree down to a
+certain depth. Guerrauoi et al. used a similar approach to ours: they made
+reasonable assumptions that simplify verification and validated the ones that
+turned out important.
+
+CEGAR~\cite{clarke2000counterexample} is similar to our technique in the spirit
+of optimizing out the high effort/low return work. It only refines the models
+when given a clear evidence of the their insufficiency. Clarke at al.
+demonstrated a very generic approach. We focus on the networking domain, and
+exploit the features inaccessible for CEGAR, like the low number of data
+structures and simplicity of the application logic.
 
@@
 
-We presented a technique for proving properties of software NFs.
-We exploit the specifics of NFs to combine the benefits of symbolic execution with formal verification.
-Our experimental results demonstrate the practicality of the approach for the development of NAT box that is guaranteed to have no memory leaks, never crash, be free of certain security vulnerabilities, yet still performs on par with NetFilter.
+We presented a technique for proving properties of software NFs. We exploit the
+specifics of NFs to combine the benefits of symbolic execution with formal
+verification. Our experimental results demonstrate the practicality of the
+approach for the development of NAT box that is guaranteed to have no memory
+leaks, never crash, be free of certain security vulnerabilities, yet still
+performs on par with NetFilter.
 

### 2016-10-28 doc/paper/paper.tex
Expand the example with the corectness proof, and more explanations.

@@
 performance is on par with other NATs.
-%\footnote{Disclaimer: we use acronym NAT in this paper to refer to a more narrow kind - Dynamic Source NAPT (network address-port translation~\cite{rfc3022}).}.
+% \footnote{Disclaimer: we use acronym NAT in this paper to refer to a more
+% narrow kind - Dynamic Source NAPT (network address-port
+% translation~\cite{rfc3022}).}.
 We view this as an important milestone on the road to complete verification of all software-based network functions.
@@
 symbolic inputs along the dataflow paths into the program's variables. When
-facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the
+facing a branch (e.g., an \code{if} or a \code{while} loop), the SEE clones the
 program state and takes each feasible path separately. Branches impose
@@
 as ``no element of \code{packets} between \code{begin} and \code{begin +
-  queue_len} has destination port equal to 9'', and are a good match for
-NF-style loops.  We added to KLEE support for havoc-ing~\cite{barnett2005boogie}
-automatically, to take advantage of such invariants and essentially reduce the
-infinite loop to a single abstract iteration that can be reasoned about in one
-pass.
+  queue_len} has \code{port == 9}'', and are a good match for NF-style loops. We
+added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically
+(replace all modified variables with unconstrained symbols), to take advantage
+of such invariants and essentially reduce the infinite loop to a single abstract
+iteration that allows reasoning in one pass.
 
@@
 sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
-\code{CAP=60} it takes 3 hours, and so on. It takes long due to another classic
+\code{CAP=60} it takes 3 hours, and so on. It takes so long due to another classic
 SE challenge, ``symbolic
@@
 
-For our example, the library provides a \code{ring} with a simple interface:
-\code{bool ring_empty()} checks whether there is a packet in the ring,
-\code{bool ring_full()} checks whether the ring has run out of space, \code{void
-  ring_push_back(struct packet* src)} adds packet \code{src} to the ring, and
-\code{void ring_pop_front(struct packet* dst)} removes a packet from the ring
+For our example, the library provides a packet ring with a simple interface:
+\code{bool ring_empty(ring*)} checks whether there is a packet in the ring,
+\code{bool ring_full(ring*)} checks whether the ring has run out of space, \code{void
+  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
+\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
 and places it in \code{dst}. We then write a symbolic model for the \code{ring}
@@
 For the symbolic execution and future validation, we need also to customize the
-general-purpose packet ring with the following definitions:
+general-purpose packet ring library with the following definitions:
 
@@
 /*@
-  fixpoint bool user_packet_property(packet p) {
-    switch(p) { case packet(port): return port != 9; }
+  fixpoint bool
+  user_property(packet p) {
+    switch(p) { case packet(port):
+                  return port != 9; }
   }
@@
 
-static bool packet_constraints(struct packet* p) {
+static bool
+packet_constraints(struct packet* p) {
   return p->port != 9;
@@
 \end{figure}
+
+Currently, we have to duplicate the packet condition in two languages (formal
+logic, and C) due to the limitations of our prototype. This is the
+application-specific part of the loop invariant given before, that we plug-in in
+order for Vigor to be able to verify the app.
 
@@
                 "popped_packet");
+  // No constraints on the packet port here.
 }
@@
 \end{figure}
-Then the \code{ASSERT} from the previous listing would fail.
+Then the \code{VIGOR_CHECK} from the previous listing would fail.
 
@@
 Here is the example of the proven specification for the function
-\code{ring_pop_front}. We do not include the certified implementation of this
-specification due to space constraints:
+\code{ring_pop_front}:
 
@@
 void ring_pop_front(struct ring* r,
-                    struct packet* p);
-/*@ requires ringp(r, ?lst,
-                   ?property, ?capacity) &*&
-             lst != nil &*&
-             packetp(p, _); @*/
-/*@ ensures ringp(r, tail(lst),
-                  property, capacity) &*&
+                    struct packet* p)
+/*@ requires ringp(r, ?user_property,
+                   ?lst, ?cap) &*&
+             lst != nil &*& packetp(p, _); @*/
+/*@ ensures ringp(r, user_property,
+                  tail(lst), cap) &*&
             packetp(p, head(lst)) &*&
-            true == property(head(lst)); @*/
+            true == user_property(head(lst)); @*/
+{
+  //@ extract_first(r);
+  struct packet* src_pkt = r->array + r->begin;
+  p->port = src_pkt->port;
+  r->len = r->len - 1;
+  r->begin = r->begin + 1;
+  if (r->cap <= r->begin) {
+    r->begin = 0;
+    //@ stitch_with_empty_overflow(r);
+  } else {
+    //@ stitch_with_empty(r);
+  }
+}
 \end{lstlisting}
-  \caption{A formal contract for the ring\_pop\_front function.}
+  \caption{A formal contract and verified implementation of the \code{ring_pop_front} function.}
   \label{lst:contract}
@@
 
-Note that \code{property(head(lst));} in the post condition ensures
+This C code is intermingled with the correctness proof annotations that help the
+formal verification engine (VeriFast) with small steps to transform the initial
+state, given in \code{requires} to the final state, specified in \code{ensures}.
+After proving the correctness of this transformation, the formal verification
+engine will use \code{requires/ensures} formulas as pre- and post- conditions
+for \code{ring_pop_front} contract. The trigraph \code{&*&} separates different
+statements that hold together. For example
+
+\begin{lstlisting}
+ringp(r, user_property, tail(lst), cap) &*&
+packetp(p, head(lst)) &*&
+true == user_property(head(lst))};
+\end{lstlisting}
+means:
+\begin{itemize}
+\item \code{r} points to a healthy ring of capacity \code{cap}, that contains
+  \code{tail(lst)} packets, that satisfy the \code{user_property}.
+\item \code{p} points to a packet \code{head(lst)}.
+\item \code{head(lst)} satisfies the \code{user_property}.
+\end{itemize}
+
+Note that \code{user_property(head(lst));} in the post condition ensures
 \code{p->port != 9} property.
@@
 
-To validate the model fit for the particular use case, SEE dumps
-all the call sequences (30 in this example) that may happen during the execution.
-Vigor checks whether the symbolic model enumerated all the possible
-outputs for each sequence.
+To validate the model fit for the particular use case, SEE dumps all the call
+sequences (30 in this example) extracted from the execution paths. Vigor
+checks whether the symbolic model has enumerated all the possible outputs for
+each call sequence.
 
@@
 \code{loop_invariant_produce} that marks the beginning of a loop iteration. It
-continues then with calls to \code{ring_full} \code{ring_empty}, noting that
-both return \code{false}. The seqeunce ends with a call to
+continues then with calls to \code{ring_full} \code{ring_empty}, noting (with
+\code{assume}) that both return \code{false}. The seqeunce ends with a call to
 \code{ring_pop_front}, formally specified above. The last line checks that the
@@
 //@ assume(ret2 == false);
-//@ close packetp(&(arg2), packet((&(arg2))->port));
+/*@ close packetp(&(arg2),
+                  packet((&(arg2))->port)); @*/
 
@@
 
-Suppose we had an underapproximate model:
+Formal verification engine then checks this sequence using the contracts
+like in the Figure~\ref{lst:contract}, and ensures the abscence of misuse. It then checks
+that the model output covers all the possible outputs specified in the contract
+for the tip function (see~\S\ref{sec:model-validation-criteria}).
 \begin{figure}[h!]
@@
 \end{figure}
-then the last line in the call sequence check will look like:
+
+Suppose we had an underapproximate model: Figure~\ref{lst:underapproximate}.
+Then the last line in the call sequence check will look like:
 \begin{lstlisting}
@@
 \end{lstlisting}
-And the formal verification will fail, because specification admits a wider
-range of possible outputs.
+And the formal verification will fail, because the contract
+(Figure~\ref{lst:contract}) specifies a wider range of possible outputs.
 

### 2016-10-29 doc/paper/paper.tex
Rephrased the validation criteria.

@@
 \usepackage{usenix,epsfig}
+\usepackage{amsmath}
 \usepackage{xcolor}
@@
 
-When symbolically executing a code path, the SEE checks whether any asserts fail.
+When symbolically executing a code path, the SEE checks whether any assert fail.
 
@@
 execution trace (a ``run''), the specification produces the output $O_s$, and
-the model produces the case output \(O_m\). Then the model is valid iff for all
-call sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).
+the model produces the output \(O_m\). Then the model is valid iff for all call
+sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).
 
@@
 \begin{enumerate}
-\item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and
+\item \(F^i\): The function input (\(F^i = fun\_name + args +
+  arg\_pointee\_vals\_before + constraints\)) consists of the function arguments
+  (possibly symbolic), and constraints bounding these arguments (specifically,
+  the transitive closure of all the constraints involving the symbols mentioned
+  in the formulas for the arguments). For a pointer argument (which must point
+  to a concrete address --- a practical limitation), the pointee must be
+  considered according to the \S\ref{sec:handling-pointers}. We use the
+  numerical value of the pointer to detect aliasing between the variables, and
+  then throw it away, since we forbid pointer comparison anyway.
+\item \(F^o\): The function output (\(F^o = ret\_val + pointee\_vals\_after +
+  constraints\)) is its return value and the pointee values for the output
+  pointer arguments, as well as the constraints for all the symbols involved
+  (transitive closure, same as for \(F^i\)).
+\item \(I\): The run input (\(I = <F^i,F^o>_{i=0..n-1} + F^i_n\)) is a set of:
+  function names, function inputs(\(F^i\)) for all the functions in the sequence
+  and function outputs(\(F^o\)) for all but the last call.
+\item \(O\): The run output (\(O = F^o_n\)) is the output of the last API
+  function call in the sequence.
+\item \(C\): A call sequence(= run), (\(C=I+O\)) consists of a run input and
   run output. It summarizes a particular interaction scenario with the model. It
@@
   calls or unrelated path conditions.
-\item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of:
-  function names, function inputs(\(Fi\)) for all the functions in the sequence
-  and function outputs(\(Fo\)) for all but the last call.
-\item \(Co\): The run output (\(Co = Fon\)) is the output of the last API
-  function call in the sequence.
-\item \(Fi\): The function input (\(Fi = fun\_name + args +
-  arg\_pointee\_vals\_before + constraints\)) consists of the function arguments
-  (possibly symbolic), and constraints bounding these arguments (specifically,
-  the transitive closure of all the constraints involving the symbols mentioned
-  in the formulas for the arguments). For a pointer argument (which must point
-  to a concrete address --- practical limitation), the pointee must be
-  considered according to the \S\ref{sec:handling-pointers}. We use the
-  numerical value of the pointer to detect aliasing between the variables, and
-  then throw it away, since we forbid pointer comparison anyway.
-\item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after +
-  constraints\)) is its return value and the pointee values for the output
-  pointer arguments, as well as the constraints for all the symbols involved
-  (transitive closure, same as above).
 \end{enumerate}
@@
 \[
-\{V_s : F_s(V_s)\} and \{V_m : F_m(V_m)\}
+  \begin{aligned}
+  &\mathrm{total~specification~state} = \{V_s : F_s(V_s)\}\\
+  &\mathrm{total~model~state} = \{V_m : F_m(V_m)\}
+  \end{aligned}
 \]
 The output is then simply a projection of the corresponding state. The
-projection (P) gives the output variables (return value and output pointers of
+projection (\(P\)) gives the output variables (return value and output pointers of
 the tip call):
@@
 \[
-O_s = \{P_s(V_s): \forall V_s F_s(V_s)\} \\
-O_m = \{P_m(V_m) : \forall V_m F_m(V_m)\}
+  \begin{aligned}
+    &O_s = \{P_s(V_s): \forall V_s F_s(V_s)\} \\
+    &O_m = \{P_m(V_m) : \forall V_m F_m(V_m)\}
+  \end{aligned}
 \]
@@
 it is easy to find appropriate values for \(V_m\) --- they happen to be some
-values from Vs. Let us call this assignment \(A\). \(A\) by construction
+values from \(Vs\). Let us call this assignment \(A\). \(A\) by construction
 satisfies the equality \(P_s(V_s)=P_m(A(V_s))\). So the condition becomes:
@@
 \]
---- a typical validity query, which we pass  to an SMT solver.
+--- which we formulate as an assertion and let the formal verification engine
+decide its validity.
 

### 2016-11-17 doc/paper/paper.tex
Refactor the solution overview section and divide it into two parts: developer experience and "under the hood"

@@
 
+\subsection{Developer experience}
+
 This section motivates and explains at a high level what it's like to use Vigor
@@
 sent; and \code{void send_packet(struct packet* src)}, which sends the packet
-stored at \code{src}. The application follows common practice: it consists of a
+stored at \code{src}. The application follows a common practice: it consists of a
 main event loop, it has no other complex loops, and it uses a ring, a popular
@@
 
-To verify that this NF does not crash, one could use the approach of Dobrescu et
-al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The
-symbolic execution (SE) technique entails using a symbolic execution engine
-(SEE) to explore all paths through the code and check whether any failure
-conditions hold --- in our case crash conditions. The inputs to the program are
-marked ``symbolic,'' meaning that they subsume all possible input values, and
-then the SEE simulates the execution of the program while propagating the
-symbolic inputs along the dataflow paths into the program's variables. When
-facing a branch (e.g., an \code{if} or a \code{while} loop), the SEE clones the
-program state and takes each feasible path separately. Branches impose
-constraints on variable values (e.g., an \code{if} statement imposes
-restrictions on the values of the condition variable along the \code{else}
-branch); the SEE collects these constraints along each path. When evaluating a
-branch condition, the SEE asks an SMT constraint solver which of the branches
-are feasible, given the constraints collected so far on that path prefix.
-
-One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the
-code above exhaustively --- once all paths are found to be crash-free, the
-crash-freedom proof is complete. To perform SE, it is necessary to provide
-symbolic models for the three network calls. Something simple suffices, such as
-\code{receive_packet} generating a new symbolic packet in \code{dst};
-\code{can_send_packet} forking program state into two states for busy and idle
-interface correspondingly; and \code{send_packet} doing nothing.
-
-Alas, symbolic execution would never terminate. The reason is ``path
-explosion'': there is a large (often infinite) number of paths to explore, so
-the SEE cannot finish. Path explosion results from cyclomatic
-complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number
-of paths, and every loop multiplies it by a factor; in our case, the infinite
-loop gives rise to an infinite number of paths.
-
-{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be
-simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These
-are annotations that inform the SEE about invariant properties of the loop, such
-as ``no element of \code{packets} between \code{begin} and \code{begin +
-  queue_len} has \code{port == 9}'', and are a good match for NF-style loops. We
-added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically
-(replace all modified variables with unconstrained symbols), to take advantage
-of such invariants and essentially reduce the infinite loop to a single abstract
-iteration that allows reasoning in one pass.
-
-This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
-sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
-\code{CAP=60} it takes 3 hours, and so on. It takes so long due to another classic
-SE challenge, ``symbolic
-indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset}: when
-indexing an array with a symbolic variable (or when dereferencing a symbolic
-pointer), the constraints on program state become extremely complex, because the
-symbolic pointer or index could be referencing many different possible
-locations. The complexity of the constraints chokes the SMT solver, and the SEE
-grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
-offsets. Alas, stateful C code almost always has some of these.
-
-{\bf Step 2: Use abstract data structures.} In Vigor, we eliminate the problem
-of symbolic indexing by encapsulating program state into well-defined data
-structures and putting them into a ``library of components''. This works well
-for NF development, because the NF code uses a relatively small set of standard
-structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
-We provide to the SEE symbolic models of these abstracted components that
-``hide'' symbolic indexes and pointers from the SEE.
-
-For our example, the library provides a packet ring with a simple interface:
-\code{bool ring_empty(ring*)} checks whether there is a packet in the ring,
-\code{bool ring_full(ring*)} checks whether the ring has run out of space, \code{void
+{\bf Step 1: Label infinite loops.} Our approach requires special handling of
+infinite loops in the application code. A developer puts a label \code{LOOP}
+into the condition in order to designate the potentially infinite loop. For a
+typical NF there should be only a handful or them (1 in our example).
+
+{\bf Step 2: Use library components. } A developer replaces every array with a
+component from Vigor library that performs the same function. For our example,
+the library provides a packet ring with a simple interface: \code{bool
+  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
+  ring_full(ring*)} checks whether the ring has run out of space, \code{void
   ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
 \code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
-and places it in \code{dst}. We then write a symbolic model for the \code{ring}
-and provide it to the SEE instead of the actual implementation. The NF now
-becomes:
+and places it in \code{dst}. Fig.\ref{lst:vigor} shows the resulting NF code.
+\code{VIGOR_CHECK} adds additional property that we would like to verify ---
+that we do not forward packets with port 9.
 
@@
 
-For the symbolic execution and future validation, we need also to customize the
-general-purpose packet ring library with the following definitions:
+{\bf Step 3: Provide invariants.} Besides usual parameters, like capacity or
+element size, most Vigor components also require simple logical invariants of
+their elements to enable the reasoning about the application code. For our
+example, the essential property of the ring is, that it contains only packets
+with a port different from 9. A developer communicates that as shown in
+Fig.~\ref{lst:customizations}.
 
@@
 Currently, we have to duplicate the packet condition in two languages (formal
-logic, and C) due to the limitations of our prototype. This is the
-application-specific part of the loop invariant given before, that we plug-in in
-order for Vigor to be able to verify the app.
-
-Now exhaustive SE of the code takes less than 100 millisec with our modified
-KLEE, regardless of the size of \code{CAP}.  The fact that the code does not
+logic, and C) due to the limitations of our prototype.
+
+{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model,
+check that the application does not crash in any execution scenario. It will
+then make sure that the chosen model is faithful to the formally proven
+implementation of the component.
+
+{\bf Step 5: Link the NF code with the verified library} of components. The
+resulting executable is guaranteed to be free of crashes, and hold the
+properties formulated in \code{VIGOR_CHECK}, subject to the assumptions listed
+in~\S\ref{sec:limitations}.
+
+\subsection{Under the hood}
+
+To verify that this NF does not crash, one could use the approach of Dobrescu et
+al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The
+symbolic execution (SE) technique entails using a symbolic execution engine
+(SEE) to explore all paths through the code and check whether any failure
+conditions hold --- in our case crash conditions. The inputs to the program are
+marked ``symbolic,'' meaning that they subsume all possible input values, and
+then the SEE simulates the execution of the program while propagating the
+symbolic inputs along the dataflow paths into the program's variables. When
+facing a branch (e.g., an \code{if} or a \code{while} loop), the SEE clones the
+program state and takes each feasible path separately. Branches impose
+constraints on variable values (e.g., an \code{if} statement imposes
+restrictions on the values of the condition variable along the \code{else}
+branch); the SEE collects these constraints along each path. When evaluating a
+branch condition, the SEE asks an SMT constraint solver which of the branches
+are feasible, given the constraints collected so far on that path prefix.
+
+One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the
+original application code (Fig.~\ref{lst:orig}) exhaustively --- once all paths
+are found to be crash-free, the crash-freedom proof is complete. To perform SE,
+it is necessary to provide symbolic models for the three network calls.
+Something simple suffices, such as \code{receive_packet} generating a new
+symbolic packet in \code{dst}; \code{can_send_packet} forking program state into
+two states for the interface being busy and idle correspondingly; and
+\code{send_packet} doing nothing.
+
+Alas, naive symbolic execution would never terminate. The reason is ``path
+explosion'': there is a large (often infinite) number of paths to explore, so
+the SEE cannot finish. Path explosion results from cyclomatic
+complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number
+of paths, and every loop multiplies it by a factor; in our case, the infinite
+loop gives rise to an infinite number of paths.
+
+Vigor transforms infinite loops using loop
+invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
+automatic havoc-ing~\cite{barnett2005boogie} (replace all variables changed in
+the loop with unconstrained symbols). The step 1 enables Vigor to reduce each
+problematic loop to a single generalized iteration that allows reasoning in one
+pass.
+
+This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
+sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
+\code{CAP=60} it takes 3 hours, and so on. Another classic SE challenge ---
+``symbolic
+indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
+the party here. When
+indexing an array with a symbolic variable (or when dereferencing a symbolic
+pointer), the constraints on program state become extremely complex, because the
+symbolic pointer or index could be referencing many different possible
+locations. The complexity of the constraints chokes the SMT solver, and the SEE
+grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
+offsets. Alas, stateful C code almost always has some of these.
+
+In Vigor, we circumvent the problem of symbolic indexing by encapsulating program
+state into well-defined data structures and putting them into a ``library of
+components'' (step 2). This works well for NF development, because the NF code
+uses a relatively small set of standard structures, such as ring buffers, hash
+tables, least-prefix-match tables, etc. We provide to the SEE symbolic models of
+these abstracted components that ``hide'' symbolic indexes and pointers from the
+SEE.
+
+After the step 2 exhaustive SE takes less than 100 millisec regardless of the
+value of \code{CAP}. The fact that the code does not
 crash on any path amounts to a proof of crash freedom. Or does it? The validity
-of the proof depends on (1)  our library implementation of \code{ring} not
+of the proof depends on our library implementation of \code{ring} not
 crashing; and (2) the symbolic model of \code{ring} accurately capturing how the
-\code{ring} implementation works ({\em Step 3}).
-
-Suppose we had an overapproximate model:
-\begin{figure}[h!]
-\begin{lstlisting}
-void ring_pop_front(struct ring* r,
-                    struct packet* p) {
-  FILL_SYMBOLIC(p, sizeof(struct packet),
-                "popped_packet");
-  // No constraints on the packet port here.
-}
-\end{lstlisting}
-  \caption{An extract from an overapproximate ring model.}
-  \label{lst:overapproximate}
-\end{figure}
-Then the \code{VIGOR_CHECK} from the previous listing would fail.
+\code{ring} implementation works.
 
@@
 the spec. This is how we enable SE to verify the high level code, while using
-separation logic to verify the library of components---two different
-verifications techniques for two different parts of the NF. Doing the library
-verification requires deep expertise in program verification, but our effort is
-amortized across all NFs that use the library. Vigor makes it possible for NF
-developers to verify their code without having to understand verification. Note
-that stating invariants for dataplane software turns out to require no
-verification expertise.
-
-Here is the example of the proven specification for the function
+separation logic to verify the library of components --- two different
+verifications techniques for two different parts of the complete application.
+Doing the library verification requires deep expertise in program verification,
+but our effort is amortized across all NFs that use the library.
+
+Vigor makes it possible for NF developers to verify their code without having to
+understand verification. Note that stating invariants(step 3) for dataplane
+software turns out to require no verification expertise.
+
+Here is an example of the proven specification for the function
 \code{ring_pop_front}:
@@
 packetp(p, head(lst)) &*&
-true == user_property(head(lst))};
+true == user_property(head(lst));
 \end{lstlisting}
@@
   \code{tail(lst)} packets, that satisfy the \code{user_property}.
-\item \code{p} points to a packet \code{head(lst)}.
-\item \code{head(lst)} satisfies the \code{user_property}.
+\item \code{p} points to a packet ``\code{head(lst)}''.
+\item ``\code{head(lst)}'' satisfies the \code{user_property}.
 \end{itemize}
@@
 Note that \code{user_property(head(lst));} in the post condition ensures
-\code{p->port != 9} property.
-
-{\bf Step 3: Validate the model used for SE.} The model must match the library
-spec. Some symbolic models work well for some code, while others work well for
-other code, so the Vigor library comes with {\em multiple} simple,
-overapproximating models for each component, and Vigor automatically chooses
-combinations of models for symbolically executing the NF code. Once it finds a
-combination of models with which the desired property (crash freedom, no memory
-leaks, absence of certain security vulnerabilities) can be proven, Vigor
+\code{p->port != 9} property (Fig.~\ref{lst:customizations}).
+
+\begin{figure}[h!]
+\begin{lstlisting}
+void ring_pop_front(struct ring* r,
+                    struct packet* p) {
+  FILL_SYMBOLIC(p, sizeof(struct packet),
+                "popped_packet");
+  ASSUME(packet_constraints(p));
+}
+\end{lstlisting}
+  \caption{A symbolic model for \code{ring_pop_front}.}
+  \label{lst:symbolic-model}
+\end{figure}
+
+Some symbolic models work well for some code, while others work well for other
+code, so the Vigor library comes with {\em multiple} simple models for each
+component, and Vigor automatically chooses combinations of models. Once it finds
+a combination of models with which the desired property (crash freedom, no
+memory leaks, absence of certain security vulnerabilities) can be proven, Vigor
 validates that the behaviors exposed by the model during verification are
@@
 this appear in \S\ref{sec:design}.
+
+A model of \code{ring_pop_front} used for this example is shown on
+Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
+exhaustive SE, but preserves the vital propertiy --- \code{packet_constraints}
+(Fig.\ref{lst:customizations}).
 
@@
 
-You can see the example call sequence below. It starts with
-\code{loop_invariant_produce} that marks the beginning of a loop iteration. It
-continues then with calls to \code{ring_full} \code{ring_empty}, noting (with
-\code{assume}) that both return \code{false}. The seqeunce ends with a call to
-\code{ring_pop_front}, formally specified above. The last line checks that the
-value returned by model (\code{arg2.port != 9}) covers all the cases prescribed
-by the specification.
-
-\begin{figure}[h!]
+You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
+It starts with \code{loop_invariant_produce} that marks the beginning of a loop
+iteration. It continues then with calls to \code{ring_full} \code{ring_empty},
+noting (with \code{assume}) that both return \code{false}. The seqeunce ends
+with a call to \code{ring_pop_front}. The last line detects underapproximation
+of the model output.
+
+\begin{figure}[h]
 \begin{lstlisting}
@@
 
-/*@ assert((false == (9 == arg2.port))); @*/
+/*@ assert(arg2.port != 9); @*/
 \end{lstlisting}
@@
 
-Formal verification engine then checks this sequence using the contracts
-like in the Figure~\ref{lst:contract}, and ensures the abscence of misuse. It then checks
-that the model output covers all the possible outputs specified in the contract
-for the tip function (see~\S\ref{sec:model-validation-criteria}).
+Formal verification engine checks this sequence using the contracts like in the
+Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g. call for
+\code{ring_pop} on a possibly empty ring. It then checks that the model output
+covers all the possible outputs specified in the contract for the tip function.
+This happens in the last line on the listing. The assert contains the
+path-constraint generated by the symbolic model. In this case the contract
+(Fig.\ref{lst:contract}) guarantees that the popped packet satisfies
+\code{user_property} that entails \code{arg2.port != 9}, hence the trace is
+ligitimate. (see~\S\ref{sec:model-validation-criteria}).
+
+Consider potential model mismatches. Suppose we had an overapproximate model ---
+Fig.~\ref{lst:overapproximate}. Then the \code{VIGOR_CHECK}
+(Fig.\ref{lst:vigor}) in the application code would fail.
+\begin{figure}[h!]
+\begin{lstlisting}
+void ring_pop_front(struct ring* r,
+                    struct packet* p) {
+  FILL_SYMBOLIC(p, sizeof(struct packet),
+                "popped_packet");
+  // No constraints on the packet port here.
+}
+\end{lstlisting}
+  \caption{An overapproximate ring model fragment.}
+  \label{lst:overapproximate}
+\end{figure}
+
 \begin{figure}[h!]
@@
 \end{lstlisting}
-  \caption{An underapproximate model fragment.}
+  \caption{An underapproximate ring model fragment.}
   \label{lst:underapproximate}
@@
 
-Suppose we had an underapproximate model: Figure~\ref{lst:underapproximate}.
-Then the last line in the call sequence check will look like:
+Suppose now we had an underapproximate model: Fig.~\ref{lst:underapproximate}.
+Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) would look like:
 \begin{lstlisting}
-/*@ assert(0 == arg2.port); @*/
+/*@ assert(arg2.port == 0); @*/
 \end{lstlisting}
-And the formal verification will fail, because the contract
-(Figure~\ref{lst:contract}) specifies a wider range of possible outputs.
-
-{\bf Step 4: Link the NF code with the verified library} of components. The
-resulting executable is guaranteed to be free of crashes, subject to the
-assumptions listed in~\S\ref{sec:limitations}.
+The assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies
+a wider range for \code{arg2.port} than just \code{0}.
 

### 2016-11-21 doc/paper/paper.tex
Luis's feedback: Expanded captions for listings; short explanation of havocing; some line number placeholders.

@@
 \end{lstlisting}
-  \caption{Original implementation of the Drop protocol}
+  \caption{Original implementation of the Drop protocol. Contains an infinite
+    loop and a latent ring implementation}
   \label{lst:orig}
@@
 and places it in \code{dst}. Fig.\ref{lst:vigor} shows the resulting NF code.
-\code{VIGOR_CHECK} adds additional property that we would like to verify ---
-that we do not forward packets with port 9.
+\code{VIGOR_CHECK} (line NN) adds additional property that we would like to
+verify --- that we do not forward packets with port 9.
 
@@
 \end{lstlisting}
-  \caption{Vigor implementation of Drop protocol.}
+  \caption{Vigor implementation of Drop protocol. \code{LOOP} labels the
+    infinite loop head and we explicitly use a ring from the Vigor library.
+    \code{VIGOR_CHECK} at the line NN adds a semantic property to be verified.}
   \label{lst:vigor}
@@
 \end{lstlisting}
-  \caption{Custom definitions capturing the desired property.}
+  \caption{Custom definitions capturing the desired property. They capture the
+    essential property of every packet in the ring.}
   \label{lst:customizations}
@@
 Currently, we have to duplicate the packet condition in two languages (formal
-logic, and C) due to the limitations of our prototype.
-
-{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model,
-check that the application does not crash in any execution scenario. It will
+logic -lines NN, and C - NN) due to the limitations of our prototype.
+
+{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model
+and check that the application does not crash in any execution scenario. It will
 then make sure that the chosen model is faithful to the formally proven
@@
 invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
-automatic havoc-ing~\cite{barnett2005boogie} (replace all variables changed in
-the loop with unconstrained symbols). The step 1 enables Vigor to reduce each
-problematic loop to a single generalized iteration that allows reasoning in one
-pass.
+automatic havoc-ing~\cite{barnett2005boogie}. Havoc-ing simply replaces all the
+variables that may change in the loop by unconstrained symbols at the beginning
+of the iteration. In our example that would be the variables \code{begin} and
+\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
+of the \code{packets} pointer because it never changes. The step 1 enables
+Vigor to reduce each problematic loop to a single generalized iteration that
+allows reasoning in one pass.
 
@@
 offsets. Alas, stateful C code almost always has some of these.
+
+In this example the symbolic indexing comes from the ring array. Once we
+replaced the infinite loop with a single iteration we have to assume (remember
+havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
+anywhere in the array.
 
@@
 \end{lstlisting}
-  \caption{A formal contract and verified implementation of the \code{ring_pop_front} function.}
+  \caption{A formal contract and verified implementation of the
+    \code{ring_pop_front} function. The comments starting with \code{@} symbol
+    are formal annotations that hold the function contract (\code{requires} -
+    precondition, \code{ensures} - postcondition) and lemmas that help the proof
+    checker to verify the state transformations in the code.}
   \label{lst:contract}
@@
 \end{lstlisting}
-  \caption{A symbolic model for \code{ring_pop_front}.}
+  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
+    a packet and adds a constraint to it.}
   \label{lst:symbolic-model}
@@
 It starts with \code{loop_invariant_produce} that marks the beginning of a loop
-iteration. It continues then with calls to \code{ring_full} \code{ring_empty},
+iteration. It continues then with calls to \code{ring_full} and \code{ring_empty},
 noting (with \code{assume}) that both return \code{false}. The seqeunce ends
@@
 \end{lstlisting}
-  \caption{A call sequence, produced by SEE.}
+  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
+    constrain the execution space.}
   \label{lst:call-sequence}
@@
 \end{lstlisting}
-  \caption{An overapproximate ring model fragment.}
+  \caption{An overapproximate ring model fragment. It does not add the
+    constraint necessary for the code to verify.}
   \label{lst:overapproximate}
@@
 \end{lstlisting}
-  \caption{An underapproximate ring model fragment.}
+  \caption{An underapproximate ring model fragment. Instead of returning a
+    constrained symbolic packet, it simply returns some particular packet, which
+    clearly can not capture what the real ring would return on its place.}
   \label{lst:underapproximate}

### 2016-11-22 doc/paper/paper.tex
bugfix in the example original code snippet (thanks, Luis!), line-number references.

@@
 \lstset{language=C,
+  escapechar=^,
   basicstyle=\footnotesize\ttfamily,
@@
   keepspaces=true,
+  columns=flexible,
   tabsize=2,
@@
 port 9 (``discard port''), and forwards the others through another interface.
+A possible implementation is on the Fig.~\ref{lst:orig}.
 
@@
 int main() {
-  struct packet packets[CAP] = {};
-  int begin = 0, queue_len = 0;
-
-  while(1) {
+  struct packet packets[CAP] = {}; ^\label{line:pkt-def}^
+  int begin = 0, queue_len = 0; ^\label{line:iter-def}^
+
+  while(1) {^\label{line:loop-begin}^
     if (queue_len < CAP) {
       int end = (begin + queue_len)%CAP;
-      if (receive_packet(&packets[end]) &&
+      if (receive_packet(&packets[end]) && ^\label{line:recv-pkt}^
           packets[end].port != 9)
@@
     }
-    if (0 < queue_len) {
-      send_packet(&packets[begin]);
+    if (0 < queue_len && can_send_packet()) { ^\label{line:can-send}^
+      send_packet(&packets[begin]); ^\label{line:send-pkt}^
       begin = (begin + 1)%CAP;
     }
+  }^\label{line:loop-end}^
+  return 0;
+}
+\end{lstlisting}
+  \caption{Original implementation of the Drop protocol. Contains an infinite
+    loop (l.\ref{line:loop-begin}-\ref{line:loop-end}) and a latent ring
+    implementation (declared on l.\ref{line:pkt-def},\ref{line:iter-def})}
+  \label{lst:orig}
+\end{figure}
+
+In order to absorb bursts, the code employs a ring buffer. The application
+interacts with the network using the following 3 functions:
+\code{receive_packet}(l.\ref{line:recv-pkt}), which non-blockingly reads an
+inbound packet and stores it at \code{packets[end]}, returning success or
+failure; \code{can_send_packet}(l.\ref{line:can-send}), which checks if a new
+packet can be sent; and \code{send_packet}(l.\ref{line:send-pkt}), which sends
+the packet stored at \code{packets[begin]}. The application follows a common
+practice: it consists of a main event loop, it has no other complex loops, and
+it uses a ring, a popular data structure in network programming.
+
+The ring keeps the state of our NF and makes many existing verification
+approaches impractical.
+
+Fig.~\ref{lst:vigor} shows the actual code that we are going to verify. The
+first two steps show how we get there. \code{VIGOR_CHECK}
+(l.~\ref{line:vigcheck}) adds additional property that we would like to verify
+--- that we do not forward packets with port 9.
+
+\begin{figure}[h!]
+\begin{lstlisting}
+int main(int argc, char** argv) {
+  struct packet p;
+  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
+  if (!r) return 1;
+  while(LOOP(1))^\label{line:loop-hd}^
+  {
+    loop_iteration_begin(&r);
+    if (!ring_full(r))
+      if (receive_packet(&p) && p.port != 9)
+        ring_push_back(r, &p);
+    if (!ring_empty(r) && can_send_packet()) {
+      ring_pop_front(r, &p);
+      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
+      send_packet(&p);
+    }
+    loop_iteration_end(&r);
   }
@@
 \end{lstlisting}
-  \caption{Original implementation of the Drop protocol. Contains an infinite
-    loop and a latent ring implementation}
-  \label{lst:orig}
-\end{figure}
-
-In order to absorb bursts, the code employs a ring buffer. The interface to the
-network is: \code{bool receive_packet(struct packet* dst)}, which non-blockingly
-reads an inbound packet and stores it at \code{dst}, returning success or
-failure; \code{bool can_send_packet()}, which checks if a new packet can be
-sent; and \code{void send_packet(struct packet* src)}, which sends the packet
-stored at \code{src}. The application follows a common practice: it consists of a
-main event loop, it has no other complex loops, and it uses a ring, a popular
-data structure in network programming.
+  \caption{Vigor implementation of Drop protocol.
+    \code{LOOP}(l.\ref{line:loop-hd}) labels the infinite loop head and we
+    explicitly use a ring(l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
+    line \ref{line:vigcheck} adds a semantic property to be verified.}
+  \label{lst:vigor}
+\end{figure}
 
@@
 infinite loops in the application code. A developer puts a label \code{LOOP}
-into the condition in order to designate the potentially infinite loop. For a
-typical NF there should be only a handful or them (1 in our example).
+(l.\label{line:loophd}) into the condition in order to designate the
+potentially infinite loop. For a typical NF there should be only a handful or
+them (1 in our example).
 
@@
 \code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
-and places it in \code{dst}. Fig.\ref{lst:vigor} shows the resulting NF code.
-\code{VIGOR_CHECK} (line NN) adds additional property that we would like to
-verify --- that we do not forward packets with port 9.
-
-\begin{figure}[h!]
-\begin{lstlisting}
-int main(int argc, char** argv) {
-  struct packet p;
-  struct ring *r = ring_create(CAP);
-  if (!r) return 1;
-  while(LOOP(1))
-  {
-    loop_iteration_begin(&r);
-    if (!ring_full(r))
-      if (receive_packet(&p) && p.port != 9)
-        ring_push_back(r, &p);
-    if (!ring_empty(r) && can_send_packet()) {
-      ring_pop_front(r, &p);
-      VIGOR_CHECK(p.port != 9);
-      send_packet(&p);
-    }
-    loop_iteration_end(&r);
-  }
-  return 0;
-}
-\end{lstlisting}
-  \caption{Vigor implementation of Drop protocol. \code{LOOP} labels the
-    infinite loop head and we explicitly use a ring from the Vigor library.
-    \code{VIGOR_CHECK} at the line NN adds a semantic property to be verified.}
-  \label{lst:vigor}
-\end{figure}
+and places it in \code{dst}.
 
@@
 
-/*@
+/*@^\label{line:vf-user-prop-from}^
   fixpoint bool
@@
   }
-  @*/
-
-static bool
+  @*/^\label{line:vf-user-prop-to}^
+
+static bool^\label{line:c-user-prop-from}^
 packet_constraints(struct packet* p) {
   return p->port != 9;
-}
+}^\label{line:c-user-prop-to}^
 \end{lstlisting}
   \caption{Custom definitions capturing the desired property. They capture the
-    essential property of every packet in the ring.}
+    essence of every packet in the ring. Currently, we have to duplicate the
+    packet condition in two languages (formal logic --
+    ll\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
+    ll\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
+    limitations of our prototype.}
   \label{lst:customizations}
 \end{figure}
-
-Currently, we have to duplicate the packet condition in two languages (formal
-logic -lines NN, and C - NN) due to the limitations of our prototype.
 
@@
 then make sure that the chosen model is faithful to the formally proven
-implementation of the component.
+implementation of the component. \S\ref{sec:under-the-hood} provides more
+details on how it is done.
 
@@
 \subsection{Under the hood}
+\label{sec:under-the-hood}
 
@@
                     struct packet* p)
-/*@ requires ringp(r, ?user_property,
+/*@ requires ringp(r, ?user_property,^\label{line:precondition-from}^
                    ?lst, ?cap) &*&
-             lst != nil &*& packetp(p, _); @*/
-/*@ ensures ringp(r, user_property,
+             lst != nil &*& packetp(p, _); @*/^\label{line:precondition-to}^
+/*@ ensures ringp(r, user_property,^\label{line:postcondition-from}^
                   tail(lst), cap) &*&
             packetp(p, head(lst)) &*&
-            true == user_property(head(lst)); @*/
+            true == user_property(head(lst)); @*/^\label{line:postcondition-to}^
 {
-  //@ extract_first(r);
+  //@ extract_first(r); ^\label{line:lemma1}^
   struct packet* src_pkt = r->array + r->begin;
@@
     r->begin = 0;
-    //@ stitch_with_empty_overflow(r);
+    //@ stitch_with_empty_overflow(r);^\label{line:lemma2}^
   } else {
-    //@ stitch_with_empty(r);
+    //@ stitch_with_empty(r);^\label{line:lemma3}^
   }
@@
     \code{ring_pop_front} function. The comments starting with \code{@} symbol
-    are formal annotations that hold the function contract (\code{requires} -
-    precondition, \code{ensures} - postcondition) and lemmas that help the proof
-    checker to verify the state transformations in the code.}
+    are formal annotations that hold the function contract
+    (\code{requires}(ll.\ref{line:precondition-from}-\ref{line:precondition-to})
+    - precondition,
+    \code{ensures}(ll.\ref{line:postcondition-from}-\ref{line:postcondition-to})
+    - postcondition) and
+    lemmas(ll.\ref{line:lemma1},\ref{line:lemma2},\ref{line:lemma3}) that help
+    the proof checker to verify the state transformations in the code.}
   \label{lst:contract}
@@
 Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
-exhaustive SE, but preserves the vital propertiy --- \code{packet_constraints}
+exhaustive SE, but preserves the vital property --- \code{packet_constraints}
 (Fig.\ref{lst:customizations}).
@@
 \code{user_property} that entails \code{arg2.port != 9}, hence the trace is
-ligitimate. (see~\S\ref{sec:model-validation-criteria}).
+legitimate. (see~\S\ref{sec:model-validation-criteria}).
+
+\subsection{Misfit symptoms}
 

### 2016-11-22 doc/paper/paper.tex
port != 9 condition jorney accross the Vigor framework.

@@
 
-The ring keeps the state of our NF and makes many existing verification
+Our NF accumulates state in the ring thus making many existing verification
 approaches impractical.
@@
     if (!ring_full(r))
-      if (receive_packet(&p) && p.port != 9)
+      if (receive_packet(&p) && p.port != 9)^\label{line:if-cond}^
         ring_push_back(r, &p);
     if (!ring_empty(r) && can_send_packet()) {
-      ring_pop_front(r, &p);
+      ring_pop_front(r, &p);^\label{line:pop-front}^
       VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
@@
     switch(p) { case packet(port):
-                  return port != 9; }
+                  return port != 9; }^\label{line:vf-user-prop}^
   }
@@
 packet_constraints(struct packet* p) {
-  return p->port != 9;
+  return p->port != 9;^\label{line:c-pkt-cnstr}^
 }^\label{line:c-user-prop-to}^
@@
 indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
-the party here. When
-indexing an array with a symbolic variable (or when dereferencing a symbolic
-pointer), the constraints on program state become extremely complex, because the
-symbolic pointer or index could be referencing many different possible
-locations. The complexity of the constraints chokes the SMT solver, and the SEE
-grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
-offsets. Alas, stateful C code almost always has some of these.
+the party here. When indexing an array with a symbolic variable (or when
+dereferencing a symbolic pointer), the constraints on program state become
+extremely complex, because the symbolic pointer or index could be referencing
+many different possible locations. The complexity of the constraints chokes the
+SMT solver, and the SEE grinds to a halt. In general, SE is not practical for
+code with symbolic pointers or offsets. Alas, stateful C code almost always has
+some of these.
 
 In this example the symbolic indexing comes from the ring array. Once we
-replaced the infinite loop with a single iteration we have to assume (remember
+replaced the infinite loop with a single iteration we have to assume (due to
 havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
@@
                 "popped_packet");
-  ASSUME(packet_constraints(p));
+  ASSUME(packet_constraints(p));^\label{line:assume}^
 }
@@
 
-A model of \code{ring_pop_front} used for this example is shown on
+The model of \code{ring_pop_front} used for this example is shown on
 Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
@@
 
-/*@ assert(arg2.port != 9); @*/
+/*@ assert(arg2.port != 9); @*/^\label{line:assert}^
 \end{lstlisting}
@@
 legitimate. (see~\S\ref{sec:model-validation-criteria}).
+
+Let us follow now the `pord != 9' condition across the framework. During
+exhaustive SE, SEE first learns the constraint in the \code{if} condition:
+Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. It however forgets it on the next
+line, because our model of \code{ring_push_back} discards the packet being
+pushed. However, this constraint arises again on the call to
+\code{ring_pop_front}(l.\ref{line:pop-front}) as
+\code{packet_constraints}(Fig.~\ref{lst:customizations},l.\ref{line:c-pkt-cnstr}),
+according to the symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied
+to the fresh packet from the ring, it satisfies the \code{VIGOR_CHECK}
+(Fig.~\ref{lst:vigor} l.\ref{line:vigcheck}).
+
+During the operation, SEE produces call sequences that also capture the said
+constraint. E.g. the line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
+reflects the condition (Fig.~\ref{lst:customizations} l.\ref{line:c-pkt-cnstr})
+that was added by the symbolic model (Fig.~\ref{lst:symbolic-model}
+l.\ref{line:assume}).
+
+Formal verification serves two purposes:
+\begin{itemize}
+\item verification of the ring implementation. During this part, it proves that
+  \code{ring_pop_front} guarantees the popped packet to satisfy the
+  \code{user_property} (Fig.~\ref{lst:contract} l.\ref{line:postcondition-to}).
+\item validation of the call sequence. Here it uses the specific user property -
+  Fig.~\ref{lst:customizations} l.\ref{line:vf-user-prop}, to check it against
+  the assertion of the \code{ring_pop_front} output:
+  Fig.~\ref{lst:call-sequence} l.\ref{line:assert}.
+\end{itemize}
 

### 2016-11-24 doc/paper/paper.tex
Another round of Luis's feedback for the Solution Overview section.

@@
 \usepackage{hyperref}
+\usepackage[all]{hypcap}        % needed to help hyperlinks direct correctly
 \definecolor{mygreen}{rgb}{0.1,0.5,0.1}
@@
   \caption{Original implementation of the Drop protocol. Contains an infinite
-    loop (l.\ref{line:loop-begin}-\ref{line:loop-end}) and a latent ring
+    loop (ll.\ref{line:loop-begin}-\ref{line:loop-end}) and a latent ring
     implementation (declared on l.\ref{line:pkt-def},\ref{line:iter-def})}
@@
 interacts with the network using the following 3 functions:
-\code{receive_packet}(l.\ref{line:recv-pkt}), which non-blockingly reads an
-inbound packet and stores it at \code{packets[end]}, returning success or
-failure; \code{can_send_packet}(l.\ref{line:can-send}), which checks if a new
-packet can be sent; and \code{send_packet}(l.\ref{line:send-pkt}), which sends
-the packet stored at \code{packets[begin]}. The application follows a common
-practice: it consists of a main event loop, it has no other complex loops, and
-it uses a ring, a popular data structure in network programming.
-
-Our NF accumulates state in the ring thus making many existing verification
-approaches impractical.
-
-Fig.~\ref{lst:vigor} shows the actual code that we are going to verify. The
-first two steps show how we get there. \code{VIGOR_CHECK}
-(l.~\ref{line:vigcheck}) adds additional property that we would like to verify
---- that we do not forward packets with port 9.
+\code{receive_packet} (l.\ref{line:recv-pkt}), which non-blockingly reads an
+inbound packet and stores it at \code{packets[end]} -- the end of the ring
+buffer, returning success or failure; \code{can_send_packet}
+(l.\ref{line:can-send}), which checks if a new packet can be sent; and
+\code{send_packet} (l.\ref{line:send-pkt}), which sends the packet stored at
+\code{packets[begin]} -- the beginning of the ring buffer. The application
+follows a common practice: it consists of a main event loop, it has no other
+complex loops, and it uses a ring, a popular data structure in network
+programming. As a result, our NF accumulates state in the ring thus making many
+existing verification approaches impractical.
+
+Fig.~\ref{lst:vigor} shows the annotated code, ready for the verification with
+Vigor. It is the result of applying the first two steps shown below.
+\code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
+would like to verify --- that we do not forward packets with port 9.
 
@@
   \caption{Vigor implementation of Drop protocol.
-    \code{LOOP}(l.\ref{line:loop-hd}) labels the infinite loop head and we
-    explicitly use a ring(l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
+    \code{LOOP} (l.\ref{line:loop-hd}) labels the infinite loop head and we
+    explicitly use a ring (l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
     line \ref{line:vigcheck} adds a semantic property to be verified.}
@@
 infinite loops in the application code. A developer puts a label \code{LOOP}
-(l.\label{line:loophd}) into the condition in order to designate the
+(l.\ref{line:loop-hd}) into the condition in order to designate the
 potentially infinite loop. For a typical NF there should be only a handful or
@@
     packet condition in two languages (formal logic --
-    ll\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
-    ll\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
+    ll.\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
+    ll.\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
     limitations of our prototype.}
@@
 code with symbolic pointers or offsets. Alas, stateful C code almost always has
-some of these.
-
-In this example the symbolic indexing comes from the ring array. Once we
-replaced the infinite loop with a single iteration we have to assume (due to
-havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
-anywhere in the array.
-
-In Vigor, we circumvent the problem of symbolic indexing by encapsulating program
-state into well-defined data structures and putting them into a ``library of
-components'' (step 2). This works well for NF development, because the NF code
-uses a relatively small set of standard structures, such as ring buffers, hash
-tables, least-prefix-match tables, etc. We provide to the SEE symbolic models of
-these abstracted components that ``hide'' symbolic indexes and pointers from the
-SEE.
+some of these. In this example the symbolic indexing comes from the ring array.
+Once we replaced the infinite loop with a single iteration we have to assume
+(due to havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may
+point anywhere in the array.
+
+In Vigor, we circumvent the problem of symbolic indexing by encapsulating
+program state into well-defined data structures and putting them into a
+``library of components'' (e.g. the ring in the step 2). This works well for NF
+development, because the NF code uses a relatively small set of standard
+structures, such as ring buffers, hash tables, least-prefix-match tables, etc.
+We provide to the SEE symbolic models of these abstracted components that
+``hide'' symbolic indexes and pointers from the SEE.
 
 After the step 2 exhaustive SE takes less than 100 millisec regardless of the
-value of \code{CAP}. The fact that the code does not
-crash on any path amounts to a proof of crash freedom. Or does it? The validity
-of the proof depends on our library implementation of \code{ring} not
-crashing; and (2) the symbolic model of \code{ring} accurately capturing how the
-\code{ring} implementation works.
+value of \code{CAP}. The fact that the code does not crash on any path amounts
+to a proof of crash freedom. Or does it? The validity of the proof depends on
+our library implementation of \code{ring} not crashing; and the symbolic model
+of \code{ring} accurately capturing how the \code{ring} implementation works.
 
@@
 software turns out to require no verification expertise.
-
-Here is an example of the proven specification for the function
-\code{ring_pop_front}:
 
@@
     are formal annotations that hold the function contract
-    (\code{requires}(ll.\ref{line:precondition-from}-\ref{line:precondition-to})
+    (\code{requires} (ll.\ref{line:precondition-from}-\ref{line:precondition-to})
     - precondition,
-    \code{ensures}(ll.\ref{line:postcondition-from}-\ref{line:postcondition-to})
+    \code{ensures} (ll.\ref{line:postcondition-from}-\ref{line:postcondition-to})
     - postcondition) and
-    lemmas(ll.\ref{line:lemma1},\ref{line:lemma2},\ref{line:lemma3}) that help
+    lemmas (ll.\ref{line:lemma1},\ref{line:lemma2},\ref{line:lemma3}) that help
     the proof checker to verify the state transformations in the code.}
@@
 
-This C code is intermingled with the correctness proof annotations that help the
-formal verification engine (VeriFast) with small steps to transform the initial
-state, given in \code{requires} to the final state, specified in \code{ensures}.
-After proving the correctness of this transformation, the formal verification
-engine will use \code{requires/ensures} formulas as pre- and post- conditions
-for \code{ring_pop_front} contract. The trigraph \code{&*&} separates different
+Figure~\ref{lst:contract} shows the formal specification and a proven
+implementation of the example function \code{ring_pop_front}. This C code is
+intermingled with the correctness proof annotations that help the formal
+verification engine (VeriFast) with small steps to transform the initial state,
+given in \code{requires} to the final state, specified in \code{ensures}. After
+proving the correctness of this transformation, the formal verification engine
+will use \code{requires/ensures} formulas as pre- and post- conditions for
+\code{ring_pop_front} contract. The trigraph \code{&*&} separates different
 statements that hold together. For example
@@
 \item \code{r} points to a healthy ring of capacity \code{cap}, that contains
-  \code{tail(lst)} packets, that satisfy the \code{user_property}.
+  \code{tail(lst)} packets, that all satisfy the \code{user_property}.
 \item \code{p} points to a packet ``\code{head(lst)}''.
@@
 You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
-It starts with \code{loop_invariant_produce} that marks the beginning of a loop
-iteration. It continues then with calls to \code{ring_full} and \code{ring_empty},
-noting (with \code{assume}) that both return \code{false}. The seqeunce ends
-with a call to \code{ring_pop_front}. The last line detects underapproximation
-of the model output.
+It starts with \code{loop_invariant_produce} that is inserted automatically at
+the beginning of a loop iteration. It continues then with calls to
+\code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
+return \code{false}. They both return false in this particular trace, in other
+traces they returned other values, but to define the context for the last
+function call (l.\ref{line:tip-call}) we must repeat the history as closely as
+possible. Vigor handles the result of the last call in the sequence very
+differently. Instead of just telling the formal verification engine what
+happened, it inserts an \code{assert}(l.\ref{line:assert}) reflecting the
+symbolic model output to check that it is consistent with the postcondition of
+the last function (Fig.~\ref{lst:contract}).
 
@@
 loop_invariant_produce(&(arg1));
-/*@ open loop_invariant(_); @*/
+/*@ open loop_invariant(_); @*/^\label{line:unexplained1}^
 bool ret1 = ring_full(arg1);
@@
 //@ assume(ret2 == false);
-/*@ close packetp(&(arg2),
-                  packet((&(arg2))->port)); @*/
-
-ring_pop_front(arg1, &(arg2));
-//@ open packetp(&(arg2), _);
+/*@ close packetp(&(arg2),^\label{line:unexplained2}^
+                  packet((&(arg2))->port)); @*/^\label{line:unexplained3}^
+
+ring_pop_front(arg1, &(arg2));^\label{line:tip-call}^
+//@ open packetp(&(arg2), _);^\label{line:unexplained4}^
 
@@
   \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
-    constrain the execution space.}
+    constrain the execution space. Unexplained
+    lines(\ref{line:unexplained1},\ref{line:unexplained2},\ref{line:unexplained3},\ref{line:unexplained4})
+    are the automatically inserted helper steps to guide the proof checker.}
   \label{lst:call-sequence}
@@
 
-Let us follow now the `pord != 9' condition across the framework. During
+Let us follow now the `port != 9' condition across the framework. During
 exhaustive SE, SEE first learns the constraint in the \code{if} condition:
-Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. It however forgets it on the next
-line, because our model of \code{ring_push_back} discards the packet being
-pushed. However, this constraint arises again on the call to
-\code{ring_pop_front}(l.\ref{line:pop-front}) as
-\code{packet_constraints}(Fig.~\ref{lst:customizations},l.\ref{line:c-pkt-cnstr}),
-according to the symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied
-to the fresh packet from the ring, it satisfies the \code{VIGOR_CHECK}
-(Fig.~\ref{lst:vigor} l.\ref{line:vigcheck}).
+Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. Here, SEE just records it in the call
+sequence for model validation, and does not use further in the SE, because our
+model of \code{ring_push_back} discards the packet being pushed. However, the
+call to \code{ring_pop_front} (l.\ref{line:pop-front}) reintroduces this
+constraint formulated in \code{packet_constraints}
+(Fig.~\ref{lst:customizations},l.\ref{line:c-pkt-cnstr}), according to the
+symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied to the fresh
+packet from the ring, it satisfies the \code{VIGOR_CHECK} (Fig.~\ref{lst:vigor}
+l.\ref{line:vigcheck}).
 
@@
 
-\subsection{Misfit symptoms}
-
-Consider potential model mismatches. Suppose we had an overapproximate model ---
-Fig.~\ref{lst:overapproximate}. Then the \code{VIGOR_CHECK}
-(Fig.\ref{lst:vigor}) in the application code would fail.
+\subsection{Model inconsystency symptoms}
+
+The component developer produces a number of model that capture the component
+semantics in more or less details for different usecases. For a particular use
+case --- a NF, a particular model may be inappropriate. An inappropriate model
+may not compromise the verification result, because Vigor detects both
+overapproximate and underapproximate models.
+
+Suppose we had an overapproximate model --- Fig.~\ref{lst:overapproximate}. Then
+the popped packet would come unconstrained and the \code{VIGOR_CHECK}
+(Fig.\ref{lst:vigor} l.\ref{line:vigcheck}) in the application code would fail.
 \begin{figure}[h!]
@@
 Suppose now we had an underapproximate model: Fig.~\ref{lst:underapproximate}.
-Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) would look like:
+Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) in
+this case would look different:
 \begin{lstlisting}
@@
 a wider range for \code{arg2.port} than just \code{0}.
+
+Different models may work well for different usecases. E.g. the overapproximate
+model is just right for a buffer, that does not drop the packets with port 9.
+The underapproximate model would be just right for a buffer that forwards only the
+packets with port equeal to 0.
 

### 2016-11-29 doc/paper/paper.tex
Some minor rewording, thanks Luis for the feedback.

@@
 \code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
-would like to verify --- that we do not forward packets with port 9.
+would like to verify: ``we do not forward packets destined to port 9''.
 
@@
 \end{lstlisting}
-  \caption{Vigor implementation of Drop protocol.
+  \caption{Annotated code of Drop buffer.
     \code{LOOP} (l.\ref{line:loop-hd}) labels the infinite loop head and we
@@
 
-{\bf Step 2: Use library components. } A developer replaces every array with a
-component from Vigor library that performs the same function. For our example,
-the library provides a packet ring with a simple interface: \code{bool
-  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
-  ring_full(ring*)} checks whether the ring has run out of space, \code{void
-  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
-\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
-and places it in \code{dst}.
+{\bf Step 2: Use library components. } A developer replaces state chunks (e.g.
+most of the arrays) with components from the Vigor library that perform the
+same function. For our example, the library provides a packet ring with a simple
+interface: \code{bool ring_empty(ring*)} checks whether there is a packet in the
+ring, \code{bool ring_full(ring*)} checks whether the ring has run out of space,
+\code{void ring_push_back(ring*, packet* src)} adds packet \code{src} to the
+ring, and \code{void ring_pop_front(ring*, packet* dst)} removes a packet from
+the ring and places it in \code{dst}.
 
 {\bf Step 3: Provide invariants.} Besides usual parameters, like capacity or
-element size, most Vigor components also require simple logical invariants of
+element size, most Vigor components also require simple logical invariants about
 their elements to enable the reasoning about the application code. For our
@@
 Vigor transforms infinite loops using loop
-invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
-automatic havoc-ing~\cite{barnett2005boogie}. Havoc-ing simply replaces all the
-variables that may change in the loop by unconstrained symbols at the beginning
-of the iteration. In our example that would be the variables \code{begin} and
-\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
-of the \code{packets} pointer because it never changes. The step 1 enables
-Vigor to reduce each problematic loop to a single generalized iteration that
-allows reasoning in one pass.
+invariants~\cite[\S~2.1]{cormen2009introduction} that allow to model all the
+iterations of a loop in a single one. To support this kind of generalization, we
+added automatic havoc-ing~\cite{barnett2005boogie} to KLEE. Havoc-ing simply
+replaces all the variables that may change in the loop by unconstrained symbols
+at the beginning of the iteration. In our example that would be the variables
+\code{begin} and \code{queue_len} and the contents of \code{packets[0..CAP]},
+but not the value of the \code{packets} pointer because it never changes. The
+step 1 enables Vigor to reduce each problematic loop to a single generalized
+iteration that allows reasoning in one pass.
 
@@
 program state into well-defined data structures and putting them into a
-``library of components'' (e.g. the ring in the step 2). This works well for NF
+``library of components'' (e.g. the ring in step 2). This works well for NF
 development, because the NF code uses a relatively small set of standard
@@
 \end{lstlisting}
-  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
-    a packet and adds a constraint to it.}
+  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically a
+    packet and adds a constraint to it.}
   \label{lst:symbolic-model}
@@
 You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
-It starts with \code{loop_invariant_produce} that is inserted automatically at
-the beginning of a loop iteration. It continues then with calls to
+It starts with \code{loop_invariant_produce} and continues then with calls to
 \code{ring_full} and \code{ring_empty}, noting (with \code{assume}) that both
@@
   \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
-    constrain the execution space. Unexplained
-    lines(\ref{line:unexplained1},\ref{line:unexplained2},\ref{line:unexplained3},\ref{line:unexplained4})
-    are the automatically inserted helper steps to guide the proof checker.}
+    constrain the execution space. The \code{loop_invariant_produce} call is
+    inserted automatically at the beginning of a loop iteration. Lines
+    \ref{line:unexplained1},\ref{line:unexplained2},\ref{line:unexplained3},\ref{line:unexplained4}
+    account for the automatically inserted helper steps to guide the proof
+    checker.}
   \label{lst:call-sequence}
@@
 
-Formal verification engine checks this sequence using the contracts like in the
-Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g. call for
-\code{ring_pop} on a possibly empty ring. It then checks that the model output
-covers all the possible outputs specified in the contract for the tip function.
-This happens in the last line on the listing. The assert contains the
+The formal verification engine checks this sequence using contracts like the one
+shown in the Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g.
+call for \code{ring_pop} on a possibly empty ring. It then checks that the model
+output covers all the possible outputs specified in the contract for the last
+function. This happens in the last line on the listing. The assert contains the
 path-constraint generated by the symbolic model. In this case the contract
@@
 
-\subsection{Model inconsystency symptoms}
+\subsection{Detecting model inconsistencies}
 

### 2016-12-01 doc/paper/paper.tex
Restructured second section according to the last project meeting.

@@
 
-\subsection{Developer experience}
+\subsection{Developer Experience}
 
@@
     loop (ll.\ref{line:loop-begin}-\ref{line:loop-end}) and a latent ring
-    implementation (declared on l.\ref{line:pkt-def},\ref{line:iter-def})}
+    implementation (declared on ll.\ref{line:pkt-def},\ref{line:iter-def})}
   \label{lst:orig}
@@
 existing verification approaches impractical.
+
+Vigor provides two things: a verified library of components and a verification
+toolchain. Each component consists of an API with formal specification, verified
+implementation and a set of symbolic models. The toolchain verifies the NF code
+and stitches it with the relevant components to produce a single reliable
+application.
 
@@
 
-\subsection{Under the hood}
+\subsection{Under The Hood}
 \label{sec:under-the-hood}
+
+\subsubsection{Symbolic Execution}
 
@@
 invariants~\cite[\S~2.1]{cormen2009introduction} that allow to model all the
-iterations of a loop in a single one. To support this kind of generalization, we
-added automatic havoc-ing~\cite{barnett2005boogie} to KLEE. Havoc-ing simply
-replaces all the variables that may change in the loop by unconstrained symbols
-at the beginning of the iteration. In our example that would be the variables
-\code{begin} and \code{queue_len} and the contents of \code{packets[0..CAP]},
-but not the value of the \code{packets} pointer because it never changes. The
-step 1 enables Vigor to reduce each problematic loop to a single generalized
-iteration that allows reasoning in one pass.
+iterations of a loop in a single one. To support this, we added automatic
+havoc-ing~\cite{barnett2005boogie} to KLEE. Havoc-ing simply replaces all the
+variables that may change in the loop by unconstrained symbols at the beginning
+of the iteration. In our example that would be the variables \code{begin} and
+\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
+of the \code{packets} pointer itself because it never changes. Step 1 enables Vigor
+to reduce each problematic loop to a single generalized iteration that allows
+reasoning in one pass.
 
@@
 ``symbolic
-indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
-the party here. When indexing an array with a symbolic variable (or when
+indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} ---
+spoils the party here. When indexing an array with a symbolic variable (or when
 dereferencing a symbolic pointer), the constraints on program state become
@@
 We provide to the SEE symbolic models of these abstracted components that
-``hide'' symbolic indexes and pointers from the SEE.
-
-After the step 2 exhaustive SE takes less than 100 millisec regardless of the
-value of \code{CAP}. The fact that the code does not crash on any path amounts
-to a proof of crash freedom. Or does it? The validity of the proof depends on
-our library implementation of \code{ring} not crashing; and the symbolic model
-of \code{ring} accurately capturing how the \code{ring} implementation works.
+``hide'' symbolic indexes and pointers from the SEE. After the step 2 exhaustive
+SE takes less than 100 millisec regardless of the value of \code{CAP}.
+
+SE is a first step on the way of a complete proof. After the exhaustive SE found
+no crashes in the application code, we need to verify for each component (e.g.
+for the ring component) that (1) the actual component implementation does not
+crash itself and follows the declared API specification, and (2) the model used
+during SE faithfully approximated the same specification.
+
+\subsubsection{Formal Verification}
 
@@
 spec for \code{ring} and a proof that the \code{ring} implementation satisfies
-the spec. This is how we enable SE to verify the high level code, while using
+the spec. This is how we enable SE to verify the NF code, while using
 separation logic to verify the library of components --- two different
-verifications techniques for two different parts of the complete application.
+verification techniques for two different parts of the complete application.
 Doing the library verification requires deep expertise in program verification,
@@
 software turns out to require no verification expertise.
+
+Figure~\ref{lst:contract} shows the formal specification and a proven
+implementation of the example function \code{ring_pop_front}. This C code is
+intermingled with the correctness proof annotations that help the formal
+verification engine (VeriFast) with small steps to transform the initial state,
+given in \code{requires} to the final state, specified in \code{ensures}. After
+proving the correctness of this transformation, the formal verification engine
+will use \code{requires/ensures} formulas as pre- and post- conditions for
+\code{ring_pop_front} contract. The trigraph \code{&*&} separates different
+statements that hold together. For example
+
+\begin{lstlisting}
+ringp(r, user_property, tail(lst), cap) &*&
+packetp(p, head(lst)) &*&
+true == user_property(head(lst));
+\end{lstlisting}
+means:
+\begin{itemize}
+\item \code{r} points to a healthy ring of capacity \code{cap}, that contains
+  \code{tail(lst)} packets, that all satisfy the \code{user_property}.
+\item \code{p} points to a packet ``\code{head(lst)}''.
+\item ``\code{head(lst)}'' satisfies the \code{user_property}.
+\end{itemize}
+
+Note that \code{user_property(head(lst));} in the post condition ensures
+\code{p->port != 9} property (Fig.~\ref{lst:customizations}).
 
@@
 
-Figure~\ref{lst:contract} shows the formal specification and a proven
-implementation of the example function \code{ring_pop_front}. This C code is
-intermingled with the correctness proof annotations that help the formal
-verification engine (VeriFast) with small steps to transform the initial state,
-given in \code{requires} to the final state, specified in \code{ensures}. After
-proving the correctness of this transformation, the formal verification engine
-will use \code{requires/ensures} formulas as pre- and post- conditions for
-\code{ring_pop_front} contract. The trigraph \code{&*&} separates different
-statements that hold together. For example
-
-\begin{lstlisting}
-ringp(r, user_property, tail(lst), cap) &*&
-packetp(p, head(lst)) &*&
-true == user_property(head(lst));
-\end{lstlisting}
-means:
-\begin{itemize}
-\item \code{r} points to a healthy ring of capacity \code{cap}, that contains
-  \code{tail(lst)} packets, that all satisfy the \code{user_property}.
-\item \code{p} points to a packet ``\code{head(lst)}''.
-\item ``\code{head(lst)}'' satisfies the \code{user_property}.
-\end{itemize}
-
-Note that \code{user_property(head(lst));} in the post condition ensures
-\code{p->port != 9} property (Fig.~\ref{lst:customizations}).
+\subsubsection{Symbolic Models}
+
+The real implementation, even the correct one will inevitably complicate SE
+of the NF code bringing back all the problems that we were so carefully
+avoiding. To make the exhaustive SE feasible, Vigor uses symbolic models,
+simplified implementation with small state that approximately mimics the real
+thing in most cases.
 
@@
                     struct packet* p) {
-  FILL_SYMBOLIC(p, sizeof(struct packet),
+  FILL_SYMBOLIC(p, sizeof(struct packet), ^\label{line:fill-symbolic}^
                 "popped_packet");
@@
   \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically a
-    packet and adds a constraint to it.}
+    packet (l.\ref{line:fill-symbolic}) and adds a constraint to it
+    (l.\ref{line:assume}).}
   \label{lst:symbolic-model}
@@
 (Fig.\ref{lst:customizations}).
+
+\subsubsection{Model Validation}
 
@@
 output covers all the possible outputs specified in the contract for the last
-function. This happens in the last line on the listing. The assert contains the
+function. This happens on the line~\ref{line:assert}. The assert contains the
 path-constraint generated by the symbolic model. In this case the contract
@@
 constraint formulated in \code{packet_constraints}
-(Fig.~\ref{lst:customizations},l.\ref{line:c-pkt-cnstr}), according to the
+(Fig.~\ref{lst:customizations}~l.\ref{line:c-pkt-cnstr}), according to the
 symbolic model - Fig.~\ref{lst:symbolic-model}. Being applied to the fresh
-packet from the ring, it satisfies the \code{VIGOR_CHECK} (Fig.~\ref{lst:vigor}
-l.\ref{line:vigcheck}).
+packet from the ring, it satisfies the \code{VIGOR_CHECK}
+(Fig.~\ref{lst:vigor}~l.\ref{line:vigcheck}).
 
@@
 
-\subsection{Detecting model inconsistencies}
-
-The component developer produces a number of model that capture the component
+\subsubsection{Detecting model inconsistencies}
+
+The component developer produces several model that capture the component
 semantics in more or less details for different usecases. For a particular use
@@
 may not compromise the verification result, because Vigor detects both
-overapproximate and underapproximate models.
+over- and under-approximations.
 
@@
 \end{lstlisting}
-The assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies
-a wider range for \code{arg2.port} than just \code{0}.
+Vigor model validation phase flags this inconsistency. Specifically, the
+assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies a
+wider range for \code{arg2.port} than just \code{0}.
 
@@
 The underapproximate model would be just right for a buffer that forwards only the
-packets with port equeal to 0.
+packets with port equal to 0.
 

### 2016-12-05 doc/paper/paper.tex
Minor corrections.

@@
 port 9 (``discard port''), and forwards the others through another interface.
-A possible implementation is on the Fig.~\ref{lst:orig}.
+A possible implementation is on Fig.~\ref{lst:orig}.
 
@@
 
-Fig.~\ref{lst:vigor} shows the annotated code, ready for the verification with
-Vigor. It is the result of applying the first two steps shown below.
-\code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds additional property that we
-would like to verify: ``we do not forward packets destined to port 9''.
+Annotated code ready for verification with Vigor is shown on
+Fig.~\ref{lst:vigor}. It is the result of applying the first two steps shown
+below. \code{VIGOR_CHECK} (l.\ref{line:vigcheck}) adds an additional property
+that we would like to verify: ``we do not forward packets destined to port 9''.
 
@@
 Vigor makes it possible for NF developers to verify their code without having to
-understand verification. Note that stating invariants(step 3) for dataplane
+understand verification. Note that stating invariants (step 3) for dataplane
 software turns out to require no verification expertise.
@@
 return \code{false}. They both return false in this particular trace, in other
-traces they returned other values, but to define the context for the last
-function call (l.\ref{line:tip-call}) we must repeat the history as closely as
-possible. Vigor handles the result of the last call in the sequence very
-differently. Instead of just telling the formal verification engine what
-happened, it inserts an \code{assert}(l.\ref{line:assert}) reflecting the
-symbolic model output to check that it is consistent with the postcondition of
-the last function (Fig.~\ref{lst:contract}).
+traces they return other values, but to define the context for the last function
+call (l.\ref{line:tip-call}) we must repeat the history as closely as possible.
+Vigor handles the result of the last call in the sequence very differently.
+Instead of just telling the formal verification engine what happened, it inserts
+an \code{assert}(l.\ref{line:assert}) reflecting the symbolic model output to
+check that it is consistent with the postcondition of the last function
+(Fig.~\ref{lst:contract}).
 
@@
 Let us follow now the `port != 9' condition across the framework. During
-exhaustive SE, SEE first learns the constraint in the \code{if} condition:
-Fig.~\ref{lst:vigor} l.\ref{line:if-cond}. Here, SEE just records it in the call
-sequence for model validation, and does not use further in the SE, because our
-model of \code{ring_push_back} discards the packet being pushed. However, the
-call to \code{ring_pop_front} (l.\ref{line:pop-front}) reintroduces this
-constraint formulated in \code{packet_constraints}
+exhaustive SE, SEE first learns the constraint in the \code{if} condition
+(Fig.~\ref{lst:vigor} l.\ref{line:if-cond}). Here, SEE just records it in the
+call sequence for model validation, and does not use it further in the SE,
+because our model of \code{ring_push_back} discards the packet being pushed.
+However, the call to \code{ring_pop_front} (l.\ref{line:pop-front}) reintroduces
+this constraint formulated in \code{packet_constraints}
 (Fig.~\ref{lst:customizations}~l.\ref{line:c-pkt-cnstr}), according to the
@@
 During the operation, SEE produces call sequences that also capture the said
-constraint. E.g. the line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
+constraint. For example, line~\ref{line:assert} on Fig.~\ref{lst:call-sequence}
 reflects the condition (Fig.~\ref{lst:customizations} l.\ref{line:c-pkt-cnstr})
@@
 
-The component developer produces several model that capture the component
-semantics in more or less details for different usecases. For a particular use
+The component developer produces several models that capture the component
+semantics in more or less details for different use cases. For a particular use
 case --- a NF, a particular model may be inappropriate. An inappropriate model
@@
 
-Suppose now we had an underapproximate model: Fig.~\ref{lst:underapproximate}.
+Suppose now we had an underapproximate model --- Fig.~\ref{lst:underapproximate}.
 Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) in
@@
 
-Different models may work well for different usecases. E.g. the overapproximate
-model is just right for a buffer, that does not drop the packets with port 9.
-The underapproximate model would be just right for a buffer that forwards only the
-packets with port equal to 0.
+Different models may work well for different usecases. For instance the
+overapproximate model is just right for a buffer, that forwards all the packets
+(including the aimed to port 9). The underapproximate model would be just right
+for a buffer that forwards only the packets with port equal to 0.
 

### 2017-06-26 doc/expiring-map.tex
publish the current vnds master snapshot

\documentclass{article}
\usepackage{dsfont}
\usepackage{amsmath}
\usepackage[margin=1.5cm]{geometry}
\renewcommand{\o}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\p}[1]{\ensuremath{\mathit{#1}}}
\newcommand{\s}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\nintset}{\ensuremath{\mathds{Z}_{\ge 0}}}
\newcommand{\moreset}[1]{\ensuremath{\mathds{Z}_{\ge #1}}}

\begin{document}

An ``expiring map'' $[K,V]$ is an abstraction that maps keys $k \in K$ to values $v \in V$ and allows individual mappings to expire at an indicated time.

\paragraph{Abstract state $S$}
\begin{itemize}
   \item \s{mapping}: a set of entries $[k,v,t]$, with each entry mapping key $k$ to value $v$ until time $t$, after which the mapping expires
   \item \s{size}: the number of unexpired keys in \s{mapping} (for now, we assume at most one entry per key)
   \item \s{maxSize}: the maximum number of keys that the map can accommodate
\end{itemize}

\paragraph{Abstract operations}
\begin{itemize}
\item $\o{constructor}(s \in \nintset$) constructs the map that holds at most $s$ entries
\item $p\gets \o{has}(k \in K)$ indicates whether key $k$ exists in the map or not
\item $(v\in V, e\in \nintset)\gets \o{get}(k \in K)$ returns the pair of value $v$ to which $k$ is mapped and time $e$ until this mapping is kept. If the key $k$ is not present, can return $fail$.
\item $x\gets \o{put}(k \in K, v \in V, e \in \nintset)$ adds to the map a mapping from $k$ to $v$ that is valid until time $e$. Returns boolean $x$ - whether the operation was successfull.
\item $x\gets \o{erase}(k \in K)$ removes the entry from the map that maps $k$ to a value. Returns boolean $x$ - whether the operation was successfull.
  %\item $\o{nop}$ does nothing, returns nothing.
\end{itemize}

We model the use of the map as a finite sequence
$\tau = <o_0, o_1, ..., o_n>$
of operations $o_t$ on the map indexed by time $t \in \nintset$.
In our model, an operation completes fully before the next one commences (i.e., there is no concurrency).
%A continuous subsequence $<o_a, o_{a+1}, ..., o_{b-1}, o_b> (0\le a\le b\le n)$ of $\tau$ is designated as $\tau_a^b$
For simplicity, we assume that each operation takes exactly one unit of time.
We use the shorthand $\o{operation}_t(...)$ to indicate the operation that occurs at time $t$.

An execution is an interleaving of operations and ``snapshots'' of the abstract state
$\epsilon = <o_0, S_0, o_1, S_1, ..., o_n, S_n>$, with
each subsequence $<S_{i-1}, o_i, S_i>$ indicating that operation $o_i$ transforms abstract state $S_{i-1}$ into $S_i$.
An accessor does not mutate state, so it leaves the abstract state unchanged, but the state can change due to continuous expiration process.
We use the shorthand $\s{component}_t$ to indicate the value of \s{component} of the state $S$ immediately after $o_t$ completes and before $o_{t+1}$ commences.

{\it Note:} Symbol $\simeq$ in the following context: $o_t\simeq ...$ means that lefthand {\it operation} ($o_t$) can be equal to the righthand operation. E.g. $\forall k \in K o_t\simeq \o{has}_t(k) \Leftrightarrow \p{const}_t$ means, that the operation number t ($o_t$) may be the $\o{has}(k)$ call (not the result of it) if and only if holds $\p{const}_t$. \underline{Attention}: $x=\o{has}_t(k)$ has different meaning then $o_t=\o{has}_t(k)$.

By default, if symbols $e, k, v, t$ are used in a formula, and not mentioned before in that formula, the universal quantifier $\forall$ is implied: $\forall k \in K, v\in V, e\in \nintset, t\in \nintset$.

\paragraph{Helper predicates}
\begin{itemize}
\item An ``empty'' map is one without any keys \\
  $\p{empty}_t \Leftrightarrow [k, v, e] \not\in \s{mapping}_t$
\item A ``const'' operation does not modify nonexpired mappings:\\
  $\p{const}_t ~\Leftrightarrow~\\ \Big([k, v, e] \in \s{mapping}_{t-1} \wedge e>t \Rightarrow [k, v, e] \in \s{mapping}_t\Big)\wedge\\ \Big([k, v, e] \not\in \s{mapping}_{t-1} \Rightarrow [k, v, e] \not\in \s{mapping}_t\Big)\wedge \s{size}_{t-1} \ge \s{size}_t$
\item An ``isolated'' operation does not modify nonexpired mappings, except for a single one:\\
  $\p{isolated}_t(k'\in K) ~\Leftrightarrow~\\ \Big(k\neq k' \wedge [k, v, e] \in \s{mapping}_{t-1} \wedge e>t \Rightarrow [k, v, e] \in \s{mapping}_t\Big)\wedge\\ \Big(k\neq k' \wedge [k, v, e] \not\in \s{mapping}_{t-1} \Rightarrow [k, v, e] \not\in \s{mapping}_t\Big)$
\end{itemize}

\paragraph{Invariants}
These predicates allways hold:
\begin{itemize}
\item The map does not keep expired items: $\forall [k, v, e] \in \s{mapping}_t \Rightarrow e > t$
\item No operation except for $\o{constructor}_0$ can change the map capacity:
  \begin{gather*}
    \forall t\in \nintset: t > 0 \Rightarrow \s{maxSize}_t=\s{maxSize}_0
  \end{gather*}
\end{itemize}

\paragraph{Semantics for accessors (may mutate $S$, do not change the nonexpired mappings)}
\begin{itemize}
\item[has]
  \begin{itemize}
  \item $\o{has}_t(k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in \s{mapping}_t \wedge o_t=\o{has}_t(k)$\\
  \item It does not change mappings:
    \begin{gather*}
      o_t \simeq \o{has}_t(k) ~\Leftrightarrow~ \p{const}_t
    \end{gather*}
  \end{itemize}
\item[get]
  \begin{itemize}
  \item One can get a mapping only if it exists in the map:
      \begin{gather}
        \o{get}_t(k \in K) \rightarrow fail ~\Leftrightarrow~ [k,v,e] \not\in \s{mapping}_t \wedge o_t=\o{get}_t(k)\label{eq:get1}\\
        (v\in V,e\in \nintset)\gets \o{get}_t(k \in K) ~\Leftrightarrow~ \exists [k,v,e] \in \s{mapping}_t \wedge o_t=\o{get}_t(k) \label{eq:get2}
      \end{gather}
  \item In anyway, it does not change existing unexpired mappings:
    \begin{gather*}
      o_t \simeq \o{get}_t(k) ~\Leftrightarrow~ \p{const}_t
    \end{gather*}
  \end{itemize}
\end{itemize}

\paragraph{Semantics for operations that mutate state}
\begin{itemize}
\item[constructor]
  \begin{itemize}
  \item The constructor forms a new empty map
    \begin{gather*}
      \o{constructor}_0(s \in \nintset) \Leftrightarrow \p{empty}_0 \wedge \big(\forall t \in \nintset \Rightarrow \s{maxSize}_t=s\big)
    \end{gather*}
  \item The constructor must be called once and only once, at the very beginning of $\tau$ (i.e., $\tau$ contains no other constructor call after the first one)
    \begin{gather*}
      o_0 \equiv \o{constructor}_0 \wedge \big(\forall t\in \moreset{1} \Rightarrow o_t \neq \o{constructor}_t\big)
    \end{gather*}
  \end{itemize}
\item[erase]
  \begin{itemize}
  \item ``Erase'' deletes an entry with a given key. If the entry does not exist in the \s{mapping}, it fails. It does not change other mappings if they are not expired.
    \begin{gather*}
      \Big(o_t \simeq \o{erase}_t(k\in K) ~\Leftrightarrow~ \p{isolated}_t(k)\Big) \wedge
    \end{gather*}
  \item If there was a such mapping for $k$, ``erase'' will remove it.
    \begin{gather}
      \Big(x=\o{erase}_t(k\in K) \wedge \exists [k, v, e] \in \s{mapping}_{t-1} ~\Leftrightarrow~\nonumber\\
      x=success \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in \s{mapping}_t\big) \wedge \s{size}_t \le \s{size}_{t-1}-1 \Big) \wedge
      \label{eq:erase1}
    \end{gather}
  \item If there was not --- it will fail.
    \begin{gather}
      \Big( x=\o{erase}_t(k\in K) \wedge [k,v,e]\not\in \s{mapping}_{t-1} ~\Leftrightarrow~ \nonumber\\
      x=fail \wedge \p{const}_t \wedge o_t=\o{erase}_t(k)\Big)
      \label{eq:erase2}
    \end{gather}
  \end{itemize}
\item[put]
  \begin{itemize}
  \item ``Put'' inserts a new mapping, if the number of unexpired elements is less than \s{maxSize} and there is no unexpired entry with the same key.
    This function does not touch any unexpired entry with key other than $k$.
    \begin{gather*}
      \Big(o_t \simeq \o{put}_t(k\in K, v\in V, e\in \moreset{t}) ~\Leftrightarrow~ \p{isolated}_t(k)\Big) \wedge
    \end{gather*}
  \item If there was no such mupping before, and there is enough room, it will succeed.
    \begin{gather}
      \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} < \s{maxSize}_{t-1} \wedge \big(\forall v'\in V, e'\in \nintset \Rightarrow [k, v', e'] \not\in \s{mapping}_{t-1}\big) \Leftrightarrow \nonumber\\
      x=success \wedge \s{size}_t \le \s{size}_{t-1}+1 \wedge [k, v, e] \in \s{mapping}_t\Big)\wedge
      \label{eq:put1}
    \end{gather}
  \item If a mapping with key $k$ allready exists, and it does not expires on $t$, put will fail.
    \begin{gather}
      \Big(x=\o{put}_t(k, v, e) \wedge \big(\exists v'\in V, e'\in \nintset: [k, v', e']\in \s{mapping}_{t-1} \wedge e > t\big) \Leftrightarrow\nonumber\\
      x=fail \wedge \p{const}_t \wedge o_t = \o{put}_t(k,v,e)\Big)\wedge
      \label{eq:put2}
    \end{gather}
  \item If a mapping with key $k$ expires on $t$, put will replace it:
    \begin{gather}
      \Big(x=\o{put}_t(k, v, e) \wedge \big(\exists v'\in V: [k, v', t]\in \s{mapping}_{t-1}\big) \Leftrightarrow\nonumber\\
      x=success \wedge \s{size}_t \le \s{size}_{t-1} \wedge [k, v, e] \in \s{mapping}_t\Big)\wedge
      \label{eq:put3}
    \end{gather}
  \item If there is no room, but some mappings expire on $t$, put will succeed:
    \begin{gather}
      \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} = \s{maxSize}_{t-1} \wedge [k, v, e] \not\in \s{mapping}_{t-1} \wedge \nonumber\\
      \big(\exists k'\in K, v'\in V: [k', v', t] \in \s{mapping}_{t-1}\big) \Leftrightarrow\nonumber\\
      x=success \wedge \s{state}_t \le \s{state}_{t-1} \wedge [k, v, t] \in \s{mapping}_t\Big)\wedge
      \label{eq:put4}
    \end{gather}
  \item If there is no room, and no mappings expire on $t$, put will fail:
    \begin{gather}
      \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} = \s{maxSize}_{t-1} \wedge \nonumber\\
      \big(\forall k'\in K, v'\in V, e' \in \nintset: [k', v', e'] \in \s{mapping}_{t-1} \Rightarrow e' > t \big) \Leftrightarrow\nonumber\\
      x=fail \wedge \p{const}_t \wedge o_t = \o{put}_t(k,v,e)\Big)
      \label{eq:put5}
    \end{gather}
  \end{itemize}

\end{itemize}

\paragraph{Properties to be proven}
\begin{itemize}
\item $\p{empty}_t \Leftrightarrow \s{size}_t=0$
\item If the map is empty, there no keys in it:
  \begin{gather*}
    \p{empty}_t \Leftrightarrow \o{has}_t(k) = false
  \end{gather*}
\item Equivalent helper predicate \p{isolated} definition
  \begin{gather*}
    \p{isolated}_t(k'\in K) ~\Leftrightarrow~\nonumber\\
    \Big(k\neq k' \wedge \o{has}_{t-1}(k) \wedge (v,e)=\o{get}_{t-1}(k) \wedge e>t \Rightarrow \o{has}_t(k) \wedge \o{get}_t(k)=(v,e)\Big)\wedge\nonumber\\
    \Big(k\neq k' \wedge \neg \o{has}_{t-1}(k) \Rightarrow \neg\o{has}_t(k)\Big)
  \end{gather*}
\item Equivalent helper predicate \p{const} definition
  \begin{gather*}
    \p{const}_t ~\Leftrightarrow~\nonumber\\
    \Big(\o{has}_{t-1}(k) \wedge (v,e)=\o{get}_{t-1}(k) \wedge e>t \Rightarrow \o{has}_t(k) \wedge \o{get}_t(k)=(v,e)\Big)\wedge\nonumber\\
    \Big(\forall k\in K: \neg \o{has}_{t-1}(k) \Rightarrow \neg\o{has}_t(k)\Big)
  \end{gather*}

\item Equivalent of \eqref{eq:get1}:
  \[
  \o{has}_t(k) \Leftrightarrow \o{get}_t(k) \not\rightarrow fail
  \]
\item If there was a mapping for $k$, ``erase'' removes it. Equivalent of \eqref{eq:erase1}:
  \begin{gather*}
    \Big(x=\o{erase}_t(k\in K) \wedge \o{has}_{t-1}(k) ~\Leftrightarrow~ x=success \wedge \neg\o{has}_t(k) \wedge \s{size}_t \le \s{size}_{t-1}-1 \Big)
  \end{gather*}
\item If there were not --- ``erase'' will fail. Equivalent of \eqref{eq:erase2}:
  \begin{gather*}
    \Big(x=\o{erase}_t(k\in K) \wedge \neg\o{has}_{t-1}(k) ~\Leftrightarrow~ x=fail\wedge \p{const}_t\Big)
  \end{gather*}

\item Equivalent of \eqref{eq:put1}
  \begin{gather*}
    \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} < \s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(k) \Leftrightarrow \nonumber\\
    x=success \wedge \s{size}_t \le \s{size}_{t-1}+1 \wedge \o{has}_t(k)\wedge\o{get}_t(k) = (v, e)\Big)
  \end{gather*}

\item If there is an unexpired mapping with the same key, put will fail. Equivalent of \eqref{eq:put2}
  \begin{gather*}
    \Big(x=\o{put}_t(k, v, e) \wedge \o{has}_{t-1}(k) \wedge \exists e'\in \nintset: (\cdot, e')=\o{get}_{t-1}(k) \wedge e' > t \Leftrightarrow\\
    x=fail \wedge \p{const}_t\Big)
  \end{gather*}

\item Equivalent of \eqref{eq:put3}. Expired mapping replaced:
  \begin{gather*}
    \Big(x=\o{put}_t(k, v, e) \wedge \o{has}_{t-1}(k) \wedge (\cdot, t)=\o{get}_{t-1}(k) \Leftrightarrow\\
    x=success \wedge \s{size}_t \le \s{size}_{t-1} \wedge \o{has}_t(k)\wedge\o{get}_t(k) = (v, e)\Big)
  \end{gather*}

\item Equivalent of \eqref{eq:put4}. If there is no room, but some mappings have expired, one still can put.
  \begin{gather*}
    \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} = \s{maxSize}_{t-1} \wedge \neg\o{has}_{t-1}(k) \wedge\\
    \big(\exists k'\in K: \o{has}_{t-1}(k')\wedge (\cdot,t)=\o{get}_{t-1}(k) \big) \Leftrightarrow\\
    x=success \wedge \o{has}_t(k)\wedge\o{get}_t(k) = (v, e) \wedge \s{size}_t \le \s{size}_{t-1}\Big)
\end{gather*}

\item Into full map, one can not put any more keys. Equivalent of \eqref{eq:put5}:
  \begin{gather*}
    \Big(x=\o{put}_t(k, v, e) \wedge \s{size}_{t-1} = \s{maxSize}_{t-1} \wedge  \\
    \big(\forall k'\in K,e'\in \nintset :\o{has}_{t-1}(k')\wedge(\cdot,e')=\o{get}_{t-1}(k') \Rightarrow e' > t \big) \Leftrightarrow\\
    x=fail \wedge size_t = size_{t-1} \wedge \p{const}_t(k)\Big)
  \end{gather*}

\item The size of the map is always properly bounded
  \begin{gather*}
    \forall t\in \nintset: t>0  \wedge \o{constructor}_0(s \in \nintset) \Rightarrow 0 \le \s{size}_t \le \s{maxSize}
  \end{gather*}

\item For any subsequence of operations other than \o{put}, size of the map may only decrease:
  \begin{gather*}
    \forall a\in \nintset,b\in \nintset: a<b \wedge \big(\forall t\in \nintset:a<t<b \Rightarrow o_t \neq \o{put}_t \big) \Rightarrow \s{size}_a \ge \s{size}_b
  \end{gather*}

\item If a subsequence of operations does not include \o{erase} for key $k\in K$, and there is a mapping, it will be kept until the end of the subsequence, or until expired:
  \begin{gather*}
    \forall a\in \nintset,b, k\in K, v\in V, e\in \moreset{b}: a < b \wedge \o{has}_a(k) \wedge (v, e)=\o{get}_a{k} \wedge e > b ~ \wedge\nonumber\\
    \big(\forall t\in \nintset: a<t<b \Rightarrow o_t\neq \o{erase}_t(k)\big) \Rightarrow\nonumber\\
    \o{has}_b(k) \wedge \o{get}_b(k) = (v,e)
  \end{gather*}

\end{itemize}

\end{document}

### 2017-06-26 doc/paper/example/loop-model.h

@@ -1,0 +1,25 @@
+#ifndef _LOOP_MODEL_H_INCLUDED_
+#define _LOOP_MODEL_H_INCLUDED_
+
+#include "user-params.h"
+
+#include "ring.h"
+
+/*@
+  predicate loop_invariant(struct ring* rp) =
+    ringp(rp, user_packet_property, _, RING_CAPACITY);
+  @*/
+
+void loop_invariant_consume(struct ring **r);
+//@ requires *r |-> ?rp &*& loop_invariant(rp);
+//@ ensures *r |-> rp;
+
+void loop_invariant_produce(struct ring **r);
+//@ requires *r |-> ?rp;
+//@ ensures *r |-> rp &*& loop_invariant(rp);
+
+void loop_iteration_begin(struct ring **r);
+void loop_iteration_end(struct ring **r);
+
+
+#endif//_LOOP_MODEL_H_INCLUDED_

### 2017-06-26 doc/paper/example/model.c

@@ -1,0 +1,25 @@
+#include "vigor.h"
+#include "packet.h"
+#include "ring.h"
+#include "loop-model.h"
+#include "user-params.h"
+
+int main(int argc, char** argv) {
+  struct packet p;
+  struct ring *r = ring_create(CAP);
+  if (!r) return 1;
+  while(LOOP(1))
+  {
+    loop_iteration_begin(&r);
+    if (!ring_full(r))
+      if (receive_packet(&p) && p.port != 9)
+        ring_push_back(r, &p);
+    if (!ring_empty(r) && can_send_packet()) {
+      ring_pop_front(r, &p);
+      VIGOR_CHECK(p.port != 9);
+      send_packet(&p);
+    }
+    loop_iteration_end(&r);
+  }
+  return 0;
+}

### 2017-06-26 doc/paper/example/orig.c

@@ -1,0 +1,23 @@
+#include "vigor.h"
+#include "packet.h"
+
+#define CAP 512
+
+int main() {
+  struct packet packets[CAP] = {};
+  int begin = 0, queue_len = 0;
+
+  while(1) {
+    if (queue_len < CAP) {
+      int end = (begin + queue_len)%CAP;
+      if (receive_packet(&packets[end]) &&
+          packets[end].port != 9)
+        queue_len += 1;
+    }
+    if (0 < queue_len && can_send_packet()) {
+      send_packet(&packets[begin]);
+      begin = (begin + 1)%CAP;
+    }
+  }
+  return 0;
+}

### 2017-06-26 doc/paper/example/ring.c

@@ -1,0 +1,386 @@
+#include <stdlib.h>
+#include "ring.h"
+
+//@ #include "lib/containers/modulo.gh"
+//@ #include "lib/containers/stdex.gh"
+
+struct ring {
+  struct packet *array;
+  int cap;
+  int begin;
+  int len;
+};
+
+
+/*@
+  predicate empty_arrayp(struct packet* arr, int len) =
+    true == ((char *)0 <= (void *)(arr)) &*&
+    true == ((void *)(arr) <= (char *)UINTPTR_MAX) &*&
+    chars((void*)arr, len*sizeof(struct packet), _);
+
+  predicate packetsp(struct packet* arr, int len,
+                     list<packet> pkts,
+                     fixpoint (packet,bool) property) =
+    switch(pkts) {
+      case nil: return len == 0;
+      case cons(h,t):
+        return true == ((void*)arr <= (void*)UINTPTR_MAX) &*&
+               true == ((void*)0 <= (void*)(arr)) &*&
+               packetp(arr, h) &*&
+               true == property(h) &*&
+               packetsp(arr+1, len-1, t, property);
+    };
+
+  predicate ring_fieldsp(struct ring* r; struct packet* arr,
+                         int begin, int len, int cap) =
+    r->begin |-> begin &*&
+    r->cap |-> cap &*&
+    r->len |-> len &*&
+    r->array |-> arr &*&
+    0 <= begin &*& begin < cap &*&
+    0 <= len &*& len <= cap &*&
+    0 < cap &*& cap < CAP_LIMIT &*&
+    malloc_block_chars((void*)arr, cap * sizeof(struct packet)) &*&
+    malloc_block_ring(r);
+
+  predicate ringp(struct ring* r,
+                  fixpoint (packet,bool) property,
+                  list<packet> packets,
+                  int cap) =
+    ring_fieldsp(r, ?arr, ?begin, ?len, cap) &*&
+    len == length(packets) &*&
+    begin + len < cap ?
+      (empty_arrayp(arr, begin) &*&
+       packetsp(arr + begin,
+                len, packets,
+                property) &*&
+       empty_arrayp(arr + begin + len,
+                    cap - len - begin)) :
+      (packetsp(arr, begin + len - cap,
+                drop(cap - begin, packets),
+                property) &*&
+       empty_arrayp(arr +
+                    (begin + len - cap),
[…]

### 2017-06-26 doc/paper/example/user-params.h

@@ -1,0 +1,19 @@
+#ifndef _USER_PARAMS_H_INCLUDED_
+#define _USER_PARAMS_H_INCLUDED_
+
+#include "packet.h"
+#include "ring.h"
+
+#define RING_CAPACITY 512
+
+/*@
+  fixpoint bool user_property(packet p) {
+    switch(p) { case packet(port): return port != 9; }
+  }
+  @*/
+
+static bool packet_constraints(struct packet* p) {
+  return p->port != 9;
+}
+
+#endif//_USER_PARAMS_H_INCLUDED_

### 2017-06-26 example/user-params.h

@@ -1,0 +1,19 @@
+#ifndef _USER_PARAMS_H_INCLUDED_
+#define _USER_PARAMS_H_INCLUDED_
+
+#include "packet.h"
+#include "ring.h"
+
+#define RING_CAPACITY 512
+
+/*@
+  fixpoint bool packet_constraints_fp(packet p) {
+    switch(p) { case packet(port): return port != 9; }
+  }
+  @*/
+
+static bool packet_constraints(struct packet* p) {
+  return p->port != 9;
+}
+
+#endif//_USER_PARAMS_H_INCLUDED_

### 2017-06-26 map-verification-attempts/bench/draw.r

@@ -1,0 +1,12 @@
+library(ggplot2)
+library(scales)
+perf_times <- read.table("export.csv", header=TRUE, sep=",")
+p <- ggplot(perf_times, aes(x=hitrate, y=val, fill=type)) + 
+  geom_bar(position=position_dodge(), stat="identity") + 
+  geom_errorbar(position=position_dodge(), aes(ymin=val-2*sigma, ymax=val+2*sigma)) +
+  facet_grid(load ~ pg,labeller = label_both) + scale_y_continuous(trans=log2_trans()) +
+  xlab("Hit rate") + ylab("execution time") +
+  ggtitle("Hash map implementations performance")
+  
+ggsave("perfplot.png", plot = p, width = 5, height=5)
+quit()

### 2017-06-26 map-verification-attempts/bench/engine.cpp

@@ -1,0 +1,130 @@
+#include <iomanip>
+#include <map>
+#include <unordered_map>
+#include <vector>
+#include <string>
+#include <chrono>
+#include <iostream>
+#include <glib.h>
+#include "../vst/map/map.c"
+
+#include "data.c"
+
+
+using namespace std;
+using namespace std::chrono;
+
+enum dimensions
+    {
+        load5 = 0, load50 = 1, load95 = 2,
+        gp003 = 0, gp1 = 1, gp30 = 2,
+        hitr5 = 0, hitr50 = 1, hitr95 = 2
+    };
+
+void pretty_print(map<string, long> results[][3][3])
+{
+    for (int l = 0; l < 3; ++l)
+    {
+        for (int g = 0; g < 3; ++g)
+            for (int h = 0; h < 3; ++h)
+            {
+                switch (l)
+                {
+                case 0: cout <<"load  5 "; break;
+                case 1: cout <<"load 50 "; break;
+                case 2: cout <<"load 95 "; break;
+                };
+                switch (g)
+                {
+                case 0: cout <<"30put->1get  "; break;
+                case 1: cout <<" 1put->1get  "; break;
+                case 2: cout <<" 1put->30get "; break;
+                };
+                switch (h)
+                {
+                case 0: cout <<"hitrate  5% "; break;
+                case 1: cout <<"hitrate 50% "; break;
+                case 2: cout <<"hitrate 95% "; break;
+                };
+                cout << " | c" <<setw(7) << results[l][g][h]["vst"]
+                     << " | s" <<setw(7) << results[l][g][h]["stl"]
+                     << " | g" <<setw(7) << results[l][g][h]["glib"] <<endl;
+            }
+        cout <<endl;
+    }
+}
+
+void dump(map<string, long> results[][3][3])
+{
+    for (int l = 0; l < 3; ++l)
+        for (int g = 0; g < 3; ++g)
+            for (int h = 0; h < 3; ++h)
+                cout << results[l][g][h]["vst"] <<"\n"
+                     << results[l][g][h]["stl"] <<"\n"
+                     << results[l][g][h]["glib"] <<endl;
+}
+
+int main(int argc, char** argv)
+{
+    const int count = 500;
+
+    int keys[100];
+    int busybits[100] = {};
+    int values[100];
+    unordered_map<int, int> m;
[…]

### 2017-06-26 map-verification-attempts/ecv/map.cpp

@@ -1,0 +1,343 @@
+/*-------------------------------------------------------*/
+/* File:    /home/necto/proj/vnds/ecv/map.cpp            */
+/* Author:  necto                                        */
+/* Created: 12:55:49 on Friday September 18th 2015 UTC   */
+/*-------------------------------------------------------*/
+
+#include "ecv.h"
+
+
+#define CAPACITY 1000000
+
+/*@ logic integer loop_l(integer k) = (k%CAPACITY + CAPACITY)%CAPACITY;
+ */
+
+/*@ lemma loop_bijection: \forall integer k;
+  @   0 <= k < CAPACITY ==> loop_l(k) == k;
+*/
+/*@ lemma loop_injection: \forall integer k;
+  @   0 <= k ==> loop_l(k + CAPACITY) == loop_l(k);
+*/
+
+/*@ predicate full(int *busybits) = 
+  @   \forall integer x; 0 <= x < CAPACITY ==> busybits[x] != 0;
+*/
+
+/*@ lemma non_full_exists: \forall int* busybits;
+  @ !full(busybits) <==> (\exists integer x; 0 <= x < CAPACITY &&
+  @                                          busybits[x] == 0);
+*/
+
+/*@ logic integer size_rec{L}(int* busybits, integer capa) = 
+  @   (capa == 0) ? 0 :
+  @   (busybits[capa-1] != 0) ? 1 + size_rec(busybits, capa - 1) :
+  @    size_rec(busybits, capa - 1);
+*/
+
+/*@ lemma size_rec_on_zero: \forall int* busybits, integer capa;
+  @   0 <= capa && busybits[capa] == 0 ==> 
+  @     size_rec(busybits, capa + 1) == size_rec(busybits, capa);
+  @
+  @ lemma size_rec_on_one: \forall int* busybits, integer capa;
+  @   0 <= capa && busybits[capa] != 0 ==>
+  @     size_rec(busybits, capa + 1) == 1 + size_rec(busybits, capa);
+*/
+
+/*@ lemma size_rec_bounds: \forall int* busybits, integer capa;
+  @   0 <= capa ==> 0 <= size_rec(busybits, capa) <= capa;
+*/
+
+/*@ logic boolean full_rec{L}(int *busybits, integer capa) = 
+  @  (capa == 0) ? \true :
+  @  (busybits[capa-1] != 0) ? full_rec(busybits, capa - 1) : \false;
+  @
+  @ lemma full_rec_full: \forall int* busybits;
[…]

### 2017-06-26 map-verification-attempts/ecv/map.h

@@ -1,0 +1,40 @@
+/*-------------------------------------------------------*/
+/* File:    /home/necto/proj/vnds/ecv/map.h              */
+/* Author:  necto                                        */
+/* Created: 09:35:21 on Tuesday September 22nd 2015 UTC  */
+/*-------------------------------------------------------*/
+
+#include "ecv.h"
+
+
+#define CAPACITY 100
+
+int loop (int k)
+  writes()
+  post(0 <= result; result < CAPACITY)
+  post((0 <= k && k < CAPACITY) => result == k)
+  post(0 <= k && k < maxof(int) => result == k%CAPACITY);
+  
+assume(forall int k :- 0 <= k && k < (maxof(int) - CAPACITY) => loop(k + CAPACITY) == loop(k));
+assert(forall k in 0..(CAPACITY-1) :- loop(k) == k);
+assume(forall k in 0..(CAPACITY-1) :- (exists int x :- loop(x) == k));
+
+/*  
+int find_empty (int busybits[CAPACITY], int start)
+  writes()
+  pre(0 <= start; start < CAPACITY)
+  pre(busybits.lim >= CAPACITY)
+  post(exists bb in busybits.all :- bb == 0 => 0 <= result && result < CAPACITY && busybits[result] == 0)
+  post(forall bb in busybits.all :- bb != 0 => result == -1);
+*/
+  
+int find_key (int busybits[CAPACITY], int keys[CAPACITY], int start, int key)
+  writes()
+  pre(busybits.lim >= CAPACITY)
+  pre(keys.lim >= CAPACITY)
+  pre(0 <= start; start < CAPACITY)
+  post((exists integer x :- 0 <= x && x < CAPACITY && busybits[x] != 0 && keys[x] == key) =>
+       0 <= result && result < CAPACITY && busybits[result] != 0 && keys[result] == key)
+  post((forall integer x :- 0 <= x && x < CAPACITY => busybits[x] == 0 || keys[x] != key) =>
+       result == -1);
+	   

### 2017-06-26 map-verification-attempts/ecv/map_internal.cpp

@@ -1,0 +1,11 @@
+/*-------------------------------------------------------*/
+/* File:    /home/necto/proj/vnds/ecv/map_internal.cpp   */
+/* Author:  necto                                        */
+/* Created: 09:35:50 on Tuesday September 22nd 2015 UTC  */
+/*-------------------------------------------------------*/
+
+#include "ecv.h"
+
+/* Add your code here... */
+
+/* End */

### 2017-06-26 map-verification-attempts/frama-c/map.c

@@ -1,0 +1,305 @@
+
+#define CAPACITY 1000000
+
+/*@ logic integer loop_l(integer k) = (k%CAPACITY + CAPACITY)%CAPACITY;
+ */
+
+/*@ lemma loop_bijection: \forall integer k;
+  @   0 <= k < CAPACITY ==> loop_l(k) == k;
+*/
+/*@ lemma loop_injection: \forall integer k;
+  @   0 <= k ==> loop_l(k + CAPACITY) == loop_l(k);
+*/
+
+/*@ predicate full(int *busybits) = 
+  @   \forall integer x; 0 <= x < CAPACITY ==> busybits[x] != 0;
+*/
+
+/*@ lemma non_full_exists: \forall int* busybits;
+  @ !full(busybits) <==> (\exists integer x; 0 <= x < CAPACITY &&
+  @                                          busybits[x] == 0);
+*/
+
+/*@ logic integer size_rec{L}(int* busybits, integer capa) = 
+  @   (capa == 0) ? 0 :
+  @   (busybits[capa-1] != 0) ? 1 + size_rec(busybits, capa - 1) :
+  @    size_rec(busybits, capa - 1);
+*/
+
+/*@ lemma size_rec_on_zero: \forall int* busybits, integer capa;
+  @   0 <= capa && busybits[capa] == 0 ==> 
+  @     size_rec(busybits, capa + 1) == size_rec(busybits, capa);
+  @
+  @ lemma size_rec_on_one: \forall int* busybits, integer capa;
+  @   0 <= capa && busybits[capa] != 0 ==>
+  @     size_rec(busybits, capa + 1) == 1 + size_rec(busybits, capa);
+*/
+
+/*@ lemma size_rec_bounds: \forall int* busybits, integer capa;
+  @   0 <= capa ==> 0 <= size_rec(busybits, capa) <= capa;
+*/
+
+/*@ logic boolean full_rec{L}(int *busybits, integer capa) = 
+  @  (capa == 0) ? \true :
+  @  (busybits[capa-1] != 0) ? full_rec(busybits, capa - 1) : \false;
+  @
+  @ lemma full_rec_full: \forall int* busybits;
+  @   full_rec(busybits, CAPACITY) <==> full(busybits);
+  @
+  @ lemma full_rec_size_rec: \forall int* busybits, integer capa;
[…]

### 2017-06-26 map-verification-attempts/frama-c/verify_map.sh

@@ -1,0 +1,1 @@
+frama-c -wp -wp-rte map.c -wp-prover why3:alt-ergo,why3:z3,coq -wp-script map.c.wp.script -then -report

### 2017-06-26 map-verification-attempts/verifast/map.c

@@ -1,0 +1,1970 @@
+//@ #include <list.gh>
+//@ #include <listex.gh>
+//@ #include "nat_seq.gh"
+//@ #include "stdex.gh"
+//@ #include "nth_prop.gh"
+
+#define CAPACITY 1000000
+
+/*@ 
+lemma void div_mod(int g, int k, int l)
+ requires g == (k % l) &*& l > 0;
+ ensures (-l <= g) &*& (g < l);
+{
+    div_rem(k, l);
+}
+@*/
+
+/*@
+lemma void div_mod_gt_0(int mod, int div, int whole)
+  requires mod == (div % whole) &*& whole > 0 &*& div >= 0;
+  ensures (0 <= mod) &*& (mod < whole);
+{
+    div_rem(div, whole);
+}
+@*/
+
+/*@
+  fixpoint int loop_l(int k) { return ((k%CAPACITY + CAPACITY)%CAPACITY); }
+  @*/
+
+int loop (int k)
+//@ requires true;
+//@ ensures 0 <= result &*& result < CAPACITY &*& result==loop_l(k);
+{
+    int g = k%CAPACITY;
+    //@ div_mod(g, k, CAPACITY);
+    int res = (g + CAPACITY)%CAPACITY;
+    //@ div_mod_gt_0(res, g + CAPACITY, CAPACITY);
+    return res;
+}
+
+/*@
+  lemma void loop_lims(int k)
+  requires true;
+  ensures 0 <= loop_l(k) &*& loop_l(k) < CAPACITY;
+  {
+    div_rem(k, CAPACITY);
+    assert(-CAPACITY <= k%CAPACITY);
+    assert(0 <= k%CAPACITY + CAPACITY);
+    div_rem((k + CAPACITY), CAPACITY);
+    assert(CAPACITY > 0);
+    div_rem(k%CAPACITY + CAPACITY, CAPACITY);
+    assert(0 <= ((k%CAPACITY + CAPACITY)%CAPACITY));
+  }
+
+  lemma void loop_bijection(int k)
+  requires 0 <= k &*& k < CAPACITY;
+  ensures loop_l(k) == k;
+  {
+    div_rem(k, CAPACITY);
+    assert(k%CAPACITY == k);
+    div_rem((k + CAPACITY), CAPACITY);
+    assert(k == ((k % CAPACITY + CAPACITY) % CAPACITY));
+  }
+  
+  lemma void loop_injection(int k)
[…]

### 2017-06-26 map-verification-attempts/verifast/memcpy.c

@@ -1,0 +1,154 @@
+#include <string.h>
+#include <stdlib.h>
+
+//@ #include <listex.gh>
+
+int foo()
+//@ requires exists<bool>(?lala) &*& (lala? exists<int>(?xa) : exists<real>(?xi)) &*& exists<fixpoint (int, int)>(?f);
+//@ ensures result == 0;
+{
+  return 0;
+}
+
+struct A {
+  int x;
+  int y;
+};
+
+/*@
+  predicate real_A(pair<int, int> v;) = INT_MIN <= fst(v) &*& fst(v) <= INT_MAX &*& INT_MIN <= snd(v) &*& snd(v) <= INT_MAX;
+  predicate A(struct A* ptr; pair<int, int> v) = ptr->x |-> ?x &*& ptr->y |-> ?y &*& real_A(pair(x,y)) &*& v == pair(x,y);
+  @*/
+
+/*@
+fixpoint list<char> chars_of_A(pair<int, int> a) {
+  switch(a){case pair<int, int>(x, y): return append(chars_of_int(x), chars_of_int(y)); }
+}
+
+fixpoint pair<int, int> A_of_chars(list<char> chs) {
+  return pair<int, int>(int_of_chars(take(4, chs)), int_of_chars(drop(4, chs)));
+}
+
+
+lemma void chars_within_limits_distr(int n, list<char> chs);
+requires true == chars_within_limits(chs);
+ensures true == chars_within_limits(chs) &*& true == chars_within_limits(take(n, chs)) &*& true == chars_within_limits(drop(n, chs));
+
+lemma void chars_within_limits_append(list<char> chs1, list<char> chs2);
+requires true == chars_within_limits(chs1) &*& true == chars_within_limits(chs2);
+ensures true == chars_within_limits(chs1) &*& true == chars_within_limits(chs2) &*& true == chars_within_limits(append(chs1, chs2));
+
+lemma void chars_of_int_within_limits(int a);
+requires INT_MIN <= a &*& a <= INT_MAX;
+ensures true == chars_within_limits(chars_of_int(a));
+
+lemma void A_to_bytes(struct A* a)
+requires A(a, ?val);
+ensures chars((void*)a, sizeof(struct A), chars_of_A(val)) &*& chars_within_limits(chars_of_A(val)) == true;
+{
+  open A_x(a,?x);
+  open A_y(a,?y);
+  
[…]

### 2017-06-26 map-verification-attempts/vst/map/map.c

@@ -1,0 +1,103 @@
+
+int loop (int k)
+{   //single "%" result ranges from -99 to 99.
+    return (k%100 + 100)%100;//ranges from 0 to 99.
+}
+
+int find_empty (int* busybits, int start, int i)
+{
+    int index = loop(1 + start + i);
+    int bb = busybits[index];
+
+    if (0 == bb)
+    {
+        return index;
+    }
+    if (0 == i)
+    {
+        return -1;
+    }
+    return find_empty(busybits, start, i - 1);
+}
+
+int find_key (int* busybits, int* keys, int start, int key, int i)
+{
+    int index = loop(1 + start + i);
+    int bb = busybits[index];
+    int k = keys[index];
+
+    if (1 == bb)
+    {
+        if (k == key)
+        {
+            return index;
+        }
+    }
+    if (i == 0)
+    {
+        return -1;
+    }
+    return find_key(busybits, keys, start, key, i - 1);
+}
+
+int get(int* busybits, int* keys, int* values, int key)
+{
+    int start = loop(key);
+    int index = find_key(busybits, keys, start, key, 99);
+
+    if (-1 == index)
+    {
+        return -1;
+    }
+    int val = values[index];
+    return val;
+}
+
+int put(int* busybits, int* keys, int* values, int key, int value)
+{
+    int start = loop(key);
+    int index = find_empty(busybits, start, 99);
+
+    if (-1 == index)
+    {
+        return -1;
+    }
+    busybits[index] = 1;
+    keys    [index] = key;
+    values  [index] = value;
+    return 0;
+}
+
+int erase(int* busybits, int* keys, int key)
+{
+    int start = loop(key);
+    int index = find_key(busybits, keys, start, key, 99);
+
+    if (-1 == index)
+    {
+        return -1;
+    }
+    busybits[index] = 0;
+    return 0;
+}
+
+int size_rec(int* busybits, int i)
+{
+    if (0 == i)
+    {
+        return 0;
+    }
+    int index = i - 1;
+    int bb = busybits[index];
+
+    if (1 == bb)
+    {
+        return 1 + size_rec(busybits, i - 1);
+    }
+    return size_rec(busybits, i - 1);
+}
+
+int size(int* busybits)
+{
+    return size_rec(busybits, 100);
+}

### 2017-06-26 map-verification-attempts/vst/samples/assoc.c

@@ -1,0 +1,30 @@
+
+int get(int* arr, int key)
+{
+    int rez = arr[key]; //using memory access (arr[key]) in an expression
+    return rez; // (return ...) is forbidden by VST.
+}
+
+void set(int* arr, int key, int val)
+{
+    arr[key] = val;
+}
+
+int hash(int key)
+{
+    if (key < 0)
+        return key%100 + 99;
+    return key%100;
+}
+
+int cGet(int* arr, int key)
+{
+    int h = hash(key);
+    return get(arr, h);
+}
+
+void cSet(int* arr, int key, int val)
+{
+    int h = hash(key);
+    set(arr, h, val);
+}

### 2017-06-26 map-verification-attempts/vst/samples/sqrt.c

@@ -1,0 +1,14 @@
+
+unsigned int guess_sqrt(unsigned int n, unsigned int guess)
+{
+    if (n == 0) return 0;
+    if (guess == 0) return 0;
+    
+    if (guess*guess <= n) return guess;
+    /*else*/ return guess_sqrt(n, guess - 1);
+}
+
+unsigned int sqrt(unsigned int n)
+{
+    return guess_sqrt(n, n);
+}

### 2017-06-26 nf/bridge/bridge_config.c

@@ -1,0 +1,101 @@
+#include <getopt.h>
+#include <inttypes.h>
+#include <string.h>
+
+#if KLEE_VERIFICATION
+#  include "lib/stubs/rte_stubs.h"
+#else
+  // DPDK needs these but doesn't include them. :|
+#  include <linux/limits.h>
+#  include <sys/types.h>
+
+#  include <rte_common.h>
+#  include <rte_ethdev.h>
+#endif
+
+#include <cmdline_parse_etheraddr.h>
+#include <cmdline_parse_ipaddr.h>
+
+#include "lib/nf_util.h"
+#include "lib/nf_log.h"
+#include "bridge_config.h"
+
+const uint32_t DEFAULT_EXP_TIME = 300;//seconds
+const uint32_t DEFAULT_CAPACITY = 128;//MAC addresses
+
+#define PARSE_ERROR(format, ...) \
+    bridge_config_cmdline_print_usage(); \
+    rte_exit(EXIT_FAILURE, format, ##__VA_ARGS__);
+
+void bridge_config_init(struct bridge_config* config,
+                        int argc, char** argv)
+{
+  // Set the default values
+  config->expiration_time = DEFAULT_EXP_TIME; //seconds
+  config->dyn_capacity = DEFAULT_CAPACITY; //MAC addresses
+  config->static_config_fname[0] = '\0';//no static filtering configuration
+
+  unsigned nb_devices = rte_eth_dev_count();
+
+  struct option long_options[] = {
+    {"expire",    required_argument,  NULL, 't'},
+    {"capacity",    required_argument,  NULL, 'c'},
+    {"config",    required_argument,  NULL, 'f'},
+    {NULL,       0,      NULL, 0  }
+  };
+
+  int opt;
+  while ((opt = getopt_long(argc, argv, "t:c:f:", long_options, NULL)) != EOF) {
+    unsigned device;
+    switch (opt) {
+    case 't':
+      config->expiration_time = nf_util_parse_int(optarg, "exp-time", 10, '\0');
+      if (config->expiration_time == 0) {
+        PARSE_ERROR("Expiration time must be strictly positive.\n");
+      }
+      break;
+
+    case 'c':
+      config->dyn_capacity = nf_util_parse_int(optarg, "capacity", 10, '\0');
+      if (config->dyn_capacity <= 0) {
+        PARSE_ERROR("Flow table size must be strictly positive.\n");
+      }
+      break;
+
[…]

### 2017-06-26 nf/bridge/bridge_config.h

@@ -1,0 +1,26 @@
+#pragma once
+
+#include <inttypes.h>
+
+// TODO can't really include rte_max_ethports here but need it :/
+
+#define CONFIG_FNAME_LEN 512
+
+struct bridge_config {
+  // Expiration time of flows in seconds
+  uint32_t expiration_time;
+
+  // Size of the dynamic filtering table
+  uint32_t dyn_capacity;
+
+  // The static configuration file name
+  char static_config_fname[CONFIG_FNAME_LEN];
+};
+
+
+void bridge_config_init(struct bridge_config* config,
+                        int argc, char** argv);
+
+void bridge_config_cmdline_print_usage(void);
+
+void bridge_print_config(struct bridge_config* config);

### 2017-06-26 nf/bridge/bridge_data.h

@@ -1,0 +1,121 @@
+#ifndef _BRIDGE_DATA_H_INCLUDED_
+#define _BRIDGE_DATA_H_INCLUDED_
+
+#include "lib/custom_ether_addr_def.h"
+
+struct Map;
+struct Vector;
+struct DoubleChain;
+
+struct StaticKey {
+  struct ether_addr addr;
+  uint8_t device;
+};
+
+struct DynamicEntry {
+  struct ether_addr addr;
+  uint8_t device;
+};
+
+struct DynamicFilterTable {
+  struct Map* map;
+  struct DoubleChain* heap;
+  struct Vector* entries;
+};
+
+struct StaticFilterTable {
+  struct Map* map;
+  struct Vector* keys;
+};
+
+/*@
+  inductive ether_addri = eaddrc(int, int, int, int, int, int);
+  inductive stat_keyi = stkc(int, ether_addri);
+  inductive dynenti = dynentc(int, ether_addri);
+
+  predicate ether_addrp(struct ether_addr* ptr; ether_addri addr) =
+    struct_ether_addr_padding(ptr) &*&
+    ptr->a |-> ?a &*&
+    ptr->b |-> ?b &*&
+    ptr->c |-> ?c &*&
+    ptr->d |-> ?d &*&
+    ptr->e |-> ?e &*&
+    ptr->f |-> ?f &*&
+    addr == eaddrc(a, b, c, d, e, f);
+
+  predicate static_keyp(struct StaticKey* ptr; stat_keyi k) =
+    struct_StaticKey_padding(ptr) &*&
+    ptr->device |-> ?device &*&
+    ether_addrp(&ptr->addr, ?addr) &*&
+    k == stkc(device, addr);
+
+  predicate dynamic_entryp(struct DynamicEntry* ptr; dynenti de) =
+    struct_DynamicEntry_padding(ptr) &*&
+    ptr->device |-> ?device &*&
+    ether_addrp(&ptr->addr, ?addr) &*&
+    de == dynentc(device, addr);
+
+  predicate dynamic_entry_barep(struct DynamicEntry* ptr, dynenti de) =
+    struct_DynamicEntry_padding(ptr) &*&
+    switch(de) { case dynentc(device,addr):
+      return ptr->device |-> device;
+    };
+
+  fixpoint bool dynentry_right_offsets(void* ptr, void* eaddr);
[…]

### 2017-06-26 nf/bridge/bridge_loop.c

@@ -1,0 +1,99 @@
+#include <klee/klee.h>
+#include "bridge_loop.h"
+#include "lib/stubs/rte-stubs-control.h"
+#include "lib/stubs/containers/double-chain-stub-control.h"
+#include "lib/stubs/containers/map-stub-control.h"
+#include "lib/stubs/containers/vector-stub-control.h"
+#include "lib/stubs/my-time-stub-control.h"
+
+
+
+
+void bridge_loop_iteration_assumptions(struct DoubleChain** dyn_heap,
+                                       struct Map** dyn_map,
+                                       struct Vector** dyn_vec,
+                                       struct Map** st_map,
+                                       struct Vector** st_vec,
+                                       uint32_t capacity,
+                                       uint32_t time) {
+  rte_reset();
+  dchain_reset(*dyn_heap, capacity);
+  map_reset(*dyn_map);
+  vector_reset(*dyn_vec);
+  map_reset(*st_map);
+}
+
+void bridge_loop_invariant_consume(struct DoubleChain** dyn_heap,
+                                   struct Map** dyn_map,
+                                   struct Vector** dyn_vec,
+                                   struct Map** st_map,
+                                   struct Vector** st_vec,
+                                   uint32_t capacity,
+                                   uint32_t time,
+                                   uint32_t dev_count) {
+  klee_trace_ret();
+  klee_trace_param_ptr(dyn_heap, sizeof(struct DoubleChain*), "dyn_heap");
+  klee_trace_param_ptr(dyn_map, sizeof(struct Map*), "dyn_map");
+  klee_trace_param_ptr(dyn_vec, sizeof(struct Vector*), "dyn_vec");
+  klee_trace_param_ptr(st_map, sizeof(struct Map*), "st_map");
+  klee_trace_param_ptr(st_vec, sizeof(struct Vector*), "st_vec");
+  klee_trace_param_i32(capacity, "capacity");
+  klee_trace_param_i32(time, "time");
+  klee_trace_param_i32(dev_count, "dev_count");
+}
+
+
+void bridge_loop_invariant_produce(struct DoubleChain** dyn_heap,
+                                   struct Map** dyn_map,
+                                   struct Vector** dyn_vec,
+                                   struct Map** st_map,
+                                   struct Vector** st_vec,
+                                   uint32_t capacity,
+                                   uint32_t* time,
+                                   uint32_t dev_count) {
+  klee_trace_ret();
[…]

### 2017-06-26 nf/bridge/bridge_loop.h

@@ -1,0 +1,120 @@
+#ifndef _BRIDGE_LOOP_H_INCLUDED_
+#define _BRIDGE_LOOP_H_INCLUDED_
+#include "lib/containers/double-chain.h"
+#include "lib/containers/map.h"
+#include "lib/containers/vector.h"
+#include "bridge_data.h"
+#include "lib/coherence.h"
+//@ #include "lib/predicates.gh"
+
+/*@
+  fixpoint bool st_entry_bound<t>(int bound, pair<t,int> p) {
+    return 0 <= snd(p) && snd(p) < bound;
+  }
+
+  predicate bridge_loop_invariant(struct DoubleChain* dyn_heap,
+                                  struct Map* dyn_map,
+                                  struct Vector* dyn_vec,
+                                  struct Map* st_map,
+                                  struct Vector* st_vec,
+                                  uint32_t capacity,
+                                  uint32_t time,
+                                  uint32_t dev_count) =
+    double_chainp(?dh, dyn_heap) &*&
+    mapp<ether_addri>(dyn_map, ether_addrp, eth_addr_hash,
+                      ?earecp,
+                      mapc(capacity, ?dm, ?daddrs)) &*&
+    vectorp<dynenti>(dyn_vec, dynamic_entryp, ?dv) &*&
+    mapp<stat_keyi>(st_map, static_keyp,
+                    st_key_hash,
+                    ?strecp,
+                    mapc(?stcap, ?sm, ?saddrs)) &*&
+    vectorp<stat_keyi>(st_vec, static_keyp, ?sv) &*&
+    0 < capacity &*&
+    length(dv) == capacity &*&
+    map_vec_chain_coherent<ether_addri, dynenti>(dm, dv, dh) &*&
+    dchain_high_fp(dh) <= time &*&
+    last_time(time) &*&
+    true == forall(sm, (st_entry_bound)(dev_count));
+  @*/
+
+void bridge_loop_invariant_consume(struct DoubleChain** dyn_heap,
+                                   struct Map** dyn_map,
+                                   struct Vector** dyn_vec,
+                                   struct Map** st_map,
+                                   struct Vector** st_vec,
+                                   uint32_t capacity,
+                                   uint32_t time,
+                                   uint32_t dev_count);
+/*@ requires *dyn_heap |-> ?dh &*&
+             *dyn_map |-> ?dm &*&
+             *dyn_vec |-> ?dv &*&
+             *st_map |-> ?sm &*&
[…]

### 2017-06-26 nf/lib/coherence.c

@@ -1,0 +1,337 @@
+#include "coherence.h"
+
+/*@
+  predicate dmap_dchain_coherent<t1,t2,vt>(dmap<t1,t2,vt> m, dchain ch) =
+    dchain_index_range_fp(ch) == dmap_cap_fp(m) &*&
+    true == subset(dchain_indexes_fp(ch), dmap_indexes_used_fp(m)) &*&
+    true == subset(dmap_indexes_used_fp(m), dchain_indexes_fp(ch));
+  @*/
+
+/*@
+
+lemma void coherent_same_indexes<t1,t2,vt>
+             (dmap<t1,t2,vt> m, dchain ch)
+requires dmap_dchain_coherent(m, ch);
+ensures dmap_dchain_coherent(m, ch) &*&
+        true == subset(dchain_indexes_fp(ch), dmap_indexes_used_fp(m)) &*&
+        true == subset(dmap_indexes_used_fp(m), dchain_indexes_fp(ch));
+{
+  open dmap_dchain_coherent(m, ch);
+  close dmap_dchain_coherent(m, ch);
+}
+@*/
+
+/*@
+
+lemma void empty_dmap_dchain_coherent<t1,t2,vt>(int len)
+requires 0 <= len;
+ensures dmap_dchain_coherent<t1,t2,vt>
+         (empty_dmap_fp<t1,t2,vt>(len), empty_dchain_fp(len, 0));
+{
+  empty_dmap_cap<t1,t2,vt>(len);
+  dmap_empty_no_indexes_used<t1,t2,vt>(len);
+  close dmap_dchain_coherent(empty_dmap_fp<t1,t2,vt>(len),
+                             empty_dchain_fp(len, 0));
+}
+
+lemma void coherent_dmap_used_dchain_allocated<t1,t2,vt>
+             (dmap<t1,t2,vt> m, dchain ch, int idx)
+requires dmap_dchain_coherent(m, ch) &*& dmap_index_used_fp(m, idx) == true;
+ensures dmap_dchain_coherent(m, ch) &*&
+        dchain_allocated_fp(ch, idx) == true;
+{
+  open dmap_dchain_coherent(m, ch);
+  dmap_index_used_inbounds(m, idx);
+  dmap_indexes_contain_index_used(m, idx);
+  mem_subset(idx, dmap_indexes_used_fp(m), dchain_indexes_fp(ch));
+  dchain_indexes_contain_index(ch, idx);
[…]

### 2017-06-26 nf/lib/coherence.h

@@ -1,0 +1,135 @@
+#ifndef _COHERENCE_H_INCLUDED_
+#define _COHERENCE_H_INCLUDED_
+#include "containers/vector.h"
+#include "containers/map.h"
+#include "containers/double-map.h"
+#include "containers/double-chain.h"
+
+
+/*@
+  predicate dmap_dchain_coherent<t1,t2,vt>(dmap<t1,t2,vt> m, dchain ch);
+  @*/
+
+/*@
+  predicate map_vec_chain_coherent<kt,vt>(list<pair<kt, int> > m,
+                                          list<vt> v, dchain ch);
+  @*/
+
+/*@
+  lemma void mvc_coherent_map_get_bounded<kt,vt>(list<pair<kt, int> > m,
+                                                 list<vt> v, dchain ch,
+                                                 kt k);
+  requires map_vec_chain_coherent<kt,vt>(m, v, ch) &*&
+           true == map_has_fp(m, k);
+  ensures 0 <= map_get_fp(m, k) &*& map_get_fp(m, k) < length(v) &*&
+          dchain_index_range_fp(ch) == length(v) &*&
+          map_vec_chain_coherent<kt,vt>(m, v, ch) &*&
+          true == dchain_allocated_fp(ch, map_get_fp(m, k));
+  @*/
+
+/*@
+  lemma void mvc_rejuvenate_preserves_coherent<kt,vt>(list<pair<kt, int> > m,
+                                                      list<vt> v, dchain ch,
+                                                      int index, uint32_t time);
+  requires map_vec_chain_coherent<kt,vt>(m, v, ch) &*&
+           true == dchain_allocated_fp(ch, index);
+  ensures map_vec_chain_coherent<kt, vt>(m, v, dchain_rejuvenate_fp(ch,
+                                                                    index,
+                                                                    time));
+  @*/
+
+/*@
+  lemma void mvc_coherent_same_len<kt,vt>(list<pair<kt, int> > m,
+                                          list<vt> v, dchain ch);
+  requires map_vec_chain_coherent<kt,vt>(m, v, ch);
+  ensures map_vec_chain_coherent<kt,vt>(m, v, ch) &*&
[…]

### 2017-06-26 nf/lib/containers/array-bat-impl.h

@@ -1,0 +1,273 @@
+#ifndef _ARRAY_BAT_IMPL_H_INCLUDED_
+#define _ARRAY_BAT_IMPL_H_INCLUDED_
+#include "array-bat.h"
+#include "../ignore.h"
+
+//@ #include "arith.gh"
+//@ #include "stdex.gh"
+
+/*@
+  predicate batcherpp(list<BATCHER_EL_TYPE> batch, struct Batcher* bat) =
+    batcherp(batch, bat) &*& length(batch) < BATCHER_CAPACITY;
+  predicate bat_arrp(struct Batcher *bp, int len,
+                     list<list<BATCHER_EL_TYPE> > bats) =
+    switch(bats) {
+      case nil: return len == 0;
+      case cons(h,t):
+        return batcherpp(h, bp) &*&
+               bat_arrp(bp+1, len-1, t);
+    };
+
+
+  predicate arrp_bat(list<list<BATCHER_EL_TYPE> > data, struct ArrayBat *arr) =
+    true == ((char *)0 <=
+             (void*)((ARRAY_BAT_EL_TYPE*)(arr->data))) &*&
+    true == ((void*)((ARRAY_BAT_EL_TYPE*)(arr->data) +
+                     ARRAY_BAT_CAPACITY) <=
+             (char *)UINTPTR_MAX) &*&
+    length(data) == ARRAY_BAT_CAPACITY &*&
+    struct_ArrayBat_padding(arr) &*&
+    bat_arrp(arr->data, ARRAY_BAT_CAPACITY, data);
+
+  predicate arrp_bat_acc(list<list<BATCHER_EL_TYPE> > data, struct ArrayBat *arr,
+                         int idx) =
+    0 <= idx &*& idx < ARRAY_BAT_CAPACITY &*&
+    true == ((char *)0 <=
+             (void*)((ARRAY_BAT_EL_TYPE*)(arr->data))) &*&
+    true == ((void*)((ARRAY_BAT_EL_TYPE*)(arr->data) +
+                     ARRAY_BAT_CAPACITY) <=
+             (char *)UINTPTR_MAX) &*&
+    length(data) == ARRAY_BAT_CAPACITY &*&
+    struct_ArrayBat_padding(arr) &*&
+    bat_arrp(arr->data, idx, take(idx, data)) &*&
+    bat_arrp((ARRAY_BAT_EL_TYPE*)(arr->data) + idx + 1,
[…]

### 2017-06-26 nf/lib/containers/array-bat-verifast.c

@@ -1,0 +1,1 @@
+array-bat-impl.h

### 2017-06-26 nf/lib/containers/array-bat.h

@@ -1,0 +1,65 @@
+#ifndef _ARRAY_BAT_H_INCLUDED_
+#define _ARRAY_BAT_H_INCLUDED_
+
+#include "../static-component-params.h"
+#include "../containers/batcher.h"
+
+#define ARRAY_BAT_EL_TYPE struct Batcher
+#define ARRAY_BAT_CAPACITY RTE_MAX_ETHPORTS
+#define ARRAY_BAT_EL_INIT batcher_init
+
+
+#ifdef KLEE_VERIFICATION
+struct ArrayBat
+{
+  char dummy;
+};
+#else//KLEE_VERIFICATION
+struct ArrayBat
+{
+  struct Batcher data[ARRAY_BAT_CAPACITY];
+};
+#endif//KLEE_VERIFICATION
+
+/*@ predicate some_batcherp(struct Batcher* p) = batcherp(_, p);
+  @*/
+
+/*@
+  //params: list<struct rte_mbuf*> , batcherp
+  predicate arrp_bat(list<list<struct rte_mbuf*> > data, struct ArrayBat *arr);
+  predicate arrp_bat_acc(list<list<struct rte_mbuf*> > data, struct ArrayBat *arr,
+                         int idx);
+
+  fixpoint ARRAY_BAT_EL_TYPE *arrp_the_missing_cell_bat(struct ArrayBat *arr,
+                                                        int idx) {
+    return (ARRAY_BAT_EL_TYPE*)(arr->data)+idx;
+  }
+
+  lemma void construct_bat_element(ARRAY_BAT_EL_TYPE *p);
+  requires p->len |-> ?l &*&
+           pointers((void*)&p->batch, ARRAY_BAT_CAPACITY, _) &*&
+           struct_Batcher_padding(p);
+  ensures batcherp(_, p);
+
+  @*/
+
+void array_bat_init(struct ArrayBat *arr_out);
+/*@ requires chars((void*)(arr_out),
+                   ARRAY_BAT_CAPACITY*sizeof(ARRAY_BAT_EL_TYPE), _) &*&
+             struct_ArrayBat_padding(arr_out); @*/
+//@ ensures arrp_bat(_, arr_out);
+
+ARRAY_BAT_EL_TYPE *array_bat_begin_access(struct ArrayBat *arr, int index);
+//@ requires arrp_bat(?lst, arr) &*& 0 <= index &*& index < ARRAY_BAT_CAPACITY;
+/*@ ensures arrp_bat_acc(lst, arr, index) &*&
[…]

### 2017-06-26 nf/lib/containers/array-rq-impl.h

@@ -1,0 +1,295 @@
+#ifndef _ARRAY_RQ_IMPL_H_INCLUDED_
+#define _ARRAY_RQ_IMPL_H_INCLUDED_
+
+#include "array-rq.h"
+
+//@ #include "stdex.gh"
+//@ #include "arith.gh"
+//@ #include "nat.gh"
+
+/*@
+  predicate rx_queue_arrp(struct lcore_rx_queue *lrq,
+                          int len, list<rx_queuei> data) =
+     switch(data) {
+       case nil: return len == 0;
+       case cons(h,t):
+         return rx_queuep(h, lrq) &*&
+                rx_queue_arrp(lrq+1,len-1,t);
+     };
+
+  predicate arrp_rq(list<rx_queuei> data, struct ArrayRq *arr) =
+    true == ((char *)0 <=
+            (void*)((ARRAY_RQ_EL_TYPE*)(arr->data))) &*&
+    true == ((void*)((ARRAY_RQ_EL_TYPE*)(arr->data) +
+                    ARRAY_RQ_CAPACITY) <=
+            (char *)UINTPTR_MAX) &*&
+    length(data) == ARRAY_RQ_CAPACITY &*&
+    struct_ArrayRq_padding(arr) &*&
+    rx_queue_arrp(arr->data, ARRAY_RQ_CAPACITY, data);
+
+  predicate arrp_rq_acc(list<rx_queuei> data, struct ArrayRq *arr, int idx) =
+    0 <= idx &*& idx < ARRAY_RQ_CAPACITY &*&
+    true == ((char *)0 <=
+             (void*)((ARRAY_RQ_EL_TYPE*)(arr->data))) &*&
+    true == ((void*)((ARRAY_RQ_EL_TYPE*)(arr->data) +
+                     ARRAY_RQ_CAPACITY) <=
+             (char *)UINTPTR_MAX) &*&
+    length(data) == ARRAY_RQ_CAPACITY &*&
+    struct_ArrayRq_padding(arr) &*&
+    rx_queue_arrp(arr->data, idx, take(idx, data)) &*&
+    rx_queue_arrp((ARRAY_RQ_EL_TYPE*)(arr->data)+idx+1,
+                  ARRAY_RQ_CAPACITY - idx - 1, drop(idx+1,data));
+  @*/
+
+static void lcore_rx_queue_init(struct lcore_rx_queue *cell)
+//@ requires chars((void*)cell, sizeof(struct lcore_rx_queue), _);
[…]

### 2017-06-26 nf/lib/containers/array-rq.h

@@ -1,0 +1,90 @@
+#ifndef _ARRAY_RQ_H_INCLUDED_
+#define _ARRAY_RQ_H_INCLUDED_
+
+#include <stdint.h>
+#include "../ignore.h"
+#include "../static-component-params.h"
+
+#define MAX_RX_QUEUE_PER_LCORE 16
+
+struct lcore_rx_queue {
+  uint8_t port_id;
+  uint8_t queue_id;
+}
+#ifdef _NO_VERIFAST_
+  __rte_cache_aligned;
+#else//_NO_VERIFAST_
+;
+#endif//_NO_VERIFAST_
+
+#define ARRAY_RQ_EL_TYPE struct lcore_rx_queue
+#define ARRAY_RQ_CAPACITY MAX_RX_QUEUE_PER_LCORE
+#define ARRAY_RQ_EL_INIT lcore_rx_queue_init
+#define ARRAY_RQ_EL_TRACE_BREAKDOWN {                                   \
+    klee_trace_ret_ptr_field(offsetof(struct lcore_rx_queue, port_id),  \
+                             sizeof(uint8_t), "port_id");               \
+    klee_trace_ret_ptr_field(offsetof(struct lcore_rx_queue, queue_id), \
+                             sizeof(uint8_t), "queue_id");              \
+  }
+#define ARRAY_RQ_EL_TRACE_EP_BREAKDOWN(ep) {                            \
+    klee_trace_extra_ptr_field(ep, offsetof(struct lcore_rx_queue, port_id), \
+                               sizeof(uint8_t), "port_id");             \
+    klee_trace_extra_ptr_field(ep, offsetof(struct lcore_rx_queue, queue_id), \
+                               sizeof(uint8_t), "queue_id");            \
+  }
+
+#ifdef KLEE_VERIFICATION
+struct ArrayRq
+{
+  char dummy;
+};
+#else
+struct ArrayRq
+{
+  ARRAY_RQ_EL_TYPE data[ARRAY_RQ_CAPACITY];
+};
+#endif//KLEE_VERIFICATION
+
+/*@
+  inductive rx_queuei = rx_queuei(int,int);
+  //VVV here was a bug in the spec: 
+  //    this line was abscent: "0 <= pi &*& pi < RTE_MAX_ETHPORTS &*&"
+  predicate rx_queuep(rx_queuei rq, struct lcore_rx_queue *lrq) =
[…]

### 2017-06-26 nf/lib/containers/array-u16-impl.h

@@ -1,0 +1,202 @@
+#include "array-u16.h"
+#include "../ignore.h"
+
+//@ #include "arith.gh"
+//@ #include "stdex.gh"
+
+/*@
+  predicate arrp_u16(list<uint16_t> data, struct ArrayU16 *arr) =
+    struct_ArrayU16_padding(arr) &*&
+    true == ((char*)0 <= (void*)arr->data) &*&
+    true == ((void*)((unsigned short*)arr->data + ARRAY_U16_CAPACITY) <=
+             (char*)UINTPTR_MAX) &*&
+    ushorts(arr->data, ARRAY_U16_CAPACITY, data);
+
+  predicate arrp_u16_acc(list<uint16_t> data, struct ArrayU16 *arr, int idx) =
+    0 <= idx &*& idx < ARRAY_U16_CAPACITY &*&
+    struct_ArrayU16_padding(arr) &*&
+    true == ((char*)0 <= (void*)arr->data) &*&
+    true == ((void*)((unsigned short*)arr->data + ARRAY_U16_CAPACITY) <=
+             (char*)UINTPTR_MAX) &*&
+    length(data) == ARRAY_U16_CAPACITY &*&
+    ushorts(arr->data, idx, take(idx, data)) &*&
+    ushorts((unsigned short *)arr->data + idx + 1, ARRAY_U16_CAPACITY - idx - 1, drop(idx+1, data));
+
+
+  @*/
+
+#ifdef KLEE_VERIFICATION
+
+#include <klee/klee.h>
+
+ARRAY_U16_EL_TYPE array_u16_model_cell;
+int array_u16_allocated_index;
+int array_u16_index_allocated;
+struct ArrayU16 *array_u16_initialized;
+int array_u16_cell_is_exposed = 1;
+
+void array_u16_init(struct ArrayU16 *arr_out)
+{
+  // No need for tracing, as it is a nested call in the
+  // formally verified domain.
+  /* klee_trace_ret(); */
+  /* klee_trace_param_i32((uint32_t)arr_out, "arr_out"); */
+  if (!array_u16_cell_is_exposed)
+    klee_allow_access(&array_u16_model_cell, sizeof(ARRAY_U16_EL_TYPE));
+  klee_make_symbolic(&array_u16_model_cell, sizeof(ARRAY_U16_EL_TYPE),
+                     "array_u16_model_cell");
+  array_u16_index_allocated = 0;
[…]

### 2017-06-26 nf/lib/containers/array-u16-verifast.c

@@ -1,0 +1,1 @@
+array-u16-impl.h

### 2017-06-26 nf/lib/containers/batcher.c

@@ -1,0 +1,105 @@
+#include "batcher.h"
+
+/*@
+  predicate valsp(BATCHER_EL_TYPE* arr, int len, list<BATCHER_EL_TYPE> lst) =
+    pointers(arr, len, lst);
+    @*/
+
+/*@
+  predicate batcherp(list<BATCHER_EL_TYPE> batch, struct Batcher* bat) =
+    bat->len |-> ?l &*&
+    0 <= l &*& l <= BATCHER_CAPACITY &*&
+    valsp(bat->batch, l, batch) &*&
+    struct_Batcher_padding(bat) &*&
+    pointers((void*)((BATCHER_EL_TYPE*)(bat->batch) + l), BATCHER_CAPACITY-l, _);
+
+  predicate batcher_accp(struct Batcher* bat, int len) =
+    bat->len |-> len &*&
+    0 <= len &*& len <= BATCHER_CAPACITY &*&
+    struct_Batcher_padding(bat) &*&
+    pointers((void*)((BATCHER_EL_TYPE*)(bat->batch) + len),
+                     BATCHER_CAPACITY-len, _);
+  @*/
+
+void batcher_init(struct Batcher* bat_out)
+/*@ requires bat_out->len |-> _ &*&
+             pointers(bat_out->batch, BATCHER_CAPACITY, _) &*&
+             struct_Batcher_padding(bat_out); @*/
+//@ ensures batcherp(nil, bat_out);
+{
+  bat_out->len = 0;
+  //@ close valsp(bat_out->batch, 0, nil);
+  //@ close batcherp(nil, bat_out);
+}
+
+void batcher_push(struct Batcher *bat, BATCHER_EL_TYPE val)
+//@ requires batcherp(?b, bat) &*& length(b) < BATCHER_CAPACITY;
+//@ ensures batcherp(append(b, cons(val, nil)), bat);
+{
+  //@ open batcherp(b, bat);
+  //@ open valsp(bat->batch, bat->len, b);
+  bat->batch[bat->len] = val;
+  ++bat->len;
+  //@ close pointers((void*)((BATCHER_EL_TYPE*)(bat->batch) + bat->len-1), 1, cons(val, nil));
+  //@ pointers_join(bat->batch);
+  //@ close valsp(bat->batch, bat->len, append(b, cons(val, nil)));
+  //@ close batcherp(append(b, cons(val, nil)), bat);
+}
+
+void batcher_take_all(struct Batcher *bat,
[…]

### 2017-06-26 nf/lib/containers/batcher.h

@@ -1,0 +1,105 @@
+#ifndef _BATCHER_H_INCLUDED_
+#define _BATCHER_H_INCLUDED_
+
+#include "../static-component-params.h"
+
+#ifndef BATCHER_EL_TYPE
+#  error "Must define the batcher element type."
+#endif//BATCHER_EL_TYPE
+
+#ifndef BATCHER_CAPACITY
+#  error "Must define the batcher capacity."
+#endif// BATCHER_CAPACITY
+
+struct Batcher
+{
+  BATCHER_EL_TYPE batch[BATCHER_CAPACITY];
+  int len;
+};
+
+/*@
+  predicate valsp(BATCHER_EL_TYPE* arr, int len, list<BATCHER_EL_TYPE> lst);
+  @*/
+
+/*@
+  predicate batcherp(list<BATCHER_EL_TYPE> batch, struct Batcher* bat);
+  predicate batcher_accp(struct Batcher* bat, int len);
+  fixpoint BATCHER_EL_TYPE* batcher_contents(struct Batcher* bat) {
+    return bat->batch;
+  }
+  @*/
+
+/**
+   In-place initialization. Initializes the preallocated memory.
+
+   @param bat_out - pointer to a preallocated memory for struct Batcher.
+*/
+void batcher_init(struct Batcher *bat_out);
+/*@ requires bat_out->len |-> _ &*&
+             pointers(bat_out->batch, BATCHER_CAPACITY, _) &*&
+             struct_Batcher_padding(bat_out); @*/
+//@ ensures batcherp(nil, bat_out);
+
+/**
+   Push an element into the batcher. The batcher must be non-full.
+
+   @param bat - a pointer to the batcher.
+   @param val - the value to push into the batcher.
+ */
+void batcher_push(struct Batcher *bat, BATCHER_EL_TYPE val);
+//@ requires batcherp(?b, bat) &*& length(b) < BATCHER_CAPACITY;
+//@ ensures batcherp(append(b, cons(val, nil)), bat);
+
+/**
+   Get access to all the elements collected in the batcher so far. When you
+   are done with processing the elements, you must call batcher_empty to prepare
+   it for new pushes.
+
+   @param bat - a pointer to the batcher.
+   @param vals_out - output pointer to the 
+   @param count_out - an output pointer to the total number of elements stored.
+*/
+void batcher_take_all(struct Batcher *bat,
+                      BATCHER_EL_TYPE **vals_out, int *count_out);
[…]

### 2017-06-26 nf/lib/containers/double-chain-impl.c

@@ -1,0 +1,2519 @@
+#include "double-chain-impl.h"
+
+//@ #include <nat.gh>
+//@ #include <listex.gh>
+//@ #include "arith.gh"
+//@ #include "stdex.gh"
+
+enum DCHAIN_ENUM {
+    ALLOC_LIST_HEAD = 0,
+    FREE_LIST_HEAD = 1,
+    INDEX_SHIFT = DCHAIN_RESERVED
+};
+
+/*@
+
+  predicate free_listp(list<dcell> cells, list<int> fl, int start, int cur) =
+    switch(fl) {
+      case nil: return nth(cur, cells) == dcell(start, start);
+      case cons(h,t):
+        return nth(cur, cells) == dcell(h, h) &*&
+               cur != h &*&
+               free_listp(cells, t, start, h);
+    };
+
+  predicate alloc_listp(list<dcell> cells, list<int> al, int start, int cur) =
+    switch(al) {
+      case nil: return nth(cur, cells) == dcell(?x, start) &*&
+                       nth(start, cells) == dcell(cur, ?y) &*&
+                       cur == start ?
+                         (x == start &*& y == cur) :
+                         true;
+      case cons(h,t):
+        return nth(cur, cells) == dcell(?x, h) &*&
+               nth(h, cells) == dcell(cur, ?y) &*&
+               cur != h &*&
+               alloc_listp(cells, t, start, h);
+    };
+
+  fixpoint list<int> shift_inds_fp(list<int> inds, int shift) {
+    switch(inds) {
+      case nil: return nil;
+      case cons(h,t): return cons(h+shift, shift_inds_fp(t, shift));
+    }
+  }
+
+  fixpoint bool dbounded(int dbound, dcell val) {
+    return 0 <= dchain_cell_get_prev(val) &&
+                dchain_cell_get_prev(val) < dbound &&
+           0 <= dchain_cell_get_next(val) &&
+                dchain_cell_get_next(val) < dbound;
+  }
+
+  fixpoint bool lbounded(int lb, int x) {
+    return lb <= x;
+  }
+
+  fixpoint bool all_engaged(list<int> al, list<int> fl, nat size) {
+    switch(size) {
+      case zero: return true;
+      case succ(n):
+        return all_engaged(al, fl, n) &&
+              ( mem(int_of_nat(n) + INDEX_SHIFT, al) ?
+               !mem(int_of_nat(n) + INDEX_SHIFT, fl) :
[…]

### 2017-06-26 nf/lib/containers/double-map.c

@@ -1,0 +1,3407 @@
+#include <stdlib.h>
+#include <stdint.h>
+#include <string.h>
+#include "double-map.h"
+
+//@ #include "arith.gh"
+
+struct DoubleMap {
+  int value_size;
+
+  uq_value_copy* cpy;
+  uq_value_destr* dstr;
+
+  uint8_t *values;
+
+  int *bbs_a;
+  void **kps_a;
+  int *khs_a;
+  int *chns_a;
+  int *inds_a;
+  map_keys_equality *eq_a;
+  map_key_hash *hsh_a;
+
+  int *bbs_b;
+  void **kps_b;
+  int *khs_b;
+  int *chns_b;
+  int *inds_b;
+  map_keys_equality *eq_b;
+  map_key_hash *hsh_b;
+
+  dmap_extract_keys *exk;
+  dmap_pack_keys *pk;
+
+  int n_vals;
+  int capacity;
+  int keys_capacity;
+};
+
+/*@
+  predicate valsp<t1,t2,vt>(void* values, int val_size,
+                            predicate (void*,vt) fvp,
+                            predicate (void*,vt) bvp,
+                            list<pair<t1,void*> > addrs1,
+                            list<pair<t2,void*> > addrs2,
+                            fixpoint (vt,t1) vk1,
+                            fixpoint (vt,t2) vk2,
+                            fixpoint (void*,void*,void*,bool) rof,
+                            int length, list<option<vt> > vals) =
+     switch(vals) {
+       case nil: return length == 0;
+       case cons(h,t):
+         return switch(h) {
+           case none: return chars(values, val_size, _) &*&
+                             valsp(values + val_size,
+                                   val_size, fvp, bvp,
+                                   addrs1, addrs2,
+                                   vk1, vk2, rof,
+                                   length-1, t);
+           case some(v): return [0.5]fvp(values, v) &*&
+                                [0.5]bvp(values, v) &*&
+                                true == map_has_fp(addrs1, vk1(v)) &*&
+                                true == map_has_fp(addrs2, vk2(v)) &*&
+                                true == rof(values,
+                                            map_get_fp(addrs1, vk1(v)),
+                                            map_get_fp(addrs2, vk2(v))) &*&
[…]

### 2017-06-26 nf/lib/containers/double-map.h

@@ -1,0 +1,730 @@
+#ifndef _DOUBLE_MAP_H_INCLUDED_
+#define _DOUBLE_MAP_H_INCLUDED_
+
+#include "map-impl.h"
+#include "map-util.h"
+
+//@ #include <nat.gh>
+//@ #include "stdex.gh"
+
+/*
+  This implementation expects keys to be the part of the value. The keys
+  are extracted with dmap_extract_keys function and are put back with
+  dmap_pack_keys.
+ */
+
+typedef void uq_value_copy/*@<K>(predicate (void*; K) vp, int size) @*/(char* dst, void* src);
+//@ requires [?fr]vp(src, ?v) &*& dst[0..size] |-> _;
+//@ ensures [fr]vp(src, v) &*& vp(dst, v);
+
+typedef void dmap_extract_keys/*@ <K1,K2,V>
+                                (predicate (void*; K1) keyp1,
+                                 predicate (void*; K2) keyp2,
+                                 predicate (void*; V) full_valp,
+                                 predicate (void*, V) bare_valp,
+                                 fixpoint (void*, void*, void*, bool)
+                                   right_offsets,
+                                 fixpoint (V,K1) vk1,
+                                 fixpoint (V,K2) vk2)
+                              @*/
+                              (void* vp, void** kpp1, void** kpp2);
+//@ requires [?fr]full_valp(vp, ?v) &*& *kpp1 |-> _ &*& *kpp2 |-> _;
+/*@ ensures [fr]bare_valp(vp, v) &*& *kpp1 |-> ?kp1 &*& *kpp2 |-> ?kp2 &*&
+            [fr]keyp1(kp1, ?k1) &*& [fr]keyp2(kp2, ?k2) &*&
+            true == right_offsets(vp, kp1, kp2) &*&
+            k1 == vk1(v) &*&
+            k2 == vk2(v); @*/
+
+//TODO: replace with pack key halves first and second,
+// because it is called two times
+typedef void dmap_pack_keys/*@ <K1,K2,V>
+                             (predicate (void*; K1) keyp1,
+                              predicate (void*; K2) keyp2,
+                              predicate (void*; V) full_valp,
+                              predicate (void*, V) bare_valp,
+                              fixpoint (void*, void*, void*, bool)
[…]

### 2017-06-26 nf/lib/containers/map-impl.c

@@ -1,0 +1,2607 @@
+#include <stdint.h>
+#include <string.h>
+
+#include "map-impl.h"
+#include "lib/ignore.h"
+
+//@ #include <list.gh>
+//@ #include <listex.gh>
+//@ #include <nat.gh>
+//@ #include "stdex.gh"
+//@ #include "nth_prop.gh"
+//@ #include "modulo.gh"
+//@ #include "map.gh"
+//@ #include "natlist.gh"
+//@ #include "chain-buckets.gh"
+
+
+/*@
+  fixpoint bool has_given_hash_fp<kt>(fixpoint (kt, int) hash,
+                                      int pos, int capacity,
+                                      pair<kt, nat> chain) {
+    return pos == loop_fp(hash(fst(chain)), capacity);
+  }
+
+  fixpoint bool key_chains_start_on_hash_fp<kt>(list<bucket<kt> > buckets,
+                                                int pos, int capacity,
+                                                fixpoint (kt, int) hash) {
+    switch(buckets) {
+      case nil: return true;
+      case cons(h,t):
+        return switch(h) { case bucket(chains):
+            return forall(chains, (has_given_hash_fp)(hash, pos, capacity)) &&
+                   key_chains_start_on_hash_fp(t, pos + 1, capacity, hash);
+          };
+    }
+  }
+  @*/
+
+static
+int loop(int k, int capacity)
+//@ requires 0 < capacity &*& 2*capacity < INT_MAX;
+/*@ ensures 0 <= result &*& result < capacity &*&
+            result == loop_fp(k, capacity); @*/
+{
+  int g = k%capacity;
+  //@ div_mod(g, k, capacity);
+  //@ assert(2*capacity< INT_MAX);
+  int res = (g + capacity)%capacity;
+  //@ div_mod_gt_0(res, g + capacity, capacity);
+  return res;
+}
+
+/*@
+  inductive hmap<kt> = hmap(list<option<kt> >, list<int>);
+
+  predicate hmapping<kt>(predicate (void*; kt) keyp,
+                         fixpoint (kt, int) hash,
+                         int capacity,
+                         int* busybits,
+                         list<void*> kps,
+                         int* k_hashes;
+                         hmap<kt> m);
+
+  fixpoint list<option<kt> > hmap_ks_fp<kt>(hmap<kt> m) {
+    switch(m) { case hmap(ks, khs): return ks; }
+  }
+
[…]

### 2017-06-26 nf/lib/containers/map-impl.h

@@ -1,0 +1,196 @@
+#ifndef _MAP_IMPL_H_INCLUDED_
+#define _MAP_IMPL_H_INCLUDED_
+
+//@ #include "stdex.gh"
+
+typedef int map_keys_equality/*@<K>(predicate (void*; K) keyp) @*/(void* k1, void* k2);
+//@ requires [?fr1]keyp(k1, ?kk1) &*& [?fr2]keyp(k2, ?kk2);
+//@ ensures [fr1]keyp(k1, kk1) &*& [fr2]keyp(k2, kk2) &*& (0 == result ? (kk1 != kk2) : (kk1 == kk2));
+
+
+/*@ predicate pred_arg4<t1,t2,t3,t4>(predicate (t1,t2,t3,t4) p) = true;
+    predicate pred_arg2<t1,t2>(predicate (t1,t2) p) = true;
+  @*/
+
+/*@
+  // map<kt> = list<pair<kt,int> >;
+
+  predicate mapping<kt>(list<pair<kt,int> > m,
+                        list<pair<kt,void*> > addrs,
+                        predicate (void*;kt) keyp,
+                        fixpoint (kt,int,bool) recp,
+                        fixpoint (kt,int) hash,
+                        int capacity,
+                        int* busybits,
+                        void** keyps,
+                        int* k_hashes,
+                        int* chns,
+                        int* values);
+
+  fixpoint list<pair<kt,vt> > empty_map_fp<kt,vt>() { return nil; }
+
+  fixpoint vt map_get_fp<kt,vt>(list<pair<kt,vt> > m, kt key) {
+    switch(m) {
+      case nil: return default_value<vt>();
+      case cons(h,t):
+        return (fst(h) == key ? snd(h) : map_get_fp(t, key));
+    }
+  }
+
+  fixpoint bool map_has_fp<kt,vt>(list<pair<kt,vt> > m, kt key) {
+    switch(m) {
+      case nil: return false;
+      case cons(h,t):
+        return (fst(h) == key) || map_has_fp(t, key);
+    }
+  }
+
+  fixpoint list<pair<kt,vt> > map_put_fp<kt,vt>(list<pair<kt,vt> > m,
+                                                kt key, vt val) {
+    return cons(pair(key,val), m);
+  }
+
+  fixpoint list<pair<kt,vt> > map_erase_fp<kt,vt>(list<pair<kt,vt> > m, kt key) {
+    switch(m) {
[…]

### 2017-06-26 nf/lib/containers/map.h

@@ -1,0 +1,88 @@
+#ifndef _MAP_H_INCLUDED_
+#define _MAP_H_INCLUDED_
+
+#include "map-impl.h"
+#include "map-util.h"
+
+struct Map;
+
+/*@
+  inductive mapi<t> = mapc(int, list<pair<t, int> >,
+                           list<pair<t, void*> >);
+
+  predicate mapp<t>(struct Map* ptr,
+                    predicate (void*;t) kp,
+                    fixpoint (t,int) hsh,
+                    fixpoint (t,int,bool) recp,
+                    mapi<t> map);
+
+  fixpoint bool nop_true<t>(t key, int val) { return true; }
+  @*/
+
+/*@
+  lemma void map_get_mem<t>(list<pair<t, int> > m, t k);
+  requires true == map_has_fp(m, k);
+  ensures true == mem(pair(k, map_get_fp(m, k)), m);
+  @*/
+
+int map_allocate/*@ <t> @*/(map_keys_equality* keq,
+                            map_key_hash* khash, int capacity,
+                            struct Map** map_out);
+/*@ requires 0 < capacity &*& capacity < CAPACITY_UPPER_LIMIT &*&
+             [_]is_map_keys_equality<t>(keq, ?kp) &*&
+             [_]is_map_key_hash<t>(khash, kp, ?hsh) &*&
+             *map_out |-> ?old_mo; @*/
+/*@ ensures result == 0 ?
+              *map_out |-> old_mo :
+              (*map_out |-> ?new_mo &*&
+               result == 1 &*&
+               mapp<t>(new_mo, kp, hsh, nop_true,
+                       mapc(capacity, nil, nil))); @*/
+
+int map_get/*@ <t> @*/(struct Map* map, void* key, int* value_out);
+/*@ requires mapp<t>(map, ?kp, ?hsh, ?recp,
+                     mapc(?capacity, ?contents, ?addrs)) &*&
+             kp(key, ?k) &*&
+             *value_out |-> ?old_v; @*/
+/*@ ensures mapp<t>(map, kp, hsh, recp,
+                    mapc(capacity, contents, addrs)) &*&
+            kp(key, k) &*&
+            map_has_fp(contents, k) ?
+              (result == 1 &*&
+               *value_out |-> ?new_v &*&
+               new_v == map_get_fp(contents, k)) :
+              (result == 0 &*&
+               *value_out |-> old_v); @*/
+
[…]

### 2017-06-26 nf/lib/containers/vector.c

@@ -1,0 +1,239 @@
+#include <stdlib.h>
+#include <stdint.h>
+#include "vector.h"
+
+//@ #include "arith.gh"
+//@ #include "stdex.gh"
+
+struct Vector {
+  char* data;
+  int elem_size;
+  int capacity;
+};
+
+/*@
+  predicate entsp<t>(void* data, int el_size,
+                     predicate (void*;t) entp,
+                     int length,
+                     list<t> vals) =
+    switch(vals) {
+      case nil:
+        return length == 0;
+      case cons(h,t):
+        return entp(data, h) &*&
+               entsp(data + el_size,
+                     el_size, entp,
+                     length - 1,
+                     t);
+    };
+
+  predicate vector_basep<t>(struct Vector* vector;
+                            int el_size,
+                            int cap,
+                            char* data) =
+    malloc_block_Vector(vector) &*&
+    vector->data |-> data &*&
+    vector->elem_size |-> el_size &*&
+    0 < el_size &*& el_size < 4096 &*&
+    vector->capacity |-> cap &*&
+    0 <= cap &*& cap < VECTOR_CAPACITY_UPPER_LIMIT &*&
+    malloc_block(data, el_size*cap) &*&
+    data + el_size*cap <= (void*)UINTPTR_MAX;
+
+
+  predicate vectorp<t>(struct Vector* vector,
+                       predicate (void*;t) entp,
+                       list<t> values) =
+    vector_basep<t>(vector, ?el_size, ?cap, ?data) &*&
+    cap == length(values) &*&
+    entsp(data, el_size, entp, cap, values);
+
+  predicate vector_accp<t>(struct Vector* vector,
+                           predicate (void*;t) entp,
+                           list<t> values,
+                           int accessed_idx,
+                           void* entry) =
+    vector_basep<t>(vector, ?el_size, ?cap, ?data) &*&
+    cap == length(values) &*&
+    0 <= accessed_idx &*& accessed_idx < cap &*&
+    entry == data + el_size*accessed_idx &*&
+    entsp(data, el_size, entp, accessed_idx, take(accessed_idx, values)) &*&
+    entsp(data + el_size*(accessed_idx + 1), el_size, entp,
[…]

### 2017-06-26 nf/lib/expirator.c

@@ -1,0 +1,141 @@
+#include <assert.h>
+#include "containers/double-chain.h"
+#include "coherence.h"
+#include "expirator.h"
+
+
+/*@
+  lemma void expire_0_indexes(dchain ch, uint32_t time)
+  requires true;
+  ensures ch == expire_n_indexes(ch, time, 0);
+  {
+    switch(ch) { case dchain(alist, size, l, h):
+      take_0(get_expired_indexes_fp(time, alist));
+    }
+  }
+  @*/
+
+
+/* @
+  lemma void erase_nothing<K1,K2,V>(dmap<K1,K2,V> m)
+  requires true;
+  ensures dmap_erase_all_fp(m, nil) == m;
+  {
+    switch(m) { case dmap(ks1, ks2, idxs):
+    }
+  }
+  @*/
+
+int expire_items/*@<K1,K2,V> @*/(struct DoubleChain* chain,
+                                 struct DoubleMap* map,
+                                 uint32_t time)
+/*@ requires dmappingp<K1,K2,V>(?m, ?kp1, ?kp2, ?hsh1, ?hsh2,
+                                ?fvp, ?bvp, ?rof, ?vsz,
+                                ?vk1, ?vk2, ?rp1, ?rp2, map) &*&
+             double_chainp(?ch, chain) &*&
+             dchain_index_range_fp(ch) < INT_MAX &*&
+             dmap_dchain_coherent<K1,K2,V>(m, ch); @*/
+/*@ ensures dmappingp<K1,K2,V>(?nm,
+                               kp1, kp2, hsh1, hsh2, fvp, bvp, rof, vsz,
+                               vk1, vk2, rp1, rp2, map) &*&
+            nm == dmap_erase_all_fp
+                               (m, dchain_get_expired_indexes_fp(ch, time),
+                                vk1, vk2) &*&
+            double_chainp(?nch, chain) &*&
+            nch == dchain_expire_old_indexes_fp(ch, time) &*&
+            dmap_dchain_coherent<K1,K2,V>(nm, nch) &*&
+            result == length(dchain_get_expired_indexes_fp(ch, time)); @*/
+{
+  int count = 0;
+  int index = -1;
+  //@ expire_0_indexes(ch, time);
+  //@ assert take(count, dchain_get_expired_indexes_fp(ch, time)) == nil;
+  //@ double_chainp_to_sorted(ch);
+  //@ dchain_expired_indexes_limited(ch, time);
[…]

### 2017-06-26 nf/lib/flow-log.c

@@ -1,0 +1,56 @@
+#include <stdio.h>
+
+#include "lib/static-component-params.h"
+
+#include "flow.h"
+
+#include "lib/nf_log.h"
+
+inline void log_ip(uint32_t addr) {
+#if !ENABLE_LOG
+  (void)addr;
+#endif//NOLOG
+  NF_DEBUG( "%d.%d.%d.%d",
+            addr&0xff, (addr>>8)&0xff,
+            (addr>>16)&0xff, (addr>>24)&0xff);
+}
+
+void log_int_key(const struct int_key *key) {
+  NF_DEBUG( "{int_src_port: %d(%d); dst_port: %d(%d);\n"
+            " int_src_ip: ",
+            key->int_src_port, rte_be_to_cpu_16(key->int_src_port),
+            key->dst_port, rte_be_to_cpu_16(key->dst_port));
+  log_ip(key->int_src_ip);
+  NF_DEBUG( "; dst_ip: ");
+  log_ip(key->dst_ip);
+  NF_DEBUG(" int_device_id: %d; protocol: %d}",
+           key->int_device_id, key->protocol);
+}
+
+void log_ext_key(const struct ext_key *key) {
+  NF_DEBUG( "{ext_src_port: %d(%d); dst_port: %d(%d);\n"
+            " ext_src_ip: ",
+            key->ext_src_port, rte_be_to_cpu_16(key->ext_src_port),
+            key->dst_port, rte_be_to_cpu_16(key->dst_port));
+  log_ip(key->ext_src_ip);
+  NF_DEBUG( "; dst_ip: ");
+  log_ip(key->dst_ip);
+  NF_DEBUG(" ext_device_id: %d; protocol: %d}",
+           key->ext_device_id, key->protocol);
+}
+
+void log_flow(const struct flow *f) {
+  NF_DEBUG( "{int_src_port: %d(%d); ext_src_port: %d(%d);\n"
+            " dst_port: %d(%d); int_src_ip: ",
+            f->int_src_port, rte_be_to_cpu_16(f->int_src_port),
+            f->ext_src_port, rte_be_to_cpu_16(f->ext_src_port),
+            f->dst_port, rte_be_to_cpu_16(f->dst_port));
+  log_ip(f->int_src_ip);
+  NF_DEBUG( ";\n ext_src_ip:");
+  log_ip(f->ext_src_ip);
+  NF_DEBUG( "; dst_ip: ");
+  log_ip(f->dst_ip);
[…]

### 2017-06-26 nf/lib/flow.c

@@ -1,0 +1,303 @@
+#include <stdbool.h>
+#include <string.h>
+#include <limits.h>
+
+#include "include_ignored_by_verifast.h"
+
+#include "flow.h"
+#include "ignore.h"
+
+#ifdef KLEE_VERIFICATION
+#  define AND(x,y) ((x)&(y))
+#  define TO_INT(x) (x)
+#else //KLEE_VERIFICATION
+#  define AND(x,y) ((x)&&(y))
+#  define TO_INT(x) (x ? 1 : 0)
+#endif //KLEE_VERIFICATION
+
+int int_key_eq(void* a, void* b)
+//@ requires [?f1]int_k_p(a, ?ak) &*& [?f2]int_k_p(b, ?bk);
+//@ ensures [f1]int_k_p(a, ak) &*& [f2]int_k_p(b, bk) &*& (0 == result ? (ak != bk) : (ak == bk));
+{
+  struct int_key* k1 = a;
+  struct int_key* k2 = b;
+  bool rez =
+    AND(k1->int_src_port  == k2->int_src_port,
+    AND(k1->dst_port      == k2->dst_port,
+    AND(k1->int_src_ip    == k2->int_src_ip,
+    AND(k1->dst_ip        == k2->dst_ip,
+    AND(k1->int_device_id == k2->int_device_id,
+       (k1->protocol      == k2->protocol))))));
+  return TO_INT(rez);
+}
+
+int ext_key_eq(void* a, void* b)
+//@ requires [?f1]ext_k_p(a, ?ak) &*& [?f2]ext_k_p(b, ?bk);
+//@ ensures [f1]ext_k_p(a, ak) &*& [f2]ext_k_p(b, bk) &*& (0 == result ? (ak != bk) : (ak == bk));
+{
+  struct ext_key* k1 = a;
+  struct ext_key* k2 = b;
+  bool rez =
+    AND(k1->ext_src_port  == k2->ext_src_port,
+    AND(k1->dst_port      == k2->dst_port,
+    AND(k1->ext_src_ip    == k2->ext_src_ip,
+    AND(k1->dst_ip        == k2->dst_ip,
+    AND(k1->ext_device_id == k2->ext_device_id,
+       (k1->protocol      == k2->protocol))))));
+  return TO_INT(rez);
+}
+
+static long long wrap(long long x)
+//@ requires true;
[…]

### 2017-06-26 nf/lib/flow.h

@@ -1,0 +1,347 @@
+#ifndef _FLOW_H_INCLUDED_
+#define _FLOW_H_INCLUDED_
+#include <stdint.h>
+
+/**
+  The "internal" key - the part of the flow ID, related to the internal network.
+  Used in a hash map to identify flows for packets coming from the internal
+  network.
+*/
+struct int_key {
+  uint16_t int_src_port;
+  uint16_t dst_port;
+  uint32_t int_src_ip;
+  uint32_t dst_ip;
+  uint8_t int_device_id;
+  uint8_t protocol;
+};
+
+/**
+  The "external" key - the part of the flow ID, related to the external network.
+  Used in a hash map to identify flows for packets coming from the external
+  network.
+*/
+struct ext_key {
+  uint16_t ext_src_port;
+  uint16_t dst_port;
+  uint32_t ext_src_ip;
+  uint32_t dst_ip;
+  uint8_t ext_device_id;
+  uint8_t protocol;
+};
+
+/**
+  The flow ID. A 7-tuple holding all elements necessary to identify both sides
+  of a flow: the internal and the external. This structure helps NAT to
+  translate between internal and external 5-tuples.
+ */
+struct flow {
+  // This incapsulation simplifies memory management, but may hurt locality.
+  struct int_key ik;//2.5x redundancy.
+  struct ext_key ek;//Can be optimized, but we do not wont that mess.
+  uint16_t int_src_port;
+  uint16_t ext_src_port;
+  uint16_t dst_port;
+  uint32_t int_src_ip;
+  uint32_t ext_src_ip;
+  uint32_t dst_ip;
+  uint8_t int_device_id;
+  uint8_t ext_device_id;
+  uint8_t protocol;
+};
+
+/**
+  Logging functions for the structures defined above.
+*/
+void log_ip(uint32_t addr);
+void log_int_key(const struct int_key *key);
+void log_ext_key(const struct ext_key *key);
+void log_flow(const struct flow *f);
+
+#ifdef KLEE_VERIFICATION
+
+#include <klee/klee.h>
+#include "lib/stubs/containers/double-map-stub-control.h"
+#include "lib/static-component-params.h"
+
+//TODO: this is dirty.
[…]

### 2017-06-26 nf/lib/nat_config.h

@@ -1,0 +1,44 @@
+#pragma once
+
+#include <inttypes.h>
+
+// TODO can't really include rte_max_ethports here but need it :/
+
+// rte_ether
+struct ether_addr;
+
+
+struct nat_config {
+	// "Main" LAN (i.e. internal) device, used for dumb forwarding
+	uint8_t lan_main_device;
+
+	// WAN device, i.e. external
+	uint8_t wan_device;
+
+	// External IP address
+	uint32_t external_addr;
+
+	// MAC addresses of devices
+	struct ether_addr device_macs[RTE_MAX_ETHPORTS];
+
+	// MAC addresses of the endpoints the devices are linked to
+	struct ether_addr endpoint_macs[RTE_MAX_ETHPORTS];
+
+	// External port at which to start allocating flows
+	// i.e. ports will be allocated in [start_port, start_port + max_flows]
+	uint16_t start_port;
+
+	// Expiration time of flows in seconds
+	uint32_t expiration_time;
+
+	// Size of the flow table
+	uint32_t max_flows;
+};
+
+
+void nat_config_init(struct nat_config* config,
+                     int argc, char** argv);
+
+void nat_config_cmdline_print_usage(void);
+
+void nat_print_config(struct nat_config* config);

### 2017-06-26 nf/lib/nf_forward.h

@@ -1,0 +1,30 @@
+#pragma once
+
+#include <inttypes.h>
+
+struct rte_mbuf;
+
+void nf_core_init(void);
+
+// TODO make the uint32_t a time_t
+int nf_core_process(uint8_t device,
+                    struct rte_mbuf* mbuf,
+                    uint32_t now);
+enum nf_core_special_results {
+  FLOOD_FRAME = -1
+};
+
+void nf_config_init(int argc, char** argv);
+void nf_config_cmdline_print_usage(void);
+void nf_print_config(void);
+
+#ifdef KLEE_VERIFICATION
+void nf_loop_iteration_begin(unsigned lcore_id,
+                             uint32_t time);
+
+void nf_add_loop_iteration_assumptions(unsigned lcore_id,
+                                       uint32_t time);
+
+void nf_loop_iteration_end(unsigned lcore_id,
+                           uint32_t time);
+#endif //KLEE_VERIFICATION

### 2017-06-26 nf/lib/nf_log.h

@@ -1,0 +1,16 @@
+#pragma once
+
+#if KLEE_VERIFICATION
+	#define NF_INFO(...)
+	#define NF_DEBUG(...)
+#else
+	#include <stdio.h>
+
+	#define NF_INFO(text, ...) printf(text "\n", ##__VA_ARGS__); fflush(stdout);
+
+	#if ENABLE_LOG
+		#define NF_DEBUG(text, ...) printf("DEBUG: " text "\n", ##__VA_ARGS__); fflush(stdout);
+	#else
+		#define NF_DEBUG(...)
+	#endif
+#endif

### 2017-06-26 nf/lib/nf_time.c

@@ -1,0 +1,10 @@
+#include <time.h>
+#include "nf_time.h"
+
+uint32_t current_time(void)
+//@ requires last_time(?x);
+//@ ensures x <= result &*& last_time(result);
+{
+  // see https://github.com/verifast/verifast/issues/35
+  return time(NULL);
+}

### 2017-06-26 nf/lib/nf_util.c

@@ -1,0 +1,115 @@
+#include <inttypes.h>
+
+#include <netinet/in.h>
+
+#ifdef KLEE_VERIFICATION
+#  include "stubs/rte_stubs.h"
+#else//KLEE_VERIFICATION
+#  include <rte_byteorder.h>
+#  include <rte_ether.h>
+#  include <rte_ip.h>
+#  include <rte_mbuf.h>
+#  include <rte_tcp.h>
+#  include <rte_udp.h>
+#endif//KLEE_VERIFICATION
+
+#include "nf_util.h"
+
+
+struct ether_hdr*
+nf_get_mbuf_ether_header(struct rte_mbuf* mbuf)
+{
+	return rte_pktmbuf_mtod(mbuf, struct ether_hdr*);
+}
+
+// TODO for consistency it'd be nice if this took an ether_hdr as argument, or if they all took rte_mbuf
+struct ipv4_hdr*
+nf_get_mbuf_ipv4_header(struct rte_mbuf* mbuf)
+{
+	struct ether_hdr* ether_header = nf_get_mbuf_ether_header(mbuf);
+	if (!RTE_ETH_IS_IPV4_HDR(mbuf->packet_type) &&
+      !(mbuf->packet_type == 0 &&
+        ether_header->ether_type == rte_cpu_to_be_16(ETHER_TYPE_IPv4))) {
+		return NULL;
+	}
+
+	return rte_pktmbuf_mtod_offset(mbuf, struct ipv4_hdr*, sizeof(struct ether_hdr));
+}
+
+struct tcpudp_hdr*
+nf_get_ipv4_tcpudp_header(struct ipv4_hdr* header)
+{
+	if (header->next_proto_id != IPPROTO_TCP &&
+      header->next_proto_id != IPPROTO_UDP) {
+		return NULL;
+	}
+
+	uint8_t offset = header->version_ihl & IPV4_HDR_IHL_MASK;
+	// TODO use offset
+	return (struct tcpudp_hdr*)(header + 1);
+}
+
+void
+nf_set_ipv4_checksum(struct ipv4_hdr* header)
+{
+	// TODO: See if can be offloaded to hardware
+	header->hdr_checksum = 0;
+
+	if (header->next_proto_id == IPPROTO_TCP) {
+		struct tcp_hdr* tcp_header = (struct tcp_hdr*)(header + 1);
+		tcp_header->cksum = 0;
+		tcp_header->cksum = rte_ipv4_udptcp_cksum(header, tcp_header);
[…]

### 2017-06-26 nf/lib/static-component-params.h

@@ -1,0 +1,122 @@
+#ifndef _STATIC_COMPONENT_PARAMS_H_INCLUDE_
+#define _STATIC_COMPONENT_PARAMS_H_INCLUDE_
+
+#include "include_ignored_by_verifast.h"
+
+#include <stdint.h>
+
+#ifdef KLEE_VERIFICATION
+#  include "lib/stubs/rte_stubs.h"
+#else//KLEE_VERIFICATION
+
+#ifdef _NO_VERIFAST_
+#  include <sys/queue.h>
+#  include <rte_common.h>
+#  include <rte_vect.h>
+#  include <rte_byteorder.h>
+#  include <rte_memory.h>
+#  include <rte_memcpy.h>
+#  include <rte_memzone.h>
+#  include <rte_eal.h>
+#  include <rte_per_lcore.h>
+#  include <rte_launch.h>
+#  include <rte_atomic.h>
+#  include <rte_cycles.h>
+#  include <rte_prefetch.h>
+#  include <rte_lcore.h>
+#  include <rte_per_lcore.h>
+#  include <rte_branch_prediction.h>
+#  include <rte_interrupts.h>
+#  include <rte_pci.h>
+#  include <rte_random.h>
+#  include <rte_debug.h>
+#  include <rte_ether.h>
+#  include <rte_ethdev.h>
+#  include <rte_ring.h>
+#  include <rte_mempool.h>
+#  include <rte_mbuf.h>
+#  include <rte_ip.h>
+#  include <rte_tcp.h>
+#  include <rte_udp.h>
+#  include <rte_string_fns.h>
+#else //_NO_VERIFAST_
+
+#include "lib/custom_ether_addr_def.h"
+
+struct rte_mempool {};
+
+struct ether_hdr {
+  struct ether_addr d_addr; /**< Destination address. */
+  struct ether_addr s_addr; /**< Source address. */
+  uint16_t ether_type;      /**< Frame type. */
+};
+
+struct tcp_hdr {
+  uint16_t src_port;  /**< TCP source port. */
+  uint16_t dst_port;  /**< TCP destination port. */
+  uint32_t sent_seq;  /**< TX data sequence number. */
+  uint32_t recv_ack;  /**< RX data acknowledgement sequence number. */
+  uint8_t  data_off;  /**< Data offset. */
+  uint8_t  tcp_flags; /**< TCP flags */
+  uint16_t rx_win;    /**< RX flow control window. */
[…]

### 2017-06-26 nf/lib/stubs/containers/batcher-stub.c

@@ -1,0 +1,77 @@
+#include "lib/containers/batcher.h"
+#include <klee/klee.h>
+
+int alloc_len;
+int was_read;
+BATCHER_EL_TYPE alloc_value[BATCHER_CAPACITY];
+struct Batcher* batcher_initialized;
+
+void batcher_init(struct Batcher* bat_out)
+{
+  // No need to trace this one, as it will be called from the inside of
+  // another traced function, specifically:
+  // array1_init -> array2_init -> batcher_init
+
+  alloc_len = klee_range(0, BATCHER_CAPACITY, "batcher_alloc_len");
+  klee_make_symbolic(alloc_value, BATCHER_CAPACITY*sizeof(BATCHER_EL_TYPE),
+                     "batcher_alloc_value");
+  was_read = 0;
+  batcher_initialized = bat_out;
+}
+
+void batcher_push(struct Batcher *bat, BATCHER_EL_TYPE val)
+{
+  klee_trace_ret();
+  klee_trace_param_just_ptr(bat, sizeof(struct Batcher), "bat");
+  klee_trace_param_just_ptr(val, sizeof(struct rte_mbuf), "val");
+
+  klee_assert(bat == batcher_initialized);
+  klee_assert(alloc_len < BATCHER_CAPACITY);
+  alloc_len += 1;
+  // Forget the value itself.
+}
+
+void batcher_take_all(struct Batcher *bat,
+                      BATCHER_EL_TYPE **vals_out,
+                      int *count_out)
+{
+  klee_trace_ret();
+  klee_trace_param_just_ptr(bat, sizeof(struct Batcher), "bat");
+  //TODO: A proper tracing here.
+  klee_trace_param_just_ptr(vals_out, sizeof(BATCHER_EL_TYPE*), "vals_out");
+  klee_trace_param_ptr(count_out, sizeof(int), "count_out");
+
+  klee_assert(bat == batcher_initialized);
+  klee_assert(was_read == 0);
+  was_read = 1;
+  *vals_out = alloc_value;
+  *count_out = alloc_len;
+  alloc_len = 0;
+}
+
+void batcher_empty(struct Batcher *bat)
+{
+  klee_trace_ret();
+  klee_trace_param_just_ptr(bat, sizeof(struct Batcher), "bat");
+  alloc_len = 0;
+}
+
+int batcher_full(struct Batcher *bat)
+{
+  klee_trace_ret();
[…]

### 2017-06-26 nf/lib/stubs/containers/double-chain-stub-control.h

@@ -1,0 +1,10 @@
+#ifndef _DOUBLE_CHAIN_STUB_CONTROL_H_INCLUDED_
+#define _DOUBLE_CHAIN_STUB_CONTROL_H_INCLUDED_
+
+#include "lib/containers/double-chain.h"
+
+void dchain_make_space(struct DoubleChain* chain);
+
+void dchain_reset(struct DoubleChain* chain, int index_range);
+
+#endif//_DOUBLE_CHAIN_STUB_CONTROL_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/containers/double-chain-stub.c

@@ -1,0 +1,103 @@
+#include <assert.h>
+#include <klee/klee.h>
+#include "lib/containers/double-chain.h"
+
+// TODO: double check that this model is enough for the NAT scenario
+
+int is_dchain_allocated;
+int out_of_space;
+int new_index;
+int is_index_allocated;
+
+struct DoubleChain* allocated_chain;
+
+int dchain_allocate(int index_range, struct DoubleChain** chain_out) {
+    klee_trace_ret();
+    klee_trace_param_i32(index_range, "index_range");
+    klee_trace_param_ptr(chain_out, sizeof(struct DoubleChain*), "chain_out");
+
+    is_dchain_allocated = klee_int("is_dchain_allocated");
+
+    if (is_dchain_allocated) {
+      klee_make_symbolic(&allocated_chain, sizeof(struct DoubleChain*),
+                         "allocated_chain_do_not_dereference");
+      *chain_out = allocated_chain;
+      new_index = klee_int("new_index");
+      klee_assume(0 <= new_index);
+      klee_assume(new_index < index_range);
+      is_index_allocated = 0;
+      out_of_space = klee_int("out_of_space");
+      return 1;
+    } else {
+      return 0;
+    }
+}
+
+int dchain_allocate_new_index(struct DoubleChain* chain, int *index_out,
+                              uint32_t time) {
+    klee_trace_ret();
+    //Deliberately trace this pointer as an integer to avoid
+    // dereference.
+    klee_trace_param_i32((uint32_t)chain, "chain");
+    klee_trace_param_ptr(index_out, sizeof(int), "index_out");
+    klee_trace_param_i32(time, "time");
+
+    klee_assert(is_dchain_allocated);
+    klee_assert(chain == allocated_chain);
+    if (out_of_space) return 0;
+    klee_assert(!is_index_allocated);
+    *index_out = new_index;
+    is_index_allocated = 1;
+    return 1;
+}
+
+int dchain_rejuvenate_index(struct DoubleChain* chain, int index,
+                            uint32_t time) {
+    klee_trace_ret();
+    //Deliberately trace this pointer as an integer to avoid
+    // dereference.
+    klee_trace_param_i32((uint32_t)chain, "chain");
+    klee_trace_param_i32(index, "index");
+    klee_trace_param_i32(time, "time");
+
+    klee_assert(is_dchain_allocated);
+    klee_assert(chain == allocated_chain);
[…]

### 2017-06-26 nf/lib/stubs/containers/double-map-stub-control.h

@@ -1,0 +1,20 @@
+#ifndef _DOUBLE_MAP_STUB_CONTROL_H_INCLUDED_
+#define _DOUBLE_MAP_STUB_CONTROL_H_INCLUDED_
+#include "str-descr.h"
+//@ #include "lib/predicates.gh"
+
+typedef int entry_condition(void* key_a, void* key_b, int index, void* value);
+
+void dmap_set_entry_condition(entry_condition* cond);
+//@ requires true;
+//@ ensures true;
+void dmap_set_layout(struct str_field_descr* key_a_fields, int key_a_count,
+                     struct str_field_descr* key_b_fields, int key_b_count,
+                     struct str_field_descr* value_fields, int value_count,
+                     struct nested_field_descr* value_nested_fields, int val_nests_count);
+
+struct DoubleMap;
+
+void dmap_reset(struct DoubleMap* map, int capacity);
+
+#endif//_DOUBLE_MAP_STUB_CONTROL_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/containers/double-map-stub.c

@@ -1,0 +1,311 @@
+#include <string.h>
+#include <assert.h>
+#include "klee/klee.h"
+#include "lib/containers/double-map.h"
+#include "double-map-stub-control.h"
+
+#define prealloc_size (256)
+
+int key_a_size_g = 0;
+int key_b_size_g = 0;
+int value_size_g = 0;
+
+//uint8_t key_a[prealloc_size];
+//uint8_t key_b[prealloc_size]; //no need for separate key storage
+uint8_t value[prealloc_size];
+int allocation_succeeded;
+int has_this_key;
+int entry_claimed = 0;
+int allocated_index;
+struct DoubleMap* allocated_map_ptr;
+
+entry_condition* ent_cond = NULL;
+struct str_field_descr key_a_fields[prealloc_size],
+  key_b_fields[prealloc_size],
+  value_fields[prealloc_size];
+struct nested_field_descr value_nests[prealloc_size];
+int key_a_field_count = 0,
+  key_b_field_count = 0,
+  value_field_count = 0,
+  value_nests_count = 0;
+
+map_keys_equality* eq_a_g = 0;
+map_keys_equality* eq_b_g = 0;
+dmap_extract_keys* dexk_g = 0;
+dmap_pack_keys* dpk_g = 0;
+
+void dmap_set_entry_condition(entry_condition* c) {
+  klee_trace_param_fptr(c, "c");
+  ent_cond = c;
+}
+
+int calculate_str_size(struct str_field_descr* descr, int len) {
+  int rez = 0;
+  int sum = 0;
+  for (int i = 0; i < len; ++i) {
+    sum += descr[i].width;
+    if (descr[i].offset + descr[i].width > rez)
+      rez = descr[i].offset + descr[i].width;
+  }
+  klee_assert(rez == sum);
+  return rez;
+}
+
+void dmap_set_layout(struct str_field_descr* key_a_field_, int key_a_count_,
+                     struct str_field_descr* key_b_field_, int key_b_count_,
+                     struct str_field_descr* value_field_, int value_count_,
+                     struct nested_field_descr* value_nests_, int value_nests_count_) {
+  klee_assert(key_a_count_ < prealloc_size);
+  klee_assert(key_b_count_ < prealloc_size);
+  klee_assert(value_count_ < prealloc_size);
+  klee_assert(value_nests_count_ < prealloc_size);
[…]

### 2017-06-26 nf/lib/stubs/containers/map-stub-control.h

@@ -1,0 +1,20 @@
+#ifndef _MAP_STUB_CONTROL_H_INCLUDED_
+#define _MAP_STUB_CONTROL_H_INCLUDED_
+
+#include "lib/containers/map.h"
+#include "str-descr.h"
+
+void map_set_layout(struct Map* map,
+                    struct str_field_descr* key_fields,
+                    int key_fields_count,
+                    struct nested_field_descr* key_nests,
+                    int nested_key_fields_count,
+                    char* key_type);
+
+typedef int map_entry_condition(void* key, int index);
+
+void map_set_entry_condition(struct Map* map, map_entry_condition* cond);
+
+void map_reset(struct Map* map);
+
+#endif//_MAP_STUB_CONTROL_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/containers/map-stub.c

@@ -1,0 +1,176 @@
+#include <stdlib.h>
+#include <string.h>
+#include "klee/klee.h"
+#include "lib/containers/map.h"
+#include "map-stub-control.h"
+
+#define PREALLOC_SIZE (256)
+
+struct Map {
+  void* keyp;
+  int has_this_key;
+  int entry_claimed;
+  int allocated_index;
+  int capacity;
+
+  int has_layout;
+  int key_size;
+  int key_field_count;
+  int nested_key_field_count;
+  map_entry_condition* ent_cond;
+  struct str_field_descr key_fields[PREALLOC_SIZE];
+  struct nested_field_descr key_nests[PREALLOC_SIZE];
+  char* key_type;
+  map_keys_equality* keq;
+};
+
+int map_allocate(map_keys_equality* keq, map_key_hash* khash,
+                 int capacity,
+                 struct Map** map_out) {
+  klee_trace_ret();
+  klee_trace_param_fptr(keq, "keq");
+  klee_trace_param_fptr(khash, "khash");
+  klee_trace_param_i32(capacity, "capacity");
+  klee_trace_param_ptr(map_out, sizeof(struct Map*), "map_out");
+  int allocation_succeeded = klee_int("map_allocation_succeeded");
+  if (allocation_succeeded) {
+    *map_out = malloc(sizeof(struct Map));
+    klee_assume(*map_out != NULL);
+    klee_make_symbolic((*map_out), sizeof(struct Map), "map");
+    klee_assert((*map_out) != NULL);
+    (*map_out)->entry_claimed = 0;
+    (*map_out)->keq = keq;
+    (*map_out)->capacity = capacity;
+    (*map_out)->has_layout = 0;
+    (*map_out)->ent_cond = 0;
+    return 1;
+  }
+  return 0;
+}
+
+void map_reset(struct Map* map) {
+  //Do not trace. This function is an internal knob of the model.
+  map->keyp = NULL;
+  map->entry_claimed = 0;
+  map->has_this_key = klee_int("map_has_this_key");
+  map->allocated_index = klee_int("map_allocated_index");
+}
+
+static
+int calculate_str_size(struct str_field_descr* descr, int len) {
+  int rez = 0;
+  int sum = 0;
+  for (int i = 0; i < len; ++i) {
+    sum += descr[i].width;
+    if (descr[i].offset + descr[i].width > rez)
+      rez = descr[i].offset + descr[i].width;
+  }
+  klee_assert(rez == sum);
+  return rez;
+}
+
[…]

### 2017-06-26 nf/lib/stubs/containers/str-descr.h

@@ -1,0 +1,25 @@
+#ifndef _STR_DESCR_H_INCLUDED_
+#define _STR_DESCR_H_INCLUDED_
+
+struct str_field_descr {
+  int offset;
+  int width;
+  char* name;
+};
+
+struct nested_field_descr {
+  int base_offset;
+  int offset;
+  int width;
+  char *name;
+};
+
+struct nested_nested_field_descr {
+  int base_base_offset;
+  int base_offset;
+  int offset;
+  int width;
+  char *name;
+};
+
+#endif//_STR_DESCR_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/containers/vector-stub-control.h

@@ -1,0 +1,16 @@
+#ifndef _VECTOR_STUB_CONTROL_H_INCLUDED_
+#define _VECTOR_STUB_CONTROL_H_INCLUDED_
+
+#include "lib/containers/vector.h"
+#include "str-descr.h"
+
+void vector_set_layout(struct Vector* vector,
+                       struct str_field_descr* value_fields,
+                       int field_count,
+                       struct nested_field_descr* val_nest_fields,
+                       int nest_field_count,
+                       char* type_tag);
+
+void vector_reset(struct Vector* vector);
+
+#endif// _VECTOR_STUB_CONTROL_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/containers/vector-stub.c

@@ -1,0 +1,135 @@
+#include "klee/klee.h"
+#include <stdlib.h>
+#include <stdint.h>
+#include <string.h>
+#include "lib/containers/vector.h"
+#include "vector-stub-control.h"
+
+#define PREALLOC_SIZE (256)
+
+struct Vector {
+  uint8_t* data;
+  int elem_size;
+  int capacity;
+  int elem_claimed;
+  int index_claimed;
+  struct str_field_descr fields[PREALLOC_SIZE];
+  struct nested_field_descr nest_fields[PREALLOC_SIZE];
+  int field_count;
+  int nested_field_count;
+  char* cell_type;
+};
+
+int vector_allocate(int elem_size, int capacity,
+                    vector_init_elem* init_elem,
+                    struct Vector** vector_out) {
+  klee_trace_ret();
+  klee_trace_param_i32(elem_size, "elem_size");
+  klee_trace_param_i32(capacity, "capacity");
+  klee_trace_param_fptr(init_elem, "init_elem");
+  klee_trace_param_ptr(vector_out, sizeof(struct Vector*), "vector_out");
+
+  int allocation_succeeded = klee_int("vector_alloc_success");
+  if (! allocation_succeeded) return 0;
+
+  *vector_out = malloc(sizeof(struct Vector));
+  klee_assume(*vector_out != NULL);
+  klee_make_symbolic(*vector_out, sizeof(struct Vector), "vector");
+  (*vector_out)->data = malloc(elem_size);
+  klee_assume((*vector_out)->data != NULL);
+  klee_make_symbolic((*vector_out)->data, elem_size, "vector_data");
+  (*vector_out)->elem_size = elem_size;
+  (*vector_out)->capacity = capacity;
+  (*vector_out)->elem_claimed = 0;
+  (*vector_out)->field_count = 0;
+  (*vector_out)->nested_field_count = 0;
+  klee_forbid_access((*vector_out)->data, elem_size, "private state");
+  return 1;
+}
+
+void vector_reset(struct Vector* vector) {
+  //Do not trace. This function is an internal knob of the model.
+  //TODO: reallocate vector->data to avoid having the same pointer?
+  klee_allow_access(vector->data, vector->elem_size);
+  klee_make_symbolic(vector->data, vector->elem_size, "vector_data_reset");
+  vector->elem_claimed = 0;
+  klee_forbid_access(vector->data, vector->elem_size, "private state");
+}
+
+void vector_set_layout(struct Vector* vector,
[…]

### 2017-06-26 nf/lib/stubs/expirator-stub.c

@@ -1,0 +1,39 @@
+#include <klee/klee.h>
+#include "lib/expirator.h"
+#include "containers/double-chain-stub-control.h"
+
+int expire_items(struct DoubleChain* chain, struct DoubleMap* map,
+                 uint32_t time) {
+  klee_trace_ret();
+  klee_trace_param_i32((uint32_t)chain, "chain");
+  klee_trace_param_i32((uint32_t)map, "map");
+  klee_trace_param_i32(time, "exp_time");
+  int nfreed = klee_int("number_of_freed_flows");
+  klee_assume(0 <= nfreed);
+  if (nfreed > 0) {
+    dchain_make_space(chain);
+  }
+  //Tell dchain model that we freed some indexes here
+  return nfreed;
+}
+
+int expire_items_single_map(struct DoubleChain* chain,
+                            struct Vector* vector,
+                            struct Map* map,
+                            entry_extract_key* eek,
+                            entry_pack_key* epk,
+                            uint32_t time) {
+  klee_trace_ret();
+  klee_trace_param_i32((uint32_t)chain, "chain");
+  klee_trace_param_i32((uint32_t)vector, "vector");
+  klee_trace_param_i32((uint32_t)map, "map");
+  klee_trace_param_fptr(eek, "eek");
+  klee_trace_param_fptr(epk, "epk");
+  klee_trace_param_i32(time, "time");
+  int nfreed = klee_int("unmber_of_freed_flows");
+  klee_assume(0 <= nfreed);
+  if (0 < nfreed) {
+    dchain_make_space(chain);
+  }
+  return nfreed;
+}

### 2017-06-26 nf/lib/stubs/my-time-stub-control.h

@@ -1,0 +1,16 @@
+#ifndef _MY_TIME_STUB_CONTROL_H_INCLUDED_
+#define _MY_TIME_STUB_CONTROL_H_INCLUDED_
+
+#include <stdint.h>
+
+//@ #include "lib/predicates.gh"
+
+uint32_t start_time(void);
+//@ requires true;
+//@ ensures last_time(result);
+
+uint32_t restart_time(void);
+
+uint32_t get_start_time(void);
+
+#endif //_MY_TIME_STUB_CONTROL_H_INCLUDED_

### 2017-06-26 nf/lib/stubs/my-time-stub.c

@@ -1,0 +1,32 @@
+#include <klee/klee.h>
+#include "lib/nf_time.h"
+#include "my-time-stub-control.h"
+
+uint32_t starting_time = 0;
+uint32_t last_time = 0;
+
+uint32_t start_time(void) {
+    klee_trace_ret();
+    starting_time = klee_int("starting_time");
+    last_time = starting_time;
+    return last_time;
+}
+
+uint32_t restart_time(void) {
+  starting_time = klee_int("restarting_time");
+  last_time = starting_time;
+  return last_time;
+}
+
+uint32_t current_time(void) {
+    klee_trace_ret();
+    uint32_t next_time = klee_int("next_time");
+    klee_assume(last_time <= next_time);
+    last_time = next_time;
+    return next_time;
+}
+
+uint32_t get_start_time_internal(void) {
+    return starting_time;
+}
+uint32_t get_start_time(void) {return get_start_time_internal();}

### 2017-06-26 nf/lib/stubs/rte_stubs.c

@@ -1,0 +1,395 @@
+#include <klee/klee.h>
+#include <string.h>
+#include "rte_stubs.h"
+#include "../flow.h"
+
+int rte_errno;
+
+uint64_t rte_get_tsc_hz(){
+    return 0;
+}
+
+unsigned rte_lcore_id(){
+    return 0;
+}
+
+uint64_t rte_rdtsc(){
+  //TODO: this prevents the simulation of sending leftover packets, (or all?)
+  return 0;
+}
+
+struct str_field_descr mbuf_descrs[] = {
+  //Do not forget about "buf_addr" -- it is a pointer that is why it is not listed here.
+  {offsetof(struct rte_mbuf, buf_physaddr), sizeof(uint64_t), "buf_physaddr"},
+  {offsetof(struct rte_mbuf, buf_len), sizeof(uint16_t), "buf_len"},
+  {offsetof(struct rte_mbuf, data_off), sizeof(uint16_t), "data_off"},
+  {offsetof(struct rte_mbuf, refcnt), sizeof(uint16_t), "refcnt"},
+  {offsetof(struct rte_mbuf, nb_segs), sizeof(uint8_t), "nb_segs"},
+  {offsetof(struct rte_mbuf, port), sizeof(uint8_t), "port"},
+  {offsetof(struct rte_mbuf, ol_flags), sizeof(uint64_t), "ol_flags"},
+  {offsetof(struct rte_mbuf, packet_type), sizeof(uint32_t), "packet_type"},
+  {offsetof(struct rte_mbuf, pkt_len), sizeof(uint32_t), "pkt_len"},
+  {offsetof(struct rte_mbuf, data_len), sizeof(uint16_t), "data_len"},
+  {offsetof(struct rte_mbuf, vlan_tci), sizeof(uint16_t), "vlan_tci"},
+  {offsetof(struct rte_mbuf, hash), sizeof(uint32_t), "hash"},
+  {offsetof(struct rte_mbuf, seqn), sizeof(uint32_t), "seqn"},
+  {offsetof(struct rte_mbuf, vlan_tci_outer), sizeof(uint16_t), "vlan_tci_outer"},
+  {offsetof(struct rte_mbuf, udata64), sizeof(uint64_t), "udata64"},
+  {offsetof(struct rte_mbuf, pool), sizeof(void*), "pool"},
+  {offsetof(struct rte_mbuf, next), sizeof(struct rte_mbuf*), "next"},
[…]

### 2017-06-26 nf/mininet/mynat.py

@@ -1,0 +1,103 @@
+#!/usr/bin/python
+
+"""
+mynat.py: network with two nats in operation:
+nat1 is the nat being tested,
+nat2 NetFilter standard reference nat
+
+
+           h0
+           |
+           s0
+           |
+    ----------------
+    |              |
+   nat1           nat2
+    |              |
+   s1              s2
+    |              |
+   h1              h2
+
+"""
+
+from mininet.topo import Topo
+from mininet.net import Mininet
+from mininet.nodelib import NAT
+from mininet.log import setLogLevel
+from mininet.cli import CLI
+from mininet.util import irange
+
+class InternetTopo(Topo):
+    "Single switch connected to n hosts."
+    def __init__(self, n=2, **opts):
+        Topo.__init__(self, **opts)
+
+        # set up inet switch
+        inetSwitch = self.addSwitch('s0')
+        # add inet host
+        inetHost = self.addHost('h0')
+        self.addLink(inetSwitch, inetHost)
+
+        # add local nets
+        inetIntf = 'nat1-eth0'
+        localIntf = 'nat1-eth1'
+        localIP = '192.168.1.1'
+        localSubnet = '192.168.1.0/24'
+        natParams = { 'ip' : '%s/24' % localIP }
+        # add NAT to topology
+        nat = self.addNode('nat1')
+        switch = self.addSwitch('s1')
+        # connect NAT to inet and local switches
+        self.addLink(nat, inetSwitch, intfName1=inetIntf)
+        self.addLink(nat, switch, intfName1=localIntf, params1=natParams)
+        # add host and connect to local switch
+        host = self.addHost('h1',
+                            ip='192.168.1.100/24',
+                            defaultRoute='via %s' % localIP)
+        self.addLink(host, switch)
+
+        inetIntf = 'nat2-eth0'
+        localIntf = 'nat2-eth1'
+        localIP = '192.168.2.1'
+        localSubnet = '192.168.2.0/24'
+        natParams = { 'ip' : '%s/24' % localIP }
+        # add NAT to topology
+        nat = self.addNode('nat2', cls=NAT, subnet=localSubnet,
+                           inetIntf=inetIntf, localIntf=localIntf)
+        switch = self.addSwitch('s2')
+        # connect NAT to inet and local switches
[…]

### 2017-06-26 nf/mininet/test.sh

@@ -1,0 +1,42 @@
+
+docker build -t vnds/nat .
+docker build -t vnds/client client-docker-img/
+
+docker network create --driver=bridge --subnet=192.168.10.0/24 vnds_int
+docker network create --driver=bridge --subnet=192.168.11.0/24 vnds_ext
+
+docker run -itd --name=vnds_c1 --net=vnds_int vnds/client /bin/bash
+docker run -itd --name=vnds_c2 --net=vnds_int vnds/client /bin/bash
+docker run -itd --name=vnds_s1 --net=vnds_ext vnds/client /bin/bash
+
+# Unfortunately, docker does not have a separate capability to mount a hugetlbfs
+docker run -itd --name=vnds_nat --net=vnds_int --privileged vnds/nat /bin/bash
+docker network connect vnds_ext vnds_nat
+
+S1_IP=`docker inspect -f '{{.NetworkSettings.Networks.vnds_ext.IPAddress}}' vnds_s1`
+NAT_INT_IP=`docker inspect -f '{{.NetworkSettings.Networks.vnds_int.IPAddress}}' vnds_nat`
+NAT_EXT_IP=`docker inspect -f '{{.NetworkSettings.Networks.vnds_ext.IPAddress}}' vnds_nat`
+NAT_INT_GATEWAY_IP=`docker inspect -f '{{.NetworkSettings.Networks.vnds_int.Gateway}}' vnds_nat`
+NAT_EXT_GATEWAY_IP=`docker inspect -f '{{.NetworkSettings.Networks.vnds_ext.Gateway}}' vnds_nat`
+
+docker exec vnds_nat ping $NAT_INT_GATEWAY_IP -c 1
+docker exec vnds_nat ping $NAT_EXT_GATEWAY_IP -c 1
+
+NAT_INT_GATEWAY_MAC=`docker exec vnds_nat sh -c "arp -n $NAT_INT_GATEWAY_IP | awk 'NR==2{print \\\$3;}'"` 
+NAT_EXT_GATEWAY_MAC=`docker exec vnds_nat sh -c "arp -n $NAT_EXT_GATEWAY_IP | awk 'NR==2{print \\\$3;}'"` 
+
+docker exec --privileged vnds_c1 ip route del default
+docker exec --privileged vnds_c1 ip route add default via $NAT_INT_IP dev eth0
+docker exec --privileged vnds_c2 ip route del default
[…]

### 2017-06-26 nf/nf_main.c

@@ -1,0 +1,260 @@
+#include <inttypes.h>
+
+// DPDK uses these but doesn't include them. :|
+#include <linux/limits.h>
+#include <sys/types.h>
+
+#ifdef KLEE_VERIFICATION
+#  include <klee/klee.h>
+#  include "lib/stubs/rte_stubs.h"
+#  include "lib/stubs/my-time-stub-control.h"
+#else//KLEE_VERIFICATION
+#  include <rte_common.h>
+#  include <rte_eal.h>
+#  include <rte_ethdev.h>
+#  include <rte_launch.h>
+#  include <rte_lcore.h>
+#  include <rte_mbuf.h>
+#  include "rte_errno.h"
+#endif//KLEE_VERIFICATION
+
+#include "lib/nf_forward.h"
+#include "lib/nf_log.h"
+#include "lib/nf_time.h"
+#include "lib/nf_util.h"
+#include <string.h>
+
+
+// --- Static config ---
+// TODO see remark in lcore_main
+// Size of batches to receive; trade-off between latency and throughput
+// Can be overriden at compile time
+//#ifndef BATCH_SIZE
+//static const uint16_t BATCH_SIZE = 32;
+//#endif
+
+// Queue sizes for receiving/transmitting packets (set to their values from l3fwd sample)
+static const uint16_t RX_QUEUE_SIZE = 128;
+static const uint16_t TX_QUEUE_SIZE = 512;
+
+// Memory pool #buffers and per-core cache size (set to their values from l3fwd sample)
+static const unsigned MEMPOOL_BUFFER_COUNT = 8192;
+static const unsigned MEMPOOL_CACHE_SIZE = 256;
+static const unsigned MEMPOOL_CLONE_COUNT = 256;
+
+static struct rte_mempool* clone_pool;
+
+// --- Initialization ---
+static int
+nf_init_device(uint8_t device, struct rte_mempool *mbuf_pool)
+{
+  int retval;
+
+  // Configure the device
+  // This is ugly code; DPDK samples use designated initializers,
+  // but those are not available in C++, and this code needs to compile
+  // both as C and C++.
+  struct rte_eth_conf device_conf;
+  memset(&device_conf, 0, sizeof(struct rte_eth_conf));
+  device_conf.rxmode.mq_mode = ETH_MQ_RX_RSS;
+  device_conf.rxmode.max_rx_pkt_len = ETHER_MAX_LEN;
[…]

### 2017-06-26 nf/testbed/analysis/plot.r

@@ -1,0 +1,35 @@
+library(plyr)
+library(ggplot2)
+
+load_exp_data <- function(sources) {
+  exp_data <- data.frame(V1=integer(), V2=integer(), medium=character())
+  for (s in sources) {
+    exp_data_i <- read.table(s[1])
+    exp_data_i["medium"] <- s[2]
+    exp_data <- rbind(exp_data, exp_data_i)
+  }
+  data.frame(abscissa=exp_data$V1, ordinata=exp_data$V2, medium=exp_data$medium)
+}
+
+plot_exp_data <- function(exp_data, titles) {
+  pd <- position_dodge(0.01)
+  summary=ddply(exp_data, c("abscissa", "medium"), summarise,
+                N=length(ordinata), mean=mean(ordinata), sd=sd(ordinata))
+  csummary=ddply(exp_data, c("medium", "abscissa", "ordinata"), nrow)
+  ggplot(summary, aes(x=abscissa, y=mean,
+                      colour=medium, group=medium)) +
+      geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
+                    width=.03, position=pd, colour="black") +
+      geom_line(position=pd) +
+      geom_point(position=pd,size=3) +
+      scale_x_continuous(trans='log10') +
+      labs(title=titles$title) +
+      xlab(titles$abscissa) +
+      ylab(titles$ordinata) +
+      geom_point(data=csummary,
+                aes(x=abscissa, y=ordinata, colour=medium,
+                    group=medium, size=V1),
+                position=pd,
+                show.legend=FALSE) +
+      scale_size(range=c(0,2),trans="log",labels=NULL)
+}

### 2017-06-26 nf/testbed/hard/1g.cfg.sh

@@ -1,0 +1,60 @@
+# Configuration file mapping real hosts, addresses and ports
+# to abstract names used in the scripts.
+
+KERN_NIC_DRIVER=igb
+DPDK_NIC_DRIVER=igb_uio
+
+# Subnets
+
+EXTERNAL_SUBNET=192.168.2.0
+SERVER_SUBNET=192.168.200.0
+
+# Middlebox
+
+MB_HOST=icnalsp3s2.epfl.ch
+
+MB_MAC_INTERNAL=00:1e:67:92:29:6d
+MB_MAC_EXTERNAL=00:1e:67:92:29:6c
+MB_MAC_TO_SRV=00:1e:67:92:29:6e
+
+MB_IP_INTERNAL=192.168.3.2
+MB_IP_EXTERNAL=192.168.2.2
+MB_IP_TO_SRV=192.168.200.2
+
+MB_DEVICE_INTERNAL=em3
+MB_DEVICE_EXTERNAL=em2
+MB_DEVICE_TO_SRV=em4
+
+MB_PCI_INTERNAL="0000:02:00.2"
+MB_PCI_EXTERNAL="0000:02:00.1"
+MB_PCI_TO_SRV="0000:02:00.3"
+
+# Tester
+
+TESTER_HOST=icnalsp3s1.epfl.ch
+
+TESTER_DEVICE_INTERNAL=em3
+TESTER_DEVICE_EXTERNAL=em2
+
+TESTER_PCI_INTERNAL=0000:02:00.2
+TESTER_PCI_EXTERNAL=0000:02:00.1
+
+TESTER_MAC_INTERNAL=00:1e:67:92:2a:bd
+TESTER_MAC_EXTERNAL=00:1e:67:92:2a:bc
+
+TESTER_IP_INTERNAL=192.168.3.5
+TESTER_IP_EXTERNAL=192.168.2.10
+TESTER_IP_FOR_STUB=192.168.200.5
+
+# Server
+
+SERVER_HOST=icnalsp3s3.epfl.ch
+
+SERVER_IP=192.168.200.10
+SERVER_MAC=00:1e:67:92:2a:2b
+SERVER_DEVICE=em4
+
+# Other
+
+export RTE_SDK=$HOME/dpdk
+export RTE_TARGET=x86_64-native-linuxapp-gcc

### 2017-06-26 nf/testbed/hard/bench-all.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+# Initialize the machines, i.e. software+scripts
+. ./init-machines.sh
+# Clean first, just in case
+. ./clean.sh
+
+VNDS_PREFIX="$HOME/vnds/nf"
+
+NOW=$(date +"%d.%m.%Y_%H_%M")
+
+MIDDLEBOXES=("$VNDS_PREFIX/vignat" "$VNDS_PREFIX/unverified-nat" "$VNDS_PREFIX/nop")
+SCENARIOS=("mg-new-flows-latency" "mg-existing-flows-latency" "mg-1p")
+
+mkdir -p $NOW
+
+for MIDDLEBOX in ${MIDDLEBOXES[@]}; do
+    # The second parameter should not matter (it doesn't for mg-* scenarios,
+    # but unfortunately it does for the old ones, so to not try this for those)
+    . ./init.sh $MIDDLEBOX "mg-1p"
+    for SCENARIO in ${SCENARIOS[@]}; do
+        . ./start-middlebox.sh $MIDDLEBOX $SCENARIO
+        CLEAN_APP_NAME=`echo "$MIDDLEBOX" | tr '/' '_'`
+        RESULTS_FILE="$NOW/$CLEAN_APP_NAME-$SCENARIO.results"
+        . ./run.sh $MIDDLEBOX $SCENARIO $RESULTS_FILE
+        . ./stop-middlebox.sh $MIDDLEBOX $SCENARIO
+    done
+done
+
+. ./clean.sh

### 2017-06-26 nf/testbed/hard/bench/run-dpdk.sh

@@ -1,0 +1,43 @@
+#!/bin/bash
+. ./config.sh
+
+# Runs a DPDK app for the specified scenario
+
+# Parameters:
+# $1: Scenario, can be "loopback" or "rr"
+# $2: Folder of the app; the app must be named "nat"
+#     and take the usual arguments e.g. "--extip"
+# $3: Additional arguments for the app
+
+if [ $1 != "loopback" -a $1 != "rr" ]; then
+    echo "[bench] Unknown scenario specified in run-dpdk" 1>&2
+    exit 1
+fi
+
+if [ -z $2 ]; then
+    echo "[bench] No app folder specified in run-dpdk" 1>&2
+    exit 2
+fi
+
+
+pushd $2 >> /dev/null
+
+echo "[bench] Building $2..."
+sudo rm build -rf
+make clean
+make
+
+echo "[bench] Running $2..."
+if [ $1 = "loopback" ]; then
+    sudo ./build/nat -c 0x01 -n 2 -- -p 0x3 --wan 0 --lan-dev 1 \
+        $3 \
+        --extip $MB_IP_EXTERNAL \
+        --eth-dest 0,$TESTER_MAC_EXTERNAL --eth-dest 1,$TESTER_MAC_INTERNAL
+else
+    sudo ./build/nat -c 0x01 -n 2 -- -p 0x7 --wan 2 --lan-dev 1 \
+        $3 \
+        --extip $MB_IP_TO_SRV \
+        --eth-dest 0,$TESTER_MAC_EXTERNAL --eth-dest 1,$TESTER_MAC_INTERNAL --eth-dest 2,$SERVER_MAC
+fi
+
+popd >> /dev/null

### 2017-06-26 nf/testbed/hard/clean/tester.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+. ~/scripts/config.sh
+
+echo "[clean] Unbinding tester interfaces from Linux..."
+sudo ifconfig $TESTER_DEVICE_INTERNAL down
+sudo ifconfig $TESTER_DEVICE_EXTERNAL down
+
+echo "[clean] Unbinding tester interfaces from DPDK..."
+sudo $RTE_SDK/tools/dpdk-devbind.py -b $KERN_NIC_DRIVER $TESTER_PCI_INTERNAL $TESTER_PCI_EXTERNAL
+
+echo "[clean] Killing pktgen on tester..."
+sudo pkill -9 pktgen
+
+echo "[clean] Killing moongen on tester..."
+sudo pkill -9 MoonGen
+sudo pkill -9 moongen
+sudo pkill -9 moon-gen

### 2017-06-26 nf/testbed/hard/init-machines.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+. ./config.sh
+
+echo "[init] Cloning scripts..."
+for h in $TESTER_HOST; do # $SERVER_HOST; do
+    rsync -q -t -r --exclude '*.log' --exclude '*.results' ./ $h:scripts
+done
+
+echo "[init] Initializing all machines..."
+ssh $TESTER_HOST 'bash ~/scripts/init-machines/tester.sh'
+#ssh $SERVER_HOST 'bash ~/scripts/init-machines/server.sh'
+. ./init-machines/middlebox.sh

### 2017-06-26 nf/testbed/hard/init-machines/tester.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+echo "[init] Initializing tester..."
+
+sudo apt-get -qq update
+
+sudo apt-get install -yqq \
+    tcpdump hping3 python-scapy git \
+    libpcap-dev libglib2.0-dev \
+    daemon iperf3 netperf liblua5.2-dev \
+    make binutils gcc \
+    bc cmake
+
+. ~/scripts/init-machines/install-dpdk.sh
+
+. ~/scripts/init-machines/install-pktgen.sh
+
+. ~/scripts/init-machines/install-moongen.sh

### 2017-06-26 nf/testbed/hard/init-network-loopback/netfilter/middlebox.sh

@@ -1,0 +1,28 @@
+#!/bin/bash
+. ./config.sh
+. ./util/functions.sh
+
+echo "[init] Configuring middlebox IPs..."
+sudo ifconfig $MB_DEVICE_EXTERNAL up
+sudo ip addr flush dev $MB_DEVICE_EXTERNAL
+sudo ip addr add $MB_IP_EXTERNAL/24 dev $MB_DEVICE_EXTERNAL
+sudo ifconfig $MB_DEVICE_INTERNAL up
+sudo ip addr flush dev $MB_DEVICE_INTERNAL
+sudo ip addr add $MB_IP_INTERNAL/24 dev $MB_DEVICE_INTERNAL
+sudo ifconfig $MB_DEVICE_TO_SRV down
+
+echo "[init] Configuring middlebox forwarding rules..."
+sudo sysctl -w net.ipv4.ip_forward=1
+sudo_append /etc/sysctl.conf "net.ipv4.ip_forward=1"
+
+sudo iptables -F FORWARD
+sudo iptables -t nat -F POSTROUTING
+sudo iptables -t nat -A POSTROUTING -o $MB_DEVICE_EXTERNAL -j MASQUERADE
+sudo iptables -A FORWARD -i $MB_DEVICE_INTERNAL -o $MB_DEVICE_EXTERNAL -m state --state NEW,RELATED,ESTABLISHED -j ACCEPT
+sudo iptables -A FORWARD -i $MB_DEVICE_INTERNAL -o $MB_DEVICE_EXTERNAL -j ACCEPT
+
+sudo arp -s $TESTER_IP_EXTERNAL $TESTER_MAC_EXTERNAL
+sudo arp -s $TESTER_IP_INTERNAL $TESTER_MAC_INTERNAL
+
+echo "[init] Unlocking software restrictions on middlebox NetFilter..."
+. ./util/netfilter-unlock.sh $MB_DEVICE_EXTERNAL

### 2017-06-26 nf/testbed/hard/init-network-rr/netfilter/middlebox.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+. ./config.sh
+. ./util/functions.sh
+
+echo "[init] Configuring middlebox IPs..."
+sudo ifconfig $MB_DEVICE_TO_SRV up
+sudo ip addr flush dev $MB_DEVICE_TO_SRV
+sudo ip addr add $MB_IP_TO_SRV/24 dev $MB_DEVICE_TO_SRV
+sudo ifconfig $MB_DEVICE_INTERNAL up
+sudo ip addr flush dev $MB_DEVICE_INTERNAL
+sudo ip addr add $MB_IP_INTERNAL/24 dev $MB_DEVICE_INTERNAL
+sudo ifconfig $MB_DEVICE_EXTERNAL up
+sudo ip addr flush dev $MB_DEVICE_EXTERNAL
+sudo ip addr add $MB_IP_EXTERNAL/24 dev $MB_DEVICE_EXTERNAL
+
+echo "[init] Configuring middlebox forwarding rules..."
+sudo sysctl -w net.ipv4.ip_forward=1
+sudo_append /etc/sysctl.conf "net.ipv4.ip_forward=1"
+
+sudo iptables -F FORWARD
+sudo iptables -t nat -F POSTROUTING
+sudo iptables -t nat -A POSTROUTING -o $MB_DEVICE_TO_SRV -j MASQUERADE
+sudo iptables -A FORWARD -i $MB_DEVICE_INTERNAL -o $MB_DEVICE_TO_SRV -m state --state NEW,RELATED,ESTABLISHED -j ACCEPT
+sudo iptables -A FORWARD -i $MB_DEVICE_INTERNAL -o $MB_DEVICE_TO_SRV -j ACCEPT
+sudo iptables -A FORWARD -i $MB_DEVICE_EXTERNAL -o $MB_DEVICE_TO_SRV -m state --state NEW,RELATED,ESTABLISHED -j ACCEPT
+sudo iptables -A FORWARD -i $MB_DEVICE_EXTERNAL -o $MB_DEVICE_TO_SRV -j ACCEPT
+
+sudo arp -s $SERVER_IP $SERVER_MAC
+sudo arp -s $TESTER_IP_INTERNAL $TESTER_MAC_INTERNAL
+sudo arp -s $TESTER_IP_EXTERNAL $TESTER_MAC_EXTERNAL
+
[…]

### 2017-06-26 nf/testbed/hard/init.sh

@@ -1,0 +1,72 @@
+#!/bin/bash
+. ./config.sh
+
+# Master script to initialize VigNAT-related programs benchmarks.
+# Can work with different implementations, including non-NATs,
+# using different scenarios.
+
+# Parameters:
+# $1: The app, either a known name or a DPDK NAT-like app.
+#     Known names: "netfilter".
+#     Otherwise, a folder name containing a DPDK NAT-like app, e.g. "~/vnds/nat"
+# $2: The scenario, one of the following:
+#     "mg-1p": Measure throughput: find the rate at which the middlebox
+#              starts loosing 1% of packets.
+#     "mg-existing-flows-latency": Measure the forwarding latency for existing
+#                                  flows.
+#     "mg-new-flows-latency": Measure the forwarding latency for new flows.
+#     "loopback": Measure throughput.
+#                 Tester and middlebox are connected together in a loop,
+#                 using 2 interfaces on each, in different subnets; server is ignored.
+#     "1p": Measure throughput.
+#           Find the point at which the middlebox starts dropping 1% of packets.
+#     "passthrough": Measure latency.
+#                    Tester sends packets to server, which pass through the middlebox;
+#                    all machines are in the same subnet.
+#     "rr": Measure latency.
+#           Tester sends packets to server, which are modified by the middlebox;
+#           there are two subnets, tester-middlebox and middlebox-server.
+#           a.k.a. request/response
+
+MIDDLEBOX=$1
+SCENARIO=$2
+
+if [ -z $MIDDLEBOX ]; then
+    echo "[bench] No app specified" 1>&2
+    exit 1
+fi
+
+if [ -z $SCENARIO ]; then
+    echo "[bench] No scenario specified" 1>&2
+    exit 2
+fi
+
+
+# Initialize the network;
+# to do that, we need to know which scenario we use, and whether we'll run a DPDK app or not.
+NETWORK_SCENARIO=$SCENARIO
+case $SCENARIO in
+    "1p"|"loopback"|"mg-1p"|"mg-existing-flows-latency"|"mg-new-flows-latency")
+        NETWORK_SCENARIO="loopback"
+        ;;
+    "passthrough")
+        NETWORK_SCENARIO="passthrough"
+        ;;
+    "rr")
+        NETWORK_SCENARIO="rr"
[…]

### 2017-06-26 nf/testbed/hard/pktgen/run-in-background.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+# Launch PktGen as a server listening 22022 in the background
+# This script contains a piece of wisdom copy&pasted from StackOverflow.
+# It is supposed to properly disconnect stdin and stdout using bahs pipe magic.
+
+cd ~/pktgen && fifo=$(mktemp -u) && mkfifo "$fifo" && (rm "$fifo" && { sudo ./app/app/x86_64-native-linuxapp-gcc/pktgen -c 0x1f -n 3 -- -P -m "[1-2:3-4].0" -f $HOME/scripts/pktgen/provision-rr.lua -G < /dev/fd/1 3&>1 > ./pktgen-log.txt <&3 3<&-; } &) 3<> "$fifo" | :

### 2017-06-26 nf/testbed/hard/pktgen/run.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+# Parameter:
+# $1: The script to run pktgen with
+
+if [ -z $1 ]; then
+    echo "[pktgen] No script given to start pktgen with"
+    exit 1
+fi
+
+cd ~/pktgen
+
+sudo ./app/app//x86_64-native-linuxapp-gcc/pktgen -c 0x1ff -n 3 -- -P -m "[1-2:3-4].0, [5-6:7-8].1" -f $1

### 2017-06-26 nf/testbed/hard/pktgen/send-command.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+# Send a command to a pktgen server running on the port 22022
+# The first argument ($1) is the command to be sent.
+
+socat - TCP4:localhost:22022 < $1

### 2017-06-26 nf/testbed/hard/start-middlebox.sh

@@ -1,0 +1,95 @@
+#!/bin/bash
+. ./config.sh
+
+# Master script to initialize VigNAT-related programs benchmarks.
+# Can work with different implementations, including non-NATs,
+# using different scenarios.
+
+# Parameters:
+# $1: The app, either a known name or a DPDK NAT-like app.
+#     Known names: "netfilter".
+#     Otherwise, a folder name containing a DPDK NAT-like app, e.g. "~/vnds/nat"
+# $2: The scenario, one of the following:
+#     "mg-1p": Measure throughput: find the rate at which the middlebox
+#              starts loosing 1% of packets.
+#     "mg-existing-flows-latency": Measure the forwarding latency for existing
+#                                  flows.
+#     "mg-new-flows-latency": Measure the forwarding latency for new flows.
+#     "loopback": Measure throughput.
+#                 Tester and middlebox are connected together in a loop,
+#                 using 2 interfaces on each, in different subnets; server is ignored.
+#     "1p": Measure throughput.
+#           Find the point at which the middlebox starts dropping 1% of packets.
+#     "passthrough": Measure latency.
+#                    Tester sends packets to server, which pass through the middlebox;
+#                    all machines are in the same subnet.
+#     "rr": Measure latency.
+#           Tester sends packets to server, which are modified by the middlebox;
+#           there are two subnets, tester-middlebox and middlebox-server.
+#           a.k.a. request/response
+
+MIDDLEBOX=$1
+SCENARIO=$2
+
+if [ -z $MIDDLEBOX ]; then
+    echo "[bench] No app specified" 1>&2
+    exit 1
+fi
+
+if [ -z $SCENARIO ]; then
+    echo "[bench] No scenario specified" 1>&2
+    exit 2
+fi
+
+CLEAN_APP_NAME=`echo "$MIDDLEBOX" | tr '/' '_'`
+LOG_FILE="bench-$CLEAN_APP_NAME-$SCENARIO-init.log"
+
+if [ -f "$LOG_FILE" ]; then
+    rm "$LOG_FILE"
+fi
+
+
+if [ "$MIDDLEBOX" = "netfilter" ]; then
+    case $SCENARIO in
+	"mg-new-flows-latency")
+	    EXPIRATION_TIME=1
+	    ;;
[…]

### 2017-06-26 nf/testbed/hard/util/netfilter-short-timeout.sh

@@ -1,0 +1,22 @@
+
+. ./util/functions.sh 
+
+TIMEOUT=$1
+
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_frag6_timeout $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_generic_timeout $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_icmp_timeout $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_icmpv6_timeout $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_max_retrans 0
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_close $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_close_wait $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_established $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_fin_wait $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_last_ack $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_max_retrans $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_syn_recv $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_syn_sent $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_time_wait $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_tcp_timeout_unacknowledged $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_udp_timeout $TIMEOUT
+sudo_overwrite /proc/sys/net/netfilter/nf_conntrack_udp_timeout_stream2 $TIMEOUT

### 2017-06-26 nf/testbed/hard/util/plot/llat.r

@@ -1,0 +1,61 @@
+library(plyr)
+library(ggplot2)
+library(scales)
+library(grid)
+
+all_data <- data.frame(V1=integer(), V2=integer(),
+                       V3=integer(), V4=integer(),
+                       V5=integer(), V6=integer(),
+                       middlebox=character())
+
+vignat <- read.table("vignat-latency-breakpoint.txt")
+vignat["middlebox"] <- "vignat"
+unverified <- read.table("unverified-latency-breakpoint.txt")
+unverified["middlebox"] <- "unverified"
+
+all_data <- rbind(all_data, vignat)
+all_data <- rbind(all_data, unverified)
+
+cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+pd <- position_dodge(1)
+
+p1 <- ggplot(all_data, aes(x=V3/1e6, y=V5/1e3,
+                          group=middlebox,
+                          color=middlebox,
+                          shape=middlebox)) +
+     geom_point(size=3) +
+     geom_line() +
+     geom_errorbar(aes(ymin=(V5-V6)/1e3, ymax=(V5+V6)/1e3), width=.01) +
+     labs(title="1 way latency, single flow") +
+     xlab("rate (Mpkt/s)") +
+     ylab(bquote("1-way latency, microsec, log scale")) +
+     theme_bw() +
+     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
+                   labels = trans_format("log10", math_format(10^.x))) +
+     #expand_limits(x=0,y=0) +
+     theme( plot.margin = unit( c(0,0,0,0) , "in" ) )
+
+p2 <- ggplot(all_data, aes(x=V3/1e6, y=V4,
+                          group=middlebox,
+                          color=middlebox,
+                          shape=middlebox)) +
+     geom_point(size=3) +
+     geom_line() +
+     xlab("rate (Mpkt/s)") +
+     ylab(bquote("packet loss, fraction, log scale")) +
+     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
+                   labels = trans_format("log10", math_format(10^.x))) +
+     theme_bw() +
+     #expand_limits(x=0,y=0) +
+     theme( plot.margin = unit( c(0,0,0,0) , "in" ) )
+
+
[…]

### 2017-06-26 nf/testbed/hard/util/plot/new-flow-lat.r

@@ -1,0 +1,49 @@
+library(plyr)
+library(ggplot2)
+library(scales)
+
+all_data <- data.frame(V1=integer(), V2=integer(),
+                       V3=integer(), V4=integer(),
+                       middlebox=character())
+
+netfilter <- read.table("netfilter-new-flows-latency.txt")
+netfilter["middlebox"] <- "NetFilter"
+vignat <- read.table("vignat-new-flows-latency.txt")
+vignat["middlebox"] <- "VigNAT"
+vignat2 <- read.table("vignat-twice-new-flows-latency.txt")
+vignat2["middlebox"] <- "VigNAT (2x bigger table)"
+unverified <- read.table("unverified-new-flows-latency.txt")
+unverified["middlebox"] <- "DPDK-unverified"
+counted <- read.table("counted-chains-new-flows-latency.txt")
+counted["middlebox"] <- "Counted-chains map"
+
+all_data <- rbind(all_data, netfilter)
+#all_data <- rbind(all_data, vignat)
+#all_data <- rbind(all_data, vignat2)
+all_data <- rbind(all_data, unverified)
+all_data <- rbind(all_data, counted)
+
+cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+pd <- position_dodge(2)
+
+p <- ggplot(all_data, aes(x=V1/1e3, y=V3/1e3,
+                          group=middlebox,
+                          color=middlebox,
+                          shape=middlebox)) +
+     geom_point(size=3,position=pd) +
+     geom_line() +
+     geom_errorbar(aes(ymin=(V3-V4)/1e3, ymax=(V3+V4)/1e3), width=.01,
+                   position=pd) +
+     labs(title="1 way latency for ~8Kpkt/s. 1000 sample new flows") +
+     xlab("# concurrent flows (K)") +
+     ylab(bquote("1-way latency, " * mu * "s")) +
+     theme_bw() +
+     expand_limits(x=0,y=1) +
+     #scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
+     #              labels = trans_format("log10", math_format(10^.x))) +
+     coord_cartesian(ylim=c(0,25)) +
+     theme( plot.margin = unit( c(0,0,0,0) , "in" ) )
+
+ggsave(filename="new-flow-latency.png")
+print(p)

### 2017-06-26 nf/testbed/hard/util/plot/thru.r

@@ -1,0 +1,40 @@
+library(plyr)
+library(ggplot2)
+
+all_data <- data.frame(V1=integer(), V2=integer(),
+                       V3=integer(), V4=integer(),
+                       V5=integer(),
+                       middlebox=character())
+
+netfilter <- read.table("netfilter-mf-find-mg-1p.txt")
+netfilter["middlebox"] <- "NetFilter"
+vignat <- read.table("vignat-mf-find-mg-1p.txt")
+vignat["middlebox"] <- "VigNAT"
+counted_chains <- read.table("counted-chains-mf-find-mg-1p.txt")
+counted_chains ["middlebox"] <- "counted-chains"
+nop <- read.table("nop-mf-find-mg-1p.txt")
+nop["middlebox"] <- "NOP"
+unverified <- read.table("unverified-mf-find-mg-1p.txt")
+unverified["middlebox"] <- "DPDK-unverified"
+
+all_data <- rbind(all_data, netfilter)
+all_data <- rbind(all_data, counted_chains)
+all_data <- rbind(all_data, vignat)
+all_data <- rbind(all_data, nop)
+all_data <- rbind(all_data, unverified)
+
+cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+p <- ggplot(all_data, aes(x=V1/1000, y=V4/1e6,
+                          group=middlebox,
+                          color=middlebox,
+                          shape=middlebox)) +
+     geom_line(size=1) +
+     labs(title="Loopback throughput, 1% packets lost") +
+     xlab("# concurrent flows (K)") +
+     ylab(bquote("Throughput Mpkt/s")) +
+     theme_bw() +
+     expand_limits(x=0,y=0)
+
+ggsave(filename="thru.png")
+print(p)

### 2017-06-26 nf/testbed/hard/util/relieve-connection-reuse.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+. ~/scripts/util/functions.sh
+
+# Sacrifices some NetFilter fidelity to reuse connections faster
+
+sudo_overwrite /proc/sys/net/ipv4/tcp_fin_timeout 5
+sudo_overwrite /proc/sys/net/ipv4/tcp_tw_reuse 1
+sudo_overwrite /proc/sys/net/ipv4/tcp_tw_recycle 1

### 2017-06-26 nf/testbed/hpingtst.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+rm -f result.txt
+
+serv=192.168.33.10
+times=100
+
+for interval in 20 30 40 45 50 55 60 70 100 150 300; do
+  for i in $(seq 1 $times); do
+    sleep 2
+    rrr=$(sudo hping3 $serv -i u$interval -k -s 3453 -n -p 80 -q -c 10000 2>&1 | tail -n3 | head -n1 | awk '{print $7;}' | sed -e 's/%//')
+    rez[$i]=$rrr
+    echo "$interval $rrr" >> result.txt
+  done
+  echo "for $interval : ${rez[@]}"
+done
+

### 2017-06-26 nf/testbed/nat.sh

@@ -1,0 +1,19 @@
+sudo apt-get update
+sudo apt-get install -y tcpdump git wget build-essential libpcap-dev linux-headers-3.13.0-86 linux-headers-3.13.0-86-generic libglib2.0-dev daemon
+
+wget http://dpdk.org/browse/dpdk/snapshot/dpdk-2.2.0.tar.gz -O dpdk.tar.gz
+tar xf dpdk.tar.gz
+mv dpdk-* dpdk
+rm dpdk.tar.gz
+
+cd dpdk
+sed -ri 's,(PMD_PCAP=).*,\1y,' config/common_linuxapp
+make config T=x86_64-native-linuxapp-gcc
+make install -j2 T=x86_64-native-linuxapp-gcc DESTDIR=.
+
+export RTE_SDK=/home/vagrant/dpdk
+export RTE_TARGET=x86_64-native-linuxapp-gcc
+
+. /nat/testbed/redeploy-nat.sh
+
+daemon -r /nat/testbed/run-nat.sh

### 2017-06-26 nf/testbed/perf/inserver.sh

@@ -1,0 +1,3 @@
+sudo apt-get update && sudo apt-get install -y tcpdump hping3 apache2
+sudo arp -s 192.168.33.2 08:00:27:00:44:71
+sudo ip route add 192.168.40.0/24 via 192.168.33.2 dev eth1

### 2017-06-26 nf/testbed/perf/simple_router.sh

@@ -1,0 +1,2 @@
+sysctl -w net.ipv4.ip_forward=1
+echo "net.ipv4.ip_forward=1" >> /etc/sysctl.conf

### 2017-06-26 nf/testbed/pktgen.sh

@@ -1,0 +1,40 @@
+apt-get update
+apt-get install -y tcpdump hping3 python-scapy wget linux-headers-3.13.0-86 linux-headers-3.13.0-86-generic libpcap-dev git
+ip route add 192.168.33.0/24 via 192.168.40.2 dev eth1
+arp -s 192.168.40.2 08:00:27:00:44:72
+#hping3 192.168.33.10 -k -p 47882 -s 0
+
+mkdir -p /mnt/huge
+chmod 777 /mnt/huge
+echo "vm.nr_hugepages=256" >> /etc/sysctl.conf
+echo "huge /mnt/huge hugetlbfs defaults 0 0" >> /etc/fstab
+# reload sysctl config without reboot
+sysctl --system
+mount huge
+# the above setting nr_hugepages does not work for some reason
+echo 256 > /sys/kernel/mm/hugepages/hugepages-2048kB/nr_hugepages
+
+cd ~vagrant
+git clone --depth 1 --branch v16.07-rc1 http://dpdk.org/git/dpdk
+git clone --depth 1 --branch pktgen-3.0.02 http://dpdk.org/git/apps/pktgen-dpdk
+
+export RTE_SDK=~vagrant/dpdk
+export RTE_TARGET=x86_64-native-linuxapp-gcc
+
+cd $RTE_SDK
+make install T=$RTE_TARGET
+
+cd ~vagrant/pktgen-dpdk
+make
+
+echo "export RTE_SDK=$RTE_SDK" >> ~vagrant/.bashrc
+echo "export RTE_TARGET=$RTE_TARGET" >> ~vagrant/.bashrc
+echo "sudo /sbin/modprobe uio" >> ~vagrant/.bashrc
+echo "sudo $PWD/setup.sh" >> ~vagrant/.bashrc
+
+. setup.sh
+/sbin/modprobe uio
+ifconfig eth1 down
+${RTE_SDK}/tools/dpdk_nic_bind.py -b igb_uio 0000:00:08.0
+
+# runstring: sudo ./app/app/x86_64-native-linuxapp-gcc/pktgen -c 0x1 -n 2 -- -P -m "0.0"

### 2017-06-26 nf/testbed/run-nat.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+EXPTIME=$1
+MAX_FLOWS=$2
+
+if [ -z $EXPTIME ]; then
+    EXPTIME=10
+fi
+
+if [ -z $MAX_FLOWS ]; then
+    MAX_FLOWS=1024
+fi
+
+sudo /nat/build/nat -c 0x01 -n 2 -- -p 0x3 --wan 0 --expire $EXPTIME --max-flows $MAX_FLOWS --starting-port 1025  --extip 192.168.33.2 --eth-dest 0,08:00:27:53:8b:38 --eth-dest 1,08:00:27:c1:13:47
+

### 2017-06-26 nf/testbed/server.sh

@@ -1,0 +1,3 @@
+sudo apt-get update && sudo apt-get install -y tcpdump hping3 apache2
+sudo arp -s 192.168.33.2 08:00:27:00:44:71
+

### 2017-06-26 nf/testbed/start-nat.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+# Start the Vigor NAT in the background (as a daemon).
+
+. ./hard/config.sh
+
+. ./redeploy-nat.sh
+daemon ./run-nat.sh $PWD

### 2017-06-26 nf/vignat/flowmanager.h

@@ -1,0 +1,23 @@
+#ifndef _FLOWMANAGER_H_INCLUDED_
+#define _FLOWMANAGER_H_INCLUDED_
+
+#include "lib/flow.h"
+
+int allocate_flowmanager(uint8_t nb_ports,
+                         uint16_t starting_port, uint32_t ext_src_ip,
+                         uint8_t ext_device_id,
+                         uint32_t expiration_time,
+                         int max_flows);
+
+int allocate_flow(struct int_key *ik, uint32_t time, struct flow* out);
+int expire_flows(uint32_t time);
+
+int get_flow_by_int_key(struct int_key* key, uint32_t time, struct flow* flow_out);
+int get_flow_by_ext_key(struct ext_key* key, uint32_t time, struct flow* flow_out);
+
+#ifdef KLEE_VERIFICATION
+struct DoubleChain** get_dchain_pp(void);
+struct DoubleMap **get_dmap_pp(void);
+#endif //KLEE_VERIFICATION
+
+#endif //_FLOWMANAGER_H_INCLUDED_

### 2017-06-26 nf/vignat/loop-stub.c

@@ -1,0 +1,98 @@
+#include <klee/klee.h>
+#include "loop.h"
+#include "lib/stubs/my-time-stub-control.h"
+#include "lib/stubs/containers/double-chain-stub-control.h"
+#include "lib/stubs/containers/double-map-stub-control.h"
+#include "lib/stubs/rte-stubs-control.h"
+
+//TODO: this is a copy-paste from array_lcc.h and other array headers
+//. Move it to a separate file.
+
+void loop_iteration_assumptions(struct DoubleMap** m, struct DoubleChain** ch,
+                                unsigned int lcore_id,
+                                uint32_t time, int max_flows, int start_port)
+{
+  rte_reset();
+  dchain_reset(*ch, max_flows);
+  dmap_reset(*m, max_flows);
+}
+
+void loop_iteration_assertions(struct DoubleMap** m, struct DoubleChain** ch,
+                               unsigned int lcore_id,
+                               uint32_t time, int max_flows, int start_port)
+{
+
+}
+
+void loop_invariant_consume(struct DoubleMap** m, struct DoubleChain** ch,
+                            unsigned int lcore_id,
+                            uint32_t time, int max_flows, int start_port)
+{
+  klee_trace_ret();
+  klee_trace_param_ptr(m, sizeof(struct DoubleMap*), "m");
+  klee_trace_param_ptr(ch, sizeof(struct DoubleChain*), "ch");
+  klee_trace_param_i32(lcore_id, "lcore_id");
+  //ARRAY_LCC_EL_TRACE_ARG_BREAKDOWN(cur_lcc);
+  klee_trace_param_i32(time, "time");
+  klee_trace_param_i32(max_flows, "max_flows");
+  klee_trace_param_i32(start_port, "start_port");
+}
+
+void loop_invariant_produce(struct DoubleMap** m, struct DoubleChain** ch,
+                            unsigned int* lcore_id,
+                            uint32_t *time, int max_flows, int start_port)
+{
+  klee_trace_ret();
+  klee_trace_param_ptr(m, sizeof(struct DoubleMap*), "m");
+  klee_trace_param_ptr(ch, sizeof(struct DoubleChain*), "ch");
+  klee_trace_param_ptr(lcore_id, sizeof(unsigned int), "lcore_id");
+  //ARRAY_LCC_EL_TRACE_ARG_BREAKDOWN(cur_lcc);
[…]

### 2017-06-26 nf/vignat/nat_forward_vignat.c

@@ -1,0 +1,193 @@
+#include <inttypes.h>
+
+#ifdef KLEE_VERIFICATION
+#  include <klee/klee.h>
+#  include "lib/stubs/rte_stubs.h"
+#  include "loop.h"
+#else
+// DPDK requires these but doesn't include them. :|
+#  include <linux/limits.h>
+#  include <sys/types.h>
+#  include <unistd.h>
+
+#  include <rte_ethdev.h>
+#  include <rte_ether.h>
+#  include <rte_ip.h>
+#  include <rte_mbuf.h>
+#endif
+
+#include "lib/flow.h"
+#include "flowmanager.h"
+
+#include "lib/nat_config.h"
+#include "lib/nf_forward.h"
+#include "lib/nf_log.h"
+#include "lib/nf_util.h"
+
+struct nat_config config;
+
+void nf_core_init()
+{
+	if (!allocate_flowmanager(rte_eth_dev_count(),
+                            config.start_port,
+                            config.external_addr,
+                            config.wan_device,
+                            config.expiration_time,
+                            config.max_flows)) {
+		rte_exit(EXIT_FAILURE, "Could not allocate flow manager");
+	}
+}
+
+int nf_core_process(uint8_t device,
+                    struct rte_mbuf* mbuf,
+                    uint32_t now)
+{
+	NF_DEBUG("It is %" PRIu32, now);
+
+	expire_flows(now);
+	NF_DEBUG("Flows have been expired");
+
+	struct ether_hdr* ether_header = nf_get_mbuf_ether_header(mbuf);
+
+	struct ipv4_hdr* ipv4_header = nf_get_mbuf_ipv4_header(mbuf);
+	if (ipv4_header == NULL) {
+		NF_DEBUG("Not IPv4, dropping");
+		return device;
+	}
+
+	struct tcpudp_hdr* tcpudp_header = nf_get_ipv4_tcpudp_header(ipv4_header);
+	if (tcpudp_header == NULL) {
+		NF_DEBUG("Not TCP/UDP, dropping");
+		return device;
+	}
+
+	NF_DEBUG("Forwarding an IPv4 packet on device %" PRIu8, device);
+
+	uint8_t dst_device;
+  int allocated = 0;
+	if (device == config.wan_device) {
+		NF_DEBUG("Device %" PRIu8 " is external", device);
+
+		struct ext_key key = {
+			.ext_src_port = tcpudp_header->dst_port, //intentionally swapped.
+			.dst_port = tcpudp_header->src_port,
[…]
